---
title: "Zapped out of existence! Home-field advantage decomposition patterns persist after removal of leaf-associated microbes"
author: "CJ Spiegel et al. 2023, in prep"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
  editor_options:
      chunk_output_type: console
---

### Project overview
For Project ‘RAD’ Alder, gamma irradiated (phyllosphere present (P)) vs. non-irradiated (phyllosphere absent (AB)) leaf packs deployed at “home” or “away” sites tests the role of phyllosphere microbial communities on home-field advantage patterns in leaf litter decomposition. We utilized a reciprocal transplant experiment between the Hoko and Sekiu rivers on the Olympic Peninsula under four treatment conditions: phyllosphere present-home (PH), phyllosphere present-away (PAW), phyllosphere absent-home (ABH), phyllosphere absent-away (ABAW) to ask the following questions: 

1) How does dispersal from the regional species pool (i.e., source communities) influence community composition on the leaf packs?
2) How does microbial community composition (alpha and beta diversity) and function differ between the leaf packs, as well as change over the course of decomposition?
3) How does alder stoichiometric and secondary chemical composition differ between the leaf packs, as well as change over the course of decomposition?
4) How does microbial metabolization differ between the leaf packs, as well as change over the course of decomposition?
5) How does particulate leaf mass loss differ between the leaf packs? 


**DATA SETS**  
This data set focuses on:  

(1) **Metabolism assay** 
(2) **Leaf mass loss** using % mass loss and HFAI
(3) **Alder C and N isotopic composition**
(4) **Alder secondary chemistry (via LC-MS/MS)** 
```{r setup chunk, setup, include = FALSE, cache=FALSE, message=FALSE, warning=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

# use pacman to load all the packages you are missing!
pacman::p_load('knitr', 'lme4', 'lmerTest', 'tidyverse', 'magrittr', 'effects', 'plyr', 'dplyr', 
               'plotrix', 'car',"gridExtra", "cowplot", "tools", "mgcv", "gratia", "tidymv", "ggthemes", "ggpubr", "scales", "visreg", "mgcViz")

Fig.formatting<-(theme_classic()) +
  theme(text=element_text(size=25),
        legend.text.align = 0,
        panel.border = element_rect(fill=NA, colour = "black", size=1),
        aspect.ratio=1, 
        axis.ticks.length=unit(0.25, "cm"),
        axis.text.y=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20), 
        axis.text.x=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20)) +
  theme(legend.key.size = unit(0.4, "cm"), legend.position = "right") +
  theme(aspect.ratio=0.5) 




```

**Note on absorbance**
Iodonitrotetrazolium chloride (INT) is utilized as a redox dye to gauge microbial activity in assays. Microbes with metabolic activity can reduce INT to a formazan compound, causing a color change from yellow (oxidized form) to red or purple (reduced form). This reduction reaction is often catalyzed by dehydrogenase enzymes present in metabolically active cells. The formazan product is  insoluble and has a distinct color (i.e., absorbance), leading to a visible change in the solution's color. The intensity of the color or absorbance is indicative of microbial metabolic activity.

We measured absorbance at 450 nm (see Blenkinsopp & Lock, 1990; Friedel et al., 1994; Jackson et al., 2013; Bhupathiraju et al., 1998). Time course measurements, or absorbance readings at different intervals (e.g., 30 mins, 2 hrs, 16 hrs, 24 hrs, 48 hrs, and 72 hrs), allowed for tracking changes in microbial metabolic activity over time.

### Import metabolism assay data
```{r}
#add data files
meta.day1.450.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_1_450_clean.csv")
meta.day1.590.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_1_590_clean.csv")


meta.day3.450.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_3_450_clean.csv")
meta.day3.590.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_3_590_clean.csv")


meta.day5.450.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_5_450_clean.csv")
meta.day5.590.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_5_590_clean.csv")

meta.day7.450.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_7_450_clean.csv")
meta.day7.590.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_7_590_clean.csv")

meta.day11.450.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_11_450_clean.csv")
meta.day11.590.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_11_590_clean.csv")

meta.day17.450.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_17_450_clean.csv")
meta.day17.590.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_17_590_clean.csv")


#merge 


meta.450.df= rbind(
  meta.day1.450.df, 
  meta.day3.450.df,
  meta.day5.450.df,
  meta.day7.450.df, 
  meta.day11.450.df, 
  meta.day17.450.df
  )

meta.590.df=rbind(
  meta.day1.590.df,
  meta.day3.590.df,
  meta.day5.590.df,
  meta.day7.590.df,
  meta.day11.590.df, 
  meta.day17.590.df
)

#remove well_id columns
meta.450.df <- meta.450.df[, !colnames(meta.450.df) %in% "well_id"]
meta.590.df <- meta.590.df[, !colnames(meta.590.df) %in% "well_id"]


#create groups based on deployment and metabolite pair

#if deploy is HOK, metabolite is HOK (home)
#if deplpy is SEK, metaboltie is SEK (home)
#if deploy is HOK, metaboltie is SEK (away)
#if deploy is SEK, metabolite is HOK (away)

meta.450.df$group <- 
  ifelse(
  meta.450.df$deploy == "HOK" & meta.450.df$metabolite == "HOK", "H",
   ifelse(
  meta.450.df$deploy == "SEK" & meta.450.df$metabolite == "SEK", "H",
   ifelse(
   meta.450.df$deploy == "HOK" & meta.450.df$metabolite == "SEK", "AW",
   ifelse(
   meta.450.df$deploy == "SEK" & meta.450.df$metabolite == "HOK", "AW", NA  # Default value if none of the conditions are met
                )
              )
            )
          )

meta.590.df$group <- 
  ifelse(
  meta.590.df$deploy == "HOK" & meta.590.df$metabolite == "HOK", "H",
   ifelse(
  meta.590.df$deploy == "SEK" & meta.590.df$metabolite == "SEK", "H",
   ifelse(
   meta.590.df$deploy == "HOK" & meta.590.df$metabolite == "SEK", "AW",
   ifelse(
   meta.590.df$deploy == "SEK" & meta.590.df$metabolite == "HOK", "AW", NA  # Default value if none of the conditions are met
                )
              )
            )
          )


#add hours rather than minutes
meta.450.df$hours <- 
  ifelse(
  meta.450.df$time == "30", "0.5",
   ifelse(
  meta.450.df$time == "120", "2",
   ifelse(
   meta.450.df$time == "960", "16",
   ifelse(
   meta.450.df$time == "1440", "24",
    ifelse(
   meta.450.df$time == "2880", "48",
   ifelse(
   meta.450.df$time == "4320", "72", NA  # Default value if none of the conditions are met
                )
              )
            )
          )
)
)

meta.590.df$hours <- 
  ifelse(
  meta.590.df$time == "30", "0.5",
   ifelse(
  meta.590.df$time == "120", "2",
   ifelse(
   meta.590.df$time == "960", "16",
   ifelse(
   meta.590.df$time == "1440", "24",
    ifelse(
   meta.590.df$time == "2880", "48",
   ifelse(
   meta.590.df$time == "4320", "72", NA  # Default value if none of the conditions are met
                )
              )
            )
          )
)
)

#reorder hours
meta.450.df <- meta.450.df %>% mutate(hours = factor(hours, levels = c("0.5", "2", "16", "24", "48", "72")))
meta.590.df <- meta.590.df %>% mutate(hours = factor(hours, levels = c("0.5", "2", "16", "24", "48", "72")))

#make colummn factors
meta.450.df$treatment=as.factor(meta.450.df$treatment)
meta.450.df$tree_id=as.factor(meta.450.df$tree_id)
meta.450.df$deploy=as.factor(meta.450.df$deploy)
meta.450.df$metabolite=as.factor(meta.450.df$metabolite)
meta.450.df$group=as.factor(meta.450.df$group)
meta.450.df$hours=factor(meta.450.df$hours)

meta.590.df$treatment=as.factor(meta.590.df$treatment)
meta.590.df$tree_id=as.factor(meta.590.df$tree_id)
meta.590.df$deploy=as.factor(meta.590.df$deploy)
meta.590.df$metabolite=as.factor(meta.590.df$metabolite)
meta.590.df$group=as.factor(meta.590.df$group)
meta.590.df$hours=factor(meta.590.df$hours)

#make pairing for 4 treatments
meta.450.df <- meta.450.df %>%
  mutate(pairing = paste(treatment, group, sep = "-"))

meta.450.df <- meta.450.df %>% 
  mutate(pairing = factor(pairing, levels = c("P-AW", "P-H", "AB-AW", "AB-H")))

#create df
meta.450.df <- meta.450.df %>%
  mutate(
    new.trt = case_when(
      treatment == "P" ~ "LAM+FLM",
      treatment == "AB" ~ "FLM-only"))

#subset by true home 
deploy.hok_only.df=meta.450.df[meta.450.df$deploy=="HOK" & meta.450.df$tree_id=="HOK",]
deploy.sek_only.df=meta.450.df[meta.450.df$deploy=="SEK" & meta.450.df$tree_id=="SEK",]
#merge
true_home.meta.df=rbind(deploy.hok_only.df, deploy.sek_only.df)

#subset by false home 
deploy.hok.df=meta.450.df[meta.450.df$deploy=="HOK" & meta.450.df$tree_id=="SEK",]
deploy.sek.df=meta.450.df[meta.450.df$deploy=="SEK" & meta.450.df$tree_id=="HOK",]
#merge
false_home.meta.df=rbind(deploy.hok.df, deploy.sek.df)


#create subsets
LAM_FL.df=true_home.meta.df%>%
  filter(new.trt=="LAM+FL")%>%
  select(tree_id, deploy, metabolite,tree_num, site_num, group, decomp_day, od_corrected_b_n, new.trt)%>%
  arrange(decomp_day)

FL_only.df=true_home.meta.df%>%
   filter(new.trt=="FLM-only")%>%
  select(tree_id, deploy, metabolite,tree_num, site_num, group, decomp_day, od_corrected_b_n, new.trt)%>%
arrange(decomp_day)
```

#### determine which absorbance wavelength to use. Various studies (as those cited above) measured the absorbance of INT at both 450 and 590 nm. This provides some flexibility for us to determine exactly which to use based on the conditions present in our system (i.e., rivers, within the wells, external conditions of the lab space, etc). Factors such as these can influence the rate of metabolization, and thus the OD values from the absorbance reader. Additionally, determining which reading time (30 mins, 2 hrs, 16 hrs, etc.) at which the OD values were measured is important. Past work has measured absorbance at 30 mins, while other recommend 2-16 hrs. However, given the specific conditions of our field experiment, it has been suggested that 24 hrs may be required for an accurate metabolic reading. Additionally, we need to include a hierarchical modeling framework to capture the nonlinear trends in the data (a priori parabolic, or unimodal, metabolization curves across decomposition day) and repeated measures for each leaf pack. 

Let's first visualize the raw OD readings at both 450 and 590 nm.
```{r}

#plot at 450 nm
meta.450.df %>%
    filter(time == "1440")%>%
  na.omit()%>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n)) +
  geom_point()+
  labs(x = "decomposition day", y = "standardized optical density (z score)")+
  scale_color_colorblind() +
  scale_fill_colorblind() +
   guides(linetype = guide_legend(override.aes = list(color = "cornflowerblue")))+
  theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )


# plot at 590 nm
meta.df %>%
    filter(nm == "590", time == "1440")%>%
  na.omit()%>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n)) +
  geom_smooth(aes(group = interaction(group, treatment), linetype = treatment, color = group, fill = group), method = 'gam', formula = y ~ s(x, k = 6), alpha = 0.1) +
  labs(x = "decomposition day", y = "standardized optical density (z score)")+
  scale_color_colorblind() +
  scale_fill_colorblind() +
   guides(linetype = guide_legend(override.aes = list(color = "cornflowerblue")))+
  theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )

```

Based on the literature, we will use an absorbance wavelength of 450 nm which is appropriate for the metabolic indicator (INT) in the assay. 
### Use 450 nm absorbance readings
```{r}
#by group and treatment using loess
meta.450.df %>%
  filter(hours == "24")%>%
  na.omit()%>%
  ggplot(aes(x = as.numeric(decomp_day), y = od_corrected_b_n)) +
  geom_smooth(aes(color = treatment, fill = treatment, linetype = group, group = interaction(group, treatment)), method = 'loess', formula = y ~ x, alpha = 0.1) +
    geom_line(aes(y = neg), color = "red", linetype = "dashed", size = 0.5, alpha = 0.5) +
    geom_ribbon(aes(ymin = neg - sd(neg), ymax = neg + sd(neg)), fill = "red", alpha = 0.2) +
  labs(x = "decomposition day", y = "optical density (450 nm)")+
  scale_x_continuous(breaks = c(1, 5, 9, 13, 17))+
  scale_color_colorblind() +
  scale_fill_colorblind() +
guides(
    linetype = guide_legend(override.aes = list(color = "cornflowerblue")),
    color = guide_legend(order = 1),  # Set order for the color legend
    fill = guide_legend(order = 1)   # Set order for the fill legend
  ) +
  theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )

#by group and treatment using gam
meta.450.df %>%
  filter(hours == "16")%>%
  na.omit()%>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n)) +
  geom_smooth(aes(color = treatment, fill = treatment, linetype = group, group = interaction(group, treatment)), method = 'gam', formula = y ~ s(x, k =4), alpha = 0.1) +
    geom_smooth(aes(y = neg), method = "gam", color = "red", fill= "red", linetype = "dashed", size = 0.5, alpha = 0.5) +
  labs(x = "decomposition day", y = "optical density (450 nm)")+
  scale_x_continuous(breaks = c(1, 5, 9, 13, 17))+
  scale_color_colorblind() +
  scale_fill_colorblind() +
guides(
    linetype = guide_legend(override.aes = list(color = "cornflowerblue")),
    color = guide_legend(order = 1),  # Set order for the color legend
    fill = guide_legend(order = 1)   # Set order for the fill legend
  ) +
  theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )

meta.450.df %>%
  filter(hours == "24")%>%
  na.omit()%>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n)) +
  geom_point(aes(y = od_corrected_b_n, group = interaction (group, treatment, tree_num)))+
  geom_smooth(aes(color = treatment, fill = treatment, linetype = group, group = interaction(group, treatment)), method = 'gam', formula = y ~ s(x, k=4), alpha = 0.1) +
      geom_smooth(aes(y = neg), method = "gam", formula = y ~ s(x, k=4), color = alpha("red", 0.5), fill= alpha("red", 0.5), linetype = "longdash", size = 0.7) +
  labs(x = "decomposition day", y = "optical density (450 nm)")+
scale_color_colorblind() +
  scale_fill_colorblind() +
guides(
    linetype = guide_legend(override.aes = list(color = "cornflowerblue")),
    color = guide_legend(order = 1),  # Set order for the color legend
    fill = guide_legend(order = 1)   # Set order for the fill legend
  ) +
  theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )


#####################

meta.450.df %>%
  filter(hours == "24") %>%
  na.omit() %>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n, group = interaction(group, treatment), color = treatment, fill = treatment)) +
  stat_summary(fun = mean, geom = "point", aes(shape = treatment), show.legend = F, size = 2.3) +
  geom_smooth(aes(color = treatment,fill = treatment, linetype=group), method = 'gam', formula = y ~ s(x, k = 6), alpha = 0.15) +
  geom_point(aes(y= neg), color = alpha("red", 0.3), shape = 15)+
  geom_line(aes(y = neg), color = alpha("red", 0.3), linetype = "longdash", size = 0.7) +
  labs(x = "decomposition day", y = "optical density (450 nm)") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  guides(
    linetype = guide_legend(override.aes = list(color = "cornflowerblue")),
    color = guide_legend(order = 1),  # Set order for the color legend
    fill = guide_legend(order = 1)   # Set order for the fill legend
  ) +
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    strip.text = element_text(size = 16)
  )


```


These are the final plots
```{r}
#####################
meta.450.df %>%
  filter(hours == "24") %>%
  na.omit() %>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n, group = interaction(group, treatment), color = treatment, fill = treatment)) +
  stat_summary(fun = mean, geom = "point", aes(shape = treatment), show.legend = T, size = 2.3) +
  geom_smooth(aes(color = treatment,fill = treatment, linetype=group), method = 'gam', formula = y ~ s(x, k = 6), alpha = 0.15) +
  geom_point(aes(y= neg,  shape = "neg"), color = alpha("red", 0.3), show.legend = F)+
  geom_line(aes(y = neg), color = alpha("red", 0.3), show.legend = F, linetype = "longdash", size = 0.7) +
  labs(x = "decomposition day", y = "optical density (450 nm)") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  guides(
    linetype = guide_legend(override.aes = list(color = "cornflowerblue")),
    color = guide_legend(order = 1),  # Set order for the color legend
    fill = guide_legend(order = 1)   # Set order for the fill legend
  ) +
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    strip.text = element_text(size = 16)
  )+
  Fig.formatting

assay.plot

ggsave("figures/assay.plot.pdf", assay.plot, width = 17, height = 8)


#4 treatment plot
trt.assay.plot=meta.450.df %>%
  filter(hours == "24") %>%
  na.omit() %>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n, color = pairing, group = pairing)) +
  stat_summary(fun = mean, geom = "point", show.legend = F, size = 2.3) +
    geom_point(aes(y= neg,  shape = "neg"), size= 2.3, color = alpha("red", 0.3), show.legend = F)+
  geom_smooth(aes(linetype= pairing, fill = pairing),method = 'gam', formula = y ~ s(x, k = 6), alpha = 0.15) +
  geom_smooth(aes(y = neg), fill = alpha ("red", 0.3), method = 'loess', formula = y ~ x, span= 0.6, color = alpha("red", 0.3), show.legend = F, linetype = "longdash") +
  labs(x = "decomposition day", y = "absorbance (OD450 nm)") +
  coord_cartesian(ylim = c(0,3))+
  scale_color_colorblind()+
  scale_fill_colorblind()+
    scale_x_discrete(expand = c(0.01, 0.01)) +  # Adjust the expand argument
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    strip.text = element_text(size = 16)
  )+
  Fig.formatting

trt.assay.plot

ggsave("figures/trt.assay.plot.pdf", trt.assay.plot, width = 15, height = 8)

meta.450.df %>%
  filter(hours == "24") %>%
  na.omit() %>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n, color = treatment, fill = treatment, group = interaction(group, treatment))) +
  stat_summary(fun = mean, geom = "point", show.legend = F, size = 2.3) +
    stat_summary(fun = mean, geom = "line", show.legend = F, size = 1) +

    geom_point(aes(y= neg,  shape = "neg"), size= 2.3, color = alpha("red", 0.3), show.legend = F)+
  geom_smooth(aes(y = neg), fill = alpha ("red", 0.3), method = 'loess', formula = y ~ x, span= 0.6, color = alpha("red", 0.3), show.legend = F, linetype = "longdash") +
  labs(x = "decomposition day", y = "absorbance (OD450 nm)") +
  coord_cartesian(ylim = c(0,3))+
  scale_color_colorblind()+
  scale_fill_colorblind()+
    scale_x_discrete(expand = c(0.01, 0.01)) +  # Adjust the expand argument
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    strip.text = element_text(size = 16)
  )+
  Fig.formatting


meta.450.df %>%
  filter(hours == "24") %>%
  na.omit() %>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n, color = pairing, group = pairing)) +
  stat_summary(fun = mean, geom = "point", show.legend = F, size = 2.3) +
    geom_point(aes(y= neg,  shape = "neg"), size= 2.3, color = alpha("red", 0.3), show.legend = F)+
  stat_summary(fun = mean, geom = "line", show.legend = F, size = 0.75) +
  geom_smooth(aes(y = neg), fill = alpha ("red", 0.3), method = 'loess', formula = y ~ x, span= 0.6, color = alpha("red", 0.3), show.legend = F, linetype = "longdash") +
  labs(x = "decomposition day", y = "absorbance (OD450 nm)") +
  coord_cartesian(ylim = c(0,3))+
  scale_color_colorblind()+
  scale_fill_colorblind()+
    scale_x_discrete(expand = c(0.01, 0.01)) +  # Adjust the expand argument
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    strip.text = element_text(size = 16)
  )+
  Fig.formatting


meta.450.df %>%
  filter(hours == "24") %>%
  na.omit() %>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n, color = pairing, group = pairing)) +
  stat_summary(fun = mean, geom = "point", show.legend = FALSE, size = 2.3) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.1, show.legend = FALSE) +  # Add error bars
  geom_point(aes(y = neg, shape = "neg"), size = 2.3, color = alpha("red", 0.3), show.legend = FALSE) +
  stat_summary(fun = mean, geom = "line", show.legend = FALSE, size = 0.75) +
  geom_smooth(aes(y = neg), fill = alpha("red", 0.3), method = 'loess', formula = y ~ x, span = 0.6, color = alpha("red", 0.3), show.legend = FALSE, linetype = "longdash") +
  labs(x = "decomposition day", y = "absorbance (OD450 nm)") +
  coord_cartesian(ylim = c(0, 3)) +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  scale_x_discrete(expand = c(0.01, 0.01)) +
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    strip.text = element_text(size = 16)
  ) +
  Fig.formatting


```


```{r}
#This pattern suggests that microbial communities from the intermediate stages of decomposition may be particularly well-equipped or adapted to metabolize the plant metabolites. 

#The hypothesis about a unimodal trend in abundance and diversity, where early and low decomposition days show low abundance and diversity, and intermediate decomposition days show high abundance and diversity, is a reasonable expectation. It aligns with the concept of priority effects, where early colonizers influence substrate conditions and, in some cases, facilitate the establishment of subsequent colonizers.

#However, the findings from the metabolization assay contradict this hypothesis. The observation that early-stage days exhibit lower metabolization despite the potential presence of key enzymes in those microbes challenges the initial hypothesis.

#This discrepancy suggests that the relationship between microbial community composition, their functional attributes (such as enzyme production), and the metabolization of plant extracts is more complex than initially anticipated. It's possible that other factors, such as the specific secondary metabolites present in the plant extracts, microbial interactions, or ecological conditions, influence the metabolic activity of the communities.


#by 4 treatments
meta.450.df %>%
  filter(hours == "24")%>%
  na.omit()%>%
  ggplot(aes(x = decomp_day, y = od_corrected_b_n)) +
  geom_smooth(aes(linetype= pairing, color = pairing, fill = pairing), method = 'gam', formula = y ~ s(x, k = 4), alpha = 0.1) +
  labs(x = "decomposition day", y = "optical density (450 nm)")+
    guides(color = guide_legend(title = "Treatment"), fill= guide_legend(title = "Treatment"), linetype = guide_legend(title = "Treatment"))+
   scale_linetype_manual(name = "Treatment",
                        values = c("P-AW" = "solid", "P-H" = "solid", "AB-AW" = "dotdash", "AB-H" = "dotdash")) +
    guides(color = guide_legend(title = "Treatment"), fill = guide_legend(title = "Treatment"), linetype = guide_legend(title = "Treatment"))+
  scale_x_continuous(breaks = c(1, 5, 9, 13, 17))+
  theme_bw() +
  scale_color_colorblind() +
  scale_fill_colorblind() +
theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )


#HFAI shows P packs to have greater incidence of HFA, while AB show greater magnitude
meta.450.df %>%
   filter(hours == "24", group == "H")%>%
  na.omit()%>%
  ggplot(aes(x = as.numeric(decomp_day), y = od_corrected_b_n)) +
  geom_smooth(aes( color = treatment, fill = treatment), method = 'loess', formula = y ~x, alpha = 0.1) +
  labs(x = "decomposition day", y = "optical density (450 nm)")+
    scale_x_continuous(breaks = c(1, 5, 9, 13, 17))+
  scale_color_colorblind() +
  scale_fill_colorblind() +
   guides(linetype = guide_legend(override.aes = list(color = "cornflowerblue")))+
  theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )

#In the context of the Homefield Advantage Index (HFAI), "incidence" and "magnitude" can be understood as follows:

#Incidence of HFAI:

#Definition: Incidence refers to the frequency or occurrence of HFAI events. In the Context of HFAI:  the incidence of HFAI is more in the "present" treatment, it means that there are more instances or occurrences where the HFAI is observed when the phyllosphere community is present. This could be interpreted as a higher frequency of home advantage events in terms of decomposition when the community is present.

#Magnitude of HFAI: Definition: Magnitude refers to the size, extent, or intensity of the HFAI events.  the magnitude is greater in the "absent" treatment, it suggests that, although the frequency might be lower, when HFAI events occur in the absence of the phyllosphere community, they have a larger impact or are more pronounced in terms of decomposition. The decomposition advantage at home (Hoko) is more significant when the community is absent, leading to a higher magnitude of the HFAI.

#In summary, the "incidence" reflects how often HFAI events occur, while the "magnitude" indicates the strength or impact of those events when they do occur. Both aspects are important for understanding the dynamics of the Homefield Advantage Index in your experimental conditions.

#Microbial communities metabolizing their home metabolite exhibit higher activity in the presence of the phyllosphere community. This is consistent with the HFai framework, where the relative advantage of home metabolite metabolization is observed in the ecological context of the phyllosphere community being present.



```


#### z score optical density
##### Day 1
```{r}
##### Day 1
#HOK 1 subset
d1.hok1_meta.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" &  meta.450.df$decomp_day == "1"& meta.450.df$hours == "24"& meta.450.df$site_num == "1", ])
# Calculate z-scores 
d1.hok1_meta.df=d1.hok1_meta.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))


#HOK 2 subset
d1.hok2_meta.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" &  meta.450.df$decomp_day == "1"& meta.450.df$hours == "24"& meta.450.df$site_num == "2", ])
# Calculate z-scores 
d1.hok2_meta.df=d1.hok2_meta.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))

#SEK 1 subset
d1.sek1_meta.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" &  meta.450.df$decomp_day == "1"& meta.450.df$hours == "24"& meta.450.df$site_num == "1", ])
# Calculate z-scores 
d1.sek1_meta.df=d1.sek1_meta.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))

#SEK 2 subset
d1.sek2_meta.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" &  meta.450.df$decomp_day == "1"& meta.450.df$hours == "24"& meta.450.df$site_num == "2", ])
# Calculate z-scores 
d1.sek2_meta.df=d1.sek2_meta.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))

#merge z scored dfs
d1.z_scored_df = rbind(d1.hok1_meta.df, d1.hok2_meta.df, d1.sek1_meta.df, d1.sek2_meta.df)

#anovas
aov_d1=aov(z_score ~ pairing, data= d1.z_scored_df )

aov_d1_od=aov(od_corrected_b_n ~ pairing, data= d1.z_scored_df )

aov_d1_od_trt=aov(od_corrected_b_n ~ treatment * group, data= d1.z_scored_df )


AIC(aov_d1, aov_d1_od,aov_d1_od_trt)
summary(aov_d1_od)
plot(aov_d1_od)

d1.z_scored_df%>%
  na.omit()%>%
  ggplot(aes(x= pairing, y =od_corrected_b_n ))+
    geom_boxplot()+
      geom_hline(aes(yintercept = mean(od_corrected_b_n), linetype = "Average"), color = alpha("red", 1), linetype = "dashed") 


############################
```


This is now looking at a "true" home scenario for the comparison between sampled communities against 'home' or 'away' metabolites. In this instance, leaves are sourced from Hoko and deployed at Hoko, or sourced from Sekiu and deployed at  Sekiu. The sampled communities are then "home" communities found at their respective sites with either leaf-associated microbes (LAMs) and free-living microbes (FLMs), FLM-only, or LAM-only. These three groups allow us to see how phyllosphere microbes (i.e., LAMS) influence metabolization of home or away metabolites.

TRUE HOME (HOK to HOK or SEK to SEK)
z scored by deployment site and decomposition day
```{r}
library(multcompView)
#filter by correct hours
true_home.meta.df=true_home.meta.df%>%
  filter(hours =="24")

#z score
##### Day 1
#HOK 1 subset
d1_hok1.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "HOK" & true_home.meta.df$site_num == "1" & true_home.meta.df$decomp_day == "1", ])
d3_hok1.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "HOK" & true_home.meta.df$site_num == "1" & true_home.meta.df$decomp_day == "3", ])
d5_hok1.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "HOK" & true_home.meta.df$site_num == "1" & true_home.meta.df$decomp_day == "5", ])
d7_hok1.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "HOK" & true_home.meta.df$site_num == "1" & true_home.meta.df$decomp_day == "7", ])
d11_hok1.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "HOK" & true_home.meta.df$site_num == "1" & true_home.meta.df$decomp_day == "11", ])
d17_hok1.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "HOK" & true_home.meta.df$site_num == "1" & true_home.meta.df$decomp_day == "17", ])
# Calculate z-scores 
d1_hok1.true_home.df=d1_hok1.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d3_hok1.true_home.df=d3_hok1.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d5_hok1.true_home.df=d5_hok1.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d7_hok1.true_home.df=d7_hok1.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d11_hok1.true_home.df=d11_hok1.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d17_hok1.true_home.df=d17_hok1.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))

#HOK 2 subset
d1_hok2.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "HOK" & true_home.meta.df$site_num == "2" & true_home.meta.df$decomp_day == "1", ])
d3_hok2.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "HOK" & true_home.meta.df$site_num == "2" & true_home.meta.df$decomp_day == "3", ])
d5_hok2.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "HOK" & true_home.meta.df$site_num == "2" & true_home.meta.df$decomp_day == "5", ])
d7_hok2.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "HOK" & true_home.meta.df$site_num == "2" & true_home.meta.df$decomp_day == "7", ])
d11_hok2.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "HOK" & true_home.meta.df$site_num == "2" & true_home.meta.df$decomp_day == "11", ])
d17_hok2.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "HOK" & true_home.meta.df$site_num == "2" & true_home.meta.df$decomp_day == "17", ])
# Calculate z-scores 
d1_hok2.true_home.df=d1_hok2.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d3_hok2.true_home.df=d3_hok2.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d5_hok2.true_home.df=d5_hok2.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d7_hok2.true_home.df=d7_hok2.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d11_hok2.true_home.df=d11_hok2.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d17_hok2.true_home.df=d17_hok2.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))

#merge all hok
all.hok.true_home.df=rbind(d1_hok1.true_home.df,
                           d3_hok1.true_home.df,
                           d5_hok1.true_home.df,
                           d7_hok1.true_home.df,
                           d11_hok1.true_home.df,
                           d17_hok1.true_home.df,
                           
                          d1_hok2.true_home.df,
                           d3_hok2.true_home.df,
                           d5_hok2.true_home.df,
                           d7_hok2.true_home.df,
                           d11_hok2.true_home.df,
                           d17_hok2.true_home.df)

#SEK 1 subset
d1_sek1.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "SEK" & true_home.meta.df$site_num == "1" & true_home.meta.df$decomp_day == "1", ])
d3_sek1.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "SEK" & true_home.meta.df$site_num == "1" & true_home.meta.df$decomp_day == "3", ])
d5_sek1.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "SEK" & true_home.meta.df$site_num == "1" & true_home.meta.df$decomp_day == "5", ])
d7_sek1.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "SEK" & true_home.meta.df$site_num == "1" & true_home.meta.df$decomp_day == "7", ])
d11_sek1.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "SEK" & true_home.meta.df$site_num == "1" & true_home.meta.df$decomp_day == "11", ])
d17_sek1.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "SEK" & true_home.meta.df$site_num == "1" & true_home.meta.df$decomp_day == "17", ])
# Calculate z-scores 
d1_sek1.true_home.df=d1_sek1.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d3_sek1.true_home.df=d3_sek1.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d5_sek1.true_home.df=d5_sek1.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d7_sek1.true_home.df=d7_sek1.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d11_sek1.true_home.df=d11_sek1.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d17_sek1.true_home.df=d17_sek1.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))


#SEK 2 subset
d1_sek2.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "SEK" & true_home.meta.df$site_num == "2" & true_home.meta.df$decomp_day == "1", ])
d3_sek2.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "SEK" & true_home.meta.df$site_num == "2" & true_home.meta.df$decomp_day == "3", ])
d5_sek2.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "SEK" & true_home.meta.df$site_num == "2" & true_home.meta.df$decomp_day == "5", ])
d7_sek2.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "SEK" & true_home.meta.df$site_num == "2" & true_home.meta.df$decomp_day == "7", ])
d11_sek2.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "SEK" & true_home.meta.df$site_num == "2" & true_home.meta.df$decomp_day == "11", ])
d17_sek2.true_home.df <- data.frame(true_home.meta.df[true_home.meta.df$deploy == "SEK" & true_home.meta.df$site_num == "2" & true_home.meta.df$decomp_day == "17", ])
# Calculate z-scores 
d1_sek2.true_home.df=d1_sek2.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d3_sek2.true_home.df=d3_sek2.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d5_sek2.true_home.df=d5_sek2.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d7_sek2.true_home.df=d7_sek2.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d11_sek2.true_home.df=d11_sek2.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d17_sek2.true_home.df=d17_sek2.true_home.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))


#merge all sek
all.sek.true_home.df=rbind(d1_sek1.true_home.df,
                           d3_sek1.true_home.df,
                           d5_sek1.true_home.df,
                           d7_sek1.true_home.df,
                           d11_sek1.true_home.df,
                           d17_sek1.true_home.df,
                           
                          d1_sek2.true_home.df,
                           d3_sek2.true_home.df,
                           d5_sek2.true_home.df,
                           d7_sek2.true_home.df,
                           d11_sek2.true_home.df,
                           d17_sek2.true_home.df)

#merge all z scored dfs
merged.true_home_z.df = rbind(all.hok.true_home.df, all.sek.true_home.df)
merged.true_home_z.df <- merged.true_home_z.df %>% 
  mutate(pairing = factor(pairing, levels = c("P-H", "P-AW", "AB-H", "AB-AW")))

# Calculate means for each decomp_day
#od readings
od_means_per_day <- merged.true_home_z.df %>%
  dplyr::group_by(decomp_day) %>%
  dplyr::summarise(mean_od = mean(od_corrected_b_n, na.rm = TRUE))

od_true.home.boxplot=merged.true_home_z.df%>%
  na.omit()%>%
  ggplot(aes(x= treatment, y = od_corrected_b_n, fill = group))+
  geom_boxplot(alpha = 0.4)+
  xlab("treatment")+
  ylab(expression("optical density (OD"[450~nm]*")"))+
  labs(fill = "hfa")+
  facet_wrap(.~decomp_day, labeller = labeller(decomp_day = c("1" = "Day 1", "3" = "Day 3", "5" = "Day 5", "7" = "Day 7", "11" = "Day 11", "17" = "Day 17"))) +
  ggtitle('"True home" comparisons')+
  scale_fill_colorblind()+
           geom_hline(data=od_means_per_day, aes(yintercept = mean_od, linetype = "Average"), color = alpha("red", 1), linetype = "dashed") +
  theme_classic()+
      theme(plot.title = element_text(face = "bold"))

ggsave("figures/od_true.home.boxplot.pdf", od_true.home.boxplot, width = 8, height = 6)

##### models by decomp day

#Day 1
d1_aov=aov(od_corrected_b_n ~ treatment*group, subset = decomp_day == "1", data = na.omit(merged.true_home_z.df))
summary(d1_aov)
d1_meta_tHSD=TukeyHSD(d1_aov, ordered = F, conf.level = 0.95)
d1_cld <- multcompLetters4(d1_aov, d1_meta_tHSD)

# table with factors and 3rd quantile
d1_meta.table <- merged.true_home_z.df%>%
  filter(decomp_day == "1")%>%
  dplyr::group_by(treatment, group) %>%
  dplyr::summarise(mean=mean(od_corrected_b_n), quant = quantile(od_corrected_b_n, probs = 0.75)) %>%
  dplyr::arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
d1_cld <- as.data.frame.list(d1_cld$`treatment:group`)
d1_meta.table$cld <- d1_cld$Letters
#plot
d1_meta.plot=merged.true_home_z.df%>%
  filter(decomp_day =="1")%>%
  na.omit()%>%
  ggplot(aes(x= interaction(treatment, group, sep ="-"), y = od_corrected_b_n, fill = interaction(treatment, group, sep ="-")))+
  geom_boxplot(alpha = 0.4)+
  coord_cartesian(ylim = c(0,4))+
  xlab("")+
  ylab(expression("optical density (OD"[450~nm]*")"))+
  labs(fill = "hfa")+
  ggtitle("Day 1")+
  scale_fill_colorblind()+
  geom_hline(aes(yintercept = mean(od_corrected_b_n), linetype = "Average"), color = alpha("red", 1), linetype = "dashed") +
  theme_classic()+
  theme(plot.title = element_text(face = "bold"))+
  geom_text(data = d1_meta.table, aes(y = quant, label = cld), color = "blue", fontface= "bold",size = 5, vjust=-2, hjust =-0.76)+
  Fig.formatting


#Day 3
d3_aov=aov(od_corrected_b_n ~ treatment*group, subset = decomp_day == "3", data = na.omit(merged.true_home_z.df))
summary(d3_aov)
d3_meta_tHSD=TukeyHSD(d3_aov, ordered = F, conf.level = 0.95)
d3_cld <- multcompLetters4(d3_aov, d3_meta_tHSD)

# table with factors and 3rd quantile
d3_meta.table <- merged.true_home_z.df%>%
  filter(decomp_day == "3")%>%
  dplyr::group_by(treatment, group) %>%
  dplyr::summarise(mean=mean(od_corrected_b_n), quant = quantile(od_corrected_b_n, probs = 0.75)) %>%
  dplyr::arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
d3_cld <- as.data.frame.list(d3_cld$`treatment:group`)
d3_meta.table$cld <- d3_cld$Letters
#plot
d3_meta.plot=merged.true_home_z.df%>%
  filter(decomp_day =="3")%>%
  na.omit()%>%
  ggplot(aes(x= interaction(treatment, group, sep ="-"), y = od_corrected_b_n, fill = interaction(treatment, group, sep ="-")))+
  geom_boxplot(alpha = 0.4)+
    coord_cartesian(ylim = c(0,4))+
  xlab("")+
  ylab("")+
  labs(fill = "hfa")+
  ggtitle("Day 3")+
  scale_fill_colorblind()+
  geom_hline(aes(yintercept = mean(od_corrected_b_n), linetype = "Average"), color = alpha("red", 1), linetype = "dashed") +
  theme_classic()+
  theme(plot.title = element_text(face = "bold"))+
  geom_text(data = d3_meta.table, aes(y = quant, label = cld), color = "blue", fontface= "bold",size = 5, vjust=-2, hjust =-0.76)+
  Fig.formatting

#Day 5
d5_aov=aov(od_corrected_b_n ~ treatment*group, subset = decomp_day == "5", data = na.omit(merged.true_home_z.df))
summary(d5_aov)
d5_meta_tHSD=TukeyHSD(d5_aov, ordered = F, conf.level = 0.95)
d5_cld <- multcompLetters4(d5_aov, d5_meta_tHSD)

# table with factors and 3rd quantile
d5_meta.table <- merged.true_home_z.df%>%
  filter(decomp_day == "5")%>%
  dplyr::group_by(treatment, group) %>%
  dplyr::summarise(mean=mean(od_corrected_b_n), quant = quantile(od_corrected_b_n, probs = 0.75)) %>%
  dplyr::arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
d5_cld <- as.data.frame.list(d5_cld$`treatment:group`)
d5_meta.table$cld <- d5_cld$Letters
#plot
d5_meta.plot=merged.true_home_z.df%>%
  filter(decomp_day == "5")%>%
  na.omit()%>%
  ggplot(aes(x= interaction(treatment, group, sep ="-"), y = od_corrected_b_n, fill = interaction(treatment, group, sep ="-")))+
  geom_boxplot(alpha = 0.4)+
    coord_cartesian(ylim = c(0,4))+
  xlab("")+
  ylab("")+
  labs(fill = "hfa")+
  ggtitle("Day 5")+
  scale_fill_colorblind()+
  geom_hline(aes(yintercept = mean(od_corrected_b_n), linetype = "Average"), color = alpha("red", 1), linetype = "dashed") +
  theme_classic()+
  theme(plot.title = element_text(face = "bold"))+
  geom_text(data = d5_meta.table, aes(y = quant, label = cld), color = "blue", fontface= "bold",size = 5, vjust=-2, hjust =-0.76)+
  Fig.formatting

d5_meta.plot

#Day 7
d7_aov=aov(od_corrected_b_n ~ treatment*group, subset = decomp_day == "7", data = na.omit(merged.true_home_z.df))
summary(d7_aov)
d7_meta_tHSD=TukeyHSD(d7_aov, ordered = F, conf.level = 0.95)
d7_cld <- multcompLetters4(d7_aov, d7_meta_tHSD)

# table with factors and 3rd quantile
d7_meta.table <- merged.true_home_z.df%>%
  filter(decomp_day == "7")%>%
  dplyr::group_by(treatment, group) %>%
  dplyr::summarise(mean=mean(od_corrected_b_n), quant = quantile(od_corrected_b_n, probs = 0.75)) %>%
  dplyr::arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
d7_cld <- as.data.frame.list(d7_cld$`treatment:group`)
d7_meta.table$cld <- d7_cld$Letters
#plot
d7_meta.plot=merged.true_home_z.df%>%
  filter(decomp_day =="7")%>%
  na.omit()%>%
  ggplot(aes(x= interaction(treatment, group, sep ="-"), y = od_corrected_b_n, fill = interaction(treatment, group, sep ="-")))+
  geom_boxplot(alpha = 0.4)+
    coord_cartesian(ylim = c(0,4))+
  xlab("treatment")+
  ylab("")+
  labs(fill = "hfa")+
  ggtitle("Day 7")+
  scale_fill_colorblind()+
  geom_hline(aes(yintercept = mean(od_corrected_b_n), linetype = "Average"), color = alpha("red", 1), linetype = "dashed") +
  theme_classic()+
  theme(plot.title = element_text(face = "bold"))+
  geom_text(data = d7_meta.table, aes(y = quant, label = cld), color = "blue", fontface= "bold",size = 5, vjust=-3, hjust =-0.76)+
  Fig.formatting

d7_meta.plot

#Day 11
d11_aov=aov(od_corrected_b_n ~ treatment*group, subset = decomp_day == "11", data = na.omit(merged.true_home_z.df))
summary(d11_aov)
d11_meta_tHSD=TukeyHSD(d11_aov, ordered = F, conf.level = 0.95)
d11_cld <- multcompLetters4(d11_aov, d11_meta_tHSD)

# table with factors and 3rd quantile
d11_meta.table <- merged.true_home_z.df%>%
  filter(decomp_day == "11")%>%
  dplyr::group_by(treatment, group) %>%
  dplyr::summarise(mean=mean(od_corrected_b_n), quant = quantile(od_corrected_b_n, probs = 0.75)) %>%
  dplyr::arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
d11_cld <- as.data.frame.list(d11_cld$`treatment:group`)
d11_meta.table$cld <- d11_cld$Letters
#plot
d11_meta.plot=merged.true_home_z.df%>%
  filter(decomp_day =="11")%>%
  na.omit()%>%
  ggplot(aes(x= interaction(treatment, group, sep ="-"), y = od_corrected_b_n, fill = interaction(treatment, group, sep ="-")))+
  geom_boxplot(alpha = 0.4)+
    coord_cartesian(ylim = c(0,4))+
  xlab("")+
  ylab("")+
  labs(fill = "hfa")+
  ggtitle("Day 11")+
  scale_fill_colorblind()+
  geom_hline(aes(yintercept = mean(od_corrected_b_n), linetype = "Average"), color = alpha("red", 1), linetype = "dashed") +
  theme_classic()+
  theme(plot.title = element_text(face = "bold"))+
  geom_text(data = d11_meta.table, aes(y = quant, label = cld), color = "blue", fontface= "bold",size = 5, vjust=-2, hjust =-0.76)+
  Fig.formatting

d11_meta.plot


#Day 17
d17_aov=aov(od_corrected_b_n ~ treatment*group, subset = decomp_day == "17", data = na.omit(merged.true_home_z.df))
summary(d17_aov)
d17_meta_tHSD=TukeyHSD(d17_aov, ordered = F, conf.level = 0.95)
d17_cld <- multcompLetters4(d17_aov, d17_meta_tHSD)

# table with factors and 3rd quantile
d17_meta.table <- merged.true_home_z.df%>%
  filter(decomp_day == "17")%>%
  dplyr::group_by(treatment, group) %>%
  dplyr::summarise(mean=mean(od_corrected_b_n), quant = quantile(od_corrected_b_n, probs = 0.75)) %>%
  dplyr::arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
d17_cld <- as.data.frame.list(d17_cld$`treatment:group`)
d17_meta.table$cld <- d17_cld$Letters
#plot
d17_meta.plot=merged.true_home_z.df%>%
  filter(decomp_day=="17")%>%
  na.omit()%>%
  ggplot(aes(x= interaction(treatment, group, sep ="-"), y = od_corrected_b_n, fill = interaction(treatment, group, sep ="-")))+
  geom_boxplot(alpha = 0.4)+
    coord_cartesian(ylim = c(0,4))+
  xlab("")+
  ylab("")+
  labs(fill = "hfa")+
  ggtitle("Day 17")+
  scale_fill_colorblind()+
  geom_hline(aes(yintercept = mean(od_corrected_b_n), linetype = "Average"), color = alpha("red", 1), linetype = "dashed") +
  theme_classic()+
  theme(plot.title = element_text(face = "bold"))+
  geom_text(data = d17_meta.table, aes(y = quant, label = cld), color = "blue", fontface= "bold",size = 5, vjust=-2, hjust =-0.76)+
  Fig.formatting


all.days_meta.plot=plot_grid(
  d1_meta.plot + theme(legend.position = "none"),
  d3_meta.plot + theme(legend.position = "none"),
  d5_meta.plot + theme(legend.position = "none"),
  d7_meta.plot + theme(legend.position = "none"),
  d11_meta.plot+ theme(legend.position = "none"),
  d17_meta.plot + theme(legend.position = "none"),
  rel_widths = c(8,8,8,8,8,8), nrow=2)

ggsave("figures/all.days_meta.plot.pdf", all.days_meta.plot, width = 22, height = 10)



#z score
z_means_per_day <- merged.true_home_z.df %>%
  dplyr::group_by(decomp_day) %>%
  dplyr::summarise(mean_z = mean(z_score, na.rm = TRUE))

library(ggthemes)
merged.true_home_z.df%>%
  na.omit()%>%
  filter(group == "H")%>%
  ggplot(aes(x= treatment, y = z_score, fill = treatment))+
  geom_boxplot(alpha = 0.4)+
  facet_wrap(.~decomp_day)+
  scale_fill_colorblind()+
  geom_hline(data=z_means_per_day, aes(yintercept = mean_z, linetype = "Average"), color = alpha("red", 1), linetype = "dashed") +
  theme_classic()

#false home
means_per_day <- false_home.meta.df %>%
  group_by(decomp_day) %>%
  summarise(mean_z = mean(z_score, na.rm = TRUE))

false_home.meta.df%>%
  na.omit()%>%
  filter(hours =="24")%>%
  ggplot(aes(x= new.trt, y = od_corrected_b_n, fill = group))+
  geom_boxplot()+
  facet_wrap(.~decomp_day)+
  scale_color_colorblind()+
           geom_hline(data= means_per_day, aes(yintercept = mean_od, linetype = "Average"), color = alpha("red", 1), linetype = "dashed") +
  theme_classic()

#find difference between LAM+FL vs FL only
lam_only.df=all.true_home.df%>%
  dplyr::group_by(tree_id,  deploy, metabolite,tree_num, site_num, group, decomp_day)%>%
  dplyr::summarize(od_corrected_b_n = sum(od_corrected_b_n[new.trt == "LAM+FL"]) - sum(od_corrected_b_n[new.trt == "FLM-only"]))%>%
  mutate(new.trt="LAM-only")%>%
  arrange(decomp_day)



AIC(d1_aov, d1_aov2)

summary(aov2)
plot(aov2)

hist(resid(aov3))

```

This is raw and z scored of all communities
```{r}
meta.450.df=meta.450.df%>%
  filter(hours =="24")
#z score
##### Day 1
#HOK 1 subset
d1_hok1.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" & meta.450.df$site_num == "1" & meta.450.df$decomp_day == "1", ])
d3_hok1.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" & meta.450.df$site_num == "1" & meta.450.df$decomp_day == "3", ])
d5_hok1.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" & meta.450.df$site_num == "1" & meta.450.df$decomp_day == "5", ])
d7_hok1.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" & meta.450.df$site_num == "1" & meta.450.df$decomp_day == "7", ])
d11_hok1.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" & meta.450.df$site_num == "1" & meta.450.df$decomp_day == "11", ])
d17_hok1.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" & meta.450.df$site_num == "1" & meta.450.df$decomp_day == "17", ])
# Calculate z-scores 
d1_hok1.df=d1_hok1.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d3_hok1.df=d3_hok1.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d5_hok1.df=d5_hok1.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d7_hok1.df=d7_hok1.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d11_hok1.df=d11_hok1.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d17_hok1.df=d17_hok1.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))


#HOK 2 subset
d1_hok2.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" & meta.450.df$site_num == "2" & meta.450.df$decomp_day == "1", ])
d3_hok2.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" & meta.450.df$site_num == "2" & meta.450.df$decomp_day == "3", ])
d5_hok2.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" & meta.450.df$site_num == "2" & meta.450.df$decomp_day == "5", ])
d7_hok2.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" & meta.450.df$site_num == "2" & meta.450.df$decomp_day == "7", ])
d11_hok2.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" & meta.450.df$site_num == "2" & meta.450.df$decomp_day == "11", ])
d17_hok2.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" & meta.450.df$site_num == "2" & meta.450.df$decomp_day == "17", ])
# Calculate z-scores 
d1_hok2.df=d1_hok2.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d3_hok2.df=d3_hok2.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d5_hok2.df=d5_hok2.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d7_hok2.df=d7_hok2.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d11_hok2.df=d11_hok2.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d17_hok2.df=d17_hok2.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))

#merge all hok
all.hok.df=rbind(d1_hok1.df,
                           d3_hok1.df,
                           d5_hok1.df,
                           d7_hok1.df,
                           d11_hok1.df,
                           d17_hok1.df,
                           
                          d1_hok2.df,
                           d3_hok2.df,
                           d5_hok2.df,
                           d7_hok2.df,
                           d11_hok2.df,
                           d17_hok2.df)

#SEK 1 subset
d1_sek1.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" & meta.450.df$site_num == "1" & meta.450.df$decomp_day == "1", ])
d3_sek1.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" & meta.450.df$site_num == "1" & meta.450.df$decomp_day == "3", ])
d5_sek1.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" & meta.450.df$site_num == "1" & meta.450.df$decomp_day == "5", ])
d7_sek1.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" & meta.450.df$site_num == "1" & meta.450.df$decomp_day == "7", ])
d11_sek1.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" & meta.450.df$site_num == "1" & meta.450.df$decomp_day == "11", ])
d17_sek1.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" & meta.450.df$site_num == "1" & meta.450.df$decomp_day == "17", ])
# Calculate z-scores 
d1_sek1.df=d1_sek1.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d3_sek1.df=d3_sek1.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d5_sek1.df=d5_sek1.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d7_sek1.df=d7_sek1.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d11_sek1.df=d11_sek1.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d17_sek1.df=d17_sek1.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))


#SEK 2 subset
d1_sek2.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" & meta.450.df$site_num == "2" & meta.450.df$decomp_day == "1", ])
d3_sek2.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" & meta.450.df$site_num == "2" & meta.450.df$decomp_day == "3", ])
d5_sek2.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" & meta.450.df$site_num == "2" & meta.450.df$decomp_day == "5", ])
d7_sek2.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" & meta.450.df$site_num == "2" & meta.450.df$decomp_day == "7", ])
d11_sek2.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" & meta.450.df$site_num == "2" & meta.450.df$decomp_day == "11", ])
d17_sek2.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" & meta.450.df$site_num == "2" & meta.450.df$decomp_day == "17", ])
# Calculate z-scores 
d1_sek2.df=d1_sek2.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d3_sek2.df=d3_sek2.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d5_sek2.df=d5_sek2.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d7_sek2.df=d7_sek2.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d11_sek2.df=d11_sek2.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))
d17_sek2.df=d17_sek2.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))


#merge all hok
all.sek.df=rbind(d1_sek1.df,
                           d3_sek1.df,
                           d5_sek1.df,
                           d7_sek1.df,
                           d11_sek1.df,
                           d17_sek1.df,
                           
                          d1_sek2.df,
                           d3_sek2.df,
                           d5_sek2.df,
                           d7_sek2.df,
                           d11_sek2.df,
                           d17_sek2.df)

#merge z scored dfs
merged_z.df = rbind(all.hok.df, all.sek.df)

 


#all combinations with meta.df
means_per_day <- meta.450.df %>%
  group_by(decomp_day) %>%
  summarise(mean_od = mean(od_corrected_b_n, na.rm = TRUE))

meta.450.df%>%
  na.omit()%>%
  ggplot(aes(x= treatment, y = od_corrected_b_n, color = group))+
  geom_boxplot()+
  facet_wrap(.~decomp_day)+
  scale_color_colorblind()+
  geom_hline(data= means_per_day, aes(yintercept = mean_od, linetype = "Average"), color = alpha("red", 1), linetype = "dashed") +
  theme_classic()

d1_all.meta_aov=aov(od_corrected_b_n~new.trt*group, subset = decomp_day == "1", data = na.omit(meta.450.df))

d1_all.meta_lmer=lmer(od_corrected_b_n~new.trt*group + (1|deploy) + (1|site_num) + (1| tree_num) + (1 | tree_id), subset = decomp_day == "1", data = na.omit(meta.450.df))

d3_all.meta_lmer=lmer(od_corrected_b_n~new.trt*group + (1|deploy) + (1|site_num) + (1| tree_num) + (1 | tree_id), subset = decomp_day == "3", data = na.omit(meta.450.df))


plot(resid(d1_all.meta_lmer))
summary(d3_all.meta_lmer)

AIC(d3_all.meta_aov, d3_all.meta_lmer)

d3_all.meta_aov=aov(od_corrected_b_n~new.trt*group, subset = decomp_day == "3", data = na.omit(meta.450.df))
TukeyHSD(d3_all.meta_aov)
summary(d3_all.meta_aov)

hist(resid(d3_all.meta_aov))
plot(resid(d3_all.meta_aov))
# z score
means_per_day <- merged_z.df %>%
  group_by(decomp_day) %>%
  summarise(mean_z = mean(z_score, na.rm = TRUE))


merged_z.df%>%
  na.omit()%>%
  filter(hours =="24")%>%
  ggplot(aes(x= new.trt, y = z_score, color = group))+
  geom_boxplot()+
  facet_wrap(.~decomp_day)+
  scale_color_colorblind()+
           geom_hline(data= means_per_day, aes(yintercept = mean_z, linetype = "Average"), color = alpha("red", 1), linetype = "dashed") +
  theme_classic()

d1_all.meta_z_aov=aov(z_score~new.trt*group, subset = decomp_day == "1", data = na.omit(merged_z.df))

d3_all.meta_z_aov=aov(z_score~new.trt*group, subset = decomp_day == "3", data = na.omit(merged_z.df))
hist(resid(d3_all.meta_z_aov))

plot(resid(d3_all.meta_z_aov))

summary(all.meta_z_aov)

AIC(d3_all.meta_aov, d3_all.meta_z_aov)

plot(resid(all.meta_z_aov))
plot(all.meta_aov)


```



##### Day 3
```{r}
##### Day 1
#HOK 1 subset
d3.hok1_meta.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" &  meta.450.df$decomp_day == "3"& meta.450.df$hours == "24"& meta.450.df$site_num == "1", ])
# Calculate z-scores 
d3.hok1_meta.df=d3.hok1_meta.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))


#HOK 2 subset
d3.hok2_meta.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" &  meta.450.df$decomp_day == "3"& meta.450.df$hours == "24"& meta.450.df$site_num == "2", ])
# Calculate z-scores 
d3.hok2_meta.df=d3.hok2_meta.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))

#SEK 1 subset
d3.sek1_meta.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" &  meta.450.df$decomp_day == "3"& meta.450.df$hours == "24"& meta.450.df$site_num == "1", ])
# Calculate z-scores 
d3.sek1_meta.df=d3.sek1_meta.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))

#SEK 2 subset
d3.sek2_meta.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" &  meta.450.df$decomp_day == "3"& meta.450.df$hours == "24"& meta.450.df$site_num == "2", ])
# Calculate z-scores 
d3.sek2_meta.df=d3.sek2_meta.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))

#merge z scored dfs
d3.z_scored_df = rbind(d3.hok1_meta.df, d3.hok2_meta.df, d3.sek1_meta.df, d3.sek2_meta.df)

#anovas
aov_d3=aov(z_score ~ pairing, data= d3.z_scored_df )

aov_d3_od=aov(od_corrected_b_n ~ pairing, data= d3.z_scored_df )

AIC(aov_d3, aov_d3_od)
summary(aov_d3_od)
plot(aov_d3_od)

d3.z_scored_df%>%
  na.omit()%>%
  ggplot(aes(x= pairing, y =od_corrected_b_n ))+
    geom_boxplot()+
      geom_hline(aes(yintercept = mean(od_corrected_b_n), linetype = "Average"), color = alpha("red", 1), linetype = "dashed") 


```



##### Day 5
```{r}
##### Day 1
#HOK 1 subset
d5.hok1_meta.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" &  meta.450.df$decomp_day == "5"& meta.450.df$hours == "24"& meta.450.df$site_num == "1", ])
# Calculate z-scores 
d5.hok1_meta.df=d5.hok1_meta.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))


#HOK 2 subset
d5.hok2_meta.df <- data.frame(meta.450.df[meta.450.df$deploy == "HOK" &  meta.450.df$decomp_day == "5"& meta.450.df$hours == "24"& meta.450.df$site_num == "2", ])
# Calculate z-scores 
d5.hok2_meta.df=d5.hok2_meta.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))

#SEK 1 subset
d5.sek1_meta.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" &  meta.450.df$decomp_day == "5"& meta.450.df$hours == "24"& meta.450.df$site_num == "1", ])
# Calculate z-scores 
d5.sek1_meta.df=d5.sek1_meta.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))

#SEK 2 subset
d5.sek2_meta.df <- data.frame(meta.450.df[meta.450.df$deploy == "SEK" &  meta.450.df$decomp_day == "5"& meta.450.df$hours == "24"& meta.450.df$site_num == "2", ])
# Calculate z-scores 
d5.sek2_meta.df=d5.sek2_meta.df%>%
  na.omit()%>%
  mutate(z_score= (od_corrected_b_n - mean(od_corrected_b_n))/sd(od_corrected_b_n))

#merge z scored dfs
d5.z_scored_df = rbind(d5.hok1_meta.df, d5.hok2_meta.df, d5.sek1_meta.df, d5.sek2_meta.df)

#anovas
aov_d5=aov(z_score ~ pairing, data= d5.z_scored_df )

aov_d5_od=aov(od_corrected_b_n ~ pairing, data= d5.z_scored_df )

lmer_d5_od=lmer(od_corrected_b_n ~ pairing +(1 | deploy)  + (1 | tree_id) + (1 | tree_num), data= d5.z_scored_df )
view(d5.z_scored_df)

AIC(aov_d5_od, lmer_d5_od)

summary(aov_d5_od)
plot(aov_d5_od)


# Plot boxplots with letters
library(ggplot2)
library(ggsignif)

# Ensure that the 'pairing' variable is a factor
d5.z_scored_df$pairing <- factor(d5.z_scored_df$pairing)

# Create the boxplot
ggplot(d5.z_scored_df, aes(x = pairing, y = od_corrected_b_n)) +
  geom_boxplot() +
  geom_hline(aes(yintercept = mean(od_corrected_b_n), linetype = "Average"), 
             color = alpha("red", 1), linetype = "dashed") +
  stat_compare_means(comparisons = tukey_result, method = "anova", label = "p.adj") +
  theme_bw()



```






```{r}
aov_d3=aov(z_score ~ pairing, data= na.omit(z_scored_df[z_scored_df$decomp_day=="3" & z_scored_df$hours=="24",]) )
summary(aov_d3)
plot(aov_d3)

aov_d5=aov(z_score ~ pairing, data= na.omit(z_scored_df[z_scored_df$decomp_day=="5" & z_scored_df$hours=="24",]) )
summary(aov_d5)
plot(aov_d5)

lmer_d5=aov(z_score ~ pairing + (1 | tree_num) + ( 1 | ), data= na.omit(z_scored_df[z_scored_df$decomp_day=="5" & z_scored_df$hours=="24",]) )
summary(lmer_d5)

AIC(aov_d5, lmer_d5)

aov_d7=aov(z_score ~ pairing, data= na.omit(z_scored_df[z_scored_df$decomp_day=="7" & z_scored_df$hours=="24",]) )
summary(aov_d7)
plot(aov_d7)

aov_d11=aov(z_score ~ pairing, data= na.omit(z_scored_df[z_scored_df$decomp_day=="11" & z_scored_df$hours=="24",]) )
summary(aov_d11)
plot(aov_d11)

aov_d17=aov(z_score ~ pairing, data= na.omit(z_scored_df[z_scored_df$decomp_day=="17" & z_scored_df$hours=="24",]) )
summary(aov_d17)

AIC(aov1, aov2)

plot(aov2)

TukeyHSD(aov2)
TukeyHSD(aov1)


z_scored_df%>%
  na.omit()%>%
ggplot(aes(x = pairing, y = z_score, color= pairing))+
    geom_boxplot()+
  scale_color_colorblind()+
    geom_hline(data = na.omit(z_scored_df), aes(yintercept = mean(z_score), linetype = "Average"), color = alpha("red",0.9), linetype = "dotted") +
  facet_wrap(.~decomp_day, scales = "free_y")+
  theme_classic()

  geom_point(aes(shape = pairing), position = position_jitter(width = 0.2)) 


```

Model exploration
**A note on GAMs**
The term "parametric" is often used to describe statistical models with a fixed, finite number of parameters. In contrast, "nonparametric" models are more flexible and can adapt to the complexity of the data without assuming a specific functional form.

Generalized Additive Models (GAMs) can be considered semiparametric. They are a type of regression model that extends generalized linear models (GLMs) by allowing for more flexible, nonlinear relationships between predictors and the response variable. The "additive" part refers to the fact that the model is built by adding together multiple smooth functions, each representing the relationship between a predictor and the response.

While GAMs introduce flexibility with the smooth functions, they still include parametric components, such as linear terms and parametric link functions. The parametric components provide a structured foundation, and the smooth functions capture additional complexity in the data.

So, we are using GAMsm that combine parametric and nonparametric elements to provide a flexible framework for modeling complex relationships in the metabolism assay data.

```{r}
lmer1 = lmer(od_corrected_b_n ~ treatment * group + ( 1 | deploy) + (1 | tree_id)+ (1|tree_num), subset = hours == "24", data = na.omit(meta.450.df))

glmer1 = lmer(od_corrected_b_n ~ treatment * group + ( 1 | deploy) + (1 | tree_id)+ (1|tree_num), subset = hours == "24", data = na.omit(meta.450.df))

AIC(lmer2, gam_trt.grp)
summary(lmer1)

plot(resid(lmer1))
qqnorm(resid(lmer1))
qqline(resid(lmer1))


lmer2 = lmer(od_corrected_b_n ~ treatment * group + ( 1 + treatment | decomp_day) + (1 | tree_num) + (1 | tree_id) + (1 | site_num) , subset = hours == "24", data = na.omit(meta.450.df))

lmer3 = lmer(od_corrected_b_n ~ treatment * group + ( 1 + treatment /group | decomp_day) + (1 | tree_num) + (1 | tree_id) + (1 | site_num) , subset = hours == "24", data = na.omit(meta.450.df))

summary(lmer3)


lmer4 = lmer(od_corrected_b_n ~ treatment * group + ( 1 + group /treatment | decomp_day) + (1 | tree_num) + (1 | tree_id) + (1 | site_num) , subset = hours == "24", data = na.omit(meta.450.df))

AIC(lmer1, lmer2, lmer3, lmer4)



```


```{r}
lmer1 = lmer(od_corrected_b_n ~ treatment * group + (1 | deploy) + (1|site_num) + (1 |tree_id) + (1|tree_num),  data = na.omit(meta.450.df[meta.450.df$decomp_day=="1" & meta.450.df$hours=="24",]))
summary(lmer1)

lmer3 = lmer(od_corrected_b_n ~ treatment * group + (1 | deploy) + (1|site_num) + (1 |tree_id) + (1|tree_num),  data = na.omit(meta.450.df[meta.450.df$decomp_day=="3" & meta.450.df$hours=="24",]))
summary(lmer3)

lmer5 = lmer(od_corrected_b_n ~ treatment * group + (1 | deploy) + (1|site_num) + (1 |tree_id) + (1|tree_num),  data = na.omit(meta.450.df[meta.450.df$decomp_day=="5" & meta.450.df$hours=="24",]))
summary(lmer5)

lmer7 = lmer(od_corrected_b_n ~ treatment * group + (1 | deploy) + (1|site_num) + (1 |tree_id) + (1|tree_num),  data = na.omit(meta.450.df[meta.450.df$decomp_day=="7" & meta.450.df$hours=="24",]))
summary(lmer7)


lmer11 = lmer(od_corrected_b_n ~ treatment * group + (1 | deploy) + (1|site_num) + (1 |tree_id) + (1|tree_num),  data = na.omit(meta.450.df[meta.450.df$decomp_day=="11" & meta.450.df$hours=="24",]))
summary(lmer11)
plot(resid(lmer11))


lmer17 = lmer(od_corrected_b_n ~ treatment * group + (1 | deploy) + (1|site_num) + (1 |tree_id) + (1|tree_num),  data = na.omit(meta.450.df[meta.450.df$decomp_day=="17" & meta.450.df$hours=="24",]))
aov(lmer17)




lmer2 = lmer(od_corrected_b_n ~ treatment * group + (1 | deploy) + (1|site_num) + (1 |tree_id) + (1|tree_num),  data = na.omit(meta.450.df[meta.450.df$decomp_day=="5" & meta.450.df$hours=="24",]))
summary(lmer2)

AIC(lmer1, glmer2)

lmer2 = lmer(z_score ~ treatment * group + (1 | decomp_day)  +  (1 | tree_id) + (1|tree_num), subset = hours == "24", data = na.omit(z_scored_df))

AIC(lmer1, lmer2)

lmer2 = lmer(od_corrected_b_n ~ pairing + (1 | decomp_day) + (1 | deploy) +  (1 | tree_id) + (1|tree_num), subset = hours == "24", data = meta.450.df)

AIC(lmer1, lmer2)
lmer3 = lmer(od_corrected_b_n ~ treatment * group + ( 1 + treatment + group | decomp_day)  + (1 | tree_id/tree_num) + (1|deploy/site_num), subset = hours == "24", data = meta.450.df)

lmer4 = lmer(od_corrected_b_n ~ treatment * group + ( treatment * group || decomp_day)  + (1 | tree_id/tree_num) , subset = hours == "24", data = meta.450.df)


gam1= gam(od_corrected_b_n ~ pairing + s(decomp_day, pairing, bs = "fs", k = 5)  +s(tree_num, k = 12, bs = 're') + s(tree_id, bs = 're') + 
                   s(site_num, bs = 're') + s(deploy, bs = 're'),  method = "REML", 
                 data = meta.450.df)
summary(gam1)
gam.check(gam1)
#REML is generally preferred if you are primarily interested in estimating fixed effects and want better estimates of variance components. It's suitable for mixed-effects models.


#testing GAMs

gam1=gam(od_corrected_b_n ~ pairing + s(decomp_day ,k =6) + s(tree_id, tree_num, bs="re") + s(deploy, site_num, bs="re"), subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam2=gam(od_corrected_b_n ~ pairing + s(decomp_day, by= pairing, k =6) + s(tree_id,bs="re") + s(tree_num, bs="re") + s(deploy,bs="re") + s(site_num, bs="re"), subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam3=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k =6) + s(decomp_day, by= group, k =6)+ s(tree_id,bs="re") + s(tree_num, bs="re") + s(deploy,bs="re") + s(site_num, bs="re"), subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))
#####
gam4=gam(od_corrected_b_n ~ treatment + group + s(decomp_day, by= treatment, k =6) + s(decomp_day, by= group, k =6)+ s(tree_id,bs="re") + s(tree_num, bs="re") + s(deploy,bs="re") + s(site_num, bs="re"), subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))
########

gam5=gam(od_corrected_b_n ~ pairing + s(decomp_day, by= pairing, k=6)  + s(tree_num, bs="re") + s(deploy,bs="re") + s(site_num, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam6=gam(od_corrected_b_n ~ pairing + s(decomp_day, by= pairing, k=6)  + s(tree_num, tree_id, bs="re") +  s(site_num, deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))


gam7=gam(od_corrected_b_n ~ treatment + group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k=6)+s(tree_num, bs="re") + s(deploy,bs="re") + s(site_num, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam8=gam(od_corrected_b_n ~ treatment + group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k=6)+s(tree_id, tree_num, bs="re")+ s(deploy, site_num, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

#########

gam9=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k=6)+s(tree_num, bs="re") + s(deploy,bs="re") + s(site_num, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam10=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k =6)+ s( tree_id, tree_num, bs="re", k= 20) + s(deploy, site_num, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam11=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k =6)+ s( tree_num, tree_id, bs="re") + s(site_num,bs="re") + s(deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam12=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k =6)+ s( site_num, tree_num, bs="re") + s(tree_id, bs="re") +  + s(deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))



AIC( gam10, gam11, gam12)

####### 2 canditate models
####### first lets look at main effects and interaction of treatment and group

# looks at interaction between main effects
tw_gam_trt.grp=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k =6)+ s( tree_num, tree_id, bs="re") + s(site_num,bs="re") + s(deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

tw_gam_trt.grp2=gam(od_corrected_b_n ~ treatment + group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k =6)+ s( tree_num, tree_id, bs="re")  + s(deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

tw_gam_trt.grp3=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, k =6)+ s( tree_num, tree_id, bs="re") + s(site_num,bs="re") + s(deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

tw_gam_trt.grp4=gam(od_corrected_b_n ~ treatment*group + s(decomp_day, treatment, bs="fs", k =6)+ s(decomp_day,group,bs="fs", k =6)+ s(tree_id, bs="re") + s(site_num, deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

summary(tw_gam_trt.grp4)

AIC(tw_gam_trt.grp, tw_gam_trt.grp2, tw_gam_trt.grp3,tw_gam_trt.grp4 )

concurvity(tw_gam_trt.grp4)

inv_gam_trt.grp=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k =6)+ s( tree_num, tree_id, bs="re") + s(site_num,bs="re") + s(deploy, bs="re"),  subset = hours == "24",family = gaussian(link = "inverse"), method = "REML", data = na.omit(meta.450.df))

log_gam_trt.grp=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k =6)+ s( tree_num, tree_id, bs="re") + s(site_num,bs="re") + s(deploy, bs="re"),  subset = hours == "24",family = gaussian(link = "log"), method = "REML", data = na.omit(meta.450.df))

gam_trt.grp=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k =6)+ s( tree_num, tree_id, bs="re") + s(site_num,bs="re") + s(deploy, bs="re"),  subset = hours == "24",family = gaussian(link = "identity"), method = "REML", data = na.omit(meta.450.df))

AIC(tw_gam_trt.grp, gam_trt.grp)

summary(tw_gam_trt.grp)
gam.check(tw_gam_trt.grp)
anova.gam(gam_trt.grp)

#plot difference for P vs AB
plot_difference(
  gam_trt.grp,
  series = decomp_day,
  difference = list(treatment = c("P", "AB"))
) + 
  Fig.formatting 
  #this is important to see where the differences in the smooths lie given that the term is significant (p = 0.002)


#plot difference for P vs AB
plot_difference(
  gam_trt.grp,
  series = decomp_day,
  difference = list(group = c("H", "AW"))
) + 
  Fig.formatting 

####### next we can look at paired main effects for pairwise comparison
#looks at individual treatments
gam_pair=gam(od_corrected_b_n ~ pairing + s(decomp_day, by= pairing, k=6) + s( tree_num, tree_id, bs="re") + s(site_num,bs="re") + s(deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))
summary(gam_pair)
gam.check(gam_pair)

AIC(gam_trt.grp, gam_pair, lmer1)
BIC(gam_trt.grp, gam_pair)
anova.gam(gam_pair)

#### in our model (gam_pair), the reference level is set to P-AW, so we can inspect the differences between smooths relative to P-AW

#plot difference for PAW and PH
paw.vs.ph.diff=plot_difference(
  gam_pair,
  series = decomp_day,
  difference = list(pairing = c("P-AW", "P-H"))
)+
   xlab("")+
  ylab("")+
      ggtitle("P-AW vs. P-H")+
  Fig.formatting

#plot difference for PAW and ABAW
paw.vs.abaw.diff=plot_difference(
  gam_pair,
  series = decomp_day,
  difference = list(pairing = c("P-AW", "AB-AW"))
)+
    xlab("")+
    ggtitle("P-AW vs. AB-AW")+
  Fig.formatting

#plot difference for PAW and ABH
paw.vs.abh.diff=plot_difference(
  gam_pair,
  series = decomp_day,
  difference = list(pairing = c("P-AW", "AB-H"))
)+
  xlab("")+
  ylab("")+
  ggtitle("P-AW vs. AB-H")+
  Fig.formatting

#plot difference for PH and ABH
ph.vs.abh.diff=plot_difference(
  gam_pair,
  series = decomp_day,
  difference = list(pairing = c("P-H", "AB-H"))
)+
  xlab("")+
  ylab("")+
  ggtitle("P-H vs. AB-H")+
  Fig.formatting

#plot difference for PH and ABAW
ph.vs.abaw.diff=plot_difference(
  gam_pair,
  series = decomp_day,
  difference = list(pairing = c("P-H", "AB-AW"))
)+
  xlab("")+
  ylab("")+
  ggtitle("P-H vs. AB-AW")+
  Fig.formatting

#plot difference for ABAW and ABH
abaw.vs.abh.diff=plot_difference(
  gam_pair,
  series = decomp_day,
  difference = list(pairing = c("AB-AW", "AB-H"))
)+
  xlab("")+
  ylab("")+
  ggtitle("AB-AW vs. AB-H")+
  Fig.formatting




pair_diffs=plot_grid(
  paw.vs.ph.diff,
  paw.vs.abaw.diff,
  paw.vs.abh.diff,
  ph.vs.abh.diff,
  ph.vs.abaw.diff,
  abaw.vs.abh.diff,
  rel_widths = c(1,1,1,1,1,1), nrow=2, ncol=3
)

pair_diffs

ggsave("figures/pair_diffs.pdf", pair_diffs, width = 16, height = 8)



```

Model 1 (gam_trt.grp):

Intercept: Represents the estimated mean response when all predictors are at their reference levels. In this case, it corresponds to the baseline condition of treatment AB and group AW.

TreatmentP: Represents the estimated change in the mean response when moving from treatment AB to treatment P while holding group constant. The estimate is 0.10648 with a standard error of 0.04688. The t-value is 2.271, and the p-value is 0.0234.

GroupH: Represents the estimated change in the mean response when moving from group AW to group H while holding treatment constant. The estimate is -0.05338 with a standard error of 0.04225. The t-value is -1.263, and the p-value is 0.2068.

TreatmentP:GroupH Interaction: Represents the additional change in the mean response when both treatment is P and group is H compared to the reference level (treatment AB, group AW). The estimate is 0.03843 with a standard error of 0.06137. The t-value is 0.626, and the p-value is 0.5314.

Interpretation:

The interaction term, TreatmentP:GroupH, provides information about whether the effect of treatment P differs when the group is H compared to the reference level (group AW).

A positive estimate (0.03843) suggests that, relative to the baseline condition (AB-AW), the mean response is higher when both treatment is P and group is H. However, this effect is not statistically significant since the p-value is 0.5314, which is greater than the commonly used significance level of 0.05.

In simpler terms, the interaction suggests that there is no significant difference in the effect of treatment P between groups H and AW. It indicates that the effect of treatment P on the mean response is consistent across both groups.


TreatmentP:GroupH Interaction Term:

The estimate (0.03843) represents the additional change in the mean response when both treatment is P and group is H compared to the reference level (AB-AW).
Interpretation:

If we consider the levels of the categorical variables:

Treatment: AB is the reference level.
Group: AW is the reference level.
The interaction term captures the additional change in the mean response when we move from the reference condition (AB-AW) to a specific condition, which is Treatment P and Group H.

However, the interpretation does not directly provide information about the situation where Treatment is AB and Group is H. This is because the interaction term explicitly involves Treatment P.

If you want to assess the difference in the mean response between Treatment AB and Treatment P specifically within Group H, you can do so by comparing the individual estimates for Treatment AB and Treatment P within Group H.

Here's a simplified interpretation:

AB-H vs. AB-AW: The effect of moving from AB to H within the same treatment (AB) is represented by the main effect of Group (GroupH). This is estimated to be -0.05338 (not statistically significant in this case).

P-H vs. AB-AW: The interaction term (TreatmentP:GroupH) captures the additional change when both Treatment is P and Group is H compared to the reference level (AB-AW). The estimate is 0.03843 (not statistically significant in this case).



------------------- % mass loss data -------------------------------

Load mass data 
```{r}
library(stringr)
library(purrr)
mass.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/leaf_mass.csv")
###check for any white spaces clean up
columns_with_whitespace <- map_lgl(mass.df, ~ any(str_detect(.x, "^\\s+|\\s+$")))
columns_with_whitespace

#correction factor adjusting for blotted and fresh dry mass
mass.df= mass.df%>%
  mutate(final_correct = (0.941*(final_mass)-0.00337))

# calculate % mass loss 
mass.df <- mass.df %>%
  mutate(perc.mass_loss = ((initial_mass - final_correct) / initial_mass)*100)

# Calculate the constant C, adding 1 for safety
C <- abs(min(mass.df$perc.mass_loss)) + 1  

#create arcsine trans column with perc.mass_loss + constant 
mass.df <- mass.df %>%
  mutate(perc.mass_loss = perc.mass_loss + C) %>%
  mutate(prop.mass_loss = perc.mass_loss/100) %>%
mutate(arcsine_trans = asin(sqrt(prop.mass_loss)))

mass.df <- mass.df %>%
  mutate(prop.mass_loss = perc.mass_loss/100) %>%
mutate(arcsine_trans = asin(sqrt(prop.mass_loss)))


#subset P treatment only for "true" HFA  
mass.present.df=mass.df %>%
  filter(treatment == "P")

#subset AB treatment only for HFA  
mass.absent.df=mass.df %>%
  filter(treatment == "AB")

#add new treatment columns
mass.df <- mass.df %>%
  mutate(
    new.trt = case_when(
      treatment == "P" ~ "LAM+FLM",
      treatment == "AB" ~ "FLM-only"))

#plot
mass.df%>%
  ggplot(aes(x=new.trt, y = perc.mass_loss, color = hfa))+
  geom_boxplot()


```


load abiotic river data 
```{r}
#load datq
temps.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/rivers/river_temps.csv")
#remove NAs
temps.df=na.omit(temps.df)


# Convert "date" to proper date format
temps.df <- temps.df %>%
  mutate(date = as.Date(date, format = "%m/%d/%y"))

# Extract date part
temps.df <- temps.df %>%
  mutate(date_only = as.Date(date))

# Group by date and calculate mean water temperature for each day
daily_mean_temp <- temps.df %>%
  group_by(site, date_only) %>%
  summarise(mean_temp = mean(temp_c),
            mean_light = mean(light_lux))

# Sum the mean daily temperatures for each site
degree_days <- daily_mean_temp %>%
  group_by(site) %>%
  summarise(total_degree_days = sum(pmax(0, mean_temp - 0)))

#k rate

k=(-(log(mass.df$final_correct/mass.df$initial_mass))/)


```


**Subset dataframes to z score by deployment site (all HOK 1, all HOK 2, etc...)**

Z scoring controls for site variability, and allows for a better comparison of treatments. But how exactly?

Z-scoring is a statistical measure that quantifies how far away a particular data point is from the mean of a dataset in terms of standard deviations (i.e., the spread of the data points from the mean). It's a way to standardize data and make it more comparable across different scales or distributions. Here's how z-scoring works:

1) calculate the mean and standard deviation: for the variable you want to z-score (in this case mass loss by deployment site), calculate its mean (average) and standard deviation. These values provide information about the central tendency (i.e., the center/midpoint of the data or a value around which the data clusters) and the spread of the data.

2) subtract the mean: for each data point, subtract the mean of the variable. This step shifts the data so that the new mean becomes 0.

3) divide by the standard deviation: after subtracting the mean, divide the result by the standard deviation. This step scales the data based on the spread of the original data, so that the new data points are expressed in terms of standard deviations from the mean.

Mathematically, the z-score (Z) for a data point (X) with mean (μ) and standard deviation (σ) is calculated using the formula:

Z = (X - μ) / σ

The resulting z-score indicates how many standard deviations the original data point is away from the mean. If the z-score is positive, it means the data point is above the mean, and if it's negative, it's below the mean.

Note: z-scoring assumes a roughly normal distribution of the data.

Below, the percent mass loss for each deployment site is z scored. Additionally, the larger dataframe has been subset to only include P treatments so that the "true" HFA can be calculated.
```{r}
#check for distribution of mass loss
hist(mass.df$perc.mass_loss)
  #roughly normal...that's good!

#####Begin z scoring
#HOK 1 subset
mass.hok1.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.hok1.df=mass.hok1.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))


#HOK 1 P subset
mass.hok1.present.df = data.frame(mass.hok1.df[mass.hok1.df$deploy == "HOK" & mass.hok1.df$treatment == "P" & mass.hok1.df$site_num == "1", ])
# Calculate z-scores 
mass.hok1.present.df=mass.hok1.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
  #HOK 1 AB subet
mass.hok1.absent.df = data.frame(mass.hok1.df[mass.hok1.df$deploy == "HOK" & mass.hok1.df$treatment == "AB" & mass.hok1.df$site_num == "1", ])



#HOK 2 subset
mass.hok2.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.hok2.df=mass.hok2.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))


#HOK 2 P subset
mass.hok2.present.df =data.frame(mass.hok2.df[mass.hok2.df$deploy == "HOK" & mass.hok2.df$treatment == "P" & mass.hok2.df$site_num == "2", ])
# Calculate z-scores 
mass.hok2.present.df=mass.hok2.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
  #HOK 2 AB subset
mass.hok2.absent.df =data.frame(mass.hok2.df[mass.hok2.df$deploy == "HOK" & mass.hok2.df$treatment == "AB" & mass.hok2.df$site_num == "2", ])

#SEK 1 subset
mass.sek1.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.sek1.df=mass.sek1.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))


#SEK 1 P subset
mass.sek1.present.df = data.frame(mass.sek1.df[mass.sek1.df$deploy == "SEK" & mass.sek1.df$treatment == "P" & mass.sek1.df$site_num == "1", ])
# Calculate z-scores 
mass.sek1.present.df=mass.sek1.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
  #SEK 1 AB subset
mass.sek1.absent.df = data.frame(mass.sek1.df[mass.sek1.df$deploy == "SEK" & mass.sek1.df$treatment == "AB" & mass.sek1.df$site_num == "1", ])


#SEK 2 subset
mass.sek2.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.sek2.df=mass.sek2.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))


#SEK 2 P subset
mass.sek2.present.df = data.frame(mass.sek2.df[mass.sek2.df$deploy == "SEK" & mass.sek2.df$treatment == "P" & mass.sek2.df$site_num == "2", ])
# Calculate z-scores 
mass.sek2.present.df=mass.sek2.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
  #SEK 2 AB subset
mass.sek2.absent.df = data.frame(mass.sek2.df[mass.sek2.df$deploy == "SEK" & mass.sek2.df$treatment == "AB" & mass.sek2.df$site_num == "2", ])

##merge all sites
mass.deploy.merged.df=rbind(mass.hok1.df, mass.hok2.df, mass.sek1.df, mass.sek2.df)

mass.deploy.merged.df=mass.deploy.merged.df %>%
  mutate(hfa = ifelse(hfa == "A", "AW", "H"))


mass.deploy.merged.df$treatment=as.factor(as.factor(mass.deploy.merged.df$treatment))
mass.deploy.merged.df$tree_id=as.factor(as.factor(mass.deploy.merged.df$tree_id))
mass.deploy.merged.df$deploy=as.factor(as.factor(mass.deploy.merged.df$deploy))
mass.deploy.merged.df$hfa=as.factor(as.factor(mass.deploy.merged.df$hfa))

#merge all P treatment sites 
mass.present.deploy.merged.df=rbind(mass.hok1.present.df, mass.hok2.present.df, mass.sek1.present.df, mass.sek2.present.df)
mass.present.deploy.merged.df$treatment=as.factor(mass.present.deploy.merged.df$treatment)
mass.present.deploy.merged.df$tree_id=as.factor(mass.present.deploy.merged.df$tree_id)
mass.present.deploy.merged.df$deploy=as.factor(mass.present.deploy.merged.df$deploy)
mass.present.deploy.merged.df$hfa=as.factor(mass.present.deploy.merged.df$hfa)

#merge all AB treatment sites 
mass.absent.deploy.merged.df=rbind(mass.hok1.absent.df, mass.hok2.absent.df, mass.sek1.absent.df, mass.sek2.absent.df)
mass.absent.deploy.merged.df$treatment=as.factor(mass.absent.deploy.merged.df$treatment)
mass.absent.deploy.merged.df$tree_id=as.factor(mass.absent.deploy.merged.df$tree_id)
mass.absent.deploy.merged.df$deploy=as.factor(mass.absent.deploy.merged.df$deploy)
mass.absent.deploy.merged.df$hfa=as.factor(mass.absent.deploy.merged.df$hfa)

```


Model and plot for z scored mass loss
```{r}
library(multcompView)

#create new site column
mass.deploy.merged.df=mass.deploy.merged.df%>%
  mutate(site = paste(deploy, site_num, sep="_"))

#make pairing for 4 treatments
mass.deploy.merged.df <- mass.deploy.merged.df %>%
  mutate(pairing = paste(treatment, hfa, sep = "-"))

mass.deploy.merged.df <- mass.deploy.merged.df %>% 
  mutate(pairing = factor(pairing, levels = c("P-H", "P-AW", "AB-H", "AB-AW")))

mass_aov1=aov(z_score ~ treatment * hfa, data = na.omit(mass.deploy.merged.df))
mass_tHSD=TukeyHSD(mass_aov1, ordered = F, conf.level = 0.95)
cld <- multcompLetters4(mass_aov1, mass_tHSD)

# table with factors and 3rd quantile
mass.table <- mass.deploy.merged.df%>%
  dplyr::group_by(treatment, hfa) %>%
  dplyr::summarise(mean=mean(z_score), quant = quantile(z_score, probs = 0.75)) %>%
  dplyr::arrange(desc(mean))

# extracting the compact letter display and adding to the Tk table
cld <- as.data.frame.list(cld$`treatment:hfa`)
mass.table$cld <- cld$Letters

z_contrast.mass.plot=mass.deploy.merged.df%>%
  ggplot(aes(x=interaction(treatment, hfa, sep = "-"), y = z_score, fill = interaction(treatment, hfa, sep = "-")))+
  geom_boxplot(alpha = 0.5)+
     geom_point(aes(shape = as.factor(site_num)), position = position_jitter(width = 0.2)) +
  xlab("")+
  ylab("% mass loss (z scored)")+
    labs(shape = "site", fill = "treatment") +  # Change "Your New Legend Title" to your desired title
  scale_fill_colorblind()+
  geom_hline(data = mass.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted", size= 1) +
    geom_text(data = mass.table, aes(y = quant, label = cld), color = "blue", fontface= "bold",size = 5, vjust=-1.3, hjust =-0.76)+
  Fig.formatting


ggsave("figures/z_contrast.mass.plot.pdf", z_contrast.mass.plot, width = 10, height = 6)

mass_aov2=aov(z_score ~ treatment * hfa, data = na.omit(mass.deploy.merged.df))



mass_lmer1=lmer(perc.mass_loss~ treatment * hfa + (1 | tree_id) + (1|tree_num), data = na.omit(mass.deploy.merged.df))
mass_lmer2=lmer(z_score~ treatment * hfa + (1 | tree_id/tree_num), data = na.omit(mass.deploy.merged.df))
mass_lm1=lm(z_score ~ treatment + hfa , data = na.omit(mass.deploy.merged.df))

AIC(aov2, lm1)
anova(lm1)
plot(resid(mass_lm1))

AIC( mass_aov2,mass_lmer1, mass_lmer2 )

plot(mass_lmer2)
plot(resid(mass_lmer2))

```



Let's first look at the P only to see if HFA is there
```{r}
###P treatment only to see HFA effects
#create a proper "site" column for the figure

#plot
ggplot(mass.deploy.merged.df, aes(x = treatment, y = z_score)) +
  geom_boxplot(alpha= 0.3) +
  geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.2)) +
  labs(x = "", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_shape_manual(values = c(7,10), name= "site") +
  geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

ggplot(mass.deploy.merged.df, aes(x = interaction(treatment,hfa), y = perc.mass_loss), group = interaction(treatment, hfa)) +
  geom_boxplot(alpha= 0.3) +
  geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.2)) +
  labs(x = "", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_shape_manual(values = c(7,10), name= "site") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

#visually, it seems like home shows overall slightly greater mass loss than away. This can be verified by some stats.


# Assuming your dataframe is mass.deployed.merged.df
t <- 19  # time in days
mt<- mass.deploy.merged.df$final_correct  # final corrected mass
m0 <- mass.deploy.merged.df$initial_mass  # initial mass

# Calculate k
k <- -1 / t * log(m_t / m_0)

# Calculate decay constant (k)
mass.deploy.merged.df$k= -1/t * log(mt/m0)



ggplot(mass.deploy.merged.df, aes(x = interaction(treatment,hfa), y = k, group = interaction(treatment, hfa), color = interaction (treatment, hfa))) +
  geom_boxplot(alpha= 0.3) +
  geom_point(aes(shape = interaction(treatment, hfa)), position = position_jitter(width = 0.2)) +
  labs(x = "treatment", y = "decomposition rate k (/day)") +
  theme(legend.position = "right") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

# Assuming you have initial and final values
initial_value = mass.deploy.merged.df$initial_mass  # Replace with your initial value
final_value <- mass.deploy.merged.df$final_correct  # Replace with your final value

# Assuming the time difference is known
time_difference <- 19  # Replace with your time difference

# Calculate decay constant (k)
k <- -log(final_value / initial_value) / time_difference
k
# Generate time values for the plot
time_values <- seq(0, time_difference, by = 1)

# Calculate N(t) using the exponential decay formula
decay_values <- initial_value * exp(-k * time_values)

# Plot the exponential decay function
plot(time_values, decay_values, type = "l", xlab = "Time", ylab = "Quantity", main = "Exponential Decay")

# Add points for the initial and final values
points(0, initial_value, col = "red", pch = 16)
points(time_difference, final_value, col = "blue", pch = 16)

# Add legend
legend("topright", legend = c("Exponential Decay", "Initial Value", "Final Value"),
       col = c("black", "red", "blue"), pch = c(NA, 16, 16))


```

Analysis for **present only** dataframe
To be analyze the mass loss data, arcsine square root transforming proportional mass loss is preferred. This are outlined below...

perform lmer and add result to P treatment only plot
Lmer 3 (see below) is the model that best fits the data
```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
#check levels of HFA factors
levels(mass.present.deploy.merged.df$hfa)

#models:
lm1 <- lm(z_score ~  hfa, data = mass.present.deploy.merged.df)
summary(lm1)
glm1 <- glm(z_score ~ hfa, data = mass.present.deploy.merged.df)
summary(lm1)

lmer1 <- lmer(z_score ~ hfa + (1 | tree_id), data= mass.present.deploy.merged.df)
summary(lmer1)

lmer2 <- lmer(z_score ~ tree_id * deploy + (1 | tree_num), data= mass.present.deploy.merged.df)
summary(lmer2)
lmer3 <- lmer(z_score ~ hfa + (1 | tree_num) + (1 | tree_id), data= mass.present.deploy.merged.df)

lmer4 <- lmer(z_score ~ hfa + (1 | tree_id /  tree_num), data= mass.present.deploy.merged.df)
summary(lmer5)

# lmer3 is best fit
lmer3 <- lmer(z_score ~ hfa + (1 | tree_num) + (1 | tree_id), data= mass.present.deploy.merged.df)
summary(lmer3)
plot(resid(lmer3))
qqnorm(resid(lmer3))
qqline(resid(lmer3))


aov1 <- aov(z_score ~ hfa, data= mass.present.deploy.merged.df)
summary(aov1)
plot(aov1)
qqnorm(aov1)
qqline(aov1)

AIC(aov1, lmer3)


```


Generic t-test for P only treatment
```{r}

# the t test will use z score values
#combined site
mass.present.deploy.merged.df$combined_site <- paste("site", mass.present.deploy.merged.df$site_num, sep = "")

stat.test.z <- mass.present.deploy.merged.df %>% 
  t_test(z_score ~ hfa, var.equal = T) %>%
  add_significance()

stat.test.arc <- mass.present.deploy.merged.df %>% 
  t_test(arcsine_trans~ hfa, var.equal = T) %>%
  add_significance()

stat.test.z

#check levels
levels(mass.present.deploy.merged.df$hfa)


# Create the box plot with original elements as ggplot from above
leaf.z.presentonly.plot=  ggboxplot(
  mass.present.deploy.merged.df, x = "hfa", y = "z_score", 
  ylab = "mass loss (z score)", xlab = "", fill="hfa", alpha= 0.4) + 
  scale_fill_manual(values = c("grey", "grey0"), guide = "none")+
  stat_pvalue_manual(stat.test.arc, y.position = 2.5, tip.length = c(0.16, 0.06)) +
  labs(subtitle = get_test_label(stat.test.arc, detailed = T)) +
    scale_x_discrete(labels = c("home", "away"))+
    geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.2))+
    scale_shape_manual(values = c(7,10), name= "site")+
    geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16),
          plot.background = element_blank()) 

leaf.z.presentonly.plot

# Create the box plot with original elements as ggplot from above
leaf.mass.presentonly.plot= ggboxplot(
  mass.present.deploy.merged.df, x = "hfa", y = "perc.mass_loss", 
  ylab = "mass loss (%)", xlab = "", fill="hfa", alpha= 0.4) + 
  scale_fill_manual(values = c("grey", "grey0"), guide = "none")+
  stat_pvalue_manual(stat.test.arc, y.position = 60, tip.length = c(0.06, 0.16)) +
  labs(subtitle = get_test_label(stat.test.arc, detailed = T)) +
    scale_x_discrete(labels = c("home", "away"))+
    geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.1))+
    scale_shape_manual(values = c(7,10), name= "site")+
    geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(perc.mass_loss), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16),
          plot.background = element_blank()) 

leaf.mass.presentonly.plot

ggsave("figures/leaf.z.presentonly.plot.pdf", leaf.z.presentonly.plot, width = 8, height = 6)


```

Analysis for **all treatment** dataframe
There are two ways to go about model exploration: using z scored mass loss or arcsine square root transforming percent mass loss. These are outlined below...

first, plot to visualize the 4 treatments
```{r}
# Combine treatment and hfa into one group
mass.deploy.merged.df$combined_group <- paste(mass.deploy.merged.df$treatment, mass.deploy.merged.df$hfa, sep = "-")

# Create a custom order for the combined groups
combined_group_order <- c("P-H", "P-A","AB-H", "AB-A")
custom_palette <- c("#2CA02C",  "#D62728", "#1F77B4","#FF7F0E")


# Plot with combined groups but no combo site
ggplot(mass.deploy.merged.df, aes(x = combined_group, y = z_score, fill = combined_group)) +
  geom_boxplot(alpha= 0.3) +
  scale_x_discrete(limits = combined_group_order)+
  geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.2), alpha = 0.7) +
  labs(x = "treatment", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_fill_manual(values = custom_palette, guide = "none") +
  scale_shape_manual(values = c(15, 16), name ="site") +
  geom_hline(data = mass.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting



```

model exploration of all treatments: ph, pa, abh, aba; lmer 7 is best fit
```{r}
lmer4 <- lmer(perc.mass_loss ~ treatment*hfa + (1| tree_id/tree_num), data=na.omit(mass.deploy.merged.df))
lmer5 <- lmer(arcsine_z_score ~ treatment*hfa + (1|tree_num) + (1|tree_id), data=mass.deploy.merged.df)

lmer5 <- lmer(arcsine_z_score ~ combined_group + (1|tree_num) + (1|tree_id), data=mass.deploy.merged.df)
lmer6 <- lmer(z_score ~ combined_group + (1|tree_num) + (1|tree_id), data=mass.deploy.merged.df)
lmer7 <- lmer(z_score ~ treatment*hfa + (1|tree_num) + (1|tree_id), data=mass.deploy.merged.df)

AIC(lmer4, lmer7)
summary(lmer7)
# Type III ANOVA for fixed effects
Anova(lmer7, type="III")

library(emmeans)
# Post-hoc tests for interactions (if significant)
interaction_emm <- emmeans(lmer7, specs = ~treatment:hfa)
contrast(interaction_emm, method = "pairwise")

# Fit your linear mixed-effects model
lmer7 <- lmer(z_score ~ treatment * hfa + (1 | tree_num) + (1 | tree_id), data = mass.deploy.merged.df)


# Calculate Cohen's d for specific group comparisons
cohen_d <- with(mass.deploy.merged.df, {
  d1 <- (mean(z_score[treatment == "P" & hfa == "H"]) - mean(z_score[treatment == "P" & hfa == "A"])) / sqrt((var(z_score[treatment == "P" & hfa == "H"]) + var(z_score[treatment == "P" & hfa == "A"])) / 2)
  d2 <- (mean(z_score[treatment == "AB" & hfa == "H"]) - mean(z_score[treatment == "AB" & hfa == "A"])) / sqrt((var(z_score[treatment == "AB" & hfa == "H"]) + var(z_score[treatment == "AB" & hfa == "A"])) / 2)
  d3 <- (mean(z_score[treatment == "P" & hfa == "H"]) - mean(z_score[treatment == "AB" & hfa == "H"])) / sqrt((var(z_score[treatment == "P" & hfa == "H"]) + var(z_score[treatment == "AB" & hfa == "H"])) / 2)
  d4 <- (mean(z_score[treatment == "P" & hfa == "A"]) - mean(z_score[treatment == "AB" & hfa == "A"])) / sqrt((var(z_score[treatment == "P" & hfa == "A"]) + var(z_score[treatment == "AB" & hfa == "A"])) / 2)
  data.frame(Group = c("P-H vs. P-A", "AB-H vs. AB-A", "P-H vs. AB-H", "P-A vs. AB-A"),
             Cohen_d = c(d1, d2, d3, d4))
})

# Print the Cohen's d values
print(cohen_d)

install.packages("MuMIn")
library(MuMIn)
r.squaredGLMM(lmer7)
```


This is the proper all treatment dataframe for nonparametric test
```{r}
install.packages("dunn.test")
library(dunn.test)
kw1 <- kruskal.test(z_score ~ combined_group, data = mass.deploy.merged.df)
kw1
# Perform post-hoc Dunn's test
posthoc <- dunn_test(z_score ~ combined_group, data = mass.deploy.merged.df, method = "bonferroni")# Create a function to format p-values and add stars for significance
format_p_value <- function(p_value) {
  if (p_value < 0.001) {
    return("***")
  } else if (p_value < 0.01) {
    return("**")
  } else if (p_value < 0.05) {
    return("*")
  } else {
    return("ns")
  }
}

# Apply the formatting function to the p-values
dunn.test.z.res$formatted_p_value <- sapply(dunn.test.z.res$p.adj, format_p_value)
##this plot works
# Create a custom order for the combined groups
combined_group_order <- c("P-H", "P-A","AB-H", "AB-A")
library(ggpubr)

custom_palette <- c("#D62728","#FF7F0E", "#2CA02C", "#1F77B4")
levels(as.factor(mass.deploy.merged.df$combined_group))

leaf.mass.all.plot=ggboxplot(
  mass.deploy.merged.df, x = "combined_group", y = "z_score", 
  ylab = "mass loss (z score)", xlab = "", fill="combined_group", alpha= 0.4) + 
    scale_x_discrete(limits = combined_group_order)+
  scale_fill_viridis_d( guide = "none")+
  stat_pvalue_manual(dunn.test.z.res, step.increase = 0.03, step.group.by = "group2", tip.length = 0.01) +
    labs( x= "treatment", y="mass loss (z score)") +
  geom_point(aes(shape = deploy), position = position_jitter(width = 0.2), alpha = 0.7) +
  scale_shape_manual(values = c(15, 16), name ="site") +
     geom_hline(data = mass.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16),
          plot.background = element_blank(),
          axis.text.x = element_text(angle = 22, hjust = 0.5, vjust = 0.5)) 

leaf.mass.all.plot

count(mass.deploy.merged.df$combined_group)

range(mass.deploy.merged.df$perc.mass_loss)

ggsave("figures/leaf.mass.all.plot.pdf", leaf.mass.all.plot, width = 8, height = 6)

```


Analysis for **absent only** dataframe
There are two ways to go about model exploration: using z scored mass loss or arcsine square root transforming percent mass loss. These are outlined below...

perform t-test and add result to AB treatment only plot
```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
#plot
mass.absent.deploy.merged.df$combined_site <- paste("site", mass.absent.deploy.merged.df$site_num, sep = "")

leaf.mass.absentonly.plot=ggplot(mass.absent.deploy.merged.df, aes(x = hfa, y = z_score)) +
  geom_boxplot(alpha= 0.3) +
  geom_point(aes(shape = as.factor(combined_site)), position = position_jitter(width = 0.2)) +
  labs(x = "", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_x_discrete(labels = c("away", "home")) +
  scale_fill_manual(values = custom_palette) +
  scale_shape_manual(values = c(7,10), name= "site") +
  geom_hline(data = mass.absent.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

leaf.mass.absentonly.plot
# Create a histogram
hist_data <- hist(mass.absent.deploy.merged.df$z_score, plot = FALSE)

# Create a density plot
density_data <- density(mass.absent.deploy.merged.df$z_score)

# Plot histogram with density overlay
hist(mass.absent.deploy.merged.df$z_score, freq = F, main = "")
lines(density_data, col = "red")
  #normal

shapiro.test(mass.absent.deploy.merged.df$z_score)
  # p= 0.39, normal

# Q-Q Plots
qqnorm(mass.absent.deploy.merged.df$z_score, main = "Q-Q Plot")
qqline(mass.absent.deploy.merged.df$z_score)
  #normal


# Levene's Test
leveneTest(z_score ~ hfa, data = mass.absent.deploy.merged.df)
  # p= 0.63, variances between H and A are equal (i.e., the homogeneity assumption of the variance is met)


# the t test will use z score values
stat.ab.test.z <- mass.absent.deploy.merged.df %>% 
  t_test(z_score ~ hfa, var.equal = T) %>%
  add_significance()
stat.ab.test.z

# Create the box plot with original elements as ggplot from above
ggboxplot(
  mass.absent.deploy.merged.df, x = "hfa", y = "z_score", 
  ylab = "mass loss (z score)", xlab = "") + 
  stat_pvalue_manual(stat.ab.test.z, tip.length = 0) +
  labs(subtitle = get_test_label(stat.ab.test.z, detailed = TRUE)) +
    scale_x_discrete(labels = c("away", "home"))+
    geom_point(aes(shape = as.factor(combined_site)), position = position_jitter(width = 0.2))+
    scale_shape_manual(values = c(7,10), name= "site")+
    geom_hline(data = mass.absent.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16)) 

leaf.mass.presentonly.plot

ggsave("figures/leaf.mass.presentonly.plot.pdf", leaf.mass.presentonly.plot, width = 10, height = 8)


```


**Exploration of nonparametric models**
Non-parametric tests are more robust to deviations from normality and heteroscedasticity, but they do have their own assumptions. For example, the Mann-Whitney U test assumes that the distributions of the two groups are similar in shape. The Kruskal-Wallis test assumes that the distributions of all groups are similar and that the differences in the distributions are due to location shifts. 


The Kruskal-Wallis test can be an alternative. This test is an extension of the Wilcoxon rank-sum test for comparing more than two independent groups.

```{r}
# Perform Kruskal-Wallis test
kw_test <- kruskal.test(z_score ~ hfa, data = mass.present.deploy.merged.df)

# Calculate effect size (Eta-squared) from Kruskal-Wallis test
H <- kw_test$statistic
df <- kw_test$parameter
N <- sum(table(mass.present.deploy.merged.df$hfa))
eta_squared <- (H - df) / (N - df)

# Print effect size
cat("Effect size (Eta-squared) for Kruskal-Wallis test:", eta_squared, "\n")

#The calculated effect size (Eta-squared) for the Kruskal-Wallis test is -0.01813634. Eta-squared is a measure of the proportion of variance explained by a factor in non-parametric tests, similar to how R-squared works in linear regression. The negative value suggests a very small effect, but it's important to note that Eta-squared can range from 0 to 1, with higher values indicating larger effects. In this case, the small negative value indicates that the factor (hfa) is not explaining much variance in the data.


```


----------------- START OF HFAI CALCULATIONS ----------------------------------
**Background information regarding Home-field Advantage:**

The Home Field Advantage Index (HFAI) derived from Ayres et al. (2009) uses the relative mass loss values (rml) of species A and B at different sites. The HFAI therefore represents the difference in the rate of litter decomposition between two different sites ("Hoko" and "Sekiu") for both species A and B. Species A and B, in this case, are alder trees of different origin. For example, species A is Hoko alder and species B is Sekiu alder. There are 10 alder trees per river, representing biological replicates. So, the HFAI quantifies the *advantage* or *disadvantage* of each "species" in their respective home environment compared to an away environment.

Interpreting HFAI values involves considering their sign and magnitude:

**Positive** HFAI Values:
Positive values indicate that both species A and B experience faster decomposition rates at their home site compared to the away site. This suggests that both species are better adapted to their home environment in terms of litter decomposition.

**Negative** HFAI Values:
Negative values indicate that both species A and B experience slower decomposition rates at their home site (HF disadvantage (HFDA)) compared to the away site. This suggests that both species might not be well-suited for decomposition in their own environment and may decompose more efficiently in a different environment.

**Magnitude** of HFAI:
The magnitude of the HFAI values indicates the strength of the advantage or disadvantage. Larger absolute values (positive or negative) suggest a greater difference in decomposition rates between the home and away sites.


Calculating HFA Index (HFAI) from Ayres et al. (2009):

There are two equations from which HFAI is derived: the relative mass loss (rml) of species decomposing at home and away site, and the scaling equation which allows for interpretaibility of the home-field advantage. Below is a brief summary of the components found in the derivations:

Aa: This represents the percent mass loss of leaf litter from species A (HOK) when it's decomposing at site a (HOK, or the area dominated by species A). This value is given as a percentage.

Ba: This represents the percent mass loss of leaf litter from species B (SEK) when it's decomposing at site a This value is also given as a percentage.

(Aa + Ba): This is the sum of the percent mass loss of leaf litter from both species A and B when they're decomposing at site a, or HOK and SEK leaf packs at HOK.

(Aa / (Aa + Ba) * 100): This is the full equation that calculates the proportion of the mass loss contributed by species A at site a, relative to the **total mass** loss by both species A and B combined at that site. This percentage value indicates how much of the total mass loss at site a is due to species A.

A similar equation is calculated for species B decomposing at site a. The variables are substituted accordingly.

Armla and Armlb; Brlma, Brmlb: These are the relative mass loss values for species A and B, respectively, when they are decomposing at site a (dominated by species A) or site b (dominated by species B). These values are expressed as percentages of total mass loss at their respective sites.

(((ARMLa + BRMLb)^2 / (ARMLb + BRMLa)^2) * 100)-100: This is the full equation that calculates the squared ratio of the combined relative mass loss of species A at site a and species B at site b (i.e., home), divided by the squared ratio of the combined relative mass loss of species A at site b and species B at site a (i.e., away).

100 is subtracted from the value obtained. This is done to measure the percentage difference in mass loss at the home site (site dominated by the species) compared to the away site (site dominated by the other species). If the value is positive, it indicates that the litter decomposes faster at its home site, showing a home-field advantage. If the value is negative, it suggests that decomposition is slower at the home site, indicating a home-field disadvantage.

The equation calculates the Home-Field Advantage Index (HFAI) by comparing the relative mass loss of litter from two species at their home and away sites. The result is expressed as a percentage, indicating how much faster (or slower) the decomposition occurs at the home site compared to the away site.


NEW: use this! This is a now a proper test dataframe with for loop!

NOTICE: THIS IS FOR SITE 1:

generate **present**, site a (HOK 1)
```{r}
########################
mass.hok1.present.df <- mass.hok1.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Create a new column combining tree_num and ayres with a period in between
mass.hok1.present.df$ayres.tree <- paste(mass.hok1.present.df$ayres, mass.hok1.present.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Armla
names_vector_armla <- character(0)
values_vector_armla <- numeric(0)
treatment_vector_armla <- character(0)


# Initialize empty vectors to store names and values for Brmla
names_vector_brmla <- character(0)
values_vector_brmla <- numeric(0)
treatment_vector_brmla <- character(0)

# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Aa_column <- paste("Aa.", X, sep = "")
    Ba_column <- paste("Ba.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    armla_result <- ((mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Aa_column, "perc.mass_loss"]) / 
                     (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Aa_column, "perc.mass_loss"] + 
                      mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Ba_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    brmla_result <- ((mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Ba_column, "perc.mass_loss"]) / 
                     (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Ba_column, "perc.mass_loss"] + 
                      mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Aa_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    armla_name <- paste("Armla", X, Y, sep = ".")
    brmla_name <- paste("Brmla", X, Y, sep = ".")

     # Extract the treatment from mass.hok1.present.df for the current combination
    treatment_name <- mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Aa_column, "treatment"]
    

    # Append the names and values to the respective vectors for Armla
    names_vector_armla <- c(names_vector_armla, armla_name)
    values_vector_armla <- c(values_vector_armla, armla_result)
    treatment_vector_armla <- c(treatment_vector_armla, treatment_name)

    
    # Append the names and values to the respective vectors for Brmla
    names_vector_brmla <- c(names_vector_brmla, brmla_name)
    values_vector_brmla <- c(values_vector_brmla, brmla_result)
    treatment_vector_brmla <- c(treatment_vector_brmla, treatment_name)


  }
}

# Create dataframes for Armla and Brmla
armla_df <- data.frame(names = names_vector_armla, values = values_vector_armla, treatment= treatment_vector_armla)
brmla_df <- data.frame(names = names_vector_brmla, values = values_vector_brmla, treatment=treatment_vector_brmla)

#merge the dataframes
armla.brmla.df_1=rbind(armla_df, brmla_df)
armla.brmla.df_1
```


NEW: use this! This is a now a proper test dataframe with for loop!
generate **absent**, site a (HOK 1)
```{r}
########################
mass.hok1.absent.df <- mass.hok1.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Create a new column combining tree_num and ayres with a period in between
mass.hok1.absent.df$ayres.tree <- paste(mass.hok1.absent.df$ayres, mass.hok1.absent.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values<- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Armla
names_vector_armla<- character(0)
values_vector_armla<- numeric(0)
treatment_vector_armla <- character(0)


# Initialize empty vectors to store names and values for Brmla
names_vector_brmla <- character(0)
values_vector_brmla <- numeric(0)
treatment_vector_brmla <- character(0)

# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Aa_column <- paste("Aa.", X, sep = "")
    Ba_column <- paste("Ba.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    armla_result <- (mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Aa_column, "perc.mass_loss"] / 
                     (mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Aa_column, "perc.mass_loss"] + 
                      mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Ba_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    brmla_result <- (mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Ba_column, "perc.mass_loss"] / 
                     (mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Ba_column, "perc.mass_loss"] + 
                      mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Aa_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    armla_name <- paste("Armla", X, Y, sep = ".")
    brmla_name <- paste("Brmla", X, Y, sep = ".")
    
    # Extract the treatment from mass.hok1.present.df for the current combination
    treatment_name <- mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Aa_column, "treatment"]
    
    
    # Append the names and values to the respective vectors for Armla
    names_vector_armla <- c(names_vector_armla, armla_name)
    values_vector_armla<- c(values_vector_armla, armla_result)
        treatment_vector_armla <- c(treatment_vector_armla, treatment_name)

    
    # Append the names and values to the respective vectors for Brmla
    names_vector_brmla <- c(names_vector_brmla, brmla_name)
    values_vector_brmla <- c(values_vector_brmla, brmla_result)
        treatment_vector_brmla <- c(treatment_vector_brmla, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
armla_ab_df <- data.frame(names = names_vector_armla, values = values_vector_armla, treatment = treatment_vector_armla)

brmla_ab_df <- data.frame(names = names_vector_brmla, values = values_vector_brmla, treatment = treatment_vector_brmla)

# View the resulting dataframes
print(armla_df)
print(brmla_df)

armla.brmla.absent.df_1=rbind(armla_ab_df, brmla_ab_df)
armla.brmla.absent.df_1

```


NEW: use this! This is a now a proper test dataframe with for loop!
generate **present**, site a (SEK 1)
```{r}
mass.sek1.present.df <- mass.sek1.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  # Create a new column combining tree_num and ayres with a period in between
mass.sek1.present.df$ayres.tree <- paste(mass.sek1.present.df$ayres, mass.sek1.present.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Brmlb
names_vector_brmlb <- character(0)
values_vector_brmlb <- numeric(0)
treatment_vector_brmlb <- character(0)

# Initialize empty vectors to store names and values for Armlb
names_vector_armlb <- character(0)
values_vector_armlb <- numeric(0)
treatment_vector_armlb <- character (0)

# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Bb_column <- paste("Bb.", X, sep = "")
    Ab_column <- paste("Ab.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    brmlb_result <- (mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Bb_column, "perc.mass_loss"] / 
                     (mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Bb_column, "perc.mass_loss"] + 
                      mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Ab_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    armlb_result <- (mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Ab_column, "perc.mass_loss"] / 
                     (mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Ab_column, "perc.mass_loss"] + 
                      mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Bb_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    brmlb_name <- paste("Brmlb", X, Y, sep = ".")
    armlb_name <- paste("Armlb", X, Y, sep = ".")
    
    # Extract the treatment from mass.hok1.present.df for the current combination
    treatment_name <- mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Ab_column, "treatment"]
    
    
    # Append the names and values to the respective vectors for Brmlb
    names_vector_brmlb <- c(names_vector_brmlb, brmlb_name)
    values_vector_brmlb <- c(values_vector_brmlb, brmlb_result)
    treatment_vector_brmlb <- c(treatment_vector_brmlb, treatment_name)
    
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_armlb <- c(names_vector_armlb, armlb_name)
    values_vector_armlb <- c(values_vector_armlb, armlb_result)
        treatment_vector_armlb <- c(treatment_vector_armlb, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
brmlb_df <- data.frame(names = names_vector_brmlb, values = values_vector_brmlb, treatment = treatment_vector_brmlb)
armlb_df <- data.frame(names = names_vector_armlb, values = values_vector_armlb, treatment = treatment_vector_armlb)

# View the resulting dataframes
print(brmlb_df)
print(armlb_df)

brmlb.armlb.df_1=rbind(brmlb_df, armlb_df)

brmlb.armlb.df_1
```

NEW: use this! This is a now a proper test dataframe with for loop!
generate **absent**, site a (SEK 1)
```{r}
mass.sek1.absent.df <- mass.sek1.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  # Create a new column combining tree_num and ayres with a period in between
mass.sek1.absent.df$ayres.tree <- paste(mass.sek1.absent.df$ayres, mass.sek1.absent.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values_3 <- 1:5
Y_values_3 <- 1:5

# Initialize empty vectors to store names and values for Brmlb
names_vector_brmlb <- character(0)
values_vector_brmlb <- numeric(0)
treatment_vector_brmlb <- character (0)

# Initialize empty vectors to store names and values for Armlb
names_vector_armlb <- character(0)
values_vector_armlb <- numeric(0)
treatment_vector_armlb <- character (0)


# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values_3) {
  for (Y in Y_values_3) {
    Bb_column <- paste("Bb.", X, sep = "")
    Ab_column <- paste("Ab.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    brmlb_result <- (mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Bb_column, "perc.mass_loss"] / 
                     (mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Bb_column, "perc.mass_loss"] + 
                      mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Ab_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    armlb_result <- (mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Ab_column, "perc.mass_loss"] / 
                     (mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Ab_column, "perc.mass_loss"] + 
                      mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Bb_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    brmlb_name <- paste("Brmlb", X, Y, sep = ".")
    armlb_name <- paste("Armlb", X, Y, sep = ".")
    
     # Extract the treatment from mass.hok1.present.df for the current combination
    treatment_name <- mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Ab_column, "treatment"]
    
    
    
    # Append the names and values to the respective vectors for Brmlb
    names_vector_brmlb <- c(names_vector_brmlb, brmlb_name)
    values_vector_brmlb <- c(values_vector_brmlb, brmlb_result)
    treatment_vector_brmlb <- c(treatment_vector_brmlb, treatment_name)
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_armlb <- c(names_vector_armlb, armlb_name)
    values_vector_armlb <- c(values_vector_armlb, armlb_result)
        treatment_vector_armlb <- c(treatment_vector_armlb, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
brmlb_ab_df <- data.frame(names = names_vector_brmlb, values = values_vector_brmlb, treatment = treatment_vector_brmlb)
armlb_ab_df <- data.frame(names = names_vector_armlb, values = values_vector_armlb, treatment = treatment_vector_armlb)


brmlb.armlb.absent.df_1=rbind(brmlb_ab_df, armlb_ab_df)
brmlb.armlb.absent.df_1

```


NEW: use this! merging HOK 1 and SEK 1 dataframe
merge **present**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site1.df=rbind(armla.brmla.df_1, brmlb.armlb.df_1)
mass.present.site1.df
# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Create an empty dataframe to store the results
result_df_1 <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla.", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla.", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb.", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb.", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.present.site1.df[mass.present.site1.df$names == armla_name, "values"] +
      mass.present.site1.df[mass.present.site1.df$names == brmlb_name, "values"])/2)/ (
      (mass.present.site1.df[mass.present.site1.df$names == armlb_name, "values"] +
      mass.present.site1.df[mass.present.site1.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
    result_name <- LETTERS[(X - 1) * 5 + Y]
    
  # Get the treatment for the current combination
    treatment_name <- mass.present.site1.df[mass.present.site1.df$names == armla_name, "treatment"]
 
# Append the name and value to the result dataframe
    result_df_1 <- rbind(result_df_1, data.frame(names = result_name, values = result, treatment= treatment_name))  }
}


# View the resulting dataframe
result_df_1
```


NEW: use this! merging HOK 1 and SEK 1 dataframe
merge **absent**, merging HOK 1 and SEK 1 dataframe
```{r}
#merge both present site 1 dataframes
mass.absent.site1.df=rbind(armla.brmla.absent.df_1, brmlb.armlb.absent.df_1)

# Create an empty dataframe to store the results
result_ab_df_1 <- data.frame(names = character(0), values = numeric(0), treatment = character (0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla.", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla.", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb.", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb.", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.absent.site1.df[mass.absent.site1.df$names == armla_name, "values"] +
      mass.absent.site1.df[mass.absent.site1.df$names == brmlb_name, "values"])/2)/ (
      (mass.absent.site1.df[mass.absent.site1.df$names == armlb_name, "values"] +
      mass.absent.site1.df[mass.absent.site1.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
result_name <- paste0(LETTERS[(X - 1) * 5 + Y], "_1")

# Get the treatment for the current combination
    treatment_name <- mass.absent.site1.df[mass.absent.site1.df$names == armla_name, "treatment"]
    

    # Append the name and value to the result dataframe
    result_ab_df_1 <- rbind(result_ab_df_1, data.frame(names = result_name, values = result, treatment= treatment_name))
  }
}


# View the resulting dataframe
result_ab_df_1


```


combine P and AB; plot check for P and AB site 1 
```{r}
#combine sites 1 and 2
hfai.allsite1.df=rbind(result_df_1, result_ab_df_1)

ggplot(hfai.allsite1.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  geom_bar(stat = "identity", position= "dodge", alpha = 0.5) +
  scale_fill_colorblind()+
  xlab("reciprocated pairs") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.x = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  )

```


NOTICE: THIS IS NOW SITE 2:

NEW: use this! This is a now a proper test dataframe with for loop!
generate **present**, site a (HOK 2)
```{r}
########################
mass.hok2.present.df <- mass.hok2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)
# Create a new column combining tree_num and ayres with a period in between
mass.hok2.present.df$ayres.tree <- paste(mass.hok2.present.df$ayres, mass.hok2.present.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Armla
names_vector_armla <- character(0)
values_vector_armla <- numeric(0)
treatment_vector_armla<- character(0)

# Initialize empty vectors to store names and values for Brmla
names_vector_brmla <- character(0)
values_vector_brmla<- numeric(0)
treatment_vector_brmla<- character(0)


# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Aa_column <- paste("Aa.", X, sep = "")
    Ba_column <- paste("Ba.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    armla_result <- (mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Aa_column, "perc.mass_loss"] / 
                     (mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Aa_column, "perc.mass_loss"] + 
                      mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Ba_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    brmla_result <- (mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Ba_column, "perc.mass_loss"] / 
                     (mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Ba_column, "perc.mass_loss"] + 
                      mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Aa_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    armla_name <- paste("Armla", X, Y, sep = ".")
    brmla_name <- paste("Brmla", X, Y, sep = ".")
    
    # Get the treatment for the current combination
    treatment_name <- mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Aa_column, "treatment"]
    
    # Append the names and values to the respective vectors for Armla
    names_vector_armla <- c(names_vector_armla, armla_name)
    values_vector_armla <- c(values_vector_armla, armla_result)
    treatment_vector_armla<- c(treatment_vector_armla, treatment_name)

    # Append the names and values to the respective vectors for Brmla
    names_vector_brmla <- c(names_vector_brmla, brmla_name)
    values_vector_brmla <- c(values_vector_brmla, brmla_result)
        treatment_vector_brmla<- c(treatment_vector_brmla, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
armla_df_2 <- data.frame(names = names_vector_armla, values = values_vector_armla, treatment = treatment_vector_armla)
brmla_df_2 <- data.frame(names = names_vector_brmla, values = values_vector_brmla, treatment = treatment_vector_brmla)



# View the resulting dataframes
print(armla_df_2)
print(brmla_df_2)

armla.brmla.df_2=rbind(armla_df_2, brmla_df_2)


```


NEW: use this! This is a now a proper test dataframe with for loop!
generate **absent**, site a (HOK 2)
```{r}
########################
mass.hok2.absent.df <- mass.hok2.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Create a new column combining tree_num and ayres with a period in between
mass.hok2.absent.df$ayres.tree <- paste(mass.hok2.absent.df$ayres, mass.hok2.absent.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Armla
names_vector_armla <- character(0)
values_vector_armla <- numeric(0)
treatment_vector_armla <- character(0)

# Initialize empty vectors to store names and values for Brmla
names_vector_brmla <- character(0)
values_vector_brmla <- numeric(0)
treatment_vector_brmla <- character(0)


# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Aa_column <- paste("Aa.", X, sep = "")
    Ba_column <- paste("Ba.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    armla_result <- (mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Aa_column, "perc.mass_loss"] / 
                     (mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Aa_column, "perc.mass_loss"] + 
                      mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Ba_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    brmla_result <- (mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Ba_column, "perc.mass_loss"] / 
                     (mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Ba_column, "perc.mass_loss"] + 
                      mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Aa_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    armla_name <- paste("Armla", X, Y, sep = ".")
    brmla_name <- paste("Brmla", X, Y, sep = ".")
    
    # create treatment 
    
    treatment_name <- mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Aa_column, "treatment"]
    
    # Append the names and values to the respective vectors for Armla
    names_vector_armla <- c(names_vector_armla, armla_name)
    values_vector_armla <- c(values_vector_armla, armla_result)
    treatment_vector_armla <- c(treatment_vector_armla, treatment_name)
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_brmla <- c(names_vector_brmla, brmla_name)
    values_vector_brmla <- c(values_vector_brmla, brmla_result)
        treatment_vector_brmla <- c(treatment_vector_brmla, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
armla_ab_df_2 <- data.frame(names = names_vector_armla, values = values_vector_armla, treatment = treatment_vector_armla)
brmla_ab_df_2 <- data.frame(names = names_vector_brmla, values = values_vector_brmla, treatment = treatment_vector_brmla)

# View the resulting dataframes
print(armla_ab_df_2)
print(armla_ab_df_2)

armla.brmla.absent.df_2=rbind(armla_ab_df_2, brmla_ab_df_2)
armla.brmla.absent.df_2

```

NEW: use this! This is a now a proper test dataframe with for loop!
generate **present**, site a (SEK 2)
```{r}
mass.sek2.present.df <- mass.sek2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  # Create a new column combining tree_num and ayres with a period in between
mass.sek2.present.df$ayres.tree <- paste(mass.sek2.present.df$ayres, mass.sek2.present.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Brmlb
names_vector_brmlb <- character(0)
values_vector_brmlb <- numeric(0)
treatment_vector_brmlb <- character(0)


# Initialize empty vectors to store names and values for Armlb
names_vector_armlb <- character(0)
values_vector_armlb <- numeric(0)
treatment_vector_armlb <- character(0)

# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Bb_column <- paste("Bb.", X, sep = "")
    Ab_column <- paste("Ab.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    brmlb_result <- (mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Bb_column, "perc.mass_loss"] / 
                     (mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Bb_column, "perc.mass_loss"] + 
                      mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Ab_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    armlb_result <- (mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Ab_column, "perc.mass_loss"] / 
                     (mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Ab_column, "perc.mass_loss"] + 
                      mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Bb_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    brmlb_name <- paste("Brmlb", X, Y, sep = ".")
    armlb_name <- paste("Armlb", X, Y, sep = ".")
    
    # create treatment
    treatment_name <- mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Bb_column, "treatment"]

    
    # Append the names and values to the respective vectors for Brmlb
    names_vector_brmlb <- c(names_vector_brmlb, brmlb_name)
    values_vector_brmlb <- c(values_vector_brmlb, brmlb_result)
    treatment_vector_brmlb<- c(treatment_vector_brmlb, treatment_name)
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_armlb <- c(names_vector_armlb, armlb_name)
    values_vector_armlb <- c(values_vector_armlb, armlb_result)
        treatment_vector_armlb<- c(treatment_vector_armlb, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
brmlb_df_2 <- data.frame(names = names_vector_brmlb, values = values_vector_brmlb, treatment = treatment_vector_brmlb)
armlb_df_2 <- data.frame(names = names_vector_armlb, values = values_vector_armlb, treatment = treatment_vector_armlb)

# View the resulting dataframes
print(brmlb_df_2)
print(armlb_df_2)

brmlb.armlb.df_2=rbind(brmlb_df_2, armlb_df_2)
brmlb.armlb.df_2

```

NEW: use this! This is a now a proper test dataframe with for loop!
generate **absent**, site a (SEK 2)
```{r}
mass.sek2.absent.df <- mass.sek2.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  # Create a new column combining tree_num and ayres with a period in between
mass.sek2.absent.df$ayres.tree <- paste(mass.sek2.absent.df$ayres, mass.sek2.absent.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values<- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Brmlb
names_vector_brmlb <- character(0)
values_vector_brmlb <- numeric(0)
treatment_vector_brmlb<- character(0)

# Initialize empty vectors to store names and values for Armlb
names_vector_armlb <- character(0)
values_vector_armlb <- numeric(0)
treatment_vector_armlb<- character(0)


# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Bb_column <- paste("Bb.", X, sep = "")
    Ab_column <- paste("Ab.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    brmlb_result <- (mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Bb_column, "perc.mass_loss"] / 
                     (mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Bb_column, "perc.mass_loss"] + 
                      mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Ab_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    armlb_result <- (mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Ab_column, "perc.mass_loss"] / 
                     (mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Ab_column, "perc.mass_loss"] + 
                      mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Bb_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    brmlb_name <- paste("Brmlb", X, Y, sep = ".")
    armlb_name <- paste("Armlb", X, Y, sep = ".")
    
    # create treatment 
    treatment_name <- mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Bb_column, "treatment"]
    
    
    # Append the names and values to the respective vectors for Brmlb
    names_vector_brmlb <- c(names_vector_brmlb, brmlb_name)
    values_vector_brmlb <- c(values_vector_brmlb, brmlb_result)
    treatment_vector_brmlb<- c(treatment_vector_brmlb, treatment_name)
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_armlb <- c(names_vector_armlb, armlb_name)
    values_vector_armlb <- c(values_vector_armlb, armlb_result)
        treatment_vector_armlb<- c(treatment_vector_armlb, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
brmlb_ab_df_2 <- data.frame(names = names_vector_brmlb, values = values_vector_brmlb, treatment = treatment_vector_brmlb)
armlb_ab_df_2 <- data.frame(names = names_vector_armlb, values = values_vector_armlb, treatment = treatment_vector_armlb)


brmlb.armlb.absent.df_2=rbind(brmlb_ab_df_2, armlb_ab_df_2)
brmlb.armlb.absent.df_2

```

NEW: use this! merging HOK 2 and SEK 2 dataframe
merge **present**, site a and b (aka HOK 2 and SEK 2) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site2.df=rbind(armla.brmla.df_2, brmlb.armlb.df_2)
mass.present.site2.df
# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Create an empty dataframe to store the results
result_df_2 <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla.", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla.", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb.", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb.", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.present.site2.df[mass.present.site2.df$names == armla_name, "values"] +
      mass.present.site2.df[mass.present.site2.df$names == brmlb_name, "values"])/2)/ (
      (mass.present.site2.df[mass.present.site2.df$names == armlb_name, "values"] +
      mass.present.site2.df[mass.present.site2.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
    result_name <- paste0(LETTERS[(X - 1) * 5 + Y], "_2")
    
  # Get the treatment for the current combination
    treatment_name <- mass.present.site2.df[mass.present.site2.df$names == armla_name, "treatment"]
 
# Append the name and value to the result dataframe
    result_df_2 <- rbind(result_df_2, data.frame(names = result_name, values = result, treatment= treatment_name))  }
}


# View the resulting dataframe

print(result_df_2)



```


NEW: use this! merging HOK 2 and SEK 2 dataframe
merge **absent**, merging HOK 2 and SEK 2 dataframe
```{r}
#merge both present site 1 dataframes
mass.absent.site2.df=rbind(armla.brmla.absent.df_2, brmlb.armlb.absent.df_2)
mass.absent.site2.df
# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Create an empty dataframe to store the results
result_ab_df_2 <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla.", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla.", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb.", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb.", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.absent.site2.df[mass.absent.site2.df$names == armla_name, "values"] +
      mass.absent.site2.df[mass.absent.site2.df$names == brmlb_name, "values"])/2)/ (
      (mass.absent.site2.df[mass.absent.site2.df$names == armlb_name, "values"] +
      mass.absent.site2.df[mass.absent.site2.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
    result_name <- paste0(LETTERS[(X - 1) * 5 + Y], "_3")
    
  # Get the treatment for the current combination
    treatment_name <- mass.absent.site2.df[mass.absent.site2.df$names == armla_name, "treatment"]
 
# Append the name and value to the result dataframe
    result_ab_df_2 <- rbind(result_ab_df_2, data.frame(names = result_name, values = result, treatment= treatment_name))  }
}


# View the resulting dataframe
result_ab_df_2

```


NEW: use this! merging all dataframes (Sites 1 and 2; includes P and AB)
merge all sites; HFAI plots
```{r}
#combine sites 1 and 2 including P and AB
hfai.allsites.df=rbind(result_df_1, result_df_2, result_ab_df_1, result_ab_df_2)
hfai.allsites.df$treatment=as.factor(hfai.allsites.df$treatment)

```

P only HFAI analysis
```{r}
####################  P ONLY ###################
#filter P treatment only for HFA effect
hfai.present.allsites.df=hfai.allsites.df%>%
  filter(treatment == "P")

# Filter the dataframe to include only positive cases
positive_present_df <- hfai.present.allsites.df[hfai.present.allsites.df$values > 0, ]

# Calculate the mean and standard error (SE)
mean_hfa <- mean(positive_present_df$values)
se_hfa <- sd(positive_present_df$values) / sqrt(nrow(positive_present_df))
# Perform a t-test
t_test_present_result <- t.test(positive_present_df$values, mu = 0)
t_test_present_result



# Format the t-test result
t_test_label <- paste("p =", formatC(t_test_present_result$p.value, format = "e", digits = 2), ",", "t-test =", round(t_test_present_result$statistic, 2))
# Format the t-test result
p_value <- t_test_present_result$p.value
formatted_p_value <- if (p_value < 0.0001) {
  "p < 0.0001"
} else {
  paste("p =", formatC(p_value, digits = 1))
}


#positive and negative values
total_positive <- sum(hfai.present.allsites.df$values > 0)
total_negative <- sum(hfai.present.allsites.df$values < 0)
# Calculate the percentage of positive values relative to negative
percentage_positive <- (total_positive / (total_positive + total_negative)) * 100
percentage_negative <- (total_negative / (total_negative + total_positive)) * 100
# 72% positive
# 28% negative
```

Plot P only
```{r}
#new HFA fig format
hfa.fig.formatting <- theme_classic() +
  theme(
    text = element_text(size = 25),
    legend.text.align = 0,
    panel.border = element_rect(fill = NA, colour = "black", size = 1),
    aspect.ratio = 1,
    axis.ticks.length.y = unit(0.25, "cm"),
    axis.text.y = element_text(margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour = "black", size = 20),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()  # Add this line to remove x-axis ticks
  ) +
  theme(legend.key.size = unit(0.4, "cm"), legend.position = "right") +
  theme(aspect.ratio = 0.4)+
  theme( plot.margin = margin(r = 1 ))


#just P treatment
hfai.present.allsites.plot=ggplot(hfai.present.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
scale_fill_manual(values = c("#E69F00"))+
geom_bar(stat = "identity", position ="dodge", alpha = 0.5) +
  xlab("") +
  ylab("% faster decomposition at home") +
  geom_vline(xintercept = 14.5, linetype = "dashed", color = "red")+
  scale_x_discrete(expand=c(0.02,0.02))+
  scale_y_continuous(limits = c(-30, 255), breaks = seq(-30, 255, by = 30))+
  hfa.fig.formatting

hfai.present.allsites.plot

ggsave("figures/hfai.present.allsites.plot.pdf", hfai.present.allsites.plot, width = 14, height = 6)

#just P treatment with coord_flip
ggplot(hfai.present.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  scale_fill_manual(values = "grey")+
  geom_bar(stat = "identity", position ="dodge", alpha = 0.5) +
  coord_flip()+
  xlab("") +
  ylab("% faster decomposition at home") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+
  geom_vline(xintercept = 14.5, linetype = "dashed", color = "red") +
scale_y_continuous(limits = c(-20, 70), breaks = seq(-20, 60, by = 10))




```

P and AB HFAI analysis
```{r}
######################### BOTH P and AB #########################################

# Filter the dataframe to include only positive cases
positive_values <- hfai.allsites.df %>%
  filter(values > 0)

#calcualte mean HFA for P
mean_positive_P <- mean(positive_values[positive_values$treatment == "P",]$values)
se_positive_P <- sd(hfai.allsites.df[hfai.allsites.df$treatment == "P",]$values) / sqrt(nrow(hfai.allsites.df[hfai.allsites.df$treatment == "P",]))
# mean = 45.09, se = 9.51

#t test for P
t_test_P <- t.test(positive_values[positive_values$treatment == "P",]$values, mu = 0)

#calcualte mean HFA for AB
mean_positive_AB <- mean(positive_values[positive_values$treatment == "AB",]$values)
se_positive_AB <- sd(hfai.allsites.df[hfai.allsites.df$treatment == "AB",]$values) / sqrt(nrow(hfai.allsites.df[hfai.allsites.df$treatment == "AB",]))
# mean = 14.22, se = 3.64

#t test for AB
t_test_AB <- t.test(positive_values[positive_values$treatment == "AB",]$values, mu = 0)

t_test_P_AB <- t.test(mean_positive_P,mean_positive_AB )

#another way to look at the means
hfai.allsites.df %>%
filter(values > 0) %>%
  summarise(mean = mean(values), se = sd(values) / sqrt(n()))

hfai.allsites.df %>%
  filter(values < 0) %>%
  summarise(mean = mean(values))

hfai.allsites.df%>%
  filter(treatment == "P", values <0)%>%
n_distinct()


# Calculate the proportion of positive values for the "P" treatment relative to all positive values
proportion_positive_P <- sum(hfai.allsites.df$treatment == "P" & hfai.allsites.df$values > 0) /
                        sum(hfai.allsites.df$values > 0)
#65% HFA for P
proportion_positive_AB <- sum(hfai.allsites.df$treatment == "AB" & hfai.allsites.df$values > 0) /
                        sum(hfai.allsites.df$values > 0)
#35% HFA for AB


# Perform a t-test for total HFA
t_test_present_result <- t.test(positive_values_subset$values, mu = 0)
t_test_present_result


##### proportion that are negative ##############

# Filter the dataframe to include only positive cases
neg_values <- hfai.allsites.df %>%
  filter(values < 0)

## Calculate the proportion of positive values for the "P" treatment relative to all positive values
proportion_neg_P <- sum(hfai.allsites.df$treatment == "P" & hfai.allsites.df$values < 0) /
                        sum(hfai.allsites.df$values < 0)

proportion_pos_P <- sum(hfai.allsites.df$treatment == "P" & hfai.allsites.df$values > 0) /
                        sum(hfai.allsites.df$values > 0)

#31% HFA for P
proportion_neg_AB <- sum(hfai.allsites.df$treatment == "AB" & hfai.allsites.df$values < 0) /
                        sum(hfai.allsites.df$values < 0)
#68% HFA for AB


mean_ab <- mean(positive_values_subset[positive_values_subset$treatment == "AB", "values"])
med_ab <- median(positive_values_subset[positive_values_subset$treatment == "AB", "values"])
se_ab <- sd(positive_values_subset[positive_values_subset$treatment == "AB", "values"]) / sqrt(sum(positive_values_subset$treatment == "AB"))

mean_p <- mean(positive_values_subset[positive_values_subset$treatment == "P", "values"])
med_p <- median(positive_values_subset[positive_values_subset$treatment == "P", "values"])
se_p <- sd(positive_values_subset[positive_values_subset$treatment == "P", "values"]) / sqrt(sum(positive_values_subset$treatment == "P"))


# Perform a t-test
t_test_result <- t.test(positive_values_subset[positive_values_subset$treatment == "P", "values"], mu = 0)
t_test_result <- t.test(positive_values_subset[positive_values_subset$treatment == "AB", "values"], mu = 0)

d <- (mean_positive_P - mean_positive_AB) / sqrt((se_positive_P^2 + se_positive_AB^2) / 2)
d

cohensD(positive_values[positive_values$treatment == "P",]$values ~ positive_values[positive_values$treatment == "AB",]$values)
#positive and negative values
total_pos_P <- sum(hfai.allsites.df$values > 0 & hfai.allsites.df$treatment =="P")
total_pos_AB <- sum(hfai.allsites.df$values > 0 & hfai.allsites.df$treatment =="AB")

total_neg_P <- sum(hfai.allsites.df$values < 0 & hfai.allsites.df$treatment =="P")

total_neg_AB <- sum(hfai.allsites.df$values < 0 & hfai.allsites.df$treatment =="AB")


total <- sum(hfai.allsites.df$values > 0)
total_neg <- sum(hfai.allsites.df$values < 0)

# Calculate the percentage of positive values relative to negative
percentage_pos_P <- (total_pos_P / ( total)) * 100
percentage_pos_AB <- (total_pos_AB / ( total)) * 100

percentage_neg_P <- (total_neg_P / ( total_neg)) * 100
percentage_eg_AB <- (total_neg_AB / ( total_neg)) * 100

percentage_negative <- (total_negative / (total_negative + total_positive)) * 100

# 53% positive
# 47% negative

library(lsr)
cohensD((hfai.allsites.df$values > 0 & hfai.allsites.df$treatment =="P") ~ (hfai.allsites.df$values > 0 & hfai.allsites.df$treatment =="AB"))

t.test((hfai.allsites.df$values > 0 & hfai.allsites.df$treatment =="P"), (hfai.allsites.df$values > 0 & hfai.allsites.df$treatment =="AB"))


```

Plot
```{r}
#all sites
View(hfai.allsites.df)

hfai.allsites.plot=hfai.allsites.df%>%
  filter(!is.na(values)) %>%
  ggplot(aes(x = reorder(names, values), y = values, fill = treatment, width = 0.85)) +
scale_fill_colorblind()+
geom_bar(stat = "identity", position ="dodge", alpha = 0.5) +
  xlab("") +
  ylab("% faster decomposition at home") +
  theme_minimal() +
  scale_x_discrete(expand = c(0.01,0.01))+
  geom_vline(xintercept = 44.5 , linetype = "dashed", color = "red")+
scale_y_continuous(limits = c(-70, 255), breaks = seq(-70, 255, by = 35)) +
  hfa.fig.formatting

hfai.allsites.plot

ggsave("figures/hfai.allsites.plot.pdf", hfai.allsites.plot, width = 14, height = 6)




```


```{r}
#both P and AB treatments
hfai.allsites.plot=hfai.allsites.df%>%
  filter(!is.na(values)) %>%
  ggplot(aes(x = reorder(names, values), y = values, fill = treatment)) +
scale_fill_colorblind()+
geom_bar(stat = "identity", position ="dodge", alpha = 0.5) +
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(plot.margin = margin(r = 0, l = 0))+
  scale_x_discrete(expand = c(0,0))+
  geom_vline(xintercept = 45.5 , linetype = "dashed", color = "red")+
scale_y_continuous(limits = c(-50, 100), breaks = seq(-50, 100, by = 20))+
  hfa.fig.formatting


hfai.allsites.plot
hfai.allsites.plot
min(hfai.allsites.df$values) 
max(hfai.allsites.df$values)


ggplot(aes(x = reorder(names, values), y = values, fill = treatment)) +
  geom_bar(stat = "identity", position ="dodge", alpha = 0.3) +
  scale_fill_manual(values = c("grey", "grey1"))+
  coord_flip()+
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 42, linetype = "dashed", color = "red")+
  scale_y_continuous(
    limits = c(-65, 255),
     breaks = seq(-65, 250, by = 65),
   
  )

hfai.allsites.plot_flipped=ggplot(hfai.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  geom_bar(stat = "identity", position ="dodge", alpha = 0.4) +
  scale_fill_manual(values = c("grey", "grey0"))+
  coord_flip()+
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 46.5, linetype = "dashed", color = "red")
hfai.allsites.plot_flipped

max(hfai.allsites.df[hfai.allsites.df$treatment =="AB" & hfai.allsites.df$values >0, "values"])

ggsave("figures/hfai.allsites.plot_flipped.pdf", hfai.allsites.plot_flipped, width = 14, height = 10)


```


Alder secondary metabolites from LC-MS
### Import LC-MS/MS data
```{r}
#installing and calling the necessary packages:
library(mixOmics)
library(ggpubr)
library(vegan)

### load negative mode
neg_data <- read_csv("/Users/Cody1/Desktop/Projects/RAD Alder/LC-MS/RAD_negative_mode_11_3_2023_quant_copy.csv") %>% arrange(Sample)
View(neg_data)
neg_metadata <- read_csv("/Users/Cody1/Desktop/Projects/RAD Alder/LC-MS/RAD_prelim_metadata_neg.csv") %>% arrange(Sample)
all(neg_data$Sample == neg_metadata$Sample)

### load positive mode
pos_data <- read_csv("/Users/Cody1/Desktop/Projects/RAD Alder/LC-MS/RAD_positive_mode_11_3_2023_quant_copy.csv") %>% arrange(Sample)
View(pos_data)
pos_metadata <- read_csv("/Users/Cody1/Desktop/Projects/RAD Alder/LC-MS/RAD_prelim_metadata_pos.csv") %>% arrange(Sample)
all(pos_data$Sample == pos_metadata$Sample)
```


Analyze positive mode
```{r}
# Calculate the interquartile range (IQR) for each feature
pos.iqr_values <- apply(pos_data[, -c(1:5)], 2, IQR)

# Determine the cutoff threshold based on the 40% filter-out percentage
pos.filter_threshold <- quantile(pos.iqr_values, 0.4)

# Identify features to be filtered out
pos.features_to_filter <- names(pos.iqr_values[pos.iqr_values > pos.filter_threshold])

# Filter out the identified features
fltr.pos_data <- pos_data[, !(names(pos_data) %in% pos.features_to_filter)]

# Define the columns to be transformed
pos.col_trans <- names(fltr.pos_data)[2:ncol(fltr.pos_data)]

# Perform Sample normalization (Normalization by median)
pos_normalized <- fltr.pos_data %>%
  mutate(across(all_of(pos.col_trans), ~./median(.)))

# Perform Log transformation (base 10)
pos_log_transformed <- pos_normalized %>%
  mutate(across(all_of(pos.col_trans), ~log10(.)))

# Perform Auto scaling (mean-centered and divided by the standard deviation of each variable)
pos_trans.df <- pos_log_transformed %>%
  mutate(across(all_of(pos.col_trans), ~scale(.)))

# Assuming your DataFrame is named df, and the columns to be melted are MS_1, MS_2, ..., MS_9
df_long <- pos_trans.df %>%
  pivot_longer(cols = starts_with("MS"), names_to = "id", values_to = "intensity")
head(df_long)

df_long %>%
filter(Sample == "30_Pos.mzML") %>%
  dplyr::summarise(number_of_ids = n_distinct(id))


# Display the head of the long-format DataFrame
head(df_long)
str(df_long)



min(df_long$intensity)


# Now, you can plot the data
# Example plot for the first MS column
filtered_df_long <- df_long[df_long$id %in% unique(df_long$id[1:30]), ]

ggplot(filtered_df_long, aes(x = intensity, y = id)) +
  geom_boxplot()

range(df_long$intensity)

```


Another way to analyze
```{r}
#using random forest data from metaboanalyst for sig MS ID
tree_rand_frst.df <- read_csv("/Users/Cody1/Desktop/Projects/RAD Alder/LC-MS/random_forest_tree.csv")
type_rand_frst.df <- read_csv("/Users/Cody1/Desktop/Projects/RAD Alder/LC-MS/random_forest_type.csv")

View(rand_frst.df)

#filter top metabolites
tree_rand_frst.df%>%
  filter(MeanDecreaseAccuracy > 0.0009)%>%
  n_distinct()
#top 50 metabolites

#filter 
type_rand_frst.df%>%
  filter(MeanDecreaseAccuracy > 0.0027)%>%
  n_distinct()
#top 40

#create new subset df
fltr.tree_rand_frst.df=tree_rand_frst.df%>%
  filter(MeanDecreaseAccuracy > 0.0009)

#create new subset df
fltr.type_rand_frst.df=type_rand_frst.df%>%
  filter(MeanDecreaseAccuracy > 0.0027)

#list them
list(fltr.tree_rand_frst.df$Feature_Index)

list(fltr.type_rand_frst.df$Feature_Index)


#merge peak data and metadata
df <- merge(pos_metadata, pos_data, by = "Sample")

# keep top 50 
tree.columns_to_keep <- c("Sample", "Attribute_type", "Attribute_TreeID", "Attribute_Sitenum", "Attribute_Treenum",
                      "MS_3998", "MS_3995", "MS_2909" ,"MS_478"  ,"MS_2911" ,"MS_2716" ,"MS_2720", "MS_2913" ,"MS_405" , "MS_2915" ,"MS_2905",
"MS_477"  ,"MS_2918", "MS_2917" ,"MS_2907", "MS_2916", "MS_2735", "MS_2887" ,"MS_678" , "MS_2910", "MS_2724" ,"MS_2728",
"MS_2726", "MS_2904" ,"MS_3808", "MS_1286", "MS_2886", "MS_2908" ,"MS_1050" ,"MS_3029", "MS_3025" ,"MS_2733" ,"MS_2884",
"MS_636" , "MS_2729" ,"MS_3039", "MS_3024", "MS_2715", "MS_2906" ,"MS_3034" ,"MS_679" , "MS_2714" ,"MS_2892", "MS_3027",
"MS_3650", "MS_2721" ,"MS_3953", "MS_2736", "MS_2717", "MS_2450")

# keep top 54 
type.columns_to_keep <- c("Sample", "Attribute_type", "Attribute_TreeID", "Attribute_Sitenum", "Attribute_Treenum",
                      "MS_1702", "MS_1167", "MS_1693", "MS_3914", "MS_3130", "MS_1700", "MS_1000" ,"MS_1701", "MS_1691", "MS_1672", "MS_1709",
"MS_1372", "MS_3915" ,"MS_1166", "MS_1215" ,"MS_1350", "MS_3774" ,"MS_1681" ,"MS_1313", "MS_1689", "MS_1315" ,"MS_4098",
"MS_1676", "MS_1007" ,"MS_1680" ,"MS_3778" ,"MS_3777" ,"MS_1699" ,"MS_1161" ,"MS_1004" ,"MS_1171", "MS_3918" ,"MS_1534",
"MS_1172", "MS_1393", "MS_1151" ,"MS_1687", "MS_1373" ,"MS_1399" ,"MS_1157")

# Remove columns not in the specified list
tree.filtered_df <- df[, c(tree.columns_to_keep)]
head(filtered_df)

type.filtered_df=df[, c(type.columns_to_keep)]
head(type.filtered_df)

# Extract the numeric columns for PCA
tree.numeric_cols <- tree.filtered_df[, 6:ncol(tree.filtered_df)]

type.numeric_cols <- type.filtered_df[, 6:ncol(type.filtered_df)]

# Run PCA
tree.pca_result <- prcomp(tree.numeric_cols, scale. = TRUE)

type.pca_result <- prcomp(type.numeric_cols, scale. = TRUE)


#tree
ggplot() +
  geom_point(data = as.data.frame(tree.pca_result$x), 
             aes(x = PC1, y = PC2, color = tree.filtered_df$Attribute_TreeID, shape = as.factor(tree.filtered_df$Attribute_Sitenum))) +
  labs(title = "Top 40: PCA Biplot",
       x = paste0("PC1 (", round(tree.pca_result$sdev[1] / sum(tree.pca_result$sdev) * 100, 2), "%)"),
       y = paste0("PC2 (", round(tree.pca_result$sdev[2] / sum(tree.pca_result$sdev) * 100, 2), "%)")) +
  theme_minimal()+
  stat_ellipse(data = as.data.frame(tree.pca_result$x), 
               aes(x = PC1, y = PC2, color = as.factor(tree.filtered_df$Attribute_TreeID), fill = as.factor(tree.filtered_df$Attribute_TreeID)),
               show.legend =F,
               geom = "polygon", 
               alpha = 0.2,
               level = 0.95) +
  theme_minimal() +
  theme(panel.grid = element_blank(), 
        panel.border = element_rect(fill = "transparent"))

#type
ggplot() +
  geom_point(data = as.data.frame(type.pca_result$x), 
             aes(x = PC1, y = PC2, color = type.filtered_df$Attribute_type)) +
  labs(title = "Top 40: PCA Biplot",
       x = paste0("PC1 (", round(type.pca_result$sdev[1] / sum(type.pca_result$sdev) * 100, 2), "%)"),
       y = paste0("PC2 (", round(type.pca_result$sdev[2] / sum(type.pca_result$sdev) * 100, 2), "%)")) +
  theme_minimal()+
  stat_ellipse(data = as.data.frame(type.pca_result$x), 
               aes(x = PC1, y = PC2, color = as.factor(type.filtered_df$Attribute_type), fill = as.factor(type.filtered_df$Attribute_type)),
               show.legend =F,
               geom = "polygon", 
               alpha = 0.2,
               level = 0.95) +
  theme_minimal() +
  theme(panel.grid = element_blank(), 
        panel.border = element_rect(fill = "transparent"))




```





Look at individual ms
```{r}

# Select relevant columns
type.filtered_df <- type.filtered_df[, c("Sample", "Attribute_type", "MS_1702")]

# Reshape data for plotting
ms_data_long <- tidyr::gather(type.filtered_df, key = "MS_Column", value = "Peak_Intensity", -c(Sample, Attribute_type))

# Plot using ggplot
ggplot(ms_data_long, aes(x = Attribute_type, y = Peak_Intensity, fill = Attribute_type)) +
  geom_boxplot(position = "identity") +
  labs(title = "Peak Intensity by Attribute Type",
       x = "Attribute Type",
       y = "Peak Intensity") +
  theme_minimal()











```




```{r}
# Assuming you have a data frame containing the loadings of MS columns in each principal component
loadings_df <- as.data.frame(pca_result$rotation[, 1:2])

type.loadings_df <- as.data.frame(type.pca_result$rotation[, 1:2])


# Assuming you have a list of MS column names
ms_columns <- c("MS_3998", "MS_3995", "MS_2909" ,"MS_478"  ,"MS_2911" ,"MS_2716" ,"MS_2720", "MS_2913" ,"MS_405" , "MS_2915" ,"MS_2905",
"MS_477"  ,"MS_2918", "MS_2917" ,"MS_2907", "MS_2916", "MS_2735", "MS_2887" ,"MS_678" , "MS_2910", "MS_2724" ,"MS_2728",
"MS_2726", "MS_2904" ,"MS_3808", "MS_1286", "MS_2886", "MS_2908" ,"MS_1050" ,"MS_3029", "MS_3025" ,"MS_2733" ,"MS_2884",
"MS_636" , "MS_2729" ,"MS_3039", "MS_3024", "MS_2715", "MS_2906" ,"MS_3034" ,"MS_679" , "MS_2714" ,"MS_2892", "MS_3027",
"MS_3650", "MS_2721" ,"MS_3953", "MS_2736", "MS_2717", "MS_2450")

type.ms_columns <- c("MS_1702", "MS_1167", "MS_1693", "MS_3914", "MS_3130", "MS_1700", "MS_1000" ,"MS_1701", "MS_1691", "MS_1672", "MS_1709",
"MS_1372", "MS_3915" ,"MS_1166", "MS_1215" ,"MS_1350", "MS_3774" ,"MS_1681" ,"MS_1313", "MS_1689", "MS_1315" ,"MS_4098",
"MS_1676", "MS_1007" ,"MS_1680" ,"MS_3778" ,"MS_3777" ,"MS_1699" ,"MS_1161" ,"MS_1004" ,"MS_1171", "MS_3918" ,"MS_1534",
"MS_1172", "MS_1393", "MS_1151" ,"MS_1687", "MS_1373" ,"MS_1399" ,"MS_1157")

# Find the corresponding loadings for MS columns
ms_loadings <- loadings_df[rownames(loadings_df) %in% ms_columns, ]

type.ms_loadings <- type.loadings_df[rownames(type.loadings_df) %in% type.ms_columns, ]


# Perform PCA
type.pca_result <- prcomp(type.loadings_df[, -c(1:5)], scale. = TRUE)

# Plot PCA
ggplot() +
  geom_point(data = as.data.frame(pca_result$x), 
             aes(x = PC1, y = PC2, color = df$Attribute_TreeID)) +
  labs(title = "PCA Biplot",
       x = paste0("PC1 (", round(pca_result$sdev[1] / sum(pca_result$sdev) * 100, 2), "%)"),
       y = paste0("PC2 (", round(pca_result$sdev[2] / sum(pca_result$sdev) * 100, 2), "%)")) +
  theme_minimal()


#looking at tree ID
ggplot() +
  geom_point(data = as.data.frame(pca_result$x), 
             aes(x = PC1, y = PC2, color = df$Attribute_type)) +
  labs(title = "Top 50 features: PCA Biplot",
       x = paste0("PC1 (", round(pca_result$sdev[1] / sum(pca_result$sdev) * 100, 2), "%)"),
       y = paste0("PC2 (", round(pca_result$sdev[2] / sum(pca_result$sdev) * 100, 2), "%)")) +
  stat_ellipse(data = as.data.frame(pca_result$x), 
               aes(x = PC1, y = PC2, color = as.factor(df$Attribute_type), fill = as.factor(df$Attribute_type)),
               show.legend =F,
               geom = "polygon", 
               alpha = 0.2,
               level = 0.95) +
  theme_minimal() +
  theme(panel.grid = element_blank(), 
        panel.border = element_rect(fill = "transparent"))

```



```{r}
# Generate negative mode PCA raw
neg_PCA_whole <- mixOmics::pca(column_to_rownames(neg_data, var = "Sample"), ncomp = 2, scale = TRUE)
neg_PCA_whole_scores<- data.frame(neg_PCA_whole$variates$X, neg_metadata)

#### negative PCA plot for treatment (can also change for tree ID)
neg_PCA_whole_scores%>%
  ggscatter(x = "PC1", y = "PC2", color = "Attribute_type", alpha = 0.6, 
            title = paste("PCA RAW -", "Attribute_type", sep = " "),
            xlab = paste("PC1 (", round(neg_PCA_whole$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(neg_PCA_whole$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = neg_PCA_whole_scores %>% group_by(Attribute_type) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = Attribute_type), size = 4, shape = 8) +
  stat_ellipse(aes(PC1, PC2, color = Attribute_type), geom = "polygon", level = 0.95, alpha = 0.2) +  # Add ellipses
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()

# PERMANOVA
neg_dist_metabolites <- vegdist(column_to_rownames(neg_data, var = "Sample"), method = "euclidean")
neg_permanova <- adonis2(neg_dist_metabolites ~ neg_PCA_whole_scores$Attribute_type, neg_PCA_whole_scores, na.action = na.omit)


###################################


# Generate positive mode PCA raw
pos_PCA_whole <- mixOmics::pca(column_to_rownames(pos_data, var = "Sample"), ncomp = 2, scale = TRUE)
pos_PCA_whole_scores <- data.frame(pos_PCA_whole$variates$X, pos_metadata)

#### positive PCA plot for treatment (can also change for tree ID)
pos_PCA_whole_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = "Attribute_type", alpha = 0.6, 
            title = paste("PCA RAW -", "Attribute_type", sep = " "),
            xlab = paste("PC1 (", round(pos_PCA_whole$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(pos_PCA_whole$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = pos_PCA_whole_scores %>% group_by(Attribute_type) %>%
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = Attribute_type), size = 4, shape = 8)+
  stat_ellipse(aes(PC1, PC2, color = as.factor(Attribute_type)), geom = "polygon", level = 0.95, alpha = 0.2) +  # Add ellipses
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()

pos_PCA_whole_scores$Attribute_Sitenum=as.factor(pos_PCA_whole_scores$Attribute_Sitenum)

pos_PCA_whole_scores %>%
  ggscatter(
    x = "PC1", y = "PC2",
    color = "Attribute_TreeID",
    shape = "Attribute_Sitenum",
    alpha = 0.6,
    title = paste("PCA RAW -", "Attribute_Sitenum", sep = " "),
    xlab = paste("PC1 (", round(pos_PCA_whole$prop_expl_var$X[1] * 100, digits = 1), "%)", sep = ""),
    ylab = paste("PC2 (", round(pos_PCA_whole$prop_expl_var$X[2] * 100, digits = 1), "%)", sep = ""),
    ggtheme = theme_classic()
  ) +geom_point(
    data = pos_PCA_whole_scores %>% 
      group_by(Attribute_TreeID, Attribute_Sitenum) %>% 
      summarise_at(vars(matches("PC")), mean),
    aes(PC1, PC2, color = Attribute_TreeID, shape = Attribute_Sitenum),
    alpha = 0.6
  )+
  stat_ellipse(aes(PC1, PC2, color = Attribute_TreeID, shape = Attribute_Sitenum), geom = "polygon", level = 0.95, alpha = 0.1) +  # Add ellipses
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()


```

Negative mode stats
```{r}

# Generate PCA relative abundance
neg_data_ra <- (column_to_rownames(neg_data, var = "Sample"))/rowSums(column_to_rownames(neg_data, var = "Sample"))
neg_PCA_whole_ra <- mixOmics::pca(neg_data_ra, ncomp = 2, scale = TRUE)
neg_PCA_whole_ra_scores <- data.frame(neg_PCA_whole_ra$variates$X, neg_metadata)

#plot
neg_PCA_whole_ra_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = "Attribute_TreeID", alpha = 0.6, 
            title = paste("PCA RA -", "Attribute_TreeID", sep = " "),
            xlab = paste("PC1 (", round(neg_PCA_whole_ra$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(neg_PCA_whole_ra$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = neg_PCA_whole_ra_scores %>% group_by(Attribute_TreeID) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = Attribute_TreeID), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) +
  stat_ellipse(aes(PC1, PC2, color = Attribute_TreeID), geom = "polygon", level = 0.95, alpha = 0.2) +  # Add ellipses
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()

neg_PCA_whole_ra_scores$Attribute_Sitenum=as.factor(neg_PCA_whole_ra_scores$Attribute_Sitenum)


# plot with site

neg_PCA_whole_ra_scores %>%
  ggscatter(
    x = "PC1", y = "PC2",
    color = "Attribute_TreeID",
    shape = "Attribute_Sitenum",
    alpha = 0.6,
    title = paste("PCA RAW -", "Attribute_Sitenum", sep = " "),
    xlab = paste("PC1 (", round(neg_PCA_whole_ra$prop_expl_var$X[1] * 100, digits = 1), "%)", sep = ""),
    ylab = paste("PC2 (", round(neg_PCA_whole_ra$prop_expl_var$X[2] * 100, digits = 1), "%)", sep = ""),
    ggtheme = theme_classic()
  ) +geom_point(
    data = neg_PCA_whole_ra_scores %>% 
      group_by(Attribute_TreeID, Attribute_Sitenum) %>% 
      summarise_at(vars(matches("PC")), mean),
    aes(PC1, PC2, color = Attribute_TreeID, shape = Attribute_Sitenum),
    alpha = 0.6
  )+
  stat_ellipse(aes(PC1, PC2, color = Attribute_TreeID, shape = Attribute_Sitenum), geom = "polygon", level = 0.95, alpha = 0.1) +  # Add ellipses
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()

library(ggforce)

neg_PCA_whole_ra_scores %>%
  ggscatter(
    x = "PC1", y = "PC2",
    color = "Attribute_TreeID",
    shape = "Attribute_Sitenum",
    alpha = 0.6,
    title = paste("PCA RAW -", "Attribute_Sitenum", sep = " "),
    xlab = paste("PC1 (", round(neg_PCA_whole_ra$prop_expl_var$X[1] * 100, digits = 1), "%)", sep = ""),
    ylab = paste("PC2 (", round(neg_PCA_whole_ra$prop_expl_var$X[2] * 100, digits = 1), "%)", sep = ""),
    ggtheme = theme_classic()
  ) +
  geom_point(
    data = neg_PCA_whole_ra_scores %>% 
      group_by(Attribute_TreeID, Attribute_Sitenum) %>% 
      summarise_at(vars(matches("PC")), mean),
    aes(PC1, PC2, color = Attribute_TreeID, shape = Attribute_Sitenum),
    alpha = 0.6
  ) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()


# PERMANOVA
neg_dist_metabolites_ra <- vegdist(neg_data_ra, method = "euclidean")
neg_permanova_ra <- adonis2(neg_dist_metabolites_ra ~ neg_PCA_whole_ra_scores$Attribute_type, neg_PCA_whole_ra_scores, na.action = na.omit)

############################

#by tree ID
adonis2(neg_dist_metabolites_ra ~ neg_PCA_whole_ra_scores$Attribute_TreeID, neg_PCA_whole_ra_scores, na.action = na.omit)

### Create PLSDA for AB vs No AB in HF diet mice
# Pairwise PLS_DA

data_clr_CT_MPS <- data_clr %>% dplyr::filter(metadata$Attribute_type != "H2O2")
metadata_clr_CT_MPS <- metadata %>% dplyr::filter(metadata$Attribute_type != "H2O2")

# Create PLSDA for AB vs No AB in HF diet mice
PLSDA_CT_MPS <- mixOmics::plsda(data_clr_CT_MPS, metadata_clr_CT_MPS$Attribute_type, ncomp = 3, scale = TRUE)
PLSDA_CT_MPS_scores <- data.frame(PLSDA_CT_MPS$variates$X, metadata_clr_CT_MPS)

numeric_matrix <- data.matrix(neg_data)
neg_data_numeric <- sapply(neg_data, function(x) as.numeric(as.character(x)))

neg_PLSDA<- mixOmics::plsda(neg_data_numeric, neg_metadata$Attribute_type, ncomp = 3, scale = TRUE)
neg_PLSDA_scores <- data.frame(neg_PLSDA$variates$X, neg_metadata)

neg_PLSDA_scores %>%
  ggscatter(x = "comp1", y = "comp2", color = "Attribute_type", alpha = 0.6, title = "PLS-DA AB vs P",
            xlab = paste("Component 1 (", round(neg_PLSDA$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("Component 2 (", round(neg_PLSDA$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            legend.title = "Group", ggtheme = theme_classic()) +
  geom_point(data = neg_PLSDA_scores %>% group_by(Attribute_type) %>% 
               summarise_at(vars(matches("comp")), mean), aes(comp1, comp2, color = Attribute_type), size = 3, shape = 8) +
  theme(plot.title = element_text(size = 10),axis.title = element_text(size = 8),
        axis.text = element_text(size = 5)) + coord_fixed()



perf_neg_plsda <- mixOmics::perf(neg_PLSDA, validation = "Mfold", progressBar = FALSE, auc = TRUE, folds = 4, nrepeat = 100) 

plot(perf_neg_plsda, legend = FALSE)

neg_loadings <- plotLoadings(neg_PLSDA, plot = FALSE, contrib = "max") %>%
  rownames_to_column() %>% dplyr::select(rowname, GroupContrib, importance)

neg_VIPs <- as.data.frame(mixOmics::vip(neg_PLSDA))
neg_VIPs_filter <- dplyr::filter(neg_VIPs, neg_VIPs$comp1 > 1)
neg_VIPs_filter$ID <- rownames(neg_VIPs_filter)
neg_VIPs_select <- neg_VIPs_filter %>% dplyr::select(ID, comp1)
neg_VIPs_Load <- neg_VIPs_select %>% left_join(neg_loadings, by = c("ID" = "rowname")) %>% arrange(desc(comp1))
neg_VIPs_Load_top <- VIPs_CT_MPS_Load %>% head(., 50)

Loadings_CT_MPS_plot <- plotLoadings(neg_loadings, plot = FALSE, contrib = "max") %>%
  arrange(GroupContrib) %>%
  rownames_to_column(var = "Feature") %>%
  dplyr::filter(Feature %in% VIPs_CT_MPS_Load_top$ID) %>%
  ggdotchart(x = "Feature", y = "importance",
             color = "GroupContrib", sorting = "ascending",  
             dot.size = 2, add = "segments",                  
             add.params = list(color = "lightgray", size = 0.1),
             title = "Loadings PLS-DA - CT v MPS") + geom_hline(yintercept = 0, linetype = 2, color = "lightgray")





```


Positive mode stats
```{r}
# PERMANOVA
pos_dist_metabolites <- vegdist(column_to_rownames(pos_data, var = "Sample"), method = "euclidean")
pos_permanova <- adonis2(pos_dist_metabolites ~ pos_PCA_whole_scores$Attribute_type, pos_PCA_whole_scores, na.action = na.omit)

# Generate PCA relative abundance
pos_data_ra <- (column_to_rownames(pos_data, var = "Sample"))/rowSums(column_to_rownames(pos_data, var = "Sample"))
pos_PCA_whole_ra <- mixOmics::pca(pos_data_ra, ncomp = 2, scale = TRUE)
pos_PCA_whole_ra_scores <- data.frame(pos_PCA_whole_ra$variates$X, pos_metadata)

PCA_whole_ra_scores_pos %>%
  ggscatter(x = "PC1", y = "PC2", color = "Attribute_TreeID", alpha = 0.6, 
            title = paste("PCA RA -", "Attribute_TreeID", sep = " "),
            xlab = paste("PC1 (", round(PCA_whole_ra_pos$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_whole_ra_pos$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_whole_ra_scores_pos %>% group_by(Attribute_TreeID) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = Attribute_TreeID), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed() +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) +
  stat_ellipse(aes(PC1, PC2, color = Attribute_TreeID), geom = "polygon", level = 0.95, alpha = 0.2)   # Add ellipses

# PERMANOVA
pos_dist_metabolites_ra <- vegdist(pos_data_ra, method = "euclidean")
pos_permanova_ra <- adonis2(pos_dist_metabolites_ra ~ pos_PCA_whole_ra_scores$Attribute_type, pos_PCA_whole_ra_scores, na.action = na.omit)

#by tree ID
adonis2(pos_dist_metabolites_ra ~ PCA_whole_ra_scores_pos$Attribute_TreeID, PCA_whole_ra_scores_pos, na.action = na.omit)


```

