---
title: "RAD Alder 2023"
author: "Cody Spiegel"
date: "`r Sys.Date()`"
output: word_document
---

### Project overview
For Project ‘RAD’ Alder, gamma irradiated vs. non-irradiated leaf packs deployed at “home” or “away” sites tests the role of phyllosphere microbial communities on home-field advantage patterns in leaf litter decomposition. We utilized a reciprocal transplant experiment between the Hoko and Sekiu rivers on the Olympic Peninsula under four treatment conditions: phyllosphere present-home (PH), phyllosphere present-away (PAW), phyllosphere absent-home (ABH), phyllosphere absent-away (ABAW) to ask the following questions: 

1) How does dispersal from the regional species pool (i.e., source communities) influence community composition on the leaf packs?
2) How does microbial community composition (alpha and beta diversity) and function differ between the leaf packs, as well as change over the course of decomposition?
3) How does alder stoichiometric and secondary chemical composition differ between the leaf packs, as well as change over the course of decomposition?
4) How does microbial metabolization differ between the leaf packs, as well as change over the course of decomposition?
5) How does particulate leaf mass loss differ between the leaf packs? 


**DATA SETS**  
This data set focuses on:  

(1) **metabolism assay** 
(2) **Leaf mass loss** using % mass loss  
(3) **Alder C and N isotopic composotion**
(4) **Alder secondary chemistry (mass speci)** 

```{r setup chunk, setup, include = FALSE, cache=FALSE, message=FALSE, warning=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

# use pacman to load all the packages you are missing!
pacman::p_load('knitr', 'lme4', 'lmerTest', 'tidyverse', 'magrittr', 'effects', 'plyr', 'dplyr', 
               'plotrix', 'car',"gridExtra", "cowplot", "tools", "mgcv", "gratia", "tidymv")

Fig.formatting<-(theme_classic()) +
  theme(text=element_text(size=25),
        legend.text.align = 0,
        panel.border = element_rect(fill=NA, colour = "black", size=1),
        aspect.ratio=1, 
        axis.ticks.length=unit(0.25, "cm"),
        axis.text.y=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20), 
        axis.text.x=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), angle = 45, colour="black", size=20)) +
  theme(legend.key.size = unit(0.4, "cm"), legend.position = "right") +
  theme(aspect.ratio=1.3) 




```


All metabolite decomp days
```{r}
meta.day1.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_1_metabolites_raw_corrected.csv")


meta.day3.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_3_metabolites_raw_corrected.csv")


meta.day5.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_5_metabolites_raw_corrected.csv")


meta.day7.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_7_metabolites_raw_corrected.csv")


meta.day11.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_11_metabolites_raw_corrected.csv")


meta.day17.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_17_metabolites_raw_corrected.csv")



#By z-scoring the raw OD values, I can effectively assess the relative changes in microbial metabolism across the wells, regardless of their absolute starting values.

#z-scoring is a reasonable choice for this situation, as it standardizes the data within each well and facilitates comparisons. Note: z-scoring assumes a roughly symmetric distribution, but the raw OD values are right skewed.

```


metabolite assay: z score by metabolite and deployment 
```{r}
#z score
##hok 1
meta.hok1.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metabolite=="HOK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])
meta.hok1.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metaboltie=="SEK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])


meta.hok1.hok.df$z_score= (meta.hok1.hok.df$od_corrected_b_n - mean(meta.hok1.hok.df$od_corrected_b_n)) / sd(meta.hok1.hok.df$od_corrected_b_n)
meta.hok1.sek.df$z_score= (meta.hok1.sek.df$od_corrected_b_n - mean(meta.hok1.sek.df$od_corrected_b_n)) / sd(meta.hok1.sek.df$od_corrected_b_n)

##sek 1
meta.sek1.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metabolite=="SEK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])
meta.sek1.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metaboltie=="HOK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])

meta.sek1.sek.df$z_score= (meta.sek1.sek.df$od_corrected_b_n - mean(meta.sek1.sek.df$od_corrected_b_n)) / sd(meta.sek1.sek.df$od_corrected_b_n)
meta.sek1.hok.df$z_score= (meta.sek1.hok.df$od_corrected_b_n - mean(meta.sek1.hok.df$od_corrected_b_n)) / sd(meta.sek1.hok.df$od_corrected_b_n)

#hok 2 
meta.hok2.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metabolite=="HOK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])
meta.hok2.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metaboltie=="SEK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])

meta.hok2.hok.df$z_score= (meta.hok2.hok.df$od_corrected_b_n - mean(meta.hok2.hok.df$od_corrected_b_n)) / sd(meta.hok2.hok.df$od_corrected_b_n)
meta.hok2.sek.df$z_score= (meta.hok2.sek.df$od_corrected_b_n - mean(meta.hok2.sek.df$od_corrected_b_n)) / sd(meta.hok2.sek.df$od_corrected_b_n)


##sek 2
meta.sek2.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metabolite=="SEK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])
meta.sek2.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metaboltie=="HOK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])

meta.sek2.sek.df$z_score= (meta.sek2.sek.df$od_corrected_b_n - mean(meta.sek2.sek.df$od_corrected_b_n)) / sd(meta.sek2.sek.df$od_corrected_b_n)
meta.sek2.hok.df$z_score= (meta.sek2.hok.df$od_corrected_b_n - mean(meta.sek2.hok.df$od_corrected_b_n)) / sd(meta.sek2.hok.df$od_corrected_b_n)

# Assuming you have datasets named meta.day1.raw.df, meta.day3.raw.df, ..., meta.day17.raw.df
# Put them into a list
metabolite_list <- list(meta.day1.raw.df, meta.day3.raw.df, meta.day5.raw.df, meta.day7.raw.df, meta.day11.raw.df, meta.day17.raw.df)

# Define the conditions for subsetting
meta.time.points <- c("30", "120", "960", "1440", "2880", "4320")
# Define the actual day numbers corresponding to each dataset
actual_day_numbers <- c(1, 3, 5, 7, 11, 17)

# Create an empty list to store subsets
subset_list <- list()

# Loop through datasets
for (i in seq_along(metabolite_list)) {
  # Access each dataset from the metabolite_list
  meta.merge <- metabolite_list[[i]]
  
  # Create a list to store subsets for each day
  day_subsets <- list()
  
  # Loop through time points
  for (time_point in meta.time.points) {
    # Subset the data based on specific conditions
    subset_data <- meta.merge[meta.merge$deploy == "HOK" &
                             meta.merge$metabolite == "HOK" &
                             meta.merge$site_num == "1" & 
                             meta.merge$time == time_point, ]
    
    # Calculate the z-scores within each subset_data
    subset_data$z_score <- (subset_data$od_corrected_b_n - mean(subset_data$od_corrected_b_n)) / sd(subset_data$od_corrected_b_n)
    
    # Store the subset with z-scores in the day_subsets list
    day_subsets[[time_point]] <- subset_data
  }
  
  # Merge all time points for each day into one dataframe
  merged_data <- bind_rows(day_subsets)
  
  # Store the merged dataframe in the subset_list
  day_number <- actual_day_numbers[i]
  subset_list[[paste0("day", day_number)]] <- merged_data
}

# Rename the list elements to indicate the day
day_names <- paste0("day", actual_day_numbers)
names(subset_list) <- day_names

#rename into new dataframes
meta.day1.df=subset_list$day1
meta.day3.df=subset_list$day3
meta.day5.df=subset_list$day5
meta.day7.df=subset_list$day7
meta.day11.df=subset_list$day11
meta.day17.df=subset_list$day17


```


Day 1: plots
```{r}
#color blind palette
library(viridis) 

day1.z.30.plot=ggplot(meta.day1.df[meta.day1.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
   scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day1.z.120.plot=ggplot(meta.day1.df[meta.day1.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.960.plot=ggplot(meta.day1.df[meta.day1.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.1440.plot=ggplot(meta.day1.df[meta.day1.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.2880.plot=ggplot(meta.day1.df[meta.day1.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day1.z.4320.plot=ggplot(meta.day1.df[meta.day1.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Create the grid with modified plots
extract.legend <- get_legend(
  day1.z.4320.plot + theme(legend.box.margin = margin(1, 1, 1, 1, "cm"),
                       legend.key.size = unit(50, "pt"),
                       legend.text = element_text(size=20),
                       legend.title= element_blank())
)


# Combine plots
all.z.day1.plot <- plot_grid(
  day1.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


all.z.day1.plot
# Set the overall size of the combined plot
ggsave("figures/all.z.day1.plot.pdf", all.z.day1.plot, width = 49, height = 10)



```


Day 3: plots
```{r}

day3.z.30.plot=ggplot(meta.day3.df[meta.day3.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
  scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day3.z.120.plot=ggplot(meta.day3.df[meta.day3.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
     scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.960.plot=ggplot(meta.day3.df[meta.day3.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
     scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.1440.plot=ggplot(meta.day3.df[meta.day3.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.2880.plot=ggplot(meta.day3.df[meta.day3.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day3.z.4320.plot=ggplot(meta.day3.df[meta.day3.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


# Combine plots
all.z.day3.plot <- plot_grid(
  day3.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day3.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day3.plot.pdf", all.z.day3.plot, width = 49, height = 10)



```



Day 5: plots
```{r}

day5.z.30.plot=ggplot(meta.day5.df[meta.day5.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day5.z.120.plot=ggplot(meta.day5.df[meta.day5.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


day5.z.960.plot=ggplot(meta.day5.df[meta.day5.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



day5.z.1440.plot=ggplot(meta.day5.df[meta.day5.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
          scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



day5.z.2880.plot=ggplot(meta.day5.df[meta.day5.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day5.z.4320.plot=ggplot(meta.day5.df[meta.day5.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Combine plots
all.z.day5.plot <- plot_grid(
  day5.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day5.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day5.plot.pdf", all.z.day5.plot, width = 49, height = 10)



```



Day 7: plots
```{r}

day7.z.30.plot=ggplot(meta.day7.df[meta.day7.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
   scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day7.z.30.plot

day7.z.120.plot=ggplot(meta.day7.df[meta.day7.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
     scale_fill_viridis(discrete = TRUE) + 
 geom_hline(data = meta.day7.df[meta.day7.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.120.plot

day7.z.960.plot=ggplot(meta.day7.df[meta.day7.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.960.plot


day7.z.1440.plot=ggplot(meta.day7.df[meta.day7.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.1440.plot

day7.z.2880.plot=ggplot(meta.day7.df[meta.day7.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.2880.plot

day7.z.4320.plot=ggplot(meta.day7.df[meta.day7.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


# Combine plots
all.z.day7.plot <- plot_grid(
  day7.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day7.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day7.plot.pdf", all.z.day7.plot, width = 49, height = 12)



```



Day 11: z scored
```{r}
day11.z.30.plot=ggplot(meta.day11.df[meta.day11.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day11.z.30.plot

day11.z.120.plot=ggplot(meta.day11.df[meta.day11.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day11.z.120.plot

day11.z.960.plot=ggplot(meta.day11.df[meta.day11.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.960.plot


day11.z.1440.plot=ggplot(meta.day11.df[meta.day11.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.1440.plot

day11.z.2880.plot=ggplot(meta.day11.df[meta.day11.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.2880.plot

day11.z.4320.plot=ggplot(meta.day11.df[meta.day11.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


# Combine plots
all.z.day11.plot <- plot_grid(
  day11.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day11.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)

# Set the overall size of the combined plot
ggsave("figures/all.z.day11.plot.pdf", all.z.day11.plot, width = 49, height = 12)



```





Day 17: z scored
```{r}

day17.z.30.plot=ggplot(meta.day17.df[meta.day17.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day17.z.30.plot

day17.z.120.plot=ggplot(meta.day17.df[meta.day17.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day17.z.120.plot

day17.z.960.plot=ggplot(meta.day17.df[meta.day17.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day17.z.960.plot


day17.z.1440.plot=ggplot(meta.day17.df[meta.day17.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


day17.z.1440.plot

day17.z.2880.plot=ggplot(meta.day17.df[meta.day17.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day17.z.2880.plot

day17.z.4320.plot=ggplot(meta.day17.df[meta.day17.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


# Combine plots
all.z.day17.plot <- plot_grid(
  day17.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day17.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
   day17.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14,14,14, 14, 3)
)

# Set the overall size of the combined plot
ggsave("figures/all.z.day17.plot.pdf", all.z.day17.plot, width = 49, height = 12)


```

All metabolite assays plots
```{r}

all.z.days.plot <- plot_grid(
  all.z.day1.plot,
  all.z.day3.plot,
  all.z.day5.plot,
  all.z.day7.plot,
  all.z.day11.plot,
  all.z.day17.plot,
  nrow = 6
)



all.z.days.plot
ggsave("figures/all.z.days.plot.pdf", all.z.days.plot, width = 49, height = 30)

```


Metabolite analysis
```{r}

#make a big metabolite dataframe

merged_df=rbind(meta.day1.raw.df,meta.day3.raw.df,meta.day5.raw.df,meta.day7.raw.df, meta.day11.raw.df)
class(merged_df$time)
merged_df$treatment=as.factor(as.character(merged_df$treatment))
merged_df$tree_id=as.factor(as.character(merged_df$tree_id))
merged_df$tree_num=as.factor(as.integer(merged_df$tree_num))
merged_df$deploy=as.factor(as.character(merged_df$deploy))
merged_df$site_num=as.factor(as.integer(merged_df$site_num))
merged_df$metabolite=as.factor(as.character(merged_df$metabolite))
merged_df$hfa=as.factor(as.character(merged_df$hfa))
merged_df$time=as.numeric(as.integer(merged_df$time))

class(merged_df$metabolite)

# Fit the linear mixed-effects model with nested random effects and fixed effect
model <- lmer(z_score ~ treatment + hfa + treatment * hfa +
              (1 | deploy/tree_id/metabolite),
              data = merged_df[merged_df$decomp_day == "7" & merged_df$time=="4320", ],
              na.action = na.exclude)
anova(model)






```
16 hr metabolite plots
```{r}

day1.16hr.plot=ggplot(meta.day1.df[meta.day1.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 1")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting +
   theme(legend.position = "none")

day1.16hr.plot

day3.16hr.plot=ggplot(meta.day3.df[meta.day3.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 3")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting +
   theme(legend.position = "none")


day5.16hr.plot=ggplot(meta.day5.df[meta.day5.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "Treatment", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 5")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting+
   theme(legend.position = "none")

day5.16hr.plot

day7.16hr.plot=ggplot(meta.day7.df[meta.day7.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "Treatment", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 5")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting+
   theme(legend.position = "none")

day11.16hr.plot=ggplot(meta.day11.df[meta.day11.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 11")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting+
  theme(legend.position = "none")

day11.16hr.plot

all.16hr.z.plots=plot_grid(
  day1.16hr.plot,
  day3.16hr.plot,
  day5.16hr.plot,
  day11.16hr.plot,
  ncol=5,
  extract.legend,
  rel_widths = c(8,8,8,8,3))


all.day16.z.plots
ggsave("figures/all.16hr.z.plots.pdf", all.16hr.z.plots, width = 26, height = 8)


```
Background information:

The Home Field Advantage Index (HFAI) derived from Ayres et al. (2009) uses the relative mass loss values (rml) of species A and B at different sites. The HFAI therefore represents the difference in the rate of litter decomposition between two different sites ("Hoko" and "Sekiu") for both species A and B. Species A and B, in this case, are alder trees of different origin. For example, species A is Hoko alder and species B is Sekiu alder. There are 10 alder trees per river, representing biological replicates. So, the HFAI quantifies the *advantage* or *disadvantage* of each "species" in their respective home environment compared to an away environment.

Interpreting HFAI values involves considering their sign and magnitude:

**Positive** HFAI Values:
Positive values indicate that both species A and B experience faster decomposition rates at their home site compared to the away site. This suggests that both species are better adapted to their home environment in terms of litter decomposition.

**Negative** HFAI Values:
Negative values indicate that both species A and B experience slower decomposition rates at their home site compared to the away site. This suggests that both species might not be well-suited for decomposition in their own environment and may decompose more efficiently in a different environment.

**Magnitude** of HFAI:
The magnitude of the HFAI values indicates the strength of the advantage or disadvantage. Larger absolute values (positive or negative) suggest a greater difference in decomposition rates between the home and away sites.


Load mass data (Ayres et al.(2009) HFA index)
```{r}
library(stringr)
library(purrr)
mass.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/leaf_mass.csv")
###check for any white spaces clean up
columns_with_whitespace <- map_lgl(mass.df2, ~ any(str_detect(.x, "^\\s+|\\s+$")))
columns_with_whitespace

#correction factor adjusting for blotted and fresh dry mass
mass.df= mass.df%>%
  mutate(final_correct = (0.941*(final_mass)-0.00337))

# calculate % mass loss 
mass.df <- mass.df %>%
  mutate(perc.mass_loss = ((initial_mass - final_correct) / initial_mass)*100)

#subset P treatment for "true" HFA only 
mass.present.df=mass.df %>%
  filter(treatment == "P")

#calculate a k constant
mass.df <- mass.df %>%
  mutate(k = -log(1 - mass_loss / 100) / (2 * 19))

```

**Subset dataframes to z score by deployment site (all HOK 1, all HOK 2, etc...)**

Z scoring controls for site variability, and allows for a better comparison of treatments. But how exactly?

Z-scoring is a statistical measure that quantifies how far away a particular data point is from the mean of a dataset in terms of standard deviations (i.e., the spread of the data points from the mean). It's a way to standardize data and make it more comparable across different scales or distributions. Here's how z-scoring works:

1) calculate the mean and standard deviation: for the variable you want to z-score (in this case mass loss by deployment site), calculate its mean (average) and standard deviation. These values provide information about the central tendency (i.e., the center/midpoint of the data or a value around which the data clusters) and the spread of the data.

2) subtract the mean: for each data point, subtract the mean of the variable. This step shifts the data so that the new mean becomes 0.

3) divide by the standard deviation: after subtracting the mean, divide the result by the standard deviation. This step scales the data based on the spread of the original data, so that the new data points are expressed in terms of standard deviations from the mean.

Mathematically, the z-score (Z) for a data point (X) with mean (μ) and standard deviation (σ) is calculated using the formula:

Z = (X - μ) / σ

The resulting z-score indicates how many standard deviations the original data point is away from the mean. If the z-score is positive, it means the data point is above the mean, and if it's negative, it's below the mean.

Note: z-scoring assumes a roughly normal distribution of the data.

Below, the percent mass loss for each deployment site is z scored. Additionally, the larger dataframe has been subset to only include P treatments so that the "true" HFA can be calculated.
```{r}
#check for distribution of mass loss
hist(mass.df$perc.mass_loss)
  #roughly normal...that's good!

#####Begin z scoring
#HOK 1 subset
mass.hok1.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.hok1.df=mass.hok1.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))

#HOK 1 P subset
mass.hok1.present.df = data.frame(mass.present.df[mass.present.df$deploy == "HOK" & mass.present.df$site_num == "1", ])
# Calculate z-scores 
mass.hok1.present.df=mass.hok1.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))

#HOK 2 subset
mass.hok2.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.hok2.df=mass.hok2.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))

#HOK 2 P subset
mass.hok2.present.df = data.frame(mass.present.df[mass.present.df$deploy == "HOK" & mass.present.df$site_num == "2", ])
# Calculate z-scores 
mass.hok2.present.df=mass.hok2.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))

#SEK 1 subset
mass.sek1.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.sek1.df=mass.sek1.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))

#SEK 1 P subset
mass.sek1.present.df = data.frame(mass.present.df[mass.present.df$deploy == "SEK" & mass.present.df$site_num == "1", ])
# Calculate z-scores 
mass.sek1.present.df=mass.sek1.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))

#SEK 2 subset
mass.sek2.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.sek2.df=mass.sek2.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))

#SEK 2 P subset
mass.sek2.present.df = data.frame(mass.present.df[mass.present.df$deploy == "SEK" & mass.present.df$site_num == "2", ])
# Calculate z-scores 
mass.sek2.present.df=mass.sek2.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))

##merge all sites
mass.deploy.merged.df=rbind(mass.hok1.df, mass.hok2.df, mass.sek1.df, mass.sek2.df)
mass.deploy.merged.df$treatment=as.factor(as.character(mass.deploy.merged.df$treatment))
mass.deploy.merged.df$tree_id=as.factor(as.character(mass.deploy.merged.df$tree_id))
mass.deploy.merged.df$tree_id=as.factor(as.character(mass.deploy.merged.df$tree_id))
mass.deploy.merged.df$deploy=as.factor(as.character(mass.deploy.merged.df$deploy))
mass.deploy.merged.df$hfa=as.factor(as.character(mass.deploy.merged.df$hfa))

#merge all P treatment sites 
mass.present.deploy.merged.df=rbind(mass.hok1.present.df, mass.hok2.present.df, mass.sek1.present.df, mass.sek2.present.df)
mass.present.deploy.merged.df$treatment=as.factor(as.character(mass.present.deploy.merged.df$treatment))
mass.present.deploy.merged.df$tree_id=as.factor(as.character(mass.present.deploy.merged.df$tree_id))
mass.present.deploy.merged.df$tree_id=as.factor(as.character(mass.present.deploy.merged.df$tree_id))
mass.present.deploy.merged.df$deploy=as.factor(as.character(mass.present.deploy.merged.df$deploy))
mass.present.deploy.merged.df$hfa=as.factor(as.character(mass.present.deploy.merged.df$hfa))
```

Let's first look at the P only to see if HFA is there
```{r}
###P treatment only to see HFA effects
#create a proper "site" column for the figure
mass.present.deploy.merged.df$combined_site <- paste("site", mass.present.deploy.merged.df$site_num, sep = "")

#plot
leaf.mass.presentonly.plot=ggplot(mass.present.deploy.merged.df, aes(x = hfa, y = z_score)) +
  geom_boxplot(alpha= 0.3) +
  geom_point(aes(shape = as.factor(combined_site)), position = position_jitter(width = 0.2)) +
  labs(x = "", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_x_discrete(labels = c("away", "home")) +
  scale_fill_manual(values = custom_palette) +
  scale_shape_manual(values = c(7,10), name= "site") +
  geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

leaf.mass.presentonly.plot
  #looks pretty good, but stats must be incorporated

#looking at away outliers
mass.present.deploy.merged.filtered.df=mass.present.deploy.merged.df %>%
  filter(hfa != "A" | (z_score >= -1.2 & z_score <= 1))
#remove the outliers
# Remove rows with ID 39 and 33
mass.present.deploy.merged.filtered.df

#replot
ggplot(mass.present.deploy.merged.filtered.df, aes(x = hfa, y = z_score)) +
  geom_boxplot(alpha= 0.3) +
  geom_point(aes(shape = as.factor(combined_site)), position = position_jitter(width = 0.2)) +
  labs(x = "treatment", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_x_discrete(labels = c("away", "home")) +
  scale_fill_manual(values = custom_palette) +
  scale_shape_manual(values = c(16,17), name= "site") +
  geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting


ggsave("figures/leaf.mass.presentonly.plot.pdf", leaf.mass.presentonly.plot, width = 8, height = 8)

```
Analysis for **present only** dataframe
There are two ways to go about model exploration: using z scored mass loss or arcsine squareroot transforming percent mass loss. These are outlined below...

First: model exploration using z scored data using parametric tests 
```{r}
library(lme4)
library(stats)

#check distribution of raw z scored mass loss
hist(mass.present.deploy.merged.df$z_score)
  #seems pretty good, but model residuals are what counts

##### Fit the models using z score
#ANOVA
m1 <- aov(z_score ~ hfa, data = mass.present.deploy.merged.df)


#LMER
m2 <- lmer(z_score ~ hfa + (1 | tree_id) , data = mass.present.deploy.merged.df)
m3 <- lmer(z_score ~ hfa + (1 | tree_id) + (1 | deploy), data = mass.present.deploy.merged.df)
m4 <- lmer(z_score ~ hfa + (1 | tree_id) + (1 | deploy / site_num), data = mass.present.deploy.merged.df)
m5 <- lmer(z_score ~ hfa + (1 | tree_id / tree_num) + (1 | deploy / site_num), data = mass.present.deploy.merged.df)

# QQ plot of residuals
qqnorm(residuals(m2))
qqline(residuals(m2))

AIC(m2, m3, m4, m5)

# Calculate Cook's distances
cook_d <- cooks.distance(m2)
# Plot Cook's distances
plot(cook_d, type = "p", pch = 19, frame.plot = FALSE,
     xlab = "Observation Number", ylab = "Cook's Distance")
abline(h = 4 / length(cook_d), col = "red", lty = 2) # Threshold for influential points


# Fit glm model
m6 <- glm(z_score ~ hfa, family = gaussian(), data = mass.present.deploy.merged.df)
summary(m6)
m6.resid= residuals(m6)
```


A t-test is more appropriate given that there are only 2 groups!
```{r}
### Visualize the data distributions
#z score
hist(mass.present.deploy.merged.df$z_score[mass.present.deploy.merged.df$hfa == "A"], main = "Group Away")
hist(mass.present.deploy.merged.df$z_score[mass.present.deploy.merged.df$hfa == "H"], main = "Group Home")

# arcsine
hist(mass.present.deploy.merged.df$arcsine_trans[mass.present.deploy.merged.df$hfa == "A"], main = "Group Away")
hist(mass.present.deploy.merged.df$arcsine_trans[mass.present.deploy.merged.df$hfa == "H"], main = "Group Home")

# rice and gaines
hist(mass.present.deploy.merged.df$rice_gaines_corrected[mass.present.deploy.merged.df$hfa == "A"], main = "Group Away")
hist(mass.present.deploy.merged.df$rice_gaines_corrected[mass.present.deploy.merged.df$hfa == "H"], main = "Group Home")


# Check normality using Q-Q plots
qqnorm(mass.present.deploy.merged.df$z_score[mass.present.deploy.merged.df$hfa == "A"])
qqline(mass.present.deploy.merged.df$z_score[mass.present.deploy.merged.df$hfa == "A"])
qqnorm(mass.present.deploy.merged.df$z_score[mass.present.deploy.merged.df$hfa == "H"])
qqline(mass.present.deploy.merged.df$z_score[mass.present.deploy.merged.df$hfa == "H"])

qqnorm(mass.present.deploy.merged.df$arcsine_trans[mass.present.deploy.merged.df$hfa == "A"])
qqline(mass.present.deploy.merged.df$arcsine_trans[mass.present.deploy.merged.df$hfa == "A"])
qqnorm(mass.present.deploy.merged.df$arcsine_trans[mass.present.deploy.merged.df$hfa == "H"])
qqline(mass.present.deploy.merged.df$arcsine_trans[mass.present.deploy.merged.df$hfa == "H"])

qqnorm(mass.present.deploy.merged.df$rice_gaines_corrected[mass.present.deploy.merged.df$hfa == "A"])
qqline(mass.present.deploy.merged.df$rice_gaines_corrected[mass.present.deploy.merged.df$hfa == "A"])
qqnorm(mass.present.deploy.merged.df$rice_gaines_corrected[mass.present.deploy.merged.df$hfa == "H"])
qqline(mass.present.deploy.merged.df$rice_gaines_corrected[mass.present.deploy.merged.df$hfa == "H"])

# Perform Shapiro-Wilk test for normality
shapiro.test(mass.present.deploy.merged.df$z_score[mass.present.deploy.merged.df$hfa == "A"])
shapiro.test(mass.present.deploy.merged.df$z_score[mass.present.deploy.merged.df$hfa == "H"])
  #both groups p > 0.05

shapiro.test(mass.present.deploy.merged.df$arcsine_trans[mass.present.deploy.merged.df$hfa == "A"])
shapiro.test(mass.present.deploy.merged.df$arcsine_trans[mass.present.deploy.merged.df$hfa == "H"])
  #fail
shapiro.test(mass.present.deploy.merged.df$rice_gaines_corrected[mass.present.deploy.merged.df$hfa == "A"])
shapiro.test(mass.present.deploy.merged.df$rice_gaines_corrected[mass.present.deploy.merged.df$hfa == "H"])


# Check equal variance
leveneTest(z_score ~ hfa, data = mass.present.deploy.merged.df)
    # p= 0.64, assumption of equal variance is met

# perform t-test with z score
m.t1 <- t.test(z_score ~ hfa, data = mass.present.deploy.merged.df)
# Fit a linear regression model (t-test equivalent)
lm1 <- lm(z_score ~ hfa, data = mass.present.deploy.merged.df)
lm1.res=residuals(lm1)
# Check assumptions
par(mfrow = c(2, 2))
plot(lm1)

# perform t-test with arcsine
m.t2 <- t.test(arcsine_trans ~ hfa, data = mass.present.deploy.merged.df)
# Fit a linear regression model (t-test equivalent)
lm2 <- lm(arcsine_trans ~ hfa, data = mass.present.deploy.merged.df)
# Check assumptions
par(mfrow = c(2, 2))
plot(lm2)

# perform t test with rice and gaines
m.t3 <- t.test(rice_gaines_corrected ~ hfa, data = mass.present.deploy.merged.df)
# Fit a linear regression model (t-test equivalent)
lm3 <- lm(rice_gaines_corrected ~ hfa, data = mass.present.deploy.merged.df)

# Check assumptions
par(mfrow = c(2, 2)) # Set up 2x2 plot grid
plot(lm3)

# Calculate AIC for each model
AIC_lm1 <- AIC(lm1)
AIC_lm2 <- AIC(lm2)
AIC_lm3 <- AIC(lm3)

AIC(lm1, lm2, lm3)

# Compare AIC values
if (AIC_lm1 < AIC_lm2 && AIC_lm1 < AIC_lm3) {
  cat("Model 1 is preferred with lower AIC:", AIC_lm1, "\n")
} else if (AIC_lm2 < AIC_lm1 && AIC_lm2 < AIC_lm3) {
  cat("Model 2 is preferred with lower AIC:", AIC_lm2, "\n")
} else {
  cat("Model 3 is preferred with lower AIC:", AIC_lm3, "\n")
}

#cohen's D
library("ggpubr")
library(rstatix)
mass.present.deploy.merged.df %>%  cohens_d(z_score ~ hfa, var.equal = TRUE, hedges.correction = T)
mass.present.deploy.merged.df %>%  cohens_d(arcsine_trans ~ hfa, var.equal = TRUE, hedges.correction = T)
mass.present.deploy.merged.df %>%  cohens_d(rice_gaines_corrected ~ hfa, var.equal = TRUE, hedges.correction = T)

#Note that, for small sample size (< 50), the Cohen’s d tends to over-inflate results. There exists a Hedge’s Corrected version of the Cohen’s d (Hedges and Olkin 1985), which reduces effect sizes for small samples by a few percentage points. Small effect size -0.23 with hedges correction. 

# the lm1 and t test will use z score values
stat.test <- mass.present.deploy.merged.df %>% 
  t_test(z_score ~ hfa) %>%
  add_significance()
stat.test

```

add t test result to P treatment only boxplot
```{r}
library(tidyverse)

# Create the box plot with original elements as ggplot from above
leaf.mass.presentonly.plot= ggboxplot(
  mass.present.deploy.merged.df, x = "hfa", y = "z_score", 
  ylab = "mass loss (z score)", xlab = "") + 
  stat_pvalue_manual(stat.test, tip.length = 0) +
  labs(subtitle = get_test_label(stat.test, detailed = TRUE)) +
    scale_x_discrete(labels = c("away", "home"))+
    geom_point(aes(shape = as.factor(combined_site)), position = position_jitter(width = 0.2))+
    scale_shape_manual(values = c(7,10), name= "site")+
    geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16)) 

leaf.mass.presentonly.plot

ggsave("figures/leaf.mass.presentonly.plot.pdf", leaf.mass.presentonly.plot, width = 10, height = 8)


```






```{r}
# Residuals vs. Fitted Values Plot
plot(m6$fitted.values, m6.resid, xlab = "Fitted Values", ylab = "Residuals",
     main = "Residuals vs. Fitted")

# Q-Q Plot
qqnorm(m6.resid)
qqline(m6.resid)

# Scale-Location Plot
plot(sqrt(abs(m6.resid)) ~ m6$fitted.values, xlab = "Fitted Values",
     ylab = "Sqrt(|Residuals|)", main = "Scale-Location")

# Calculate AIC for each model
AIC_m1 <- AIC(m1)
AIC_m2 <- AIC(m2)
AIC_m3 = AIC (m3)

AIC(m1, m2, m3, m4, m5, m6)

# Compare AIC values
lowest_AIC <- min(AIC_m1, AIC_m2, AIC_m3)

if (lowest_AIC == AIC_m1) {
  cat("Model 1 (aov) is preferred with the lowest AIC:", AIC_m1, "\n")
} else if (lowest_AIC == AIC_m2) {
  cat("Model 2 (lmer) is preferred with the lowest AIC:", AIC_m2, "\n")
} else if (lowest_AIC == AIC_m3) {
  cat("Model 3 (glm) is preferred with the lowest AIC:", AIC_m3, "\n")
} 
    #Model 3 and (aov) is preferred with the lowest AIC: 114.7048

### among the models list, model 1 has the lowest AIC and warrants further investigation
# Check assumptions of model 1
par(mfrow = c(2, 2)) # Set up 2x2 plot grid
plot(resid(lmer_model))
#The "Residuals vs Fitted" in the top left panel displays the residuals (eij= γij - γ̂ij) on the y-axis and the fitted values (γ̂ij) on the x-axis. This allows you to see if the variability of the observations differs across the groups because all observations in the same group get the same fitted value. In this plot, the points seem to have fairly similar spreads at the fitted values for the 2 groups. The "Scale-Location" plot in the lower left panel has the same x-axis but the y-axis contains the square-root of the absolute value of the standardized residuals. The absolute value transforms all the residuals into a magnitude scale (removing direction) and the square-root helps you see differences in variability more accurately. The usage is similar in the two plots - you want to assess whether it appears that the groups have somewhat similar or noticeably different amounts of variability. If you see a clear funnel shape in the Residuals vs Fitted or an increase or decrease in the edge of points in the Scale-Location plot, that may indicate a violation of the constant variance assumption. Remember that some variation across the groups is expected and is ok...


```

Mode 1 (aov) further exploration
```{r}
##### exploration of model 1

# Linearity and Homoscedasticity
plot(m1)

# Normality of Residuals - Q-Q plot
qqnorm(m1$residuals)
qqline(m1$residuals)

# Normality of Residuals - Shapiro-Wilk test
shapiro.test(m1$residuals)
  # p=0.41; does not violate normality assumption

#residual distribution
eij=residuals(m1)
hist(eij,main="Histogram of residuals")
plot(density(eij),main="Density plot of residuals",ylab="Density",xlab="Residuals")

```
**Exploration of nonparametric models**
Non-parametric tests are more robust to deviations from normality and heteroscedasticity, but they do have their own assumptions. For example, the Mann-Whitney U test assumes that the distributions of the two groups are similar in shape. The Kruskal-Wallis test assumes that the distributions of all groups are similar and that the differences in the distributions are due to location shifts. 

Scenario 1: comparing two treatments and the data seems to cluster at the extremes, the Mann-Whitney U test (also known as the Wilcoxon rank-sum test) could be a suitable choice. This test is used to compare the distribution of a continuous variable between two groups, i.e., comparing the "home" and "away" treatments (HFA).

```{r}
# Visual inspection of box plots
boxplot(z_score ~ hfa, data = mass.present.deploy.merged.df)

# Perform Mann-Whitney U test
mw_test <- wilcox.test(z_score ~ hfa, data = mass.present.deploy.merged.df)

# Calculate effect size (r) from U statistic
n1 <- sum(mass.present.deploy.merged.df$hfa == "H")
n2 <- sum(mass.present.deploy.merged.df$hfa == "A")
U <- mw_test$statistic
r <- (2 * U) / (n1 * n2) - 1

# Print effect size
cat("Effect size (r) for Mann-Whitney U test:", r, "\n")

#The calculated effect size (r) for the Mann-Whitney U test is -0.1. Effect size values range from -1 to 1, with 0 indicating no effect and values closer to -1 or 1 indicating stronger effects. In this case, a value of -0.1 suggests a small negative effect. Keep in mind that effect size interpretation can vary depending on the context and the field of study.

```
Scenario 2: the Kruskal-Wallis test can be an alternative. This test is an extension of the Wilcoxon rank-sum test for comparing more than two independent groups.

```{r}
# Perform Kruskal-Wallis test
kw_test <- kruskal.test(z_score ~ hfa, data = mass.present.deploy.merged.df)

# Calculate effect size (Eta-squared) from Kruskal-Wallis test
H <- kw_test$statistic
df <- kw_test$parameter
N <- sum(table(mass.present.deploy.merged.df$hfa))
eta_squared <- (H - df) / (N - df)

# Print effect size
cat("Effect size (Eta-squared) for Kruskal-Wallis test:", eta_squared, "\n")

#The calculated effect size (Eta-squared) for the Kruskal-Wallis test is -0.01813634. Eta-squared is a measure of the proportion of variance explained by a factor in non-parametric tests, similar to how R-squared works in linear regression. The negative value suggests a very small effect, but it's important to note that Eta-squared can range from 0 to 1, with higher values indicating larger effects. In this case, the small negative value indicates that the factor (hfa) is not explaining much variance in the data.


```



Model exploration using percent mass loss data
Arcsine sqaure root transformations help to stabilize the variance in proportional data which can sometimes "clump" around the extremes (i.e., 0 and 1). The transformation essentially spreads the data, adjusting for potential heteroscedasticity (or non-normal distributions) which violates most statistical models. Ideally, after being transformed the data will no longer violate assumptions of homoscedasticity.

```{r}
##### Fit the ANOVA model using percent mass loss
#create proportional mass loss column
mass.present.deploy.merged.df=mass.present.deploy.merged.df%>%
  mutate(dec.mass_loss = perc.mass_loss/100)

#find min of mass loss
min_value= min(mass.present.deploy.merged.df$dec.mass_loss)
    #min = -0.358

#find max of mass loss
max_value= max(mass.present.deploy.merged.df$dec.mass_loss)
    #max = 0.258

# Perform min-max scaling
mass.present.deploy.merged.df$scaled_values <- (mass.present.deploy.merged.df$dec.mass_loss - min_value) / (max_value - min_value)

#apply arcsine transformation and rice and gaines 
mass.present.deploy.merged.df=mass.present.deploy.merged.df%>%
  mutate(arcsine_trans = asin(sqrt(scaled_values)))%>%
  mutate(rice_gaines_corrected = asin(sqrt(scaled_values)) - (3/8) * sin(2 * asin(sqrt(scaled_values))))


#new model
m2 <- aov(arcsine_trans ~ hfa, data = mass.present.deploy.merged.df)

# Print summary of the ANOVA
summary(m2)
Anova(m2)

# Visualize the residuals using a QQ plot
qqnorm(m2$residuals)
qqline(m2$residuals)

qqnorm(mass.present.deploy.merged.df$rice_gaines_corrected)
qqline(mass.present.deploy.merged.df$rice_gaines_corrected)

shapiro.test(m2$residuals) 


# Create a histogram of the residuals
hist(m2$residuals)


```


leaf % mass loss with z score by deployment site plot
```{r}
# Combine treatment and hfa into one group
mass.deploy.merged.df$combined_group <- paste(mass.deploy.merged.df$treatment, mass.deploy.merged.df$hfa, sep = "-")

# Create a custom order for the combined groups
combined_group_order <- c("AB-A", "AB-H", "P-A","P-H")

# Plot with combined groups but no combo site
leaf.mass.bothtreatments.plot=ggplot(mass.deploy.merged.df, aes(x = combined_group, y = z_score, fill = combined_group)) +
  geom_boxplot(alpha= 0.3) +
  scale_x_discrete(limits = combined_group_order)+
  geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.2), alpha = 0.7) +
  labs(x = "treatment", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_fill_manual(values = custom_palette, guide = "none") +
  scale_shape_manual(values = c(15, 16), name ="site") +
  geom_hline(data = mass.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting
leaf.mass.combo.group.zsite.plot

# Define your custom color palette
custom_palette.hfa <- c("#1F77B4", "#FF7F0E")

ggplot(mass.present.deploy.merged.df, aes(x = combined_site, y = z_score)) +
  geom_boxplot(alpha= 0.3) +
  geom_point(aes(shape = interaction(hfa)), position = position_jitter(width = 0.2)) +
  labs(x = "treatment", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_fill_manual(values = custom_palette) +
  scale_shape_manual(values = c(15, 16)) +
  geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting
```





Subset dataframes to z score by deployment site and treatment
```{r}
#P HOK 1 subset
mass.p.hok1.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$treatment == "P" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.p.hok1.df=mass.p.hok1.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#AB HOK 1 subset
mass.ab.hok1.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$treatment == "AB" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.ab.hok1.df=mass.ab.hok1.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#P HOK 2 subset
mass.p.hok2.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$treatment == "P" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.p.hok2.df=mass.p.hok2.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#AB HOK 2 subset
mass.ab.hok2.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$treatment == "AB" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.ab.hok2.df=mass.ab.hok2.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#P SEK 1 subset
mass.p.sek1.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$treatment == "P" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.p.sek1.df=mass.p.sek1.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#AB SEK 1 subset
mass.ab.sek1.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$treatment == "AB" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.ab.sek1.df=mass.ab.sek1.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#P SEK 2 subset
mass.p.sek2.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$treatment == "P" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.p.sek2.df=mass.p.sek2.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#AB SEK 2 subset
mass.ab.sek2.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$treatment == "AB" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.ab.sek2.df=mass.ab.sek2.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))



##merge all sites
mass.deploy.treat.merged.df=rbind(mass.p.hok1.df, mass.ab.hok1.df, mass.p.hok2.df, mass.ab.hok2.df, mass.p.sek1.df, mass.ab.sek1.df, mass.p.sek2.df, mass.ab.sek2.df)
mass.deploy.treat.merged.df$treatment=as.factor(as.character(mass.deploy.treat.merged.df$treatment))
mass.deploy.treat.merged.df$tree_id=as.factor(as.character(mass.deploy.treat.merged.df$tree_id))
mass.deploy.treat.merged.df$tree_id=as.factor(as.character(mass.deploy.treat.merged.df$tree_id))
mass.deploy.treat.merged.df$deploy=as.factor(as.character(mass.deploy.treat.merged.df$deploy))
mass.deploy.treat.merged.df$hfa=as.factor(as.character(mass.deploy.treat.merged.df$hfa))

```


leaf % mass loss as z score by deployment site AND treatment
```{r}
# Combine treatment and hfa into one group
mass.deploy.treat.merged.df$combined_group <- paste(mass.deploy.treat.merged.df$treatment, mass.deploy.treat.merged.df$hfa, sep = "-")
mass.deploy.treat.merged.df$combined_site <- paste("site", mass.deploy.treat.merged.df$site_num, sep = "")


# Create a custom order for the combined groups
combined_group_order <- c("AB-A", "AB-H", "P-A","P-H")


custom_palette <- c("#1F77B4", "#FF7F0E", "#2CA02C", "#D62728")

custom_palette <- c("#1F77B4", "#FF7F0E", "#2CA02C", "#D62728", "#9467BD", "#8C564B", "#E377C2", "#7F7F7F")


# Plot with combined groups and site 
leaf.mass.combo.site.plot=ggplot(mass.deploy.treat.merged.df, aes(x = combined_group, y = z_score, fill= interaction(combined_site, combined_group))) +
  geom_boxplot(alpha = 0.3) +
  scale_x_discrete(limits = combined_group_order) +
  geom_point(aes(shape = interaction(combined_site, combined_group)), position = position_jitter(width = 0.2), alpha = 0.5) +
  labs(x = "treatment", y = "mass loss (z score)") +
  theme(legend.position = "none",
axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(0,1,2,3,4,5,6,7), name = "") +
   scale_fill_manual(values = custom_palette, name = "")+
  geom_hline(data = mass.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  Fig.formatting
leaf.mass.combo.site.plot

# Plot with combined groups but no combo site 
leaf.mass.combo.group.plot=ggplot(mass.deploy.treat.merged.df, aes(x = combined_group, y = z_score, fill= combined_group)) +
  geom_boxplot(alpha = 0.3) +
  scale_x_discrete(limits = combined_group_order) +
  geom_point(aes(shape = combined_group), position = position_jitter(width = 0.2), alpha = 0.5) +
  labs(x = "treatment", y = "mass loss (z score)") +
  theme(legend.position = "none",
axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(21, 22, 23, 24), name = "") +
   scale_fill_manual(values = custom_palette, name = "")+
  geom_hline(data = mass.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  Fig.formatting

leaf.mass.combo.group.plot

#trying to shade with different shapes
ggplot(mass.deploy.treat.merged.df, aes(x = combined_group, y = z_score, fill = combined_group)) +
  geom_boxplot_pattern(
    aes(pattern = combined_group),
    colour = "black",
    pattern_fill = "black",
    pattern_angle = 45,
    pattern_density = 0.1,
    pattern_spacing = 0.01,
    position = position_dodge2(preserve = 'single'),
    alpha = 0.2  # Adjust the alpha value to control transparency
  ) +
    geom_point(aes(shape = interaction(site_num, combined_group)), position = position_jitter(width = 0.2), alpha = 0.5)+
  geom_boxplot(alpha = 0.3) +
  scale_x_discrete(limits = combined_group_order) +
  labs(x = "Treatment Group", y = "Mass Loss (z score)") +
  theme(legend.position = "bottom",  # Move legend to the bottom
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16, 17, 18, 19, 20, 21, 22, 23)) +
  scale_fill_manual(values = custom_palette) +
  geom_hline(data = mass.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  guides(fill = guide_legend(override.aes = list(shape = c(16, 17, 18, 19, 20, 21, 22, 23)))) +  # Specify the shape order
  Fig.formatting



#basic HFA
ggplot(mass.deploy.treat.merged.df, aes(x = combined_group, y = z_score)) +
  geom_boxplot(alpha = 0.3) +
  scale_x_discrete(limits = combined_group_order) +
  geom_point(aes(shape = interaction(site_num, hfa)), position = position_jitter(width = 0.2), alpha = 0.5) +
  labs(x = "Treatment Group", y = "Mass Loss (z score)") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.title = "Treatment") +
  scale_shape_manual(values = c(16, 17, 18, 19)) +
  geom_hline(data = mass.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  Fig.formatting


 leaf.mass.z.plot
ggsave("figures/leaf.mass.combo.group.plot.pdf", leaf.mass.combo.group.plot, width = 8, height = 6)



```
analysis for z score by treatment and site
```{r}

#RICE & GAINES CORRECTION
save<-aggregate(as.numeric(z.new.deal1), list(new.decomp$LOCATION_CODE), mean) #use the z standardized data
save
save$Rank<-c(2,2,2,1,1)                                   
save #(Away River=2,Away River =2,SameRiveraway=2, SameRiverHOme=1, Home=1)
cor.test(x=save$x, y=save$Rank, method = 'spearman',exact=FALSE)
#RsPc= 0.8660254*(1-0.03953) = 0.831, so p ~ 0.01* significant ???
RSPC<- (-0.5773503)*(1-2.723e-06)









```


maybe try z scoring by site and using the z score percent mass loss for HFAI?


generate **present**, site a (HOK 1)
```{r}
mass.p.hoko1.df=mass.df%>%
  filter(treatment == "P" & deploy=="HOK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.p.hoko1.df <- mass.p.hoko1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.p.hoko1.df$mass_loss[mass.p.hoko1.df$tree_id == "HOK" & mass.p.hoko1.df$tree_num == tree_num]
    sek_mass_loss <- mass.p.hoko1.df$mass_loss[mass.p.hoko1.df$tree_id == "SEK" & mass.p.hoko1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}

##check the calculations (Armla.1-Armla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss)) * 100


##check the calculations (Brmla.1-Brmla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss)) * 100


# Add the rml values to the dataframe
mass.p.hoko1.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.p.hoko1.df$ayres.tree <- paste(mass.p.hoko1.df$ayres, mass.p.hoko1.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.p.hoko1.df$ayres.eq <- ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **present**, site b (SEK 1)
```{r}
mass.p.sek1.df=mass.df%>%
  filter(treatment == "P" & deploy=="SEK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.p.sek1.df <- mass.p.sek1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)


###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.p.sek1.df$mass_loss[mass.p.sek1.df$tree_id == "HOK" & mass.p.sek1.df$tree_num == tree_num]
    sek_mass_loss <- mass.p.sek1.df$mass_loss[mass.p.sek1.df$tree_id == "SEK" & mass.p.sek1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}

##check the calculations (Armla.1-Armla.5)
((mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.1",]$mass_loss) /
  (mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.1",]$mass_loss +
   mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Ba.1",]$mass_loss)) * 100

((mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.2",]$mass_loss) /
  (mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.2",]$mass_loss +
   mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Ba.2",]$mass_loss)) * 100


((mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.3",]$mass_loss) /
  (mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.3",]$mass_loss +
   mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Ba.3",]$mass_loss)) * 100

((mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.4",]$mass_loss) /
  (mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.4",]$mass_loss +
   mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Ba.4",]$mass_loss)) * 100


((mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.5",]$mass_loss) /
  (mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.5",]$mass_loss +
   mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Ba.5",]$mass_loss)) * 100


##check the calculations (Brmla.1-Brmla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss)) * 100

# Add the rml values to the dataframe
mass.p.sek1.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.p.sek1.df$ayres.tree <- paste(mass.p.sek1.df$ayres, mass.p.sek1.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.p.sek1.df$ayres.eq <- ifelse(mass.p.sek1.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.p.sek1.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **present**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.p.site1.df=rbind(mass.p.hoko1.df, mass.p.sek1.df)

# Create an empty data frame to store the results
hfai.df <- data.frame(hfai = numeric(0))
# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.p.site1.df[mass.p.site1.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.p.site1.df[mass.p.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.p.site1.df[mass.p.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.p.site1.df[mass.p.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.p.site1.df[mass.p.site1.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = LETTERS[i], treatment = treatment))
}

#HFAI (this is a stop and check to verify that values are correct!)
(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```

## Now we'll tackle absent:
generate **absent**, site a (HOK 1)
```{r}
mass.ab.hoko1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="HOK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.hoko1.df <- mass.ab.hoko1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "HOK" & mass.ab.hoko1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "SEK" & mass.ab.hoko1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}

##check the calculations (Armla.1-Armla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss)) * 100


##check the calculations (Brmla.1-Brmla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss)) * 100

# Add the rml values to the dataframe
mass.ab.hoko1.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.hoko1.df$ayres.tree <- paste(mass.ab.hoko1.df$ayres, mass.ab.hoko1.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.hoko1.df$ayres.eq <- ifelse(mass.ab.hoko1.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **absent**, site b (SEK 1)
```{r}
mass.ab.sek1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="SEK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.sek1.df <- mass.ab.sek1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)


###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "HOK" & mass.ab.sek1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "SEK" & mass.ab.sek1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.sek1.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.sek1.df$ayres.tree <- paste(mass.ab.sek1.df$ayres, mass.ab.sek1.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.sek1.df$ayres.eq <- ifelse(mass.ab.sek1.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.p.sek1.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **asbent**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.ab.site1.df=rbind(mass.ab.hoko1.df, mass.ab.sek1.df)

# Calculate the last letter index based on the existing data frame
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = next_label, treatment = treatment))
}


#HFAI (this is a stop and check to verify that values are correct!)
(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```

plot check
```{r}

ggplot(hfai.df, aes(x = label, y = hfai)) +
  geom_bar(stat = "identity",
           show.legend = FALSE) + # Remove the legend
  facet_grid(~treatment)+
  xlab("Group") +
  ylab("Value")+
  coord_flip()




```

generate **present**, site a (HOK 2)
```{r}
mass.p.hoko2.df=mass.df%>%
  filter(treatment == "P" & deploy=="HOK" & site_num=="2")
#now create a new column to have some reference to the Ayres et al. (2009) paper (species A= hoko trees, species B = sekiu trees)
mass.p.hoko2.df <- mass.p.hoko2.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 2-1 with SEK 2-1; this follows similar logic as Ayres et al. (2009) Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.p.hoko2.df$mass_loss[mass.p.hoko2.df$tree_id == "HOK" & mass.p.hoko2.df$tree_num == tree_num]
    sek_mass_loss <- mass.p.hoko2.df$mass_loss[mass.p.hoko2.df$tree_id == "SEK" & mass.p.hoko2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.p.hoko2.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.p.hoko2.df$ayres.tree <- paste(mass.p.hoko2.df$ayres, mass.p.hoko2.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.p.hoko2.df$ayres.eq <- ifelse(mass.p.hoko2.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **present**, site b (SEK 2)
```{r}
mass.p.sek2.df=mass.df%>%
  filter(treatment == "P" & deploy=="SEK" & site_num=="2")
#now create a new column to have some reference to the Ayres et al. (2009) paper (species A= hoko trees, species B = sekiu trees)
mass.p.sek2.df <- mass.p.sek2.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)


###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 2-1 with SEK 2-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.p.sek2.df$mass_loss[mass.p.sek2.df$tree_id == "HOK" & mass.p.sek2.df$tree_num == tree_num]
    sek_mass_loss <- mass.p.sek2.df$mass_loss[mass.p.sek2.df$tree_id == "SEK" & mass.p.sek2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.p.sek2.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.p.sek2.df$ayres.tree <- paste(mass.p.sek2.df$ayres, mass.p.sek2.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.p.sek2.df$ayres.eq <- ifelse(mass.p.sek2.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.p.sek1.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **present**, site a and b (aka HOK 2 and SEK 2) dataframes 
```{r}
#merge both present site 1 dataframes
mass.p.site2.df=rbind(mass.p.hoko2.df, mass.p.sek2.df)

# Calculate the last letter index based on the existing data frame
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.p.site2.df[mass.p.site2.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.p.site2.df[mass.p.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.p.site2.df[mass.p.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.p.site2.df[mass.p.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.p.site2.df[mass.p.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = next_label, treatment = treatment))
}

#HFAI (this is a stop and check to verify that values are correct!)
(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```

## Now we'll tackle absent:
generate **absent**, site a (HOK 2)
```{r}
mass.ab.hoko2.df=mass.df%>%
  filter(treatment == "AB" & deploy=="HOK" & site_num=="2")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.hoko2.df <- mass.ab.hoko2.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.hoko2.df$mass_loss[mass.ab.hoko2.df$tree_id == "HOK" & mass.ab.hoko2.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.hoko2.df$mass_loss[mass.ab.hoko2.df$tree_id == "SEK" & mass.ab.hoko2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.hoko2.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.hoko2.df$ayres.tree <- paste(mass.ab.hoko2.df$ayres, mass.ab.hoko2.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.hoko2.df$ayres.eq <- ifelse(mass.ab.hoko2.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **absent**, site b (SEK 1)
```{r}
mass.ab.sek2.df=mass.df%>%
  filter(treatment == "AB" & deploy=="SEK" & site_num=="2")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.sek2.df <- mass.ab.sek2.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.sek2.df$mass_loss[mass.ab.sek2.df$tree_id == "HOK" & mass.ab.sek2.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.sek2.df$mass_loss[mass.ab.sek2.df$tree_id == "SEK" & mass.ab.sek2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.sek2.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.sek2.df$ayres.tree <- paste(mass.ab.sek2.df$ayres, mass.ab.sek2.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.sek2.df$ayres.eq <- ifelse(mass.ab.sek2.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.p.sek1.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **asbent**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.ab.site2.df=rbind(mass.ab.hoko2.df, mass.ab.sek2.df)

# Calculate the last letter index based on the existing data frame
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = next_label, treatment = treatment))
}

#HFAI (this is a stop and check to verify that values are correct!)
(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```

another plot check
```{r}
ggplot(hfai.df, aes(x = reorder(label, hfai), y = hfai, fill = treatment)) +
  geom_bar(stat = "identity") +
  xlab("") +
  ylab("Home-field advantage (% faster decomposition at home)") +
  coord_flip() +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  )

# Filter out outliers
filtered.hfai.data <- hfai.df %>%
  filter(hfai >= -300, hfai <= 300)

#final plot looks good
  #no color, but shading with slight opacity
hfai.plot=ggplot(filtered.hfai.data, aes(x = reorder(label, hfai), y = hfai)) +
  geom_bar(stat = "identity", fill = "white", color = "black") +
  geom_col_pattern(
    aes(pattern = treatment),
    colour = "black",
    pattern_fill = "black",
    pattern_angle = 45,
    pattern_density = 0.1,
    pattern_spacing = 0.01,
    position = position_dodge2(preserve = 'single'),
    alpha = 0.2  # Adjust the alpha value to control transparency
  ) +
  scale_pattern_manual(
    values = c("none", "stripe"),
    guide = guide_legend(override.aes = list(fill = "grey70")) # <- make lighter
  ) +
  xlab("") +
  ylab("Home-field advantage (% faster decomposition at home)") +
coord_flip() +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  )

hfai.plot
ggsave("figures/hfai.plot.pdf", hfai.plot, width = 26, height = 8)

```


















```{r}
########### Next we'll try out absent, hoko 1
mass.ab.hoko1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="HOK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (Home= Aa, Away= Bb)
mass.ab.hoko1.df <- mass.ab.hoko1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "HOK" & mass.ab.hoko1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "SEK" & mass.ab.hoko1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.hoko1.df$rml <- rml_values







########### Next we'll try out absent, sekiu 1
mass.ab.sek1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="SEK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (Home= Aa, Away= Bb)
mass.ab.sek1.df <- mass.ab.sek1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "HOK" & mass.ab.sek1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "SEK" & mass.ab.sek1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.sek1.df$rml <- rml_values


#merge dataframes
mass.site1.df=rbind(mass.p.hoko1.df, mass.ab.hoko1.df, mass.p.sek1.df, mass.ab.sek1.df )



```




Ayres HFA index
```{r}
#the calculations for arml and brml are different, dependent on "relative to"

##Site 1: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site a (Hoko) 
site1.armla.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="1" & mass.df$deploy == "HOK", ])
#mass loss calculation
site1.armla.df= site1.armla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.armla.df <- site1.armla.df %>%
  mutate(rml = ((mass_loss[tree_id == "HOK" & deploy == "HOK"]) / 
                 (mass_loss[tree_id == "HOK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "HOK"])) * 100)

##Site 2: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site a (Hoko) 
site2.armla.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="2" & mass.df$deploy == "HOK", ])
#mass loss calculation
site2.armla.df= site2.armla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.armla.df <- site2.armla.df %>%
  mutate(rml = (mass_loss[tree_id == "HOK" & deploy == "HOK"] / 
                 (mass_loss[tree_id == "HOK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "HOK"])) * 100)

##Site 1: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site b (Sekiu) 
site1.armlb.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="2" & mass.df$deploy == "SEK", ])
#mass loss calculation
site1.armlb.df= site1.armlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.armlb.df <- site1.armlb.df %>%
  mutate(rml = (mass_loss[tree_id == "HOK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "HOK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "SEK"])) * 100)

##Site 2: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site b (Sekiu) 
site2.armlb.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="2" & mass.df$deploy == "SEK", ])
#mass loss calculation
site2.armlb.df= site2.armlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.armlb.df <- site2.armlb.df %>%
  mutate(rml = (mass_loss[tree_id == "HOK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "HOK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "SEK"])) * 100)

##Site 1: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site a (Hoko) 
site1.brmla.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="1" & mass.df$deploy == "HOK", ])
#mass loss calculation
site1.brmla.df= site1.brmla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.brmla.df <- site1.brmla.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "HOK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "HOK"])) * 100)

##Site 2: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site a (Hoko) 
site2.brmla.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="2" & mass.df$deploy == "HOK", ])
#mass loss calculation
site2.brmla.df= site2.brmla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.brmla.df <- site2.brmla.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "HOK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "HOK"])) * 100)

##Site 1: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site b (Sekiu) 
site1.brmlb.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="1" & mass.df$deploy == "SEK", ])
#mass loss calculation
site1.brmlb.df= site1.brmlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.brmlb.df <- site1.brmlb.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "SEK"])) * 100)

##Site 2: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site b (Sekiu) 
site2.brmlb.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="2" & mass.df$deploy == "SEK", ])
#mass loss calculation
site2.brmlb.df= site2.brmlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.brmlb.df <- site2.brmlb.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "SEK"])) * 100)

#merge dataframes
site1.hfa_index.df=rbind(site1.armla.df, site1.armlb.df, site1.brmla.df, site1.brmlb.df)
#HFA index calculation
site1.hfa_index.df <- site1.hfa_index.df %>%
  mutate(hfai = (((site1.armla.df$rml + site1.brmlb.df$rml) / 2) / ((site1.armlb.df$rml + site1.brmla.df$rml) / 2)) * 100 - 100)

#merge dataframes
site2.hfa_index.df=rbind(site2.armla.df, site2.armlb.df, site2.brmla.df, site2.brmlb.df)
#HFA index calculation
site2.hfa_index.df <- site2.hfa_index.df %>%
  mutate(hfai = (((site2.armla.df$rml + site2.brmlb.df$rml) / 2) / ((site2.armlb.df$rml + site2.brmla.df$rml) / 2)) * 100 - 100)

# Create the horizontal bar plot
ggplot(hfa_index.df, aes(x = hfai, y = interaction(deploy, hfa))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(title = "HFA Index Comparison",
       x = "HFA Index") +
  theme_minimal() +
  theme(legend.position = "none") +
  geom_vline(xintercept = 0, color = "black", size = 1) +
  scale_fill_manual(values = c("a" = "blue", "b" = "red")) 



```







