---
title: "RAD Alder 2023"
author: "Cody Spiegel"
date: "`r Sys.Date()`"
output: word_document
---

### Project overview
For Project ‘RAD’ Alder, leaf packs are gamma irradiated or not, and deployed at “home” or “away” sites to test the role of  phyllosphere microbial communities on home-field advantage patterns in leaf litter decomposition between rivers on WA Olympic Peninsula. Using a reciprocal transplant experiment of leaf packs under four treatment conditions: phyllosphere present-home (PH), phyllosphere present-away (PAW), phyllosphere absent-home (ABH), phyllosphere absent-away (ABAW), the following questions are asked: 1) Do the source communities differ between the four treatments, 2) Do community metrics differ between the four treatments?, 3) Does plant compound metabolization differ between the four treatments?, 4) Does decomposition differ between the four treatments? 



**DATA SETS**  
This data set is among 3 to be generated for the project and focuses on:  

(1) **dissolved organic carbon** (DOC) and  
(2) **Fluorescence and UV-vis absorbance** indices  
(3) **DOC Degradation** from either photo- or microbial decomposition 
(4) **burning effects on plant material decomposition** using % mass loss

```{r setup chunk, setup, include = FALSE, cache=FALSE, message=FALSE, warning=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

# use pacman to load all the packages you are missing!
pacman::p_load('knitr', 'lme4', 'lmerTest', 'tidyverse', 'magrittr', 'effects', 'plyr', 'dplyr', 
               'plotrix', 'car',"gridExtra", "cowplot", "tools", "mgcv", "gratia", "tidymv")

Fig.formatting<-(theme_classic()) +
  theme(text=element_text(size=25),
        legend.text.align = 0,
        panel.border = element_rect(fill=NA, colour = "black", size=1),
        aspect.ratio=1, 
        axis.ticks.length=unit(0.25, "cm"),
        axis.text.y=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20), 
        axis.text.x=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20)) +
  theme(legend.key.size = unit(0.4, "cm")) +
  theme(aspect.ratio=1.3) 




```


All metabolite decomp days
```{r}
meta.day1.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_1_metabolites_corrected.csv")
meta.day3.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_3_metabolites_corrected.csv")
meta.day5.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_5_metabolites_corrected.csv")
meta.day7.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_7_metabolites_corrected.csv")

meta.day1.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_1_metabolites_raw_corrected.csv")
meta.day3.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_3_metabolites_raw_corrected.csv")
meta.day5.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_5_metabolites_raw_corrected.csv")
meta.day7.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_7_metabolites_raw_corrected.csv")

```

Metabolic Assay
```{r}
meta.day1.df=read.csv("/Users/Cody1/Downloads/'RAD' Alder Metadata - day_1_metabolites_corrected.csv")

aggra.day1.df <- meta.day1.df %>%
  distinct(treatment, tree_id, deploy, metabolite, hfa) %>%
  group_by(treatment, tree_id, deploy, metabolite, hfa) %>%
  mutate(mean.od = mean(meta.day1.df$od_corrected, na.rm = TRUE))


ggplot(meta.day1.df, aes(x = time, y = od_corrected, color = hfa, group = interaction(tree_id, metabolite, deploy))) +
  geom_line()+
  facet_grid(~treatment) +
  labs(x = "Time", y = "Mean OD value") +
  theme_minimal() +
  theme(legend.position = "bottom")

library(RColorBrewer)

ggplot(meta.day1.df, aes(x = treatment, y = od_corrected, fill = treatment)) +
  geom_boxplot() +
  facet_grid(time~tree_id  + deploy +metabolite+hfa) +
  scale_fill_manual(values = brewer.pal(n = 3, name = "Dark2"))+
  labs(x = "Treatment", y = "OD Corrected") +
  theme_minimal() +
  theme(legend.position = "bottom")


###visualize Day 1
ggplot(meta.day1.df, aes(x = treatment, y = od_corrected, fill = treatment)) +
  geom_boxplot() +
  facet_grid(time ~ deploy + metabolite + hfa) +
  scale_fill_manual(values = brewer.pal(n = 3, name = "Dark2")) +
  labs(x = "Treatment", y = "OD Corrected") +
  theme_minimal() +
  theme(legend.position = "bottom")

###############################

# Calculate the overall average od_corrected value
overall_avg <- mean(meta.day1.df[meta.day1.df$time == "1440", ]$od_corrected)

ggplot(meta.day1.df[meta.day1.df$time == "1440", ], aes(x = hfa, y = od_corrected, group = interaction(tree_id, deploy, metabolite), fill = interaction(interaction(tree_id, deploy, metabolite)))) +
  geom_boxplot() +
  facet_grid(time ~ treatment) +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting +
  theme(legend.position = "bottom") +

  # Add a horizontal reference line at the overall average
  geom_hline(yintercept = overall_avg,
             linetype = "dotted",
             color = "red",
             size=1.3) 

#interpret the boxplots relative to the horizontal reference line. Boxplots above the line indicate greater od_corrected values, while boxplots below the line indicate lesser od_corrected values.

ggplot(meta.day1.df[meta.day1.df$time=="1440",], aes(x = treatment, y = od_corrected, group= interaction(tree_id,deploy,metabolite), fill = interaction (interaction(tree_id, deploy, metabolite)))) +
  geom_boxplot() +
  facet_grid(time~hfa)+
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "bottom")+
  geom_hline(data = meta.day1.df[meta.day1.df$time=="1440",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted") 

###### THIS ONE


ggplot(meta.day1.df[meta.day1.df$time=="1440",], aes(x = hfa, y = od_corrected, group= interaction(tree_id,deploy,metabolite), fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~treatment)+
  scale_fill_manual(values = color_palette)+
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "bottom")+
  geom_hline(data = meta.day1.df[meta.day1.df$time=="1440",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")
```



Day 1: z scored
```{r}
mean_od_raw.day1 <- mean(meta.day1.raw.df$od_raw, na.rm=T)
sd_od_raw.day1 <- sd(meta.day1.raw.df$od_raw, na.rm=T)
meta.day1.raw.df$z_score <- (meta.day1.raw.df$od_raw - mean_od_raw.day1) / sd_od_raw.day1


day1.z.30.plot=ggplot(meta.day1.raw.df[meta.day1.raw.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
  coord_cartesian(ylim = c(-1.5,4))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day1.raw.df[meta.day1.raw.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day1.z.30.plot

day1.z.120.plot=ggplot(meta.day1.raw.df[meta.day1.raw.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
    coord_cartesian(ylim = c(-1.5,4))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day1.raw.df[meta.day1.raw.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.960.plot=ggplot(meta.day1.raw.df[meta.day1.raw.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
    coord_cartesian(ylim = c(-1.5,4))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day1.raw.df[meta.day1.raw.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.1440.plot=ggplot(meta.day1.raw.df[meta.day1.raw.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
    coord_cartesian(ylim = c(-1.5,4))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day1.raw.df[meta.day1.raw.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.2880.plot=ggplot(meta.day1.raw.df[meta.day1.raw.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
    coord_cartesian(ylim = c(-1.5,4))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day1.raw.df[meta.day1.raw.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day1.z.4320.plot=ggplot(meta.day1.raw.df[meta.day1.raw.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
    coord_cartesian(ylim = c(-1.5,4))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day1.raw.df[meta.day1.raw.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Create the grid with modified plots
extract.legend <- get_legend(
  day1.30.plot + theme(legend.box.margin = margin(0, 0, 0, 1, "cm"),
                       legend.key.size = unit(100, "pt"),
                       legend.text = element_text(size=20),
                       legend.title= element_blank())
)

# Combine plots
all.z.day1.plot <- plot_grid(
  day1.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  extract.legend,
  ncol = 7,
  rel_widths = c(8, 8, 8, 8, 8, 8, 3)
)


all.z.day1.plot
# Set the overall size of the combined plot
ggsave("figures/all.z.day1.plot.pdf", all.z.day1.plot, width = 49, height = 12)



```






Best looking figs: Day 1 metabolites
```{r}
########## THIS ONE TOO (JUST P VS AB)
# Define a colorblind-friendly palette
color_palette <- brewer.pal(n = 3, name = "Set1")


day1.30.plot=ggplot(meta.day1.df[meta.day1.df$time=="30",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
   scale_fill_manual(values = color_palette)+
  coord_cartesian(ylim = c(0.25,3.15))+
 geom_hline(data = meta.day1.df[meta.day1.df$time=="30",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day1.30.plot

day1.120.plot=ggplot(meta.day1.df[meta.day1.df$time=="120",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
   scale_fill_manual(values = color_palette)+
  coord_cartesian(ylim = c(0.25,3.15))+
 geom_hline(data = meta.day1.df[meta.day1.df$time=="120",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.960.plot=ggplot(meta.day1.df[meta.day1.df$time=="960",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
  coord_cartesian(ylim = c(0.25,3.15))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day1.df[meta.day1.df$time=="960",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.1440.plot=ggplot(meta.day1.df[meta.day1.df$time=="1440",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
  coord_cartesian(ylim = c(0.25,3.15))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day1.df[meta.day1.df$time=="1440",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.2880.plot=ggplot(meta.day1.df[meta.day1.df$time=="2880",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
   coord_cartesian(ylim = c(0.25,3.15))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day1.df[meta.day1.df$time=="2880",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day1.4320.plot=ggplot(meta.day1.df[meta.day1.df$time=="4320",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
  coord_cartesian(ylim = c(0.25,3.15))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day1.df[meta.day1.df$time=="4320",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Create the grid with modified plots
extract.legend <- get_legend(
  day1.30.plot + theme(legend.box.margin = margin(0, 0, 0, 1, "cm"),
                       legend.key.size = unit(100, "pt"),
                       legend.text = element_text(size=20),
                       legend.title= element_blank())
)

# Combine plots
all.day1.plot <- plot_grid(
  day1.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  extract.legend,
  ncol = 7,
  rel_widths = c(8, 8, 8, 8, 8, 8, 3)
)


all.day1.plot <- all.day1.plot + labs(caption = "OD corrected values (raw OD - blanks) for the P and AB treatments with HFA. The horizontal dotted line represents the average OD value specifically for each time, providing a reference for comparison.")


all.day1.plot
# Set the overall size of the combined plot
ggsave("figures/all.day1.plot.pdf", all.day1.plot, width = 49, height = 12)


```


Best looking figs: Day 3 metabolites
```{r}
########## THIS ONE TOO (JUST P VS AB)
# Define a colorblind-friendly palette
color_palette <- brewer.pal(n = 3, name = "Set1")


day3.30.plot=ggplot(meta.day3.df[meta.day3.df$time=="30",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
  coord_cartesian(ylim = c(0.15,2.5))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day3.df[meta.day3.df$time=="30",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day3.30.plot

day3.120.plot=ggplot(meta.day3.df[meta.day3.df$time=="120",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
   coord_cartesian(ylim = c(0.15,2.5))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day3.df[meta.day3.df$time=="120",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.960.plot=ggplot(meta.day3.df[meta.day3.df$time=="960",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
   coord_cartesian(ylim = c(0.15,2.5))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day3.df[meta.day3.df$time=="960",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.1440.plot=ggplot(meta.day3.df[meta.day3.df$time=="1440",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
   coord_cartesian(ylim = c(0.15,2.5))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day3.df[meta.day3.df$time=="1440",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.2880.plot=ggplot(meta.day3.df[meta.day3.df$time=="2880",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
   coord_cartesian(ylim = c(0.15,2.5))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day3.df[meta.day3.df$time=="2880",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day3.4320.plot=ggplot(meta.day3.df[meta.day3.df$time=="4320",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
   coord_cartesian(ylim = c(0.15,2.5))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day3.df[meta.day3.df$time=="4320",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


# Combine plots
all.day3.plot <- plot_grid(
  day3.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day3.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day3.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day3.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day3.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day3.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  extract.legend,
  ncol = 7,
  rel_widths = c(8, 8, 8, 8, 8, 8, 3)
)


# Set the overall size of the combined plot
ggsave("figures/all.day3.plot.pdf", all.day3.plot, width = 49, height = 12)


```


Best looking figs: Day 5 metabolites
```{r}
########## THIS ONE TOO (JUST P VS AB)
# Define a colorblind-friendly palette
color_palette <- brewer.pal(n = 3, name = "Set1")


day5.30.plot=ggplot(meta.day5.df[meta.day5.df$time=="30",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
   coord_cartesian(ylim = c(0.15,2))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day5.df[meta.day5.df$time=="30",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.30.plot

day5.120.plot=ggplot(meta.day5.df[meta.day5.df$time=="120",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
  coord_cartesian(ylim = c(0.15,2))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day5.df[meta.day5.df$time=="120",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day5.960.plot=ggplot(meta.day5.df[meta.day5.df$time=="960",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
  coord_cartesian(ylim = c(0.15,2))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day5.df[meta.day5.df$time=="960",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day5.1440.plot=ggplot(meta.day5.df[meta.day5.df$time=="1440",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
   coord_cartesian(ylim = c(0.15,2))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day5.df[meta.day5.df$time=="1440",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day5.2880.plot=ggplot(meta.day5.df[meta.day5.df$time=="2880",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
   coord_cartesian(ylim = c(0.15,2))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day5.df[meta.day5.df$time=="2880",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.4320.plot=ggplot(meta.day5.df[meta.day5.df$time=="4320",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
   coord_cartesian(ylim = c(0.15,2))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day5.df[meta.day5.df$time=="4320",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Create the grid with modified plots
extract.legend <- get_legend(
  day3.30.plot + theme(legend.box.margin = margin(0, 0, 0, 1, "cm"),
                       legend.key.size = unit(100, "pt"),
                       legend.text = element_text(size=20),
                       legend.title= element_blank())
)

# Combine plots
all.day5.plot <- plot_grid(
  day5.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day5.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day5.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day5.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day5.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day5.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  extract.legend,
  ncol = 7,
  rel_widths = c(8, 8, 8, 8, 8, 8, 3)
)



all.day5.plot
# Set the overall size of the combined plot
ggsave("figures/all.day5.plot.pdf", all.day5.plot, width = 49, height = 12)


```



Best looking figs: Day 7 metabolites
```{r}
########## THIS ONE TOO (JUST P VS AB)
# Define a colorblind-friendly palette
color_palette <- brewer.pal(n = 3, name = "Set1")


day7.30.plot=ggplot(meta.day7.df[meta.day7.df$time=="30",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
  coord_cartesian(ylim = c(0.15, 2))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day7.df[meta.day7.df$time=="30",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day7.30.plot

day7.120.plot=ggplot(meta.day7.df[meta.day7.df$time=="120",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
    coord_cartesian(ylim = c(0.15, 2))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day7.df[meta.day7.df$time=="120",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.120.plot


day7.960.plot=ggplot(meta.day7.df[meta.day7.df$time=="960",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
    coord_cartesian(ylim = c(0.15, 2))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day7.df[meta.day7.df$time=="960",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.960.plot

day7.1440.plot=ggplot(meta.day7.df[meta.day7.df$time=="1440",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
    coord_cartesian(ylim = c(0.15, 2))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day7.df[meta.day7.df$time=="1440",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.1440.plot

day7.2880.plot=ggplot(meta.day7.df[meta.day7.df$time=="2880",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
    coord_cartesian(ylim = c(0.15, 2))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day7.df[meta.day7.df$time=="2880",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day7.2880.plot

day7.4320.plot=ggplot(meta.day7.df[meta.day7.df$time=="4320",], aes(x = treatment, y = od_corrected, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "right")+
    coord_cartesian(ylim = c(0.15, 2))+
   scale_fill_manual(values = color_palette)+
 geom_hline(data = meta.day7.df[meta.day7.df$time=="4320",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.4320.plot

# Create the grid with modified plots
extract.legend <- get_legend(
  day3.30.plot + theme(legend.box.margin = margin(0, 0, 0, 1, "cm"),
                       legend.key.size = unit(100, "pt"),
                       legend.text = element_text(size=20),
                       legend.title= element_blank())
)

# Combine plots
all.day7.plot <- plot_grid(
  day7.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day7.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day7.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day7.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day7.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day7.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  extract.legend,
  ncol = 7,
  rel_widths = c(8, 8, 8, 8, 8, 8, 3)
)



all.day7.plot
# Set the overall size of the combined plot
ggsave("figures/all.day7.plot.pdf", all.day7.plot, width = 49, height = 12)


```




```{r}

# Fit ANCOVA model with random effect
model <- lmer(od_corrected ~ treatment + hfa + (1 | deploy), data = meta.day1.df)

# Extract model summary
summary(model)




```


```{r}
# Filter the dataframe based on the treatments you want to compare
filtered_df <- meta.day1.long.df[c("Phyllosphere", "Pack.ID", "Deployment", "Metabolite", "HFA", "Time..min.", "OD_corrected", "OD_value")]


# Calculate the mean OD_corrected values for each treatment and time point
mean_values <- aggregate(. ~ Phyllosphere + Pack.ID + Deployment + Metabolite + HFA + Time..min.+ OD_corrected, data = filtered_df, FUN = mean)

# Plot the mean OD_corrected values against the time points
ggplot(mean_values, aes(x = HFA, y = OD_value, fill=Phyllosphere, linetype= Phyllosphere)) +
  geom_boxplot(aes(group = interaction(Pack.ID, Deployment))) +
  facet_grid(Metabolite~OD_corrected)+
  ggtitle("Mean OD_value at each OD_corrected for different treatments") +
  theme_minimal() 




funique(filtered_data$HFA)
unique(filtered_data$Well.ID)
unique(meta.day1.long.df$Phyllosphere)
unique(meta.day1.long.df$Well.ID)
filtered_data <- meta.day1.long.df[meta.day1.long.df$HFA != "" &
                                    meta.day1.long.df$Well.ID != "" &
                                    meta.day1.long.df$Phyllosphere %in% c("P", "AB"), ]



```


```{r}
# Calculate the mean OD_corrected values for each treatment and time point
mean_values <- aggregate(. ~ Phyllosphere + Tree_ID + Deployment + Metabolite + HFA+ Time..min., data = filtered_df, FUN = mean)

# Plot the mean OD_corrected values against the time points

filtered_mean_values <- mean_values[mean_values$Deployment == "HOK 1", ]

ggplot(filtered_mean_values, aes(x = HFA , y = OD_value)) +
  geom_boxplot(aes(color = interaction(Phyllosphere, Tree_ID, Deployment))) +
  labs(x = "Time (min)", y = "Mean OD_corrected") +
  facet_grid(~Metabolite)+
  ggtitle("Mean OD_corrected at each time point for different treatments") +
  ggtitle("Deployment Site HOK 1")+
Fig.formatting

# Plot the mean OD_corrected values against the time points
ggplot(mean_values, aes(x = Time..min., y = OD_30_corrected)) +  # Update with the desired time point column
  geom_line(aes(color = interaction(Phyllosphere, Tree_ID, Deployment, Metabolite))) +
  labs(x = "Time (min)", y = "Mean OD_corrected") +
  ggtitle("Mean OD_corrected at each time point for different treatments") +
  theme_minimal()


```


```{r}
meta.day1.df$Tree_ID <- gsub("[0-9-]", "", meta.day1.df$Tree_ID)
meta.day1.df$Deployment <- gsub("[0-9]", "", meta.day1.df$Deployment )

library(tidyr)

# Pivot the specific columns to long format
meta.day1.long.df=pivot_longer(meta.day1.df, cols = matches("^OD_\\d+_corrected$"), names_to = "OD_corrected", values_to = "OD_value")

# Calculate average OD_value based on groups
avg.day1.long.df<- meta.day1.long.df %>%
  distinct(Phyllosphere, Tree_ID, Deployment, Metabolite, HFA, OD_corrected, OD_value,) %>%
  mutate(across(.cols = everything(), .fns = mean, na.rm = TRUE, .names = "avg_{.col}"))

avg.day1.long.df$OD_corrected=as.factor(as.character(avg.day1.long.df$OD_corrected))
avg.day1.long.df$OD_corrected <- factor(avg.day1.long.df$OD_corrected, 
                                        levels = c("OD_30_corrected", "OD_2_corrected", "OD_16_corrected", "OD_24_corrected", "OD_48_corrected", "OD_72_corrected"))


avg.day1.long.df %>%
  na.omit() %>%
  group_by(Phyllosphere, Tree_ID, Deployment, Metabolite, HFA, OD_corrected, .keep_all = TRUE) %>%
  summarise(avg_OD_value = mean(OD_value, na.rm = TRUE), .groups = "drop")


levels(avg.day1.long.df$OD_corrected)
```


```{r}
filtered_data <- avg.day1.long.df[avg.day1.long.df$Phyllosphere %in% c("P", "AB"), ]
filtered_data <- avg.day1.long.df[filtered_data$HFA %in% c("A", "H"), ]

ggplot(avg.day1.long.df[avg.day1.long.df$Deployment == "HOK", ], aes(x = OD_corrected, y = OD_value, group = interaction(Tree_ID, Deployment, Metabolite, HFA), fill = interaction(Tree_ID, Deployment, Metabolite, HFA))) +
  geom_boxplot() +
  facet_grid(~Phyllosphere)+
  xlab("Time") +
  ylab("Optical density (OD)") +
  Fig.formatting +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

unique(avg.day1.long.df$HFA)
unique(avg.day1.long.df$Phyllosphere)
unique(filtered_data$Tree_ID)

filtered_data <- avg.day1.long.df[avg.day1.long.df$HFA != "", ]
filtered_data$Tree_ID <- trimws(filtered_data$Tree_ID)

unique(filtered_data$HFA)

```


```{r}
# Violin plot
ggplot(meta.df[(meta.df$Deployment=="HOK 2"),], aes(x = HFA, y = Corrected_OD, fill = Treatment)) +
  geom_violin() +
  facet_grid(Decomp_day ~ Time_min) +
  xlab("ID") +
  ylab("OD Value")


met.hok2.plot=ggplot(meta.df[(meta.df$Deployment=="HOK 2"),], aes(x=Time_min, y=Corrected_OD, color=Treatment, group=Well_ID, linetype=Treatment))+
geom_line()+
  scale_color_manual(values = c("purple", "mediumseagreen")) + 
  scale_fill_manual(values = c("purple", "mediumseagreen")) +
  facet_grid(HFA~Decomp_day, scales= "free_x")+
  ggtitle("HOK 2") +
  ylab("Optical Density (590 nm)") +
  Fig.formatting

met.hok2.plot

met.hok1.plot=ggplot(meta.df[(meta.df$Deployment=="HOK 1"),], aes(x=Time_min, y=Corrected_OD, color=Treatment, group=Well_ID, linetype=Treatment))+
geom_line()+
  scale_color_manual(values = c("purple", "mediumseagreen")) + 
  scale_fill_manual(values = c("purple", "mediumseagreen")) +
  facet_grid(HFA~Decomp_day, scales= "free_x")+
  ggtitle("HOK 1") +
  ylab("Optical Density (590 nm)") +
  Fig.formatting

met.hok1.plot

met.sek1.plot=ggplot(meta.df[(meta.df$Deployment=="SEK 1"),], aes(x=Time_min, y=Corrected_OD, color=Treatment, group=Well_ID, linetype=Treatment))+
geom_line()+
  scale_color_manual(values = c("purple", "mediumseagreen")) + 
  scale_fill_manual(values = c("purple", "mediumseagreen")) +
  facet_grid(HFA~Decomp_day, scales= "free_x")+
  ggtitle("SEK 1") +
  ylab("Optical Density (590 nm)") +
  Fig.formatting

met.sek2.plot=ggplot(meta.df[(meta.df$Deployment=="SEK 2"),], aes(x=Time_min, y=Corrected_OD, color=Treatment, group=Well_ID, linetype=Treatment))+
geom_line()+
  scale_color_manual(values = c("purple", "mediumseagreen")) + 
  scale_fill_manual(values = c("purple", "mediumseagreen")) +
  facet_grid(HFA~Decomp_day, scales= "free_x")+
  ggtitle("SEK 2") +
  ylab("Optical Density (590 nm)") +
  Fig.formatting

met.sek2.plot
```


```{r}
##Hoko only
hoko.subset1 <- meta.df[meta.df$deployment == "HOK 1", ]
hoko.subset2 <- meta.df[meta.df$deployment == "HOK 2", ]

hoko_subset.df <- rbind(hoko.subset1, hoko.subset2)
color_palette <- c("#1f77b4", "#ff7f0e")


met.hoko1.plot <- ggplot(hoko_subset.df[(hoko_subset.df$deployment=="HOK 1"),], aes(x = Time, y = OD, color = Treatment, group = Pack_ID, linetype = Treatment)) +
geom_line()+
  scale_color_manual(values = color_palette) +
  scale_fill_manual(values = color_palette) +
  facet_grid(HFA~Decomp) +
  ggtitle("Deployed at HOK 1") +
  ylab("Optical Density (590 nm)") +
  Fig.formatting


# Subset the data for Away and Deployment == "HOK 1"
hoko.present.subset1 <- hoko_subset.df[hoko_subset.df$HFA == "H" & hoko_subset.df$deployment == "HOK 1", ]

# Create the bar plot
hok1.home=ggplot(hoko.present.subset1, aes(x = Time, y = OD, color = Treatment, linetype=Treatment, group = Pack_ID)) +
 geom_line()+
  scale_color_manual(values = color_palette) +
  scale_fill_manual(values = color_palette) +
  facet_grid(~Decomp) +
  ggtitle("Home packs deployed at HOK 1") +
  ylab("Optical Density (590 nm)") +
  Fig.formatting

# Subset the data for Home and Deployment == "HOK 1"
hoko.away.subset1 <- hoko_subset.df[hoko_subset.df$HFA == "A" & hoko_subset.df$deployment == "HOK 1", ]

hok1.away=ggplot(hoko.away.subset1, aes(x = Time, y = OD, color = Treatment, linetype=Treatment, group = Pack_ID)) +
 geom_line()+
  scale_color_manual(values = color_palette) +
  scale_fill_manual(values = color_palette) +
  facet_grid(~Decomp) +
  ggtitle("Away packs deployed at HOK 1") +
  ylab("Optical Density (590 nm)") +
  Fig.formatting

plot_grid(hok1.away, hok1.home, nrow =2)
```


```{r}
met.hoko1.boxplot

met.hoko2.plot <- ggplot(hoko_subset.df[(hoko_subset.df$deployment=="HOK 2"),], aes(x = Time, y = OD, color = Treatment, group = Pack_ID, linetype = Treatment)) +
geom_line()+
  scale_color_manual(values = color_palette) +
  scale_fill_manual(values = color_palette) +
  facet_grid(HFA~Decomp) +
  ggtitle("Deployed at HOK 2") +
  ylab("Optical Density (590 nm)") +
  Fig.formatting

met.hoko2.plot

met.allhoko<-plot_grid(
  met.hoko1.plot+ theme(legend.position = "none"),
  met.hoko2.plot+ theme(legend.position = "none"), extract.legend,
  rel_widths = c(8,8,3), nrow=2)

met.allhoko

##SEKIU only
sekiu.subset1 <- aggregate.df[aggregate.df$deployment == "SEK 1", ]
sekiu.subset2 <- aggregate.df[aggregate.df$deployment == "SEK 2", ]
sekiu_subset.df <- rbind(sekiu.subset1, sekiu.subset2)


met.sek1.plot <- ggplot(sekiu_subset.df[(sekiu_subset.df$deployment=="SEK 1"),], aes(x = Time, y = OD, color = Treatment, group = Pack_ID, linetype = Treatment)) +
geom_line()+
  scale_color_manual(values = color_palette) +
  scale_fill_manual(values = color_palette) +
  facet_grid(HFA~Decomp) +
  ggtitle("Deployed at SEK 1") +
  ylab("Optical Density (590 nm)") +
  Fig.formatting

met.sek1.plot


met.sek2.plot <- ggplot(sekiu_subset.df[(sekiu_subset.df$deployment=="SEK 2"),], aes(x = Time, y = OD, color = Treatment, group = Pack_ID, linetype = Treatment)) +
geom_line()+
  scale_color_manual(values = color_palette) +
  scale_fill_manual(values = color_palette) +
  facet_grid(HFA~Decomp) +
  ggtitle("Deployed at SEK 2") +
  ylab("Optical Density (590 nm)") +
  Fig.formatting

met.sek2.plot

extract.legend <- get_legend(
  # create some space to the left of the legend
  met.sek1.plot + theme(legend.box.margin = margin(0, 0, 0, 10)))

met.alltimes<-plot_grid(
  met.hoko1.plot,
  met.hoko2.plot,
  met.sek1.plot,
  met.sek2.plot,
  rel_widths = c(8,8,8,8,3), nrow=2, ncol=2)

met.alltimes

```

