---
title: "RAD Alder 2023"
author: "Cody Spiegel"
date: "`r Sys.Date()`"
output: word_document
---

### Project overview
For Project ‘RAD’ Alder, gamma irradiated vs. non-irradiated leaf packs deployed at “home” or “away” sites tests the role of phyllosphere microbial communities on home-field advantage patterns in leaf litter decomposition. We utilized a reciprocal transplant experiment between the Hoko and Sekiu rivers on the Olympic Peninsula under four treatment conditions: phyllosphere present-home (PH), phyllosphere present-away (PAW), phyllosphere absent-home (ABH), phyllosphere absent-away (ABAW) to ask the following questions: 

1) How does dispersal from the regional species pool (i.e., source communities) influence community composition on the leaf packs?
2) How does microbial community composition (alpha and beta diversity) and function differ between the leaf packs, as well as change over the course of decomposition?
3) How does alder stoichiometric and secondary chemical composition differ between the leaf packs, as well as change over the course of decomposition?
4) How does microbial metabolization differ between the leaf packs, as well as change over the course of decomposition?
5) How does particulate leaf mass loss differ between the leaf packs? 


**DATA SETS**  
This data set is among 3 to be generated for the project and focuses on:  

(1) **dissolved organic carbon** (DOC) and  
(2) **Fluorescence and UV-vis absorbance** indices  
(3) **DOC Degradation** from either photo- or microbial decomposition 
(4) **burning effects on plant material decomposition** using % mass loss

```{r setup chunk, setup, include = FALSE, cache=FALSE, message=FALSE, warning=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

# use pacman to load all the packages you are missing!
pacman::p_load('knitr', 'lme4', 'lmerTest', 'tidyverse', 'magrittr', 'effects', 'plyr', 'dplyr', 
               'plotrix', 'car',"gridExtra", "cowplot", "tools", "mgcv", "gratia", "tidymv")

Fig.formatting<-(theme_classic()) +
  theme(text=element_text(size=25),
        legend.text.align = 0,
        panel.border = element_rect(fill=NA, colour = "black", size=1),
        aspect.ratio=1, 
        axis.ticks.length=unit(0.25, "cm"),
        axis.text.y=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20), 
        axis.text.x=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20)) +
  theme(legend.key.size = unit(0.4, "cm")) +
  theme(aspect.ratio=1.3) 




```


All metabolite decomp days
```{r}
meta.day1.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_1_metabolites_raw_corrected.csv")


meta.day3.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_3_metabolites_raw_corrected.csv")


meta.day5.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_5_metabolites_raw_corrected.csv")


meta.day7.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_7_metabolites_raw_corrected.csv")


meta.day11.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_11_metabolites_raw_corrected.csv")


meta.day17.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_17_metabolites_raw_corrected.csv")



#By z-scoring the raw OD values, I can effectively assess the relative changes in microbial metabolism across the wells, regardless of their absolute starting values.

#z-scoring is a reasonable choice for this situation, as it standardizes the data within each well and facilitates comparisons. Note: z-scoring assumes a roughly symmetric distribution, but the raw OD values are right skewed.

```


metabolite assay: z score by metabolite and deployment 
```{r}
#z score
##hok 1
meta.hok1.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metabolite=="HOK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])
meta.hok1.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metaboltie=="SEK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])


meta.hok1.hok.df$z_score= (meta.hok1.hok.df$od_corrected_b_n - mean(meta.hok1.hok.df$od_corrected_b_n)) / sd(meta.hok1.hok.df$od_corrected_b_n)
meta.hok1.sek.df$z_score= (meta.hok1.sek.df$od_corrected_b_n - mean(meta.hok1.sek.df$od_corrected_b_n)) / sd(meta.hok1.sek.df$od_corrected_b_n)

##sek 1
meta.sek1.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metabolite=="SEK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])
meta.sek1.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metaboltie=="HOK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])

meta.sek1.sek.df$z_score= (meta.sek1.sek.df$od_corrected_b_n - mean(meta.sek1.sek.df$od_corrected_b_n)) / sd(meta.sek1.sek.df$od_corrected_b_n)
meta.sek1.hok.df$z_score= (meta.sek1.hok.df$od_corrected_b_n - mean(meta.sek1.hok.df$od_corrected_b_n)) / sd(meta.sek1.hok.df$od_corrected_b_n)

#hok 2 
meta.hok2.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metabolite=="HOK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])
meta.hok2.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metaboltie=="SEK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])

meta.hok2.hok.df$z_score= (meta.hok2.hok.df$od_corrected_b_n - mean(meta.hok2.hok.df$od_corrected_b_n)) / sd(meta.hok2.hok.df$od_corrected_b_n)
meta.hok2.sek.df$z_score= (meta.hok2.sek.df$od_corrected_b_n - mean(meta.hok2.sek.df$od_corrected_b_n)) / sd(meta.hok2.sek.df$od_corrected_b_n)


##sek 2
meta.sek2.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metabolite=="SEK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])
meta.sek2.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metaboltie=="HOK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])

meta.sek2.sek.df$z_score= (meta.sek2.sek.df$od_corrected_b_n - mean(meta.sek2.sek.df$od_corrected_b_n)) / sd(meta.sek2.sek.df$od_corrected_b_n)
meta.sek2.hok.df$z_score= (meta.sek2.hok.df$od_corrected_b_n - mean(meta.sek2.hok.df$od_corrected_b_n)) / sd(meta.sek2.hok.df$od_corrected_b_n)

# Assuming you have datasets named meta.day1.raw.df, meta.day3.raw.df, ..., meta.day17.raw.df
# Put them into a list
metabolite_list <- list(meta.day1.raw.df, meta.day3.raw.df, meta.day5.raw.df, meta.day7.raw.df, meta.day11.raw.df, meta.day17.raw.df)

# Define the conditions for subsetting
meta.time.points <- c("30", "120", "960", "1440", "2880", "4320")
# Define the actual day numbers corresponding to each dataset
actual_day_numbers <- c(1, 3, 5, 7, 11, 17)

# Create an empty list to store subsets
subset_list <- list()

# Loop through datasets
for (i in seq_along(metabolite_list)) {
  # Access each dataset from the metabolite_list
  meta.merge <- metabolite_list[[i]]
  
  # Create a list to store subsets for each day
  day_subsets <- list()
  
  # Loop through time points
  for (time_point in meta.time.points) {
    # Subset the data based on specific conditions
    subset_data <- meta.merge[meta.merge$deploy == "HOK" &
                             meta.merge$metabolite == "HOK" &
                             meta.merge$site_num == "1" & 
                             meta.merge$time == time_point, ]
    
    # Calculate the z-scores within each subset_data
    subset_data$z_score <- (subset_data$od_corrected_b_n - mean(subset_data$od_corrected_b_n)) / sd(subset_data$od_corrected_b_n)
    
    # Store the subset with z-scores in the day_subsets list
    day_subsets[[time_point]] <- subset_data
  }
  
  # Merge all time points for each day into one dataframe
  merged_data <- bind_rows(day_subsets)
  
  # Store the merged dataframe in the subset_list
  day_number <- actual_day_numbers[i]
  subset_list[[paste0("day", day_number)]] <- merged_data
}

# Rename the list elements to indicate the day
day_names <- paste0("day", actual_day_numbers)
names(subset_list) <- day_names

#rename into new dataframes
meta.day1.df=subset_list$day1
meta.day3.df=subset_list$day3
meta.day5.df=subset_list$day5
meta.day7.df=subset_list$day7
meta.day11.df=subset_list$day11
meta.day17.df=subset_list$day17


```


Day 1: plots
```{r}
#color blind palette
library(viridis) 

day1.z.30.plot=ggplot(meta.day1.df[meta.day1.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
   scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day1.z.120.plot=ggplot(meta.day1.df[meta.day1.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.960.plot=ggplot(meta.day1.df[meta.day1.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.1440.plot=ggplot(meta.day1.df[meta.day1.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.2880.plot=ggplot(meta.day1.df[meta.day1.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day1.z.4320.plot=ggplot(meta.day1.df[meta.day1.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Create the grid with modified plots
extract.legend <- get_legend(
  day1.z.4320.plot + theme(legend.box.margin = margin(1, 1, 1, 1, "cm"),
                       legend.key.size = unit(50, "pt"),
                       legend.text = element_text(size=20),
                       legend.title= element_blank())
)


# Combine plots
all.z.day1.plot <- plot_grid(
  day1.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


all.z.day1.plot
# Set the overall size of the combined plot
ggsave("figures/all.z.day1.plot.pdf", all.z.day1.plot, width = 49, height = 10)



```


Day 3: plots
```{r}

day3.z.30.plot=ggplot(meta.day3.df[meta.day3.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
  scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day3.z.120.plot=ggplot(meta.day3.df[meta.day3.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
     scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.960.plot=ggplot(meta.day3.df[meta.day3.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
     scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.1440.plot=ggplot(meta.day3.df[meta.day3.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.2880.plot=ggplot(meta.day3.df[meta.day3.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day3.z.4320.plot=ggplot(meta.day3.df[meta.day3.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


# Combine plots
all.z.day3.plot <- plot_grid(
  day3.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day3.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day3.plot.pdf", all.z.day3.plot, width = 49, height = 10)



```



Day 5: plots
```{r}

day5.z.30.plot=ggplot(meta.day5.df[meta.day5.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day5.z.120.plot=ggplot(meta.day5.df[meta.day5.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


day5.z.960.plot=ggplot(meta.day5.df[meta.day5.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



day5.z.1440.plot=ggplot(meta.day5.df[meta.day5.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
          scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



day5.z.2880.plot=ggplot(meta.day5.df[meta.day5.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day5.z.4320.plot=ggplot(meta.day5.df[meta.day5.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Combine plots
all.z.day5.plot <- plot_grid(
  day5.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day5.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day5.plot.pdf", all.z.day5.plot, width = 49, height = 10)



```



Day 7: plots
```{r}

day7.z.30.plot=ggplot(meta.day7.df[meta.day7.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
   scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day7.z.30.plot

day7.z.120.plot=ggplot(meta.day7.df[meta.day7.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
     scale_fill_viridis(discrete = TRUE) + 
 geom_hline(data = meta.day7.df[meta.day7.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.120.plot

day7.z.960.plot=ggplot(meta.day7.df[meta.day7.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.960.plot


day7.z.1440.plot=ggplot(meta.day7.df[meta.day7.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.1440.plot

day7.z.2880.plot=ggplot(meta.day7.df[meta.day7.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.2880.plot

day7.z.4320.plot=ggplot(meta.day7.df[meta.day7.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


# Combine plots
all.z.day7.plot <- plot_grid(
  day7.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day7.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day7.plot.pdf", all.z.day7.plot, width = 49, height = 12)



```



Day 11: z scored
```{r}
day11.z.30.plot=ggplot(meta.day11.df[meta.day11.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day11.z.30.plot

day11.z.120.plot=ggplot(meta.day11.df[meta.day11.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day11.z.120.plot

day11.z.960.plot=ggplot(meta.day11.df[meta.day11.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.960.plot


day11.z.1440.plot=ggplot(meta.day11.df[meta.day11.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.1440.plot

day11.z.2880.plot=ggplot(meta.day11.df[meta.day11.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.2880.plot

day11.z.4320.plot=ggplot(meta.day11.df[meta.day11.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


# Combine plots
all.z.day11.plot <- plot_grid(
  day11.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day11.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)

# Set the overall size of the combined plot
ggsave("figures/all.z.day11.plot.pdf", all.z.day11.plot, width = 49, height = 12)



```





Day 17: z scored
```{r}

day17.z.30.plot=ggplot(meta.day17.df[meta.day17.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day17.z.30.plot

day17.z.120.plot=ggplot(meta.day17.df[meta.day17.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day17.z.120.plot

day17.z.960.plot=ggplot(meta.day17.df[meta.day17.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day17.z.960.plot


day17.z.1440.plot=ggplot(meta.day17.df[meta.day17.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


day17.z.1440.plot

day17.z.2880.plot=ggplot(meta.day17.df[meta.day17.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day17.z.2880.plot

day17.z.4320.plot=ggplot(meta.day17.df[meta.day17.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


# Combine plots
all.z.day17.plot <- plot_grid(
  day17.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day17.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
   day17.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14,14,14, 14, 3)
)

# Set the overall size of the combined plot
ggsave("figures/all.z.day17.plot.pdf", all.z.day17.plot, width = 49, height = 12)


```

All metabolite assays plots

```{r}

all.z.days.plot <- plot_grid(
  all.z.day1.plot,
  all.z.day3.plot,
  all.z.day5.plot,
  all.z.day7.plot,
  all.z.day11.plot,
  all.z.day17.plot,
  nrow = 6
)



all.z.days.plot
ggsave("figures/all.z.days.plot.pdf", all.z.days.plot, width = 49, height = 30)

```


Metabolite analysis

```{r}

#make a big metabolite dataframe

merged_df=rbind(meta.day1.raw.df,meta.day3.raw.df,meta.day5.raw.df,meta.day7.raw.df, meta.day11.raw.df)
class(merged_df$time)
merged_df$treatment=as.factor(as.character(merged_df$treatment))
merged_df$tree_id=as.factor(as.character(merged_df$tree_id))
merged_df$tree_num=as.factor(as.integer(merged_df$tree_num))
merged_df$deploy=as.factor(as.character(merged_df$deploy))
merged_df$site_num=as.factor(as.integer(merged_df$site_num))
merged_df$metabolite=as.factor(as.character(merged_df$metabolite))
merged_df$hfa=as.factor(as.character(merged_df$hfa))
merged_df$time=as.numeric(as.integer(merged_df$time))

class(merged_df$metabolite)

# Fit the linear mixed-effects model with nested random effects and fixed effect
model <- lmer(z_score ~ treatment + hfa + treatment * hfa +
              (1 | deploy/tree_id/metabolite),
              data = merged_df[merged_df$decomp_day == "7" & merged_df$time=="4320", ],
              na.action = na.exclude)
anova(model)






```
Day 16 metabolite plots
```{r}

day1.16hr.plot=ggplot(meta.day1.df[meta.day1.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 1")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting +
   theme(legend.position = "none")

day1.16hr.plot

day3.16hr.plot=ggplot(meta.day3.df[meta.day3.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 3")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting +
   theme(legend.position = "none")


day5.16hr.plot=ggplot(meta.day5.df[meta.day5.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "Treatment", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 5")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting+
   theme(legend.position = "none")

day5.16hr.plot

day11.16hr.plot=ggplot(meta.day11.df[meta.day11.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 11")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting+
  theme(legend.position = "none")

day11.16hr.plot

all.16hr.z.plots=plot_grid(
  day1.16hr.plot,
  day3.16hr.plot,
  day5.16hr.plot,
  day11.16hr.plot,
  ncol=5,
  extract.legend,
  rel_widths = c(8,8,8,8,3))


all.day16.z.plots
ggsave("figures/all.16hr.z.plots.pdf", all.16hr.z.plots, width = 26, height = 8)


```

Leaf mass data config
```{r}
mass.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/leaf_mass.csv")

#correction factor
mass.df$final_correct<-0.941*(mass.df$final_mass)-0.00337 #correction

#z score across deployment site (mass within hok 1, mass within hok 2)

##HOK 1
mass.hok1.df<- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$site_num == "1", ])
mass.hok1.df$z_score= (mass.hok1.df$final_correct - mean(mass.hok1.df$final_correct)) / sd(mass.hok1.df$final_correct)
#percent remaining
mass.hok1.df$perc.remain=(mass.hok1.df$final_correct / mass.hok1.df$initial_mass)*100
#z score perc. remaining
mass.hok1.df$z_score_perc=(mass.hok1.df$perc.remain - mean(mass.hok1.df$perc.remain))/sd(mass.hok1.df$perc.remain)
# Apply the asin(sqrt) transformation
mass.hok1.df$arc_sin_remain= asin(sqrt(mass.hok1.df$perc.remain / max(mass.hok1.df$perc.remain)))
# Calculate percent mass loss
mass.hok1.df$percent_loss <- (mass.hok1.df$initial_mass - mass.hok1.df$final_correct) / mass.hok1.df$initial_mass * 100
#z score percent loss
mass.hok1.df$z_score_loss=(mass.hok1.df$percent_loss - mean(mass.hok1.df$percent_loss))/sd(mass.hok1.df$percent_loss)



##HOK 2
mass.hok2.df<- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$site_num == "2", ])
mass.hok2.df$z_score= (mass.hok2.df$final_correct - mean(mass.hok2.df$final_correct)) / sd(mass.hok2.df$final_correct)
#percent remaining
mass.hok2.df$perc.remain=(mass.hok2.df$final_correct / mass.hok2.df$initial_mass)*100
#z score perc. remaining
mass.hok2.df$z_score_perc=(mass.hok2.df$perc.remain - mean(mass.hok2.df$perc.remain))/sd(mass.hok2.df$perc.remain)
# Apply the asin(sqrt) transformation
mass.hok2.df$arc_sin_remain= asin(sqrt(mass.hok2.df$perc.remain / max(mass.hok2.df$perc.remain)))
# Calculate percent mass loss
mass.hok2.df$percent_loss <- (mass.hok2.df$initial_mass - mass.hok2.df$final_correct) / mass.hok2.df$initial_mass * 100
#z score percent loss
mass.hok2.df$z_score_loss=(mass.hok2.df$percent_loss - mean(mass.hok2.df$percent_loss))/sd(mass.hok2.df$percent_loss)


##SEK 1
mass.sek1.df<- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$site_num == "1", ])
mass.sek1.df$z_score= (mass.sek1.df$final_correct - mean(mass.sek1.df$final_correct)) / sd(mass.sek1.df$final_correct)
#percent remaining
mass.sek1.df$perc.remain=(mass.sek1.df$final_correct / mass.sek1.df$initial_mass)*100
#z score perc. remaining
mass.sek1.df$z_score_perc=(mass.sek1.df$perc.remain - mean(mass.sek1.df$perc.remain))/sd(mass.sek1.df$perc.remain)
# Apply the asin(sqrt) transformation
mass.sek1.df$arc_sin_remain= asin(sqrt(mass.sek1.df$perc.remain / max(mass.sek1.df$perc.remain)))
# Calculate percent mass loss
mass.sek1.df$percent_loss <- (mass.sek1.df$initial_mass - mass.sek1.df$final_correct) / mass.sek1.df$initial_mass * 100
#z score percent loss
mass.sek1.df$z_score_loss=(mass.sek1.df$percent_loss - mean(mass.sek1.df$percent_loss))/sd(mass.sek1.df$percent_loss)


##SEK 2
mass.sek2.df<- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$site_num == "2", ])
mass.sek2.df$z_score= (mass.sek2.df$final_correct - mean(mass.sek2.df$final_correct)) / sd(mass.sek2.df$final_correct)
#percent remaining
mass.sek2.df$perc.remain=(mass.sek2.df$final_correct / mass.sek2.df$initial_mass)*100
#z score perc. remaining
mass.sek2.df$z_score_perc=(mass.sek2.df$perc.remain - mean(mass.sek2.df$perc.remain))/sd(mass.sek2.df$perc.remain)
# Apply the asin(sqrt) transformation
mass.sek2.df$arc_sin_remain= asin(sqrt(mass.sek2.df$perc.remain / max(mass.sek2.df$perc.remain)))
# Calculate percent mass loss
mass.sek2.df$percent_loss <- (mass.sek2.df$initial_mass - mass.sek2.df$final_correct) / mass.sek2.df$initial_mass * 100
#z score percent loss
mass.sek2.df$z_score_loss=(mass.sek2.df$percent_loss - mean(mass.sek2.df$percent_loss))/sd(mass.sek2.df$percent_loss)


##merge all sites
mass.merged.df=rbind(mass.hok1.df, mass.hok2.df, mass.sek1.df, mass.sek2.df)
mass.merged.df$treatment=as.factor(as.character(mass.merged.df$treatment))
mass.merged.df$tree_id=as.factor(as.character(mass.merged.df$tree_id))
mass.merged.df$tree_id=as.factor(as.character(mass.merged.df$tree_id))
mass.merged.df$deploy=as.factor(as.character(mass.merged.df$deploy))
mass.merged.df$hfa=as.factor(as.character(mass.merged.df$hfa))
```

Leaf mass plots
```{r}
#plots
#z score plot
leaf.mass.z.plot=
  ggplot(mass.merged.df, aes(x = treatment, y = z_score_loss, fill = factor(hfa))) +
  geom_boxplot() +
  facet_grid(~deploy) +
  labs(x = "treatment", y = "mass loss (%)") +
  Fig.formatting +
  theme(legend.position = "right") +
  scale_fill_viridis(discrete = TRUE) +
  geom_hline(data = mass.merged.df,
             aes(yintercept = mean(z_score_loss),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

#raw precent loss plot
leaf.mass.plot=
  ggplot(mass.merged.df, aes(x = treatment, y = percent_loss, fill = factor(hfa))) +
  geom_boxplot() +
  facet_grid(~deploy) +
  labs(x = "treatment", y = "mass loss (%)") +
  Fig.formatting +
  theme(legend.position = "right") +
  scale_fill_viridis(discrete = TRUE) +
  geom_hline(data = mass.merged.df,
             aes(yintercept = mean(percent_loss),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting
leaf.mass.plot

#no deploy facet, z plot
no.facet.mass.z.loss.plot=ggplot(mass.merged.df, aes(x = treatment, y = z_score_loss, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "treatment", y = "mass loss (%)") +
  Fig.formatting +
  theme(legend.position = "right") +
  scale_fill_viridis(discrete = TRUE) +
  geom_hline(data = mass.merged.df,
             aes(yintercept = mean(z_score_loss),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

#facet by origin and deployment
origin.deploy.mass.z.loss.plot=ggplot(mass.merged.df, aes(x = treatment, y = z_score_loss, fill = factor(hfa))) +
  geom_boxplot() +
  facet_grid(~interaction(deploy, tree_id)) +
  labs(x = "treatment", y = "mass loss (%)") +
  Fig.formatting +
  theme(legend.position = "right") +
  scale_fill_viridis(discrete = TRUE) +
  geom_hline(data = mass.merged.df,
             aes(yintercept = mean(z_score_loss),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting




ggsave("figures/origin.deploy.mass.z.loss.plot.pdf", origin.deploy.mass.z.loss.plot, width = 10, height = 6)



```







