---
title: "RAD Alder 2023"
author: "Cody Spiegel"
date: "`r Sys.Date()`"
output: word_document
---

### Project overview
For Project ‘RAD’ Alder, leaf packs are gamma irradiated or not, and deployed at “home” or “away” sites to test the role of  phyllosphere microbial communities on home-field advantage patterns in leaf litter decomposition between rivers on WA Olympic Peninsula. Using a reciprocal transplant experiment of leaf packs under four treatment conditions: phyllosphere present-home (PH), phyllosphere present-away (PAW), phyllosphere absent-home (ABH), phyllosphere absent-away (ABAW), the following questions are asked: 1) Do the source communities differ between the four treatments, 2) Do community metrics differ between the four treatments?, 3) Does plant compound metabolization differ between the four treatments?, 4) Does decomposition differ between the four treatments? 



**DATA SETS**  
This data set is among 3 to be generated for the project and focuses on:  

(1) **dissolved organic carbon** (DOC) and  
(2) **Fluorescence and UV-vis absorbance** indices  
(3) **DOC Degradation** from either photo- or microbial decomposition 
(4) **burning effects on plant material decomposition** using % mass loss

```{r setup chunk, setup, include = FALSE, cache=FALSE, message=FALSE, warning=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

# use pacman to load all the packages you are missing!
pacman::p_load('knitr', 'lme4', 'lmerTest', 'tidyverse', 'magrittr', 'effects', 'plyr', 'dplyr', 
               'plotrix', 'car',"gridExtra", "cowplot", "tools", "mgcv", "gratia", "tidymv")

Fig.formatting<-(theme_classic()) +
  theme(text=element_text(size=25),
        legend.text.align = 0,
        panel.border = element_rect(fill=NA, colour = "black", size=1),
        aspect.ratio=1, 
        axis.ticks.length=unit(0.25, "cm"),
        axis.text.y=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20), 
        axis.text.x=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20)) +
  theme(legend.key.size = unit(0.4, "cm")) +
  theme(aspect.ratio=1.3) 




```


All metabolite decomp days
```{r}
meta.day1.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_1_metabolites_corrected.csv")
meta.day3.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_3_metabolites_corrected.csv")
meta.day5.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_5_metabolites_corrected.csv")
meta.day7.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_7_metabolites_corrected.csv")

meta.day1.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_1_metabolites_raw_corrected.csv")
meta.day3.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_3_metabolites_raw_corrected.csv")
meta.day5.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_5_metabolites_raw_corrected.csv")
meta.day7.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_7_metabolites_raw_corrected.csv")
meta.day11.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_11_metabolites_raw_corrected.csv")
meta.day17.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_17_metabolites_raw_corrected.csv")


#By z-scoring the raw OD values, I can effectively assess the relative changes in microbial metabolism across the wells, regardless of their absolute starting values.

#z-scoring is a reasonable choice for this situation, as it standardizes the data within each well and facilitates comparisons. Note: z-scoring assumes a roughly symmetric distribution, but the raw OD values are right skewed.

```

Metabolic Assay
```{r}
meta.day1.df=read.csv("/Users/Cody1/Downloads/'RAD' Alder Metadata - day_1_metabolites_corrected.csv")

aggra.day1.df <- meta.day1.df %>%
  distinct(treatment, tree_id, deploy, metabolite, hfa) %>%
  group_by(treatment, tree_id, deploy, metabolite, hfa) %>%
  mutate(mean.od = mean(meta.day1.df$od_corrected, na.rm = TRUE))


ggplot(meta.day1.df, aes(x = time, y = od_corrected, color = hfa, group = interaction(tree_id, metabolite, deploy))) +
  geom_line()+
  facet_grid(~treatment) +
  labs(x = "Time", y = "Mean OD value") +
  theme_minimal() +
  theme(legend.position = "bottom")

library(RColorBrewer)

ggplot(meta.day1.df, aes(x = treatment, y = od_corrected, fill = treatment)) +
  geom_boxplot() +
  facet_grid(time~tree_id  + deploy +metabolite+hfa) +
  scale_fill_manual(values = brewer.pal(n = 3, name = "Dark2"))+
  labs(x = "Treatment", y = "OD Corrected") +
  theme_minimal() +
  theme(legend.position = "bottom")


###visualize Day 1
ggplot(meta.day1.df, aes(x = treatment, y = od_corrected, fill = treatment)) +
  geom_boxplot() +
  facet_grid(time ~ deploy + metabolite + hfa) +
  scale_fill_manual(values = brewer.pal(n = 3, name = "Dark2")) +
  labs(x = "Treatment", y = "OD Corrected") +
  theme_minimal() +
  theme(legend.position = "bottom")

###############################

# Calculate the overall average od_corrected value
overall_avg <- mean(meta.day1.df[meta.day1.df$time == "1440", ]$od_corrected)

ggplot(meta.day1.df[meta.day1.df$time == "1440", ], aes(x = hfa, y = od_corrected, group = interaction(tree_id, deploy, metabolite), fill = interaction(interaction(tree_id, deploy, metabolite)))) +
  geom_boxplot() +
  facet_grid(time ~ treatment) +
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting +
  theme(legend.position = "bottom") +

  # Add a horizontal reference line at the overall average
  geom_hline(yintercept = overall_avg,
             linetype = "dotted",
             color = "red",
             size=1.3) 

#interpret the boxplots relative to the horizontal reference line. Boxplots above the line indicate greater od_corrected values, while boxplots below the line indicate lesser od_corrected values.

ggplot(meta.day1.df[meta.day1.df$time=="1440",], aes(x = treatment, y = od_corrected, group= interaction(tree_id,deploy,metabolite), fill = interaction (interaction(tree_id, deploy, metabolite)))) +
  geom_boxplot() +
  facet_grid(time~hfa)+
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "bottom")+
  geom_hline(data = meta.day1.df[meta.day1.df$time=="1440",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted") 

###### THIS ONE


ggplot(meta.day1.df[meta.day1.df$time=="1440",], aes(x = hfa, y = od_corrected, group= interaction(tree_id,deploy,metabolite), fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~treatment)+
  scale_fill_manual(values = color_palette)+
  labs(x = "Treatment", y = "OD Corrected") +
  Fig.formatting+
  theme(legend.position = "bottom")+
  geom_hline(data = meta.day1.df[meta.day1.df$time=="1440",],
             aes(yintercept = mean(od_corrected),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")
```



Day 1: z scored
```{r}
mean_od_raw.day1 <- mean(meta.day1.raw.df$od_corrected, na.rm=T)
sd_od_raw.day1 <- sd(meta.day1.raw.df$od_corrected, na.rm=T)
meta.day1.raw.df$z_score <- (meta.day1.raw.df$od_corrected - mean_od_raw.day1) / sd_od_raw.day1

#color blind palette
library(viridis) 

day1.z.30.plot=ggplot(meta.day1.raw.df[meta.day1.raw.df$time=="30",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2, 3.4))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
   scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.raw.df[meta.day1.raw.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day1.z.30.plot

day1.z.120.plot=ggplot(meta.day1.raw.df[meta.day1.raw.df$time=="120",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2, 3.4))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.raw.df[meta.day1.raw.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.960.plot=ggplot(meta.day1.raw.df[meta.day1.raw.df$time=="960",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2, 3.4))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.raw.df[meta.day1.raw.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.1440.plot=ggplot(meta.day1.raw.df[meta.day1.raw.df$time=="1440",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2, 3.4))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.raw.df[meta.day1.raw.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.2880.plot=ggplot(meta.day1.raw.df[meta.day1.raw.df$time=="2880",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2, 3.4))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.raw.df[meta.day1.raw.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day1.z.4320.plot=ggplot(meta.day1.raw.df[meta.day1.raw.df$time=="4320",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2, 3.4))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.raw.df[meta.day1.raw.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Create the grid with modified plots
extract.legend <- get_legend(
  day1.z.4320.plot + theme(legend.box.margin = margin(1, 1, 1, 1, "cm"),
                       legend.key.size = unit(50, "pt"),
                       legend.text = element_text(size=20),
                       legend.title= element_blank())
)


# Combine plots
all.z.day1.plot <- plot_grid(
  day1.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


all.z.day1.plot
# Set the overall size of the combined plot
ggsave("figures/all.z.day1.plot.pdf", all.z.day1.plot, width = 49, height = 10)



```


Day 3: z scored
```{r}
mean_od_raw.day3 <- mean(meta.day3.raw.df$od_corrected, na.rm=T)
sd_od_raw.day3 <- sd(meta.day3.raw.df$od_corrected, na.rm=T)
meta.day3.raw.df$z_score <- (meta.day3.raw.df$od_corrected - mean_od_raw.day3) / sd_od_raw.day3


day3.z.30.plot=ggplot(meta.day3.raw.df[meta.day3.raw.df$time=="30",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2, 3))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
  scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.raw.df[meta.day3.raw.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day3.z.30.plot

day3.z.120.plot=ggplot(meta.day3.raw.df[meta.day3.raw.df$time=="120",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
   coord_cartesian(ylim = c(-1.2, 3))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
     scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.raw.df[meta.day3.raw.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.960.plot=ggplot(meta.day3.raw.df[meta.day3.raw.df$time=="960",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
   coord_cartesian(ylim = c(-1.2, 3))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
     scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.raw.df[meta.day3.raw.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.1440.plot=ggplot(meta.day3.raw.df[meta.day3.raw.df$time=="1440",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
   coord_cartesian(ylim = c(-1.2, 3))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.raw.df[meta.day3.raw.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.2880.plot=ggplot(meta.day3.raw.df[meta.day3.raw.df$time=="2880",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
   coord_cartesian(ylim = c(-1.2, 3))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.raw.df[meta.day3.raw.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day3.z.4320.plot=ggplot(meta.day3.raw.df[meta.day3.raw.df$time=="4320",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
   coord_cartesian(ylim = c(-1.2, 3))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.raw.df[meta.day3.raw.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


# Combine plots
all.z.day3.plot <- plot_grid(
  day3.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day3.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day3.plot.pdf", all.z.day3.plot, width = 49, height = 12)



```



Day 5: z scored
```{r}
mean_od_raw.day5 <- mean(meta.day5.raw.df$od_corrected, na.rm=T)
sd_od_raw.day5 <- sd(meta.day5.raw.df$od_corrected, na.rm=T)
meta.day5.raw.df$z_score <- (meta.day5.raw.df$od_corrected - mean_od_raw.day5) / sd_od_raw.day5


day5.z.30.plot=ggplot(meta.day5.raw.df[meta.day5.raw.df$time=="30",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2,3.75))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.raw.df[meta.day5.raw.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.30.plot

day5.z.120.plot=ggplot(meta.day5.raw.df[meta.day5.raw.df$time=="120",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2,3.75))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.raw.df[meta.day5.raw.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day5.z.120.plot

day5.z.960.plot=ggplot(meta.day5.raw.df[meta.day5.raw.df$time=="960",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2,3.75))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.raw.df[meta.day5.raw.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day5.z.960.plot


day5.z.1440.plot=ggplot(meta.day5.raw.df[meta.day5.raw.df$time=="1440",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2,3.75))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
          scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.raw.df[meta.day5.raw.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day5.z.1440.plot

day5.z.2880.plot=ggplot(meta.day5.raw.df[meta.day5.raw.df$time=="2880",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2,3.75))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.raw.df[meta.day5.raw.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.2880.plot

day5.z.4320.plot=ggplot(meta.day5.raw.df[meta.day5.raw.df$time=="4320",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2,3.75))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.raw.df[meta.day5.raw.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Combine plots
all.z.day5.plot <- plot_grid(
  day5.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day5.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day5.plot.pdf", all.z.day5.plot, width = 49, height = 12)



```



Day 7: z scored
```{r}
mean_od_raw.day7 <- mean(meta.day7.raw.df$od_corrected, na.rm=T)
sd_od_raw.day7 <- sd(meta.day7.raw.df$od_corrected, na.rm=T)
meta.day7.raw.df$z_score <- (meta.day7.raw.df$od_corrected - mean_od_raw.day7) / sd_od_raw.day7

day7.z.30.plot=ggplot(meta.day7.raw.df[meta.day7.raw.df$time=="30",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~metabolite)+
  coord_cartesian(ylim = c(-1.2,4.25))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
   scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.raw.df[meta.day7.raw.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day7.z.30.plot

day7.z.120.plot=ggplot(meta.day7.raw.df[meta.day7.raw.df$time=="120",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
    facet_grid(~metabolite)+
    coord_cartesian(ylim = c(-1.2,4.25))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+ 
     scale_fill_viridis(discrete = TRUE) + 
 geom_hline(data = meta.day7.raw.df[meta.day7.raw.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.120.plot

day7.z.960.plot=ggplot(meta.day7.raw.df[meta.day7.raw.df$time=="960",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
    facet_grid(~metabolite)+
    coord_cartesian(ylim = c(-1.2,4.25))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.raw.df[meta.day7.raw.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.960.plot


day7.z.1440.plot=ggplot(meta.day7.raw.df[meta.day7.raw.df$time=="1440",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
    facet_grid(~metabolite)+
    coord_cartesian(ylim = c(-1.2,4.25))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.raw.df[meta.day7.raw.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.1440.plot

day7.z.2880.plot=ggplot(meta.day7.raw.df[meta.day7.raw.df$time=="2880",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
    facet_grid(~metabolite)+
    coord_cartesian(ylim = c(-1.2,4.25))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.raw.df[meta.day7.raw.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.2880.plot

day7.z.4320.plot=ggplot(meta.day7.raw.df[meta.day7.raw.df$time=="4320",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
    facet_grid(~metabolite)+
    coord_cartesian(ylim = c(-1.2,4.25))+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.raw.df[meta.day7.raw.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


# Combine plots
all.z.day7.plot <- plot_grid(
  day7.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day7.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day7.plot.pdf", all.z.day7.plot, width = 49, height = 12)



```



Day 11: z scored
```{r}
mean_od_raw.day11 <- mean(meta.day11.raw.df$od_corrected, na.rm=T)
sd_od_raw.day11 <- sd(meta.day11.raw.df$od_corrected, na.rm=T)
meta.day11.raw.df$z_score <- (meta.day11.raw.df$od_corrected - mean_od_raw.day11) / sd_od_raw.day11

day11.z.30.plot=ggplot(meta.day11.raw.df[meta.day11.raw.df$time=="30",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~metabolite)+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.raw.df[meta.day11.raw.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day11.z.30.plot

day11.z.120.plot=ggplot(meta.day11.raw.df[meta.day11.raw.df$time=="120",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
    facet_grid(~metabolite)+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.raw.df[meta.day11.raw.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day11.z.120.plot

day11.z.960.plot=ggplot(meta.day11.raw.df[meta.day11.raw.df$time=="960",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.raw.df[meta.day11.raw.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.960.plot


day11.z.1440.plot=ggplot(meta.day11.raw.df[meta.day11.raw.df$time=="1440",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.raw.df[meta.day11.raw.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.1440.plot

day11.z.2880.plot=ggplot(meta.day11.raw.df[meta.day11.raw.df$time=="2880",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~metabolite)+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.raw.df[meta.day11.raw.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.2880.plot

day11.z.4320.plot=ggplot(meta.day11.raw.df[meta.day11.raw.df$time=="4320",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~metabolite)+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.raw.df[meta.day11.raw.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


# Combine plots
all.z.day11.plot <- plot_grid(
  day11.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day11.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)

# Set the overall size of the combined plot
ggsave("figures/all.z.day11.plot.pdf", all.z.day11.plot, width = 49, height = 12)



```





Day 17: z scored
```{r}
mean_od_raw.day17 <- mean(meta.day17.raw.df$od_corrected, na.rm=T)
sd_od_raw.day17 <- sd(meta.day17.raw.df$od_corrected, na.rm=T)
meta.day17.raw.df$z_score <- (meta.day17.raw.df$od_corrected - mean_od_raw.day17) / sd_od_raw.day17

day17.z.30.plot=ggplot(meta.day17.raw.df[meta.day17.raw.df$time=="30",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~metabolite)+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.raw.df[meta.day17.raw.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day17.z.30.plot

day17.z.120.plot=ggplot(meta.day17.raw.df[meta.day17.raw.df$time=="120",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
    facet_grid(~metabolite)+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.raw.df[meta.day17.raw.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day17.z.120.plot

day17.z.960.plot=ggplot(meta.day17.raw.df[meta.day17.raw.df$time=="960",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.raw.df[meta.day17.raw.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day17.z.960.plot


day17.z.1440.plot=ggplot(meta.day17.raw.df[meta.day17.raw.df$time=="1440",], aes(x = hfa, y = z_score, fill = factor(treatment))) +
  geom_boxplot() +
   facet_grid(~metabolite)+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.raw.df[meta.day17.raw.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


day17.z.1440.plot

day11.z.2880.plot=ggplot(meta.day11.raw.df[meta.day11.raw.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  facet_grid(~deploy)+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
    coord_cartesian(ylim = c(-1.25,4))+
 geom_hline(data = meta.day11.raw.df[meta.day11.raw.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.2880.plot

day11.z.4320.plot=ggplot(meta.day11.raw.df[meta.day11.raw.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  facet_grid(~deploy)+
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  Fig.formatting+
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
    coord_cartesian(ylim = c(-1.25,4))+
 geom_hline(data = meta.day11.raw.df[meta.day11.raw.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


# Combine plots
all.z.day17.plot <- plot_grid(
  day17.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day17.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 5,
  rel_widths = c(14, 14, 14,14, 3)
)

# Set the overall size of the combined plot
ggsave("figures/all.z.day17.plot.pdf", all.z.day17.plot, width = 49, height = 12)



```

All metabolite assays plots

```{r}

all.z.days.plot <- plot_grid(
  all.z.day1.plot,
  all.z.day3.plot,
  all.z.day5.plot,
  all.z.day7.plot,
  all.z.day11.plot,
  all.z.day17.plot,
  nrow = 6
)



all.z.days.plot
ggsave("figures/all.z.days.plot.pdf", all.z.days.plot, width = 49, height = 30)

```


Metabolite analysis

```{r}

#make a big metabolite dataframe

merged_df=rbind(meta.day1.raw.df,meta.day3.raw.df,meta.day5.raw.df,meta.day7.raw.df, meta.day11.raw.df)
class(merged_df$time)
merged_df$treatment=as.factor(as.character(merged_df$treatment))
merged_df$tree_id=as.factor(as.character(merged_df$tree_id))
merged_df$tree_num=as.factor(as.integer(merged_df$tree_num))
merged_df$deploy=as.factor(as.character(merged_df$deploy))
merged_df$site_num=as.factor(as.integer(merged_df$site_num))
merged_df$metabolite=as.factor(as.character(merged_df$metabolite))
merged_df$hfa=as.factor(as.character(merged_df$hfa))
merged_df$time=as.numeric(as.integer(merged_df$time))

class(merged_df$metabolite)

# Fit the linear mixed-effects model with nested random effects and fixed effect
model <- lmer(z_score ~ treatment + hfa + treatment * hfa +
              (1 | deploy/tree_id/metabolite),
              data = merged_df[merged_df$decomp_day == "7" & merged_df$time=="4320", ],
              na.action = na.exclude)
anova(model)






```

Leaf mass

```{r}
mass.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/leaf_mass.csv")

#correction factor
mass.df$final_correct<-0.941*(mass.df$final_mass)-0.00337 #correction

#z score
mass.df$z_score= (mass.df$final_correct - mean(mass.df$final_correct)) / sd(mass.df$final_correct)

#percent remaining
mass.df$perc.remain=(mass.df$final_correct / mass.df$initial_mass)*100

#z score perc. remaining
mass.df$z_score_perc=(mass.df$perc.remain - mean(mass.df$perc.remain))/sd(mass.df$perc.remain)

#plot
ggplot(mass.df, aes(x = hfa, y = z_score_perc, fill = factor(treatment))) +
  geom_boxplot() +
  facet_grid(~factor(deloy))+
  labs(x = "Treatment", y = "Z score (% mass remaining)") +
  Fig.formatting+
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = mass.df,
             aes(yintercept = mean(z_score_perc),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting



```





