---
title: "Zapped out of existence! Home-field advantage decomposition patterns persist after removal of leaf-associated microbes"
author: "CJ Spiegel et al. 2023, in prep"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
  editor_options:
      chunk_output_type: console
---

### Project overview
For Project ‘RAD’ Alder, gamma irradiated (phyllosphere present (P)) vs. non-irradiated (phyllosphere absent (AB)) leaf packs deployed at “home” or “away” sites tests the role of phyllosphere microbial communities on home-field advantage patterns in leaf litter decomposition. We utilized a reciprocal transplant experiment between the Hoko and Sekiu rivers on the Olympic Peninsula under four treatment conditions: phyllosphere present-home (PH), phyllosphere present-away (PAW), phyllosphere absent-home (ABH), phyllosphere absent-away (ABAW) to ask the following questions: 

1) How does dispersal from the regional species pool (i.e., source communities) influence community composition on the leaf packs?
2) How does microbial community composition (alpha and beta diversity) and function differ between the leaf packs, as well as change over the course of decomposition?
3) How does alder stoichiometric and secondary chemical composition differ between the leaf packs, as well as change over the course of decomposition?
4) How does microbial metabolization differ between the leaf packs, as well as change over the course of decomposition?
5) How does particulate leaf mass loss differ between the leaf packs? 


**DATA SETS**  
This data set focuses on:  

(1) **Metabolism assay** 
(2) **Leaf mass loss** using % mass loss and HFAI
(3) **Alder C and N isotopic composition**
(4) **Alder secondary chemistry (via LC-MS/MS)** 
```{r setup chunk, setup, include = FALSE, cache=FALSE, message=FALSE, warning=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

# use pacman to load all the packages you are missing!
pacman::p_load('knitr', 'lme4', 'lmerTest', 'tidyverse', 'magrittr', 'effects', 'plyr', 'dplyr', 
               'plotrix', 'car',"gridExtra", "cowplot", "tools", "mgcv", "gratia", "tidymv", "ggthemes", "ggpubr", "scales", "visreg", "mgcViz")

Fig.formatting<-(theme_classic()) +
  theme(text=element_text(size=25),
        legend.text.align = 0,
        panel.border = element_rect(fill=NA, colour = "black", size=1),
        aspect.ratio=1, 
        axis.ticks.length=unit(0.25, "cm"),
        axis.text.y=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20), 
        axis.text.x=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20)) +
  theme(legend.key.size = unit(0.4, "cm"), legend.position = "right") +
  theme(aspect.ratio=0.5) 


```

**Note on absorbance**
Iodonitrotetrazolium chloride (INT) is utilized as a redox dye to gauge microbial activity in assays. Microbes with metabolic activity can reduce INT to a formazan compound, causing a color change from yellow (oxidized form) to red or purple (reduced form). This reduction reaction is often catalyzed by dehydrogenase enzymes present in metabolically active cells. The formazan product is  insoluble and has a distinct color (i.e., absorbance), leading to a visible change in the solution's color. The intensity of the color or absorbance is indicative of microbial metabolic activity.

We measured absorbance at 450 nm (see Blenkinsopp & Lock, 1990; Friedel et al., 1994; Jackson et al., 2013; Bhupathiraju et al., 1998). Time course measurements, or absorbance readings at different intervals (e.g., 30 mins, 2 hrs, 16 hrs, 24 hrs, 48 hrs, and 72 hrs), allowed for tracking changes in microbial metabolic activity over time.

### Import metabolism assay data
```{r}
#add data files
meta.day1.450.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_1_450_clean.csv")
meta.day1.590.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_1_590_clean.csv")


meta.day3.450.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_3_450_clean.csv")
meta.day3.590.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_3_590_clean.csv")


meta.day5.450.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_5_450_clean.csv")
meta.day5.590.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_5_590_clean.csv")

meta.day7.450.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_7_450_clean.csv")
meta.day7.590.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_7_590_clean.csv")

meta.day11.450.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_11_450_clean.csv")
meta.day11.590.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_11_590_clean.csv")

meta.day17.450.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_17_450_clean.csv")
meta.day17.590.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_17_590_clean.csv")


#merge 


meta.450.df= rbind(
  meta.day1.450.df, 
  meta.day3.450.df,
  meta.day5.450.df,
  meta.day7.450.df, 
  meta.day11.450.df, 
  meta.day17.450.df
  )

meta.590.df=rbind(
  meta.day1.590.df,
  meta.day3.590.df,
  meta.day5.590.df,
  meta.day7.590.df,
  meta.day11.590.df, 
  meta.day17.590.df
)

#remove well_id columns
meta.450.df <- meta.450.df[, !colnames(meta.450.df) %in% "well_id"]
meta.590.df <- meta.590.df[, !colnames(meta.590.df) %in% "well_id"]


#create groups based on deployment and metabolite pair

#if deploy is HOK, metabolite is HOK (home)
#if deplpy is SEK, metaboltie is SEK (home)
#if deploy is HOK, metaboltie is SEK (away)
#if deploy is SEK, metabolite is HOK (away)

meta.450.df$group <- 
  ifelse(
  meta.450.df$deploy == "HOK" & meta.450.df$metabolite == "HOK", "H",
   ifelse(
  meta.450.df$deploy == "SEK" & meta.450.df$metabolite == "SEK", "H",
   ifelse(
   meta.450.df$deploy == "HOK" & meta.450.df$metabolite == "SEK", "AW",
   ifelse(
   meta.450.df$deploy == "SEK" & meta.450.df$metabolite == "HOK", "AW", NA  # Default value if none of the conditions are met
                )
              )
            )
          )

meta.590.df$group <- 
  ifelse(
  meta.590.df$deploy == "HOK" & meta.590.df$metabolite == "HOK", "H",
   ifelse(
  meta.590.df$deploy == "SEK" & meta.590.df$metabolite == "SEK", "H",
   ifelse(
   meta.590.df$deploy == "HOK" & meta.590.df$metabolite == "SEK", "AW",
   ifelse(
   meta.590.df$deploy == "SEK" & meta.590.df$metabolite == "HOK", "AW", NA  # Default value if none of the conditions are met
                )
              )
            )
          )


#add hours rather than minutes
meta.450.df$hours <- 
  ifelse(
  meta.450.df$time == "30", "0.5",
   ifelse(
  meta.450.df$time == "120", "2",
   ifelse(
   meta.450.df$time == "960", "16",
   ifelse(
   meta.450.df$time == "1440", "24",
    ifelse(
   meta.450.df$time == "2880", "48",
   ifelse(
   meta.450.df$time == "4320", "72", NA  # Default value if none of the conditions are met
                )
              )
            )
          )
)
)

meta.590.df$hours <- 
  ifelse(
  meta.590.df$time == "30", "0.5",
   ifelse(
  meta.590.df$time == "120", "2",
   ifelse(
   meta.590.df$time == "960", "16",
   ifelse(
   meta.590.df$time == "1440", "24",
    ifelse(
   meta.590.df$time == "2880", "48",
   ifelse(
   meta.590.df$time == "4320", "72", NA  # Default value if none of the conditions are met
                )
              )
            )
          )
)
)

#reorder hours
meta.450.df <- meta.450.df %>% mutate(hours = factor(hours, levels = c("0.5", "2", "16", "24", "48", "72")))
meta.590.df <- meta.590.df %>% mutate(hours = factor(hours, levels = c("0.5", "2", "16", "24", "48", "72")))

#make colummn factors
meta.450.df$treatment=as.factor(meta.450.df$treatment)
meta.450.df$tree_id=as.factor(meta.450.df$tree_id)
meta.450.df$deploy=as.factor(meta.450.df$deploy)
meta.450.df$metabolite=as.factor(meta.450.df$metabolite)
meta.450.df$group=as.factor(meta.450.df$group)
meta.450.df$hours=factor(meta.450.df$hours)

meta.590.df$treatment=as.factor(meta.590.df$treatment)
meta.590.df$tree_id=as.factor(meta.590.df$tree_id)
meta.590.df$deploy=as.factor(meta.590.df$deploy)
meta.590.df$metabolite=as.factor(meta.590.df$metabolite)
meta.590.df$group=as.factor(meta.590.df$group)
meta.590.df$hours=factor(meta.590.df$hours)

#make pairing for 4 treatments
meta.450.df <- meta.450.df %>%
  mutate(pairing = paste(treatment, group, sep = "-"))

meta.450.df <- meta.450.df %>% 
  mutate(pairing = factor(pairing, levels = c("P-AW", "P-H", "AB-AW", "AB-H")))

class(meta.450.df$decomp_day)
```

#### determine which absorbance wavelength to use. Various studies (as those cited above) measured the absorbance of INT at both 450 and 590 nm. This provides some flexibility for us to determine exactly which to use based on the conditions present in our system (i.e., rivers, within the wells, external conditions of the lab space, etc). Factors such as these can influence the rate of metabolization, and thus the OD values from the absorbance reader. Additionally, determining which reading time (30 mins, 2 hrs, 16 hrs, etc.) at which the OD values were measured is important. Past work has measured absorbance at 30 mins, while other recommend 2-16 hrs. However, given the specific conditions of our field experiment, it has been suggested that 24 hrs may be required for an accurate metabolic reading. Additionally, we need to include a hierarchical modeling framework to capture the nonlinear trends in the data (a priori parabolic, or unimodal, metabolization curves across decomposition day) and repeated measures for each leaf pack. 

Let's first visualize the raw OD readings at both 450 and 590 nm.
```{r}

#plot at 450 nm
meta.df %>%
    filter(nm == "450", time == "1440")%>%
  na.omit()%>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n)) +
  geom_smooth(aes(group = interaction(group, treatment), linetype = treatment, color = group, fill = group), method = 'gam', formula = y ~ s(x, k = 4), alpha = 0.1) +
  labs(x = "decomposition day", y = "standardized optical density (z score)")+
  scale_color_colorblind() +
  scale_fill_colorblind() +
   guides(linetype = guide_legend(override.aes = list(color = "cornflowerblue")))+
  theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )


# plot at 590 nm
meta.df %>%
    filter(nm == "590", time == "1440")%>%
  na.omit()%>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n)) +
  geom_smooth(aes(group = interaction(group, treatment), linetype = treatment, color = group, fill = group), method = 'gam', formula = y ~ s(x, k = 6), alpha = 0.1) +
  labs(x = "decomposition day", y = "standardized optical density (z score)")+
  scale_color_colorblind() +
  scale_fill_colorblind() +
   guides(linetype = guide_legend(override.aes = list(color = "cornflowerblue")))+
  theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )

```

Based on the literature, we will use an absorbance wavelength of 450 nm which is appropriate for the metabolic indicator (INT) in the assay. 
### Use 450 nm absorbance readings
```{r}
#by group and treatment using loess
meta.450.df %>%
  filter(hours == "24")%>%
  na.omit()%>%
  ggplot(aes(x = as.numeric(decomp_day), y = od_corrected_b_n)) +
  geom_smooth(aes(color = treatment, fill = treatment, linetype = group, group = interaction(group, treatment)), method = 'loess', formula = y ~ x, alpha = 0.1) +
    geom_line(aes(y = neg), color = "red", linetype = "dashed", size = 0.5, alpha = 0.5) +
    geom_ribbon(aes(ymin = neg - sd(neg), ymax = neg + sd(neg)), fill = "red", alpha = 0.2) +
  labs(x = "decomposition day", y = "optical density (450 nm)")+
  scale_x_continuous(breaks = c(1, 5, 9, 13, 17))+
  scale_color_colorblind() +
  scale_fill_colorblind() +
guides(
    linetype = guide_legend(override.aes = list(color = "cornflowerblue")),
    color = guide_legend(order = 1),  # Set order for the color legend
    fill = guide_legend(order = 1)   # Set order for the fill legend
  ) +
  theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )

#by group and treatment using gam
meta.450.df %>%
  filter(hours == "16")%>%
  na.omit()%>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n)) +
  geom_smooth(aes(color = treatment, fill = treatment, linetype = group, group = interaction(group, treatment)), method = 'gam', formula = y ~ s(x, k =4), alpha = 0.1) +
    geom_smooth(aes(y = neg), method = "gam", color = "red", fill= "red", linetype = "dashed", size = 0.5, alpha = 0.5) +
  labs(x = "decomposition day", y = "optical density (450 nm)")+
  scale_x_continuous(breaks = c(1, 5, 9, 13, 17))+
  scale_color_colorblind() +
  scale_fill_colorblind() +
guides(
    linetype = guide_legend(override.aes = list(color = "cornflowerblue")),
    color = guide_legend(order = 1),  # Set order for the color legend
    fill = guide_legend(order = 1)   # Set order for the fill legend
  ) +
  theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )

meta.450.df %>%
  filter(hours == "24")%>%
  na.omit()%>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n)) +
  geom_point(aes(y = od_corrected_b_n, group = interaction (group, treatment, tree_num)))+
  geom_smooth(aes(color = treatment, fill = treatment, linetype = group, group = interaction(group, treatment)), method = 'gam', formula = y ~ s(x, k=4), alpha = 0.1) +
      geom_smooth(aes(y = neg), method = "gam", formula = y ~ s(x, k=4), color = alpha("red", 0.5), fill= alpha("red", 0.5), linetype = "longdash", size = 0.7) +
  labs(x = "decomposition day", y = "optical density (450 nm)")+
scale_color_colorblind() +
  scale_fill_colorblind() +
guides(
    linetype = guide_legend(override.aes = list(color = "cornflowerblue")),
    color = guide_legend(order = 1),  # Set order for the color legend
    fill = guide_legend(order = 1)   # Set order for the fill legend
  ) +
  theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )


#####################

meta.450.df %>%
  filter(hours == "24") %>%
  na.omit() %>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n, group = interaction(group, treatment), color = treatment, fill = treatment)) +
  stat_summary(fun = mean, geom = "point", aes(shape = treatment), show.legend = F, size = 2.3) +
  geom_smooth(aes(color = treatment,fill = treatment, linetype=group), method = 'gam', formula = y ~ s(x, k = 6), alpha = 0.15) +
  geom_point(aes(y= neg), color = alpha("red", 0.3), shape = 15)+
  geom_line(aes(y = neg), color = alpha("red", 0.3), linetype = "longdash", size = 0.7) +
  labs(x = "decomposition day", y = "optical density (450 nm)") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  guides(
    linetype = guide_legend(override.aes = list(color = "cornflowerblue")),
    color = guide_legend(order = 1),  # Set order for the color legend
    fill = guide_legend(order = 1)   # Set order for the fill legend
  ) +
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    strip.text = element_text(size = 16)
  )


```


These are the final plots
```{r}
#####################
#this is the final plot
assay.plot=meta.450.df %>%
  filter(hours == "24") %>%
  na.omit() %>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n, group = interaction(group, treatment), color = treatment, fill = treatment)) +
  stat_summary(fun = mean, geom = "point", aes(shape = treatment), show.legend = T, size = 2.3) +
  geom_smooth(aes(color = treatment,fill = treatment, linetype=group), method = 'gam', formula = y ~ s(x, k = 6), alpha = 0.15) +
  geom_point(aes(y= neg,  shape = "neg"), color = alpha("red", 0.3), show.legend = F)+
  geom_line(aes(y = neg), color = alpha("red", 0.3), show.legend = F, linetype = "longdash", size = 0.7) +
  labs(x = "decomposition day", y = "optical density (450 nm)") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  guides(
    linetype = guide_legend(override.aes = list(color = "cornflowerblue")),
    color = guide_legend(order = 1),  # Set order for the color legend
    fill = guide_legend(order = 1)   # Set order for the fill legend
  ) +
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    strip.text = element_text(size = 16)
  )+
  Fig.formatting

assay.plot

ggsave("figures/assay.plot.pdf", assay.plot, width = 17, height = 8)


#4 treatment plot
trt.assay.plot=meta.450.df %>%
  filter(hours == "24") %>%
  na.omit() %>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n, color = pairing, group = pairing)) +
  stat_summary(fun = mean, geom = "point", show.legend = F, size = 2.3) +
    geom_point(aes(y= neg,  shape = "neg"), size= 2.3, color = alpha("red", 0.3), show.legend = F)+
  geom_smooth(aes(linetype= pairing, fill = pairing),method = 'gam', formula = y ~ s(x, k = 6), alpha = 0.15) +
  geom_smooth(aes(y = neg), fill = alpha ("red", 0.3), method = 'loess', formula = y ~ x, span= 0.6, color = alpha("red", 0.3), show.legend = F, linetype = "longdash") +
  labs(x = "decomposition day", y = "absorbance (OD450 nm)") +
  coord_cartesian(ylim = c(0,3))+
  scale_color_colorblind()+
  scale_fill_colorblind()+
    scale_x_discrete(expand = c(0.01, 0.01)) +  # Adjust the expand argument
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    strip.text = element_text(size = 16)
  )+
  Fig.formatting

trt.assay.plot

ggsave("figures/trt.assay.plot.pdf", trt.assay.plot, width = 15, height = 8)

meta.450.df$decomp_day=as.numeric(meta.450.df$decomp_day)

meta.450.df %>%
  filter(hours == "24") %>%
  na.omit() %>%
  ggplot(aes(x = as.factor(decomp_day), y = od_corrected_b_n, color = pairing, group = pairing)) +
  stat_summary(fun = mean, geom = "point", show.legend = F, size = 2.3) +
    geom_point(aes(y= neg,  shape = "neg"), size= 2.3, color = alpha("red", 0.3), show.legend = F)+
  geom_smooth(aes(linetype= pairing, fill = pairing), method = 'gam', formula = y ~ s(x, k = 6), alpha = 0.15) +
  geom_smooth(aes(y = neg), fill = alpha ("red", 0.3), method = 'loess', formula = y ~ x, span= 0.6, color = alpha("red", 0.3), show.legend = F, linetype = "longdash") +
  labs(x = "decomposition day", y = "absorbance (450 nm)") +
  coord_cartesian(ylim = c(0,3))+
  scale_x_discrete(breaks = c(1,3,5,7,11,17)) +
      scale_x_discrete(expand = c(0.01, 0.01)) +
  scale_color_colorblind()+
  scale_fill_colorblind()+
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    strip.text = element_text(size = 16)
  )+
  Fig.formatting


```


```{r}
#This pattern suggests that microbial communities from the intermediate stages of decomposition may be particularly well-equipped or adapted to metabolize the plant metabolites. 

#The hypothesis about a unimodal trend in abundance and diversity, where early and low decomposition days show low abundance and diversity, and intermediate decomposition days show high abundance and diversity, is a reasonable expectation. It aligns with the concept of priority effects, where early colonizers influence substrate conditions and, in some cases, facilitate the establishment of subsequent colonizers.

#However, the findings from the metabolization assay contradict this hypothesis. The observation that early-stage days exhibit lower metabolization despite the potential presence of key enzymes in those microbes challenges the initial hypothesis.

#This discrepancy suggests that the relationship between microbial community composition, their functional attributes (such as enzyme production), and the metabolization of plant extracts is more complex than initially anticipated. It's possible that other factors, such as the specific secondary metabolites present in the plant extracts, microbial interactions, or ecological conditions, influence the metabolic activity of the communities.



#by 4 treatments
meta.450.df %>%
  filter(hours == "24")%>%
  na.omit()%>%
  ggplot(aes(x = decomp_day, y = od_corrected_b_n)) +
  geom_smooth(aes(linetype= pairing, color = pairing, fill = pairing), method = 'gam', formula = y ~ s(x, k = 4), alpha = 0.1) +
  labs(x = "decomposition day", y = "optical density (450 nm)")+
    guides(color = guide_legend(title = "Treatment"), fill= guide_legend(title = "Treatment"), linetype = guide_legend(title = "Treatment"))+
   scale_linetype_manual(name = "Treatment",
                        values = c("P-AW" = "solid", "P-H" = "solid", "AB-AW" = "dotdash", "AB-H" = "dotdash")) +
    guides(color = guide_legend(title = "Treatment"), fill = guide_legend(title = "Treatment"), linetype = guide_legend(title = "Treatment"))+
  scale_x_continuous(breaks = c(1, 5, 9, 13, 17))+
  theme_bw() +
  scale_color_colorblind() +
  scale_fill_colorblind() +
theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )


#HFAI shows P packs to have greater incidence of HFA, while AB show greater magnitude
meta.450.df %>%
   filter(hours == "24", group == "H")%>%
  na.omit()%>%
  ggplot(aes(x = as.numeric(decomp_day), y = od_corrected_b_n)) +
  geom_smooth(aes( color = treatment, fill = treatment), method = 'loess', formula = y ~x, alpha = 0.1) +
  labs(x = "decomposition day", y = "optical density (450 nm)")+
    scale_x_continuous(breaks = c(1, 5, 9, 13, 17))+
  scale_color_colorblind() +
  scale_fill_colorblind() +
   guides(linetype = guide_legend(override.aes = list(color = "cornflowerblue")))+
  theme_bw()+
  theme(
    axis.text = element_text(size = 12),      # Adjust text size for axis labels
    axis.title = element_text(size = 14),     # Adjust text size for axis titles
    legend.text = element_text(size = 12),    # Adjust text size for legend labels
    legend.title = element_text(size = 14),   # Adjust text size for legend title
    strip.text = element_text(size = 16)      # Adjust text size for facet labels
  )

#In the context of the Homefield Advantage Index (HFAI), "incidence" and "magnitude" can be understood as follows:

#Incidence of HFAI:

#Definition: Incidence refers to the frequency or occurrence of HFAI events. In the Context of HFAI:  the incidence of HFAI is more in the "present" treatment, it means that there are more instances or occurrences where the HFAI is observed when the phyllosphere community is present. This could be interpreted as a higher frequency of home advantage events in terms of decomposition when the community is present.

#Magnitude of HFAI: Definition: Magnitude refers to the size, extent, or intensity of the HFAI events.  the magnitude is greater in the "absent" treatment, it suggests that, although the frequency might be lower, when HFAI events occur in the absence of the phyllosphere community, they have a larger impact or are more pronounced in terms of decomposition. The decomposition advantage at home (Hoko) is more significant when the community is absent, leading to a higher magnitude of the HFAI.

#In summary, the "incidence" reflects how often HFAI events occur, while the "magnitude" indicates the strength or impact of those events when they do occur. Both aspects are important for understanding the dynamics of the Homefield Advantage Index in your experimental conditions.

#Microbial communities metabolizing their home metabolite exhibit higher activity in the presence of the phyllosphere community. This is consistent with the HFai framework, where the relative advantage of home metabolite metabolization is observed in the ecological context of the phyllosphere community being present.



```


#### z score optical density
```{r}
###########################
#z score by deployment and decomposition day and metabolite
deployments <- c("HOK", "SEK")
site_numbers <- c("1", "2")

# Initialize an empty data frame to store the z-scored data
z_scored_df <- data.frame()

# Loop through decomposition days, deployments, site numbers, and metabolites
  for (deploy in deployments) {
    for (site in site_numbers) {
        subset <- meta.450.df %>%
          filter(deploy == deploy, site_num == site)

        # Calculate z-scores for the subset
        subset <- subset %>%
          mutate(z_score = (od_corrected_b_n - mean(od_corrected_b_n, na.rm = TRUE)) / sd(od_corrected_b_n, na.rm = TRUE))


# Combine the z-scored subset into the z_scored_df
        z_scored_df <- bind_rows(z_scored_df, subset)
      }
  }

z_scored_mean_df=aggregate(z_score ~ treatment + tree_id + deploy + metabolite  + site_num + tree_num + decomp_day + group, data = z_scored_d_d_h_df, mean, na.rm=T)

```

Model exploration
**A note on GAMs**
The term "parametric" is often used to describe statistical models with a fixed, finite number of parameters. In contrast, "nonparametric" models are more flexible and can adapt to the complexity of the data without assuming a specific functional form.

Generalized Additive Models (GAMs) can be considered semiparametric. They are a type of regression model that extends generalized linear models (GLMs) by allowing for more flexible, nonlinear relationships between predictors and the response variable. The "additive" part refers to the fact that the model is built by adding together multiple smooth functions, each representing the relationship between a predictor and the response.

While GAMs introduce flexibility with the smooth functions, they still include parametric components, such as linear terms and parametric link functions. The parametric components provide a structured foundation, and the smooth functions capture additional complexity in the data.

So, we are using GAMsm that combine parametric and nonparametric elements to provide a flexible framework for modeling complex relationships in the metabolism assay data.

```{r}
lmer1 = lmer(od_corrected_b_n ~ treatment * group + (1 | decomp_day) + (1 | deploy) +  (1 | tree_id) + (1|tree_num), subset = hours == "24", data = meta.450.df)

lmer2 = lmer(od_corrected_b_n ~ pairing + (1 | decomp_day) + (1 | deploy) +  (1 | tree_id) + (1|tree_num), subset = hours == "24", data = meta.450.df)

lmer3 = lmer(od_corrected_b_n ~ treatment * group + ( 1 + treatment + group | decomp_day)  + (1 | tree_id/tree_num) + (1|deploy/site_num), subset = hours == "24", data = meta.450.df)

lmer4 = lmer(od_corrected_b_n ~ treatment * group + ( treatment * group || decomp_day)  + (1 | tree_id/tree_num) , subset = hours == "24", data = meta.450.df)


gam1= gam(od_corrected_b_n ~ pairing + s(decomp_day, pairing, bs = "fs", k = 5)  +s(tree_num, k = 12, bs = 're') + s(tree_id, bs = 're') + 
                   s(site_num, bs = 're') + s(deploy, bs = 're'),  method = "REML", 
                 data = meta.450.df)
summary(gam1)
gam.check(gam1)
#REML is generally preferred if you are primarily interested in estimating fixed effects and want better estimates of variance components. It's suitable for mixed-effects models.


#testing GAMs

gam1=gam(od_corrected_b_n ~ pairing + s(decomp_day ,k =6) + s(tree_id, tree_num, bs="re") + s(deploy, site_num, bs="re"), subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam2=gam(od_corrected_b_n ~ pairing + s(decomp_day, by= pairing, k =6) + s(tree_id,bs="re") + s(tree_num, bs="re") + s(deploy,bs="re") + s(site_num, bs="re"), subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam3=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k =6) + s(decomp_day, by= group, k =6)+ s(tree_id,bs="re") + s(tree_num, bs="re") + s(deploy,bs="re") + s(site_num, bs="re"), subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))
#####
gam4=gam(od_corrected_b_n ~ treatment + group + s(decomp_day, by= treatment, k =6) + s(decomp_day, by= group, k =6)+ s(tree_id,bs="re") + s(tree_num, bs="re") + s(deploy,bs="re") + s(site_num, bs="re"), subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))
########

gam5=gam(od_corrected_b_n ~ pairing + s(decomp_day, by= pairing, k=6)  + s(tree_num, bs="re") + s(deploy,bs="re") + s(site_num, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam6=gam(od_corrected_b_n ~ pairing + s(decomp_day, by= pairing, k=6)  + s(tree_num, tree_id, bs="re") +  s(site_num, deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))


gam7=gam(od_corrected_b_n ~ treatment + group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k=6)+s(tree_num, bs="re") + s(deploy,bs="re") + s(site_num, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam8=gam(od_corrected_b_n ~ treatment + group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k=6)+s(tree_id, tree_num, bs="re")+ s(deploy, site_num, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

#########

gam9=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k=6)+s(tree_num, bs="re") + s(deploy,bs="re") + s(site_num, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam10=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k =6)+ s( tree_id, tree_num, bs="re", k= 20) + s(deploy, site_num, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam11=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k =6)+ s( tree_num, tree_id, bs="re") + s(site_num,bs="re") + s(deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))

gam12=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k =6)+ s( site_num, tree_num, bs="re") + s(tree_id, bs="re") +  + s(deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))



AIC( gam10, gam11, gam12)

####### 2 canditate models
####### first lets look at main effects and interaction of treatment and group

# looks at interaction between main effects
gam_trt.grp=gam(od_corrected_b_n ~ treatment * group + s(decomp_day, by= treatment, k=6)  + s(decomp_day, by= group, k =6)+ s( tree_num, tree_id, bs="re") + s(site_num,bs="re") + s(deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))
summary(gam_trt.grp)
gam.check(gam_trt.grp)
anova.gam(gam_trt.grp)

#plot difference for P vs AB
plot_difference(
  gam_trt.grp,
  series = decomp_day,
  difference = list(treatment = c("P", "AB"))
) + 
  Fig.formatting 
  #this is important to see where the differences in the smooths lie given that the term is significant (p = 0.002)


#plot difference for P vs AB
plot_difference(
  gam_trt.grp,
  series = decomp_day,
  difference = list(group = c("H", "AW"))
) + 
  Fig.formatting 

####### next we can look at paired main effects for pairwise comparison
#looks at individual treatments
gam_pair=gam(od_corrected_b_n ~ pairing + s(decomp_day, by= pairing, k=6) + s( tree_num, tree_id, bs="re") + s(site_num,bs="re") + s(deploy, bs="re"),  subset = hours == "24",family = tw(link = "identity"), method = "REML", data = na.omit(meta.450.df))
summary(gam_pair)
gam.check(gam_pair)

AIC(gam_trt.grp, gam_pair)
BIC(gam_trt.grp, gam_pair)
anova.gam(gam_pair)

#### in our model (gam_pair), the reference level is set to P-AW, so we can inspect the differences between smooths relative to P-AW

#plot difference for PAW and PH
paw.vs.ph.diff=plot_difference(
  gam_pair,
  series = decomp_day,
  difference = list(pairing = c("P-AW", "P-H"))
)+
   xlab("")+
  ylab("")+
      ggtitle("P-AW vs. P-H")+
  Fig.formatting

#plot difference for PAW and ABAW
paw.vs.abaw.diff=plot_difference(
  gam_pair,
  series = decomp_day,
  difference = list(pairing = c("P-AW", "AB-AW"))
)+
    xlab("")+
    ggtitle("P-AW vs. AB-AW")+
  Fig.formatting

#plot difference for PAW and ABH
paw.vs.abh.diff=plot_difference(
  gam_pair,
  series = decomp_day,
  difference = list(pairing = c("P-AW", "AB-H"))
)+
  xlab("")+
  ylab("")+
  ggtitle("P-AW vs. AB-H")+
  Fig.formatting

#plot difference for PH and ABH
ph.vs.abh.diff=plot_difference(
  gam_pair,
  series = decomp_day,
  difference = list(pairing = c("P-H", "AB-H"))
)+
  xlab("")+
  ylab("")+
  ggtitle("P-H vs. AB-H")+
  Fig.formatting

#plot difference for PH and ABAW
ph.vs.abaw.diff=plot_difference(
  gam_pair,
  series = decomp_day,
  difference = list(pairing = c("P-H", "AB-AW"))
)+
  xlab("")+
  ylab("")+
  ggtitle("P-H vs. AB-AW")+
  Fig.formatting

#plot difference for ABAW and ABH
abaw.vs.abh.diff=plot_difference(
  gam_pair,
  series = decomp_day,
  difference = list(pairing = c("AB-AW", "AB-H"))
)+
  xlab("")+
  ylab("")+
  ggtitle("AB-AW vs. AB-H")+
  Fig.formatting




pair_diffs=plot_grid(
  paw.vs.ph.diff,
  paw.vs.abaw.diff,
  paw.vs.abh.diff,
  ph.vs.abh.diff,
  ph.vs.abaw.diff,
  abaw.vs.abh.diff,
  rel_widths = c(1,1,1,1,1,1), nrow=2, ncol=3
)

pair_diffs

ggsave("figures/pair_diffs.pdf", pair_diffs, width = 16, height = 8)



```

Model 1 (gam_trt.grp):

Intercept: Represents the estimated mean response when all predictors are at their reference levels. In this case, it corresponds to the baseline condition of treatment AB and group AW.

TreatmentP: Represents the estimated change in the mean response when moving from treatment AB to treatment P while holding group constant. The estimate is 0.10648 with a standard error of 0.04688. The t-value is 2.271, and the p-value is 0.0234.

GroupH: Represents the estimated change in the mean response when moving from group AW to group H while holding treatment constant. The estimate is -0.05338 with a standard error of 0.04225. The t-value is -1.263, and the p-value is 0.2068.

TreatmentP:GroupH Interaction: Represents the additional change in the mean response when both treatment is P and group is H compared to the reference level (treatment AB, group AW). The estimate is 0.03843 with a standard error of 0.06137. The t-value is 0.626, and the p-value is 0.5314.

Interpretation:

The interaction term, TreatmentP:GroupH, provides information about whether the effect of treatment P differs when the group is H compared to the reference level (group AW).

A positive estimate (0.03843) suggests that, relative to the baseline condition (AB-AW), the mean response is higher when both treatment is P and group is H. However, this effect is not statistically significant since the p-value is 0.5314, which is greater than the commonly used significance level of 0.05.

In simpler terms, the interaction suggests that there is no significant difference in the effect of treatment P between groups H and AW. It indicates that the effect of treatment P on the mean response is consistent across both groups.


TreatmentP:GroupH Interaction Term:

The estimate (0.03843) represents the additional change in the mean response when both treatment is P and group is H compared to the reference level (AB-AW).
Interpretation:

If we consider the levels of the categorical variables:

Treatment: AB is the reference level.
Group: AW is the reference level.
The interaction term captures the additional change in the mean response when we move from the reference condition (AB-AW) to a specific condition, which is Treatment P and Group H.

However, the interpretation does not directly provide information about the situation where Treatment is AB and Group is H. This is because the interaction term explicitly involves Treatment P.

If you want to assess the difference in the mean response between Treatment AB and Treatment P specifically within Group H, you can do so by comparing the individual estimates for Treatment AB and Treatment P within Group H.

Here's a simplified interpretation:

AB-H vs. AB-AW: The effect of moving from AB to H within the same treatment (AB) is represented by the main effect of Group (GroupH). This is estimated to be -0.05338 (not statistically significant in this case).

P-H vs. AB-AW: The interaction term (TreatmentP:GroupH) captures the additional change when both Treatment is P and Group is H compared to the reference level (AB-AW). The estimate is 0.03843 (not statistically significant in this case).


**Background information regarding Home-field Advantage:**

The Home Field Advantage Index (HFAI) derived from Ayres et al. (2009) uses the relative mass loss values (rml) of species A and B at different sites. The HFAI therefore represents the difference in the rate of litter decomposition between two different sites ("Hoko" and "Sekiu") for both species A and B. Species A and B, in this case, are alder trees of different origin. For example, species A is Hoko alder and species B is Sekiu alder. There are 10 alder trees per river, representing biological replicates. So, the HFAI quantifies the *advantage* or *disadvantage* of each "species" in their respective home environment compared to an away environment.

Interpreting HFAI values involves considering their sign and magnitude:

**Positive** HFAI Values:
Positive values indicate that both species A and B experience faster decomposition rates at their home site compared to the away site. This suggests that both species are better adapted to their home environment in terms of litter decomposition.

**Negative** HFAI Values:
Negative values indicate that both species A and B experience slower decomposition rates at their home site (HF disadvantage (HFDA)) compared to the away site. This suggests that both species might not be well-suited for decomposition in their own environment and may decompose more efficiently in a different environment.

**Magnitude** of HFAI:
The magnitude of the HFAI values indicates the strength of the advantage or disadvantage. Larger absolute values (positive or negative) suggest a greater difference in decomposition rates between the home and away sites.


Load mass data 
```{r}
library(stringr)
library(purrr)
mass.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/leaf_mass.csv")
###check for any white spaces clean up
columns_with_whitespace <- map_lgl(mass.df, ~ any(str_detect(.x, "^\\s+|\\s+$")))
columns_with_whitespace

#correction factor adjusting for blotted and fresh dry mass
mass.df= mass.df%>%
  mutate(final_correct = (0.941*(final_mass)-0.00337))

# calculate % mass loss 
mass.df <- mass.df %>%
  mutate(perc.mass_loss = ((initial_mass - final_correct) / initial_mass)*100)

# Calculate the constant C, adding 1 for safety
C <- abs(min(mass.df$perc.mass_loss)) + 1  

#create new column with perc.mass_loss + constant
mass.df <- mass.df %>%
  mutate(perc.mass_loss = perc.mass_loss + C) %>%
  mutate(prop.mass_loss = perc.mass_loss/100) %>%
mutate(arcsine_trans = asin(sqrt(prop.mass_loss)))


#subset P treatment only for "true" HFA  
mass.present.df=mass.df %>%
  filter(treatment == "P")

#subset AB treatment only for HFA  
mass.absent.df=mass.df %>%
  filter(treatment == "AB")


```

**Subset dataframes to z score by deployment site (all HOK 1, all HOK 2, etc...)**

Z scoring controls for site variability, and allows for a better comparison of treatments. But how exactly?

Z-scoring is a statistical measure that quantifies how far away a particular data point is from the mean of a dataset in terms of standard deviations (i.e., the spread of the data points from the mean). It's a way to standardize data and make it more comparable across different scales or distributions. Here's how z-scoring works:

1) calculate the mean and standard deviation: for the variable you want to z-score (in this case mass loss by deployment site), calculate its mean (average) and standard deviation. These values provide information about the central tendency (i.e., the center/midpoint of the data or a value around which the data clusters) and the spread of the data.

2) subtract the mean: for each data point, subtract the mean of the variable. This step shifts the data so that the new mean becomes 0.

3) divide by the standard deviation: after subtracting the mean, divide the result by the standard deviation. This step scales the data based on the spread of the original data, so that the new data points are expressed in terms of standard deviations from the mean.

Mathematically, the z-score (Z) for a data point (X) with mean (μ) and standard deviation (σ) is calculated using the formula:

Z = (X - μ) / σ

The resulting z-score indicates how many standard deviations the original data point is away from the mean. If the z-score is positive, it means the data point is above the mean, and if it's negative, it's below the mean.

Note: z-scoring assumes a roughly normal distribution of the data.

Below, the percent mass loss for each deployment site is z scored. Additionally, the larger dataframe has been subset to only include P treatments so that the "true" HFA can be calculated.
```{r}
#check for distribution of mass loss
hist(mass.df$perc.mass_loss)
  #roughly normal...that's good!

#####Begin z scoring
#HOK 1 subset
mass.hok1.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.hok1.df=mass.hok1.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.hok1.df=mass.hok1.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))


#HOK 1 P subset
mass.hok1.present.df = data.frame(mass.present.df[mass.present.df$deploy == "HOK" & mass.present.df$site_num == "1", ])
# Calculate z-scores 
mass.hok1.present.df=mass.hok1.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.hok1.present.df=mass.hok1.present.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))



#HOK 2 subset
mass.hok2.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.hok2.df=mass.hok2.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.hok2.df=mass.hok2.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))


#HOK 2 P subset
mass.hok2.present.df = data.frame(mass.present.df[mass.present.df$deploy == "HOK" & mass.present.df$site_num == "2", ])
# Calculate z-scores 
mass.hok2.present.df=mass.hok2.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.hok2.present.df=mass.hok2.present.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))



#SEK 1 subset
mass.sek1.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.sek1.df=mass.sek1.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.sek1.df=mass.sek1.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))


#SEK 1 P subset
mass.sek1.present.df = data.frame(mass.present.df[mass.present.df$deploy == "SEK" & mass.present.df$site_num == "1", ])
# Calculate z-scores 
mass.sek1.present.df=mass.sek1.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.sek1.present.df=mass.sek1.present.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))


#SEK 2 subset
mass.sek2.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.sek2.df=mass.sek2.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.sek2.df=mass.sek2.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))


#SEK 2 P subset
mass.sek2.present.df = data.frame(mass.present.df[mass.present.df$deploy == "SEK" & mass.present.df$site_num == "2", ])
# Calculate z-scores 
mass.sek2.present.df=mass.sek2.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.sek2.present.df=mass.sek2.present.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))



##merge all sites
mass.deploy.merged.df=rbind(mass.hok1.df, mass.hok2.df, mass.sek1.df, mass.sek2.df)
mass.deploy.merged.df$treatment=as.factor(as.character(mass.deploy.merged.df$treatment))
mass.deploy.merged.df$tree_id=as.factor(as.character(mass.deploy.merged.df$tree_id))
mass.deploy.merged.df$tree_id=as.factor(as.character(mass.deploy.merged.df$tree_id))
mass.deploy.merged.df$deploy=as.factor(as.character(mass.deploy.merged.df$deploy))
mass.deploy.merged.df$hfa=as.factor(as.character(mass.deploy.merged.df$hfa))

#merge all P treatment sites 
mass.present.deploy.merged.df=rbind(mass.hok1.present.df, mass.hok2.present.df, mass.sek1.present.df, mass.sek2.present.df)
mass.present.deploy.merged.df$treatment=as.factor(as.character(mass.present.deploy.merged.df$treatment))
mass.present.deploy.merged.df$tree_id=as.factor(as.character(mass.present.deploy.merged.df$tree_id))
mass.present.deploy.merged.df$tree_id=as.factor(as.character(mass.present.deploy.merged.df$tree_id))
mass.present.deploy.merged.df$deploy=as.factor(as.character(mass.present.deploy.merged.df$deploy))
mass.present.deploy.merged.df$hfa=as.factor(as.character(mass.present.deploy.merged.df$hfa))

```

Let's first look at the P only to see if HFA is there
```{r}
###P treatment only to see HFA effects
#create a proper "site" column for the figure

#plot
ggplot(mass.present.deploy.merged.df, aes(x = hfa, y = z_score)) +
  geom_boxplot(alpha= 0.3) +
  geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.2)) +
  labs(x = "", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_x_discrete(labels = c("away", "home")) +
  scale_shape_manual(values = c(7,10), name= "site") +
  geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

```


Analysis for **present only** dataframe
To be analyze the mass loss data, arcsine square root transforming proportional mass loss is preferred. This are outlined below...

perform lmer and add result to P treatment only plot
Lmer 3 (see below) is the model that best fits the data
```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
#relevel HFA factors
levels(mass.present.deploy.merged.df$hfa)
mass.present.deploy.merged.df$hfa=factor(mass.present.deploy.merged.df$hfa, levels = c("H", "A"))
levels(mass.present.deploy.merged.df$hfa)
names(mass.deploy.merged.df)

#models:
lm1 <- lm(z_score ~  hfa, data = mass.present.deploy.merged.df)
lm2 <- lm(z_score ~ factor(hfa, levels = c("A", "H")), data = mass.present.deploy.merged.df)
lmer1 <- lmer(z_score ~ hfa + (1 | tree_id), data= mass.present.deploy.merged.df)
lmer2 <- lmer(z_score ~ tree_id * deploy + (1 | tree_num), data= mass.present.deploy.merged.df)

lmer3 <- lmer(z_score ~ hfa + (1 | tree_num) + (1 | tree_id), data= mass.present.deploy.merged.df)

lmer4 <- lmer(z_score ~ hfa + (1 | tree_id /  tree_num), data= mass.present.deploy.merged.df)
summary(lmer5)

# lmer3 is best fit
lmer3 <- lmer(z_score ~ hfa + (1 | tree_num) + (1 | tree_id), data= mass.present.deploy.merged.df)

lmer10 <- lmer(arcsine_trans_scaled ~ hfa + (1 | deploy) + (1 | tree_num) + (1 | tree_id), data= mass.present.deploy.merged.df)

summary(lmer10)
plot(residuals(lmer5))
plot(fitted(lmer5), residuals(lmer5))
shapiro.test(residuals(lmer5))

qqnorm(residuals(lmer5))
qqline(residuals(lmer5))

AIC(lmer3, lmer10)


```


Generic t-test for P only treatment
```{r}

# the t test will use z score values
#combined site
mass.present.deploy.merged.df$combined_site <- paste("site", mass.present.deploy.merged.df$site_num, sep = "")

stat.test.z <- mass.present.deploy.merged.df %>% 
  t_test(z_score ~ hfa, var.equal = T) %>%
  add_significance()

stat.test.arc <- mass.present.deploy.merged.df %>% 
  t_test(arcsine_trans~ hfa, var.equal = T) %>%
  add_significance()

stat.test.z

#check levels
levels(mass.present.deploy.merged.df$hfa)
mass.present.deploy.merged.df$hfa=factor(mass.present.deploy.merged.df$hfa, levels = c("H", "A"))
levels(mass.present.deploy.merged.df$hfa)


# Create the box plot with original elements as ggplot from above
leaf.z.presentonly.plot= ggboxplot(
  mass.present.deploy.merged.df, x = "hfa", y = "z_score", 
  ylab = "mass loss (z score)", xlab = "", fill="hfa", alpha= 0.4) + 
  scale_fill_manual(values = c("grey", "grey0"), guide = "none")+
  stat_pvalue_manual(stat.test.arc, y.position = 2.5, tip.length = c(0.16, 0.06)) +
  labs(subtitle = get_test_label(stat.test.arc, detailed = T)) +
    scale_x_discrete(labels = c("home", "away"))+
    geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.2))+
    scale_shape_manual(values = c(7,10), name= "site")+
    geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16),
          plot.background = element_blank()) 

leaf.z.presentonly.plot

# Create the box plot with original elements as ggplot from above
leaf.mass.presentonly.plot= ggboxplot(
  mass.present.deploy.merged.df, x = "hfa", y = "perc.mass_loss", 
  ylab = "mass loss (%)", xlab = "", fill="hfa", alpha= 0.4) + 
  scale_fill_manual(values = c("grey", "grey0"), guide = "none")+
  stat_pvalue_manual(stat.test.arc, y.position = 60, tip.length = c(0.06, 0.16)) +
  labs(subtitle = get_test_label(stat.test.arc, detailed = T)) +
    scale_x_discrete(labels = c("home", "away"))+
    geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.1))+
    scale_shape_manual(values = c(7,10), name= "site")+
    geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(perc.mass_loss), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16),
          plot.background = element_blank()) 

leaf.mass.presentonly.plot

ggsave("figures/leaf.z.presentonly.plot.pdf", leaf.z.presentonly.plot, width = 8, height = 6)


```
Analysis for **all treatment** dataframe
There are two ways to go about model exploration: using z scored mass loss or arcsine square root transforming percent mass loss. These are outlined below...

first, plot to visualize the 4 treatments
```{r}
# Combine treatment and hfa into one group
mass.deploy.merged.df$combined_group <- paste(mass.deploy.merged.df$treatment, mass.deploy.merged.df$hfa, sep = "-")

# Create a custom order for the combined groups
combined_group_order <- c("P-H", "P-A","AB-H", "AB-A")
custom_palette <- c("#2CA02C",  "#D62728", "#1F77B4","#FF7F0E")


# Plot with combined groups but no combo site
ggplot(mass.deploy.merged.df, aes(x = combined_group, y = z_score, fill = combined_group)) +
  geom_boxplot(alpha= 0.3) +
  scale_x_discrete(limits = combined_group_order)+
  geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.2), alpha = 0.7) +
  labs(x = "treatment", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_fill_manual(values = custom_palette, guide = "none") +
  scale_shape_manual(values = c(15, 16), name ="site") +
  geom_hline(data = mass.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting



```

model exploration of all treatments: ph, pa, abh, aba; lmer 7 is best fit
```{r}
lmer4 <- lmer(arcsine_z_score ~ treatment*hfa + (1|tree_num) + (1|tree_id), data=mass.deploy.merged.df)
lmer5 <- lmer(arcsine_z_score ~ combined_group + (1|tree_num) + (1|tree_id), data=mass.deploy.merged.df)
lmer6 <- lmer(z_score ~ combined_group + (1|tree_num) + (1|tree_id), data=mass.deploy.merged.df)
lmer7 <- lmer(z_score ~ treatment*hfa + (1|tree_num) + (1|tree_id), data=mass.deploy.merged.df)

AIC(lmer4, lmer7)
summary(lmer7)
# Type III ANOVA for fixed effects
Anova(lmer7, type="III")

library(emmeans)
# Post-hoc tests for interactions (if significant)
interaction_emm <- emmeans(lmer7, specs = ~treatment:hfa)
contrast(interaction_emm, method = "pairwise")

# Fit your linear mixed-effects model
lmer7 <- lmer(z_score ~ treatment * hfa + (1 | tree_num) + (1 | tree_id), data = mass.deploy.merged.df)


# Calculate Cohen's d for specific group comparisons
cohen_d <- with(mass.deploy.merged.df, {
  d1 <- (mean(z_score[treatment == "P" & hfa == "H"]) - mean(z_score[treatment == "P" & hfa == "A"])) / sqrt((var(z_score[treatment == "P" & hfa == "H"]) + var(z_score[treatment == "P" & hfa == "A"])) / 2)
  d2 <- (mean(z_score[treatment == "AB" & hfa == "H"]) - mean(z_score[treatment == "AB" & hfa == "A"])) / sqrt((var(z_score[treatment == "AB" & hfa == "H"]) + var(z_score[treatment == "AB" & hfa == "A"])) / 2)
  d3 <- (mean(z_score[treatment == "P" & hfa == "H"]) - mean(z_score[treatment == "AB" & hfa == "H"])) / sqrt((var(z_score[treatment == "P" & hfa == "H"]) + var(z_score[treatment == "AB" & hfa == "H"])) / 2)
  d4 <- (mean(z_score[treatment == "P" & hfa == "A"]) - mean(z_score[treatment == "AB" & hfa == "A"])) / sqrt((var(z_score[treatment == "P" & hfa == "A"]) + var(z_score[treatment == "AB" & hfa == "A"])) / 2)
  data.frame(Group = c("P-H vs. P-A", "AB-H vs. AB-A", "P-H vs. AB-H", "P-A vs. AB-A"),
             Cohen_d = c(d1, d2, d3, d4))
})

# Print the Cohen's d values
print(cohen_d)

install.packages("MuMIn")
library(MuMIn)
r.squaredGLMM(lmer7)
```


This is the proper all treatment dataframe for nonparametric test
```{r}
install.packages("dunn.test")
library(dunn.test)
kw1 <- kruskal.test(z_score ~ combined_group, data = mass.deploy.merged.df)
kw1
# Perform post-hoc Dunn's test
posthoc <- dunn_test(z_score ~ combined_group, data = mass.deploy.merged.df, method = "bonferroni")# Create a function to format p-values and add stars for significance
format_p_value <- function(p_value) {
  if (p_value < 0.001) {
    return("***")
  } else if (p_value < 0.01) {
    return("**")
  } else if (p_value < 0.05) {
    return("*")
  } else {
    return("ns")
  }
}

# Apply the formatting function to the p-values
dunn.test.z.res$formatted_p_value <- sapply(dunn.test.z.res$p.adj, format_p_value)
##this plot works
# Create a custom order for the combined groups
combined_group_order <- c("P-H", "P-A","AB-H", "AB-A")
library(ggpubr)

custom_palette <- c("#D62728","#FF7F0E", "#2CA02C", "#1F77B4")
levels(as.factor(mass.deploy.merged.df$combined_group))

leaf.mass.all.plot=ggboxplot(
  mass.deploy.merged.df, x = "combined_group", y = "z_score", 
  ylab = "mass loss (z score)", xlab = "", fill="combined_group", alpha= 0.4) + 
    scale_x_discrete(limits = combined_group_order)+
  scale_fill_viridis_d( guide = "none")+
  stat_pvalue_manual(dunn.test.z.res, step.increase = 0.03, step.group.by = "group2", tip.length = 0.01) +
    labs( x= "treatment", y="mass loss (z score)") +
  geom_point(aes(shape = deploy), position = position_jitter(width = 0.2), alpha = 0.7) +
  scale_shape_manual(values = c(15, 16), name ="site") +
     geom_hline(data = mass.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16),
          plot.background = element_blank(),
          axis.text.x = element_text(angle = 22, hjust = 0.5, vjust = 0.5)) 

leaf.mass.all.plot

count(mass.deploy.merged.df$combined_group)

range(mass.deploy.merged.df$perc.mass_loss)

ggsave("figures/leaf.mass.all.plot.pdf", leaf.mass.all.plot, width = 8, height = 6)

```


Analysis for **absent only** dataframe
There are two ways to go about model exploration: using z scored mass loss or arcsine square root transforming percent mass loss. These are outlined below...

perform t-test and add result to AB treatment only plot
```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
#plot
mass.absent.deploy.merged.df$combined_site <- paste("site", mass.absent.deploy.merged.df$site_num, sep = "")

leaf.mass.absentonly.plot=ggplot(mass.absent.deploy.merged.df, aes(x = hfa, y = z_score)) +
  geom_boxplot(alpha= 0.3) +
  geom_point(aes(shape = as.factor(combined_site)), position = position_jitter(width = 0.2)) +
  labs(x = "", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_x_discrete(labels = c("away", "home")) +
  scale_fill_manual(values = custom_palette) +
  scale_shape_manual(values = c(7,10), name= "site") +
  geom_hline(data = mass.absent.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

leaf.mass.absentonly.plot
# Create a histogram
hist_data <- hist(mass.absent.deploy.merged.df$z_score, plot = FALSE)

# Create a density plot
density_data <- density(mass.absent.deploy.merged.df$z_score)

# Plot histogram with density overlay
hist(mass.absent.deploy.merged.df$z_score, freq = F, main = "")
lines(density_data, col = "red")
  #normal

shapiro.test(mass.absent.deploy.merged.df$z_score)
  # p= 0.39, normal

# Q-Q Plots
qqnorm(mass.absent.deploy.merged.df$z_score, main = "Q-Q Plot")
qqline(mass.absent.deploy.merged.df$z_score)
  #normal


# Levene's Test
leveneTest(z_score ~ hfa, data = mass.absent.deploy.merged.df)
  # p= 0.63, variances between H and A are equal (i.e., the homogeneity assumption of the variance is met)


# the t test will use z score values
stat.ab.test.z <- mass.absent.deploy.merged.df %>% 
  t_test(z_score ~ hfa, var.equal = T) %>%
  add_significance()
stat.ab.test.z

# Create the box plot with original elements as ggplot from above
ggboxplot(
  mass.absent.deploy.merged.df, x = "hfa", y = "z_score", 
  ylab = "mass loss (z score)", xlab = "") + 
  stat_pvalue_manual(stat.ab.test.z, tip.length = 0) +
  labs(subtitle = get_test_label(stat.ab.test.z, detailed = TRUE)) +
    scale_x_discrete(labels = c("away", "home"))+
    geom_point(aes(shape = as.factor(combined_site)), position = position_jitter(width = 0.2))+
    scale_shape_manual(values = c(7,10), name= "site")+
    geom_hline(data = mass.absent.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16)) 

leaf.mass.presentonly.plot

ggsave("figures/leaf.mass.presentonly.plot.pdf", leaf.mass.presentonly.plot, width = 10, height = 8)


```


**Exploration of nonparametric models**
Non-parametric tests are more robust to deviations from normality and heteroscedasticity, but they do have their own assumptions. For example, the Mann-Whitney U test assumes that the distributions of the two groups are similar in shape. The Kruskal-Wallis test assumes that the distributions of all groups are similar and that the differences in the distributions are due to location shifts. 


The Kruskal-Wallis test can be an alternative. This test is an extension of the Wilcoxon rank-sum test for comparing more than two independent groups.

```{r}
# Perform Kruskal-Wallis test
kw_test <- kruskal.test(z_score ~ hfa, data = mass.present.deploy.merged.df)

# Calculate effect size (Eta-squared) from Kruskal-Wallis test
H <- kw_test$statistic
df <- kw_test$parameter
N <- sum(table(mass.present.deploy.merged.df$hfa))
eta_squared <- (H - df) / (N - df)

# Print effect size
cat("Effect size (Eta-squared) for Kruskal-Wallis test:", eta_squared, "\n")

#The calculated effect size (Eta-squared) for the Kruskal-Wallis test is -0.01813634. Eta-squared is a measure of the proportion of variance explained by a factor in non-parametric tests, similar to how R-squared works in linear regression. The negative value suggests a very small effect, but it's important to note that Eta-squared can range from 0 to 1, with higher values indicating larger effects. In this case, the small negative value indicates that the factor (hfa) is not explaining much variance in the data.


```


----------------- START OF HFAI CALCULATIONS ----------------------------------

Calculating HFA Index (HFAI) from Ayres et al. (2009):

There are two equations from which HFAI is dervied: the relative mass loss (rml) of species decomposing at home and away site, and the scaling equation which allows for interpretaibility of the home-field advantage. Below is a brief summary of the components found in the derivaitons:

Aa: This represents the percent mass loss of leaf litter from species A (HOK) when it's decomposing at site a (HOK, or the area dominated by species A). This value is given as a percentage.

Ba: This represents the percent mass loss of leaf litter from species B (SEK) when it's decomposing at site a This value is also given as a percentage.

(Aa + Ba): This is the sum of the percent mass loss of leaf litter from both species A and B when they're decomposing at site a, or HOK and SEK leaf packs at HOK.

(Aa / (Aa + Ba) * 100): This is the full equation that calculates the proportion of the mass loss contributed by species A at site a, relative to the **total mass** loss by both species A and B combined at that site. This percentage value indicates how much of the total mass loss at site a is due to species A.

A similar equation is calculated for species B decomposing at site a. The variables are substituted accordingly.

Armla and Armlb; Brlma, Brmlb: These are the relative mass loss values for species A and B, respectively, when they are decomposing at site a (dominated by species A) or site b (dominated by species B). These values are expressed as percentages of total mass loss at their respective sites.

(((ARMLa + BRMLb)^2 / (ARMLb + BRMLa)^2) * 100)-100: This is the full equation that calculates the squared ratio of the combined relative mass loss of species A at site a and species B at site b (i.e., home), divided by the squared ratio of the combined relative mass loss of species A at site b and species B at site a (i.e., away).

100 is subtracted from the value obtained. This is done to measure the percentage difference in mass loss at the home site (site dominated by the species) compared to the away site (site dominated by the other species). If the value is positive, it indicates that the litter decomposes faster at its home site, showing a home-field advantage. If the value is negative, it suggests that decomposition is slower at the home site, indicating a home-field disadvantage.

The equation calculates the Home-Field Advantage Index (HFAI) by comparing the relative mass loss of litter from two species at their home and away sites. The result is expressed as a percentage, indicating how much faster (or slower) the decomposition occurs at the home site compared to the away site.


NEW: use this! This is a now a proper test dataframe with for loop!

NOTICE: THIS IS FOR SITE 1:

generate **present**, site a (HOK 1)
```{r}
########################
mass.hok1.present.df <- mass.hok1.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Create a new column combining tree_num and ayres with a period in between
mass.hok1.present.df$ayres.tree <- paste(mass.hok1.present.df$ayres, mass.hok1.present.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Armla
names_vector_armla <- character(0)
values_vector_armla <- numeric(0)
treatment_vector_armla <- character(0)


# Initialize empty vectors to store names and values for Brmla
names_vector_brmla <- character(0)
values_vector_brmla <- numeric(0)
treatment_vector_brmla <- character(0)

# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Aa_column <- paste("Aa.", X, sep = "")
    Ba_column <- paste("Ba.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    armla_result <- (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Aa_column, "perc.mass_loss"] / 
                     (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Aa_column, "perc.mass_loss"] + 
                      mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Ba_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    brmla_result <- (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Ba_column, "perc.mass_loss"] / 
                     (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Ba_column, "perc.mass_loss"] + 
                      mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Aa_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    armla_name <- paste("Armla", X, Y, sep = ".")
    brmla_name <- paste("Brmla", X, Y, sep = ".")

     # Extract the treatment from mass.hok1.present.df for the current combination
    treatment_name <- mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Aa_column, "treatment"]
    

    # Append the names and values to the respective vectors for Armla
    names_vector_armla <- c(names_vector_armla, armla_name)
    values_vector_armla <- c(values_vector_armla, armla_result)
    treatment_vector_armla <- c(treatment_vector_armla, treatment_name)

    
    # Append the names and values to the respective vectors for Brmla
    names_vector_brmla <- c(names_vector_brmla, brmla_name)
    values_vector_brmla <- c(values_vector_brmla, brmla_result)
    treatment_vector_brmla <- c(treatment_vector_brmla, treatment_name)


  }
}

# Create dataframes for Armla and Brmla
armla_df <- data.frame(names = names_vector_armla, values = values_vector_armla, treatment= treatment_vector_armla)
brmla_df <- data.frame(names = names_vector_brmla, values = values_vector_brmla, treatment=treatment_vector_brmla)

#merge the dataframes
armla.brmla.df_1=rbind(armla_df, brmla_df)
armla.brmla.df_1
```


NEW: use this! This is a now a proper test dataframe with for loop!
generate **absent**, site a (HOK 1)
```{r}
########################
mass.hok1.absent.df <- mass.hok1.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Create a new column combining tree_num and ayres with a period in between
mass.hok1.absent.df$ayres.tree <- paste(mass.hok1.absent.df$ayres, mass.hok1.absent.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values<- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Armla
names_vector_armla<- character(0)
values_vector_armla<- numeric(0)
treatment_vector_armla <- character(0)


# Initialize empty vectors to store names and values for Brmla
names_vector_brmla <- character(0)
values_vector_brmla <- numeric(0)
treatment_vector_brmla <- character(0)

# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Aa_column <- paste("Aa.", X, sep = "")
    Ba_column <- paste("Ba.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    armla_result <- (mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Aa_column, "perc.mass_loss"] / 
                     (mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Aa_column, "perc.mass_loss"] + 
                      mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Ba_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    brmla_result <- (mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Ba_column, "perc.mass_loss"] / 
                     (mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Ba_column, "perc.mass_loss"] + 
                      mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Aa_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    armla_name <- paste("Armla", X, Y, sep = ".")
    brmla_name <- paste("Brmla", X, Y, sep = ".")
    
    # Extract the treatment from mass.hok1.present.df for the current combination
    treatment_name <- mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Aa_column, "treatment"]
    
    
    # Append the names and values to the respective vectors for Armla
    names_vector_armla <- c(names_vector_armla, armla_name)
    values_vector_armla<- c(values_vector_armla, armla_result)
        treatment_vector_armla <- c(treatment_vector_armla, treatment_name)

    
    # Append the names and values to the respective vectors for Brmla
    names_vector_brmla <- c(names_vector_brmla, brmla_name)
    values_vector_brmla <- c(values_vector_brmla, brmla_result)
        treatment_vector_brmla <- c(treatment_vector_brmla, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
armla_ab_df <- data.frame(names = names_vector_armla, values = values_vector_armla, treatment = treatment_vector_armla)

brmla_ab_df <- data.frame(names = names_vector_brmla, values = values_vector_brmla, treatment = treatment_vector_brmla)

# View the resulting dataframes
print(armla_df)
print(brmla_df)

armla.brmla.absent.df_1=rbind(armla_ab_df, brmla_ab_df)
armla.brmla.absent.df_1

```


NEW: use this! This is a now a proper test dataframe with for loop!
generate **present**, site a (SEK 1)
```{r}
mass.sek1.present.df <- mass.sek1.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  # Create a new column combining tree_num and ayres with a period in between
mass.sek1.present.df$ayres.tree <- paste(mass.sek1.present.df$ayres, mass.sek1.present.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Brmlb
names_vector_brmlb <- character(0)
values_vector_brmlb <- numeric(0)
treatment_vector_brmlb <- character(0)

# Initialize empty vectors to store names and values for Armlb
names_vector_armlb <- character(0)
values_vector_armlb <- numeric(0)
treatment_vector_armlb <- character (0)

# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Bb_column <- paste("Bb.", X, sep = "")
    Ab_column <- paste("Ab.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    brmlb_result <- (mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Bb_column, "perc.mass_loss"] / 
                     (mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Bb_column, "perc.mass_loss"] + 
                      mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Ab_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    armlb_result <- (mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Ab_column, "perc.mass_loss"] / 
                     (mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Ab_column, "perc.mass_loss"] + 
                      mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Bb_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    brmlb_name <- paste("Brmlb", X, Y, sep = ".")
    armlb_name <- paste("Armlb", X, Y, sep = ".")
    
    # Extract the treatment from mass.hok1.present.df for the current combination
    treatment_name <- mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Ab_column, "treatment"]
    
    
    # Append the names and values to the respective vectors for Brmlb
    names_vector_brmlb <- c(names_vector_brmlb, brmlb_name)
    values_vector_brmlb <- c(values_vector_brmlb, brmlb_result)
    treatment_vector_brmlb <- c(treatment_vector_brmlb, treatment_name)
    
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_armlb <- c(names_vector_armlb, armlb_name)
    values_vector_armlb <- c(values_vector_armlb, armlb_result)
        treatment_vector_armlb <- c(treatment_vector_armlb, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
brmlb_df <- data.frame(names = names_vector_brmlb, values = values_vector_brmlb, treatment = treatment_vector_brmlb)
armlb_df <- data.frame(names = names_vector_armlb, values = values_vector_armlb, treatment = treatment_vector_armlb)

# View the resulting dataframes
print(brmlb_df)
print(armlb_df)

brmlb.armlb.df_1=rbind(brmlb_df, armlb_df)

brmlb.armlb.df_1
```

NEW: use this! This is a now a proper test dataframe with for loop!
generate **absent**, site a (SEK 1)
```{r}
mass.sek1.absent.df <- mass.sek1.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  # Create a new column combining tree_num and ayres with a period in between
mass.sek1.absent.df$ayres.tree <- paste(mass.sek1.absent.df$ayres, mass.sek1.absent.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values_3 <- 1:5
Y_values_3 <- 1:5

# Initialize empty vectors to store names and values for Brmlb
names_vector_brmlb <- character(0)
values_vector_brmlb <- numeric(0)
treatment_vector_brmlb <- character (0)

# Initialize empty vectors to store names and values for Armlb
names_vector_armlb <- character(0)
values_vector_armlb <- numeric(0)
treatment_vector_armlb <- character (0)


# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values_3) {
  for (Y in Y_values_3) {
    Bb_column <- paste("Bb.", X, sep = "")
    Ab_column <- paste("Ab.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    brmlb_result <- (mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Bb_column, "perc.mass_loss"] / 
                     (mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Bb_column, "perc.mass_loss"] + 
                      mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Ab_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    armlb_result <- (mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Ab_column, "perc.mass_loss"] / 
                     (mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Ab_column, "perc.mass_loss"] + 
                      mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Bb_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    brmlb_name <- paste("Brmlb", X, Y, sep = ".")
    armlb_name <- paste("Armlb", X, Y, sep = ".")
    
     # Extract the treatment from mass.hok1.present.df for the current combination
    treatment_name <- mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Ab_column, "treatment"]
    
    
    
    # Append the names and values to the respective vectors for Brmlb
    names_vector_brmlb <- c(names_vector_brmlb, brmlb_name)
    values_vector_brmlb <- c(values_vector_brmlb, brmlb_result)
    treatment_vector_brmlb <- c(treatment_vector_brmlb, treatment_name)
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_armlb <- c(names_vector_armlb, armlb_name)
    values_vector_armlb <- c(values_vector_armlb, armlb_result)
        treatment_vector_armlb <- c(treatment_vector_armlb, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
brmlb_ab_df <- data.frame(names = names_vector_brmlb, values = values_vector_brmlb, treatment = treatment_vector_brmlb)
armlb_ab_df <- data.frame(names = names_vector_armlb, values = values_vector_armlb, treatment = treatment_vector_armlb)


brmlb.armlb.absent.df_1=rbind(brmlb_ab_df, armlb_ab_df)
brmlb.armlb.absent.df_1

```


NEW: use this! merging HOK 1 and SEK 1 dataframe
merge **present**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site1.df=rbind(armla.brmla.df_1, brmlb.armlb.df_1)
mass.present.site1.df
# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Create an empty dataframe to store the results
result_df_1 <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla.", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla.", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb.", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb.", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.present.site1.df[mass.present.site1.df$names == armla_name, "values"] +
      mass.present.site1.df[mass.present.site1.df$names == brmlb_name, "values"])/2)/ (
      (mass.present.site1.df[mass.present.site1.df$names == armlb_name, "values"] +
      mass.present.site1.df[mass.present.site1.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
    result_name <- LETTERS[(X - 1) * 5 + Y]
    
  # Get the treatment for the current combination
    treatment_name <- mass.present.site1.df[mass.present.site1.df$names == armla_name, "treatment"]
 
# Append the name and value to the result dataframe
    result_df_1 <- rbind(result_df_1, data.frame(names = result_name, values = result, treatment= treatment_name))  }
}


# View the resulting dataframe
result_df_1
```


NEW: use this! merging HOK 1 and SEK 1 dataframe
merge **absent**, merging HOK 1 and SEK 1 dataframe
```{r}
#merge both present site 1 dataframes
mass.absent.site1.df=rbind(armla.brmla.absent.df_1, brmlb.armlb.absent.df_1)

# Create an empty dataframe to store the results
result_ab_df_1 <- data.frame(names = character(0), values = numeric(0), treatment = character (0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla.", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla.", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb.", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb.", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.absent.site1.df[mass.absent.site1.df$names == armla_name, "values"] +
      mass.absent.site1.df[mass.absent.site1.df$names == brmlb_name, "values"])/2)/ (
      (mass.absent.site1.df[mass.absent.site1.df$names == armlb_name, "values"] +
      mass.absent.site1.df[mass.absent.site1.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
result_name <- paste0(LETTERS[(X - 1) * 5 + Y], "_1")

# Get the treatment for the current combination
    treatment_name <- mass.absent.site1.df[mass.absent.site1.df$names == armla_name, "treatment"]
    

    # Append the name and value to the result dataframe
    result_ab_df_1 <- rbind(result_ab_df_1, data.frame(names = result_name, values = result, treatment= treatment_name))
  }
}


# View the resulting dataframe
result_ab_df_1


```


combine P and AB; plot check for P and AB site 1 
```{r}
#combine sites 1 and 2
hfai.allsite1.df=rbind(result_df_1, result_ab_df_1)

ggplot(hfai.allsite1.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  geom_bar(stat = "identity", position= "dodge", alpha = 0.5) +
  scale_fill_viridis_d()+
  xlab("reciprocated pairs") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.x = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  )


hfai.allsite1.df$values %>%
  n_distinct()



```


NOTICE: THIS IS NOW SITE 2:

NEW: use this! This is a now a proper test dataframe with for loop!
generate **present**, site a (HOK 2)
```{r}
########################
mass.hok2.present.df <- mass.hok2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)
# Create a new column combining tree_num and ayres with a period in between
mass.hok2.present.df$ayres.tree <- paste(mass.hok2.present.df$ayres, mass.hok2.present.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Armla
names_vector_armla <- character(0)
values_vector_armla <- numeric(0)
treatment_vector_armla<- character(0)

# Initialize empty vectors to store names and values for Brmla
names_vector_brmla <- character(0)
values_vector_brmla<- numeric(0)
treatment_vector_brmla<- character(0)


# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Aa_column <- paste("Aa.", X, sep = "")
    Ba_column <- paste("Ba.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    armla_result <- (mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Aa_column, "perc.mass_loss"] / 
                     (mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Aa_column, "perc.mass_loss"] + 
                      mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Ba_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    brmla_result <- (mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Ba_column, "perc.mass_loss"] / 
                     (mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Ba_column, "perc.mass_loss"] + 
                      mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Aa_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    armla_name <- paste("Armla", X, Y, sep = ".")
    brmla_name <- paste("Brmla", X, Y, sep = ".")
    
    # Get the treatment for the current combination
    treatment_name <- mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Aa_column, "treatment"]
    
    # Append the names and values to the respective vectors for Armla
    names_vector_armla <- c(names_vector_armla, armla_name)
    values_vector_armla <- c(values_vector_armla, armla_result)
    treatment_vector_armla<- c(treatment_vector_armla, treatment_name)

    # Append the names and values to the respective vectors for Brmla
    names_vector_brmla <- c(names_vector_brmla, brmla_name)
    values_vector_brmla <- c(values_vector_brmla, brmla_result)
        treatment_vector_brmla<- c(treatment_vector_brmla, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
armla_df_2 <- data.frame(names = names_vector_armla, values = values_vector_armla, treatment = treatment_vector_armla)
brmla_df_2 <- data.frame(names = names_vector_brmla, values = values_vector_brmla, treatment = treatment_vector_brmla)



# View the resulting dataframes
print(armla_df_2)
print(brmla_df_2)

armla.brmla.df_2=rbind(armla_df_2, brmla_df_2)


```


NEW: use this! This is a now a proper test dataframe with for loop!
generate **absent**, site a (HOK 2)
```{r}
########################
mass.hok2.absent.df <- mass.hok2.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Create a new column combining tree_num and ayres with a period in between
mass.hok2.absent.df$ayres.tree <- paste(mass.hok2.absent.df$ayres, mass.hok2.absent.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Armla
names_vector_armla <- character(0)
values_vector_armla <- numeric(0)
treatment_vector_armla <- character(0)

# Initialize empty vectors to store names and values for Brmla
names_vector_brmla <- character(0)
values_vector_brmla <- numeric(0)
treatment_vector_brmla <- character(0)


# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Aa_column <- paste("Aa.", X, sep = "")
    Ba_column <- paste("Ba.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    armla_result <- (mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Aa_column, "perc.mass_loss"] / 
                     (mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Aa_column, "perc.mass_loss"] + 
                      mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Ba_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    brmla_result <- (mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Ba_column, "perc.mass_loss"] / 
                     (mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Ba_column, "perc.mass_loss"] + 
                      mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Aa_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    armla_name <- paste("Armla", X, Y, sep = ".")
    brmla_name <- paste("Brmla", X, Y, sep = ".")
    
    # create treatment 
    
    treatment_name <- mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Aa_column, "treatment"]
    
    # Append the names and values to the respective vectors for Armla
    names_vector_armla <- c(names_vector_armla, armla_name)
    values_vector_armla <- c(values_vector_armla, armla_result)
    treatment_vector_armla <- c(treatment_vector_armla, treatment_name)
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_brmla <- c(names_vector_brmla, brmla_name)
    values_vector_brmla <- c(values_vector_brmla, brmla_result)
        treatment_vector_brmla <- c(treatment_vector_brmla, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
armla_ab_df_2 <- data.frame(names = names_vector_armla, values = values_vector_armla, treatment = treatment_vector_armla)
brmla_ab_df_2 <- data.frame(names = names_vector_brmla, values = values_vector_brmla, treatment = treatment_vector_brmla)

# View the resulting dataframes
print(armla_ab_df_2)
print(armla_ab_df_2)

armla.brmla.absent.df_2=rbind(armla_ab_df_2, brmla_ab_df_2)
armla.brmla.absent.df_2

```

NEW: use this! This is a now a proper test dataframe with for loop!
generate **present**, site a (SEK 2)
```{r}
mass.sek2.present.df <- mass.sek2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  # Create a new column combining tree_num and ayres with a period in between
mass.sek2.present.df$ayres.tree <- paste(mass.sek2.present.df$ayres, mass.sek2.present.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Brmlb
names_vector_brmlb <- character(0)
values_vector_brmlb <- numeric(0)
treatment_vector_brmlb <- character(0)


# Initialize empty vectors to store names and values for Armlb
names_vector_armlb <- character(0)
values_vector_armlb <- numeric(0)
treatment_vector_armlb <- character(0)

# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Bb_column <- paste("Bb.", X, sep = "")
    Ab_column <- paste("Ab.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    brmlb_result <- (mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Bb_column, "perc.mass_loss"] / 
                     (mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Bb_column, "perc.mass_loss"] + 
                      mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Ab_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    armlb_result <- (mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Ab_column, "perc.mass_loss"] / 
                     (mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Ab_column, "perc.mass_loss"] + 
                      mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Bb_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    brmlb_name <- paste("Brmlb", X, Y, sep = ".")
    armlb_name <- paste("Armlb", X, Y, sep = ".")
    
    # create treatment
    treatment_name <- mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Bb_column, "treatment"]

    
    # Append the names and values to the respective vectors for Brmlb
    names_vector_brmlb <- c(names_vector_brmlb, brmlb_name)
    values_vector_brmlb <- c(values_vector_brmlb, brmlb_result)
    treatment_vector_brmlb<- c(treatment_vector_brmlb, treatment_name)
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_armlb <- c(names_vector_armlb, armlb_name)
    values_vector_armlb <- c(values_vector_armlb, armlb_result)
        treatment_vector_armlb<- c(treatment_vector_armlb, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
brmlb_df_2 <- data.frame(names = names_vector_brmlb, values = values_vector_brmlb, treatment = treatment_vector_brmlb)
armlb_df_2 <- data.frame(names = names_vector_armlb, values = values_vector_armlb, treatment = treatment_vector_armlb)

# View the resulting dataframes
print(brmlb_df_2)
print(armlb_df_2)

brmlb.armlb.df_2=rbind(brmlb_df_2, armlb_df_2)
brmlb.armlb.df_2

```

NEW: use this! This is a now a proper test dataframe with for loop!
generate **absent**, site a (SEK 2)
```{r}
mass.sek2.absent.df <- mass.sek2.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  # Create a new column combining tree_num and ayres with a period in between
mass.sek2.absent.df$ayres.tree <- paste(mass.sek2.absent.df$ayres, mass.sek2.absent.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values<- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Brmlb
names_vector_brmlb <- character(0)
values_vector_brmlb <- numeric(0)
treatment_vector_brmlb<- character(0)

# Initialize empty vectors to store names and values for Armlb
names_vector_armlb <- character(0)
values_vector_armlb <- numeric(0)
treatment_vector_armlb<- character(0)


# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Bb_column <- paste("Bb.", X, sep = "")
    Ab_column <- paste("Ab.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    brmlb_result <- (mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Bb_column, "perc.mass_loss"] / 
                     (mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Bb_column, "perc.mass_loss"] + 
                      mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Ab_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    armlb_result <- (mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Ab_column, "perc.mass_loss"] / 
                     (mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Ab_column, "perc.mass_loss"] + 
                      mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Bb_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    brmlb_name <- paste("Brmlb", X, Y, sep = ".")
    armlb_name <- paste("Armlb", X, Y, sep = ".")
    
    # create treatment 
    treatment_name <- mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Bb_column, "treatment"]
    
    
    # Append the names and values to the respective vectors for Brmlb
    names_vector_brmlb <- c(names_vector_brmlb, brmlb_name)
    values_vector_brmlb <- c(values_vector_brmlb, brmlb_result)
    treatment_vector_brmlb<- c(treatment_vector_brmlb, treatment_name)
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_armlb <- c(names_vector_armlb, armlb_name)
    values_vector_armlb <- c(values_vector_armlb, armlb_result)
        treatment_vector_armlb<- c(treatment_vector_armlb, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
brmlb_ab_df_2 <- data.frame(names = names_vector_brmlb, values = values_vector_brmlb, treatment = treatment_vector_brmlb)
armlb_ab_df_2 <- data.frame(names = names_vector_armlb, values = values_vector_armlb, treatment = treatment_vector_armlb)


brmlb.armlb.absent.df_2=rbind(brmlb_ab_df_2, armlb_ab_df_2)
brmlb.armlb.absent.df_2

```

NEW: use this! merging HOK 2 and SEK 2 dataframe
merge **present**, site a and b (aka HOK 2 and SEK 2) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site2.df=rbind(armla.brmla.df_2, brmlb.armlb.df_2)
mass.present.site2.df
# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Create an empty dataframe to store the results
result_df_2 <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla.", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla.", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb.", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb.", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.present.site2.df[mass.present.site2.df$names == armla_name, "values"] +
      mass.present.site2.df[mass.present.site2.df$names == brmlb_name, "values"])/2)/ (
      (mass.present.site2.df[mass.present.site2.df$names == armlb_name, "values"] +
      mass.present.site2.df[mass.present.site2.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
    result_name <- paste0(LETTERS[(X - 1) * 5 + Y], "_2")
    
  # Get the treatment for the current combination
    treatment_name <- mass.present.site2.df[mass.present.site2.df$names == armla_name, "treatment"]
 
# Append the name and value to the result dataframe
    result_df_2 <- rbind(result_df_2, data.frame(names = result_name, values = result, treatment= treatment_name))  }
}


# View the resulting dataframe

print(result_df_2)



```


NEW: use this! merging HOK 2 and SEK 2 dataframe
merge **absent**, merging HOK 2 and SEK 2 dataframe
```{r}
#merge both present site 1 dataframes
mass.absent.site2.df=rbind(armla.brmla.absent.df_2, brmlb.armlb.absent.df_2)
mass.absent.site2.df
# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Create an empty dataframe to store the results
result_ab_df_2 <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla.", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla.", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb.", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb.", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.absent.site2.df[mass.absent.site2.df$names == armla_name, "values"] +
      mass.absent.site2.df[mass.absent.site2.df$names == brmlb_name, "values"])/2)/ (
      (mass.absent.site2.df[mass.absent.site2.df$names == armlb_name, "values"] +
      mass.absent.site2.df[mass.absent.site2.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
    result_name <- paste0(LETTERS[(X - 1) * 5 + Y], "_3")
    
  # Get the treatment for the current combination
    treatment_name <- mass.absent.site2.df[mass.absent.site2.df$names == armla_name, "treatment"]
 
# Append the name and value to the result dataframe
    result_ab_df_2 <- rbind(result_ab_df_2, data.frame(names = result_name, values = result, treatment= treatment_name))  }
}


# View the resulting dataframe
result_ab_df_2

```


NEW: use this! merging all dataframes (Sites 1 and 2; includes P and AB)
merge all sites; HFAI plots
```{r}
#combine sites 1 and 2 including P and AB
hfai.allsites.df=rbind(result_df_1, result_df_2, result_ab_df_1, result_ab_df_2)
hfai.allsites.df$treatment=as.factor(hfai.allsites.df$treatment)


####################  P ONLY ###################
#filter P treatment only for HFA effect
hfai.present.allsites.df=hfai.allsites.df%>%
  filter(treatment == "P")


# Filter the dataframe to include only positive cases
positive_present_df <- hfai.present.allsites.df[hfai.present.allsites.df$values > 0, ]

# Calculate the mean and standard error (SE)
mean_hfa <- mean(positive_present_df$values)
se_hfa <- sd(positive_present_df$values) / sqrt(nrow(positive_present_df))
# Perform a t-test
t_test_present_result <- t.test(positive_present_df$values, mu = 0)
t_test_present_result



# Format the t-test result
t_test_label <- paste("p =", formatC(t_test_present_result$p.value, format = "e", digits = 2), ",", "t-test =", round(t_test_present_result$statistic, 2))
# Format the t-test result
p_value <- t_test_present_result$p.value
formatted_p_value <- if (p_value < 0.0001) {
  "p < 0.0001"
} else {
  paste("p =", formatC(p_value, digits = 1))
}


#positive and negative values
total_positive <- sum(hfai.present.allsites.df$values > 0)
total_negative <- sum(hfai.present.allsites.df$values < 0)
# Calculate the percentage of positive values relative to negative
percentage_positive <- (total_positive / (total_positive + total_negative)) * 100
percentage_negative <- (total_negative / (total_negative + total_positive)) * 100
# 68% positive
# 32% negative

#just P treatment
#just P treatment with coord_flip
hfai.present.allsites.plot=ggplot(hfai.present.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  scale_fill_manual(values = "grey")+
  geom_bar(stat = "identity", position ="dodge", alpha = 0.6) +
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.x = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 14.5, linetype = "dashed", color = "red") +
scale_y_continuous(limits = c(-20, 70), breaks = seq(-20, 80, by = 20))
hfai.present.allsites.plot

#just P treatment with coord_flip
hfai.present.allsites.plot_flipped=ggplot(hfai.present.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  scale_fill_manual(values = "grey")+
  geom_bar(stat = "identity", position ="dodge", alpha = 0.6) +
  coord_flip()+
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 14.5, linetype = "dashed", color = "red") +
scale_y_continuous(limits = c(-20, 70), breaks = seq(-20, 70, by = 10))



hfai.present.allsites.plot_flipped=ggplot(hfai.present.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  scale_fill_manual(values = "grey")+
  geom_bar(stat = "identity", position ="dodge", alpha = 0.6) +
  coord_flip()+
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 16.5, linetype = "dashed", color = "red")

hfai.present.allsites.plot_flipped
ggsave("figures/hfai.present.allsites.plot_flipped.pdf", hfai.present.allsites.plot_flipped, width = 14, height = 10)


```


```{r}
######################### BOTH P and AB #########################################

# Filter the dataframe to include only positive cases

mean_positive_P <- mean(hfai.allsites.df[hfai.allsites.df$treatment == "P" & hfai.allsites.df$values > 0, "values"])

positive_values <- hfai.allsites.df %>%
  filter(values > 0)

mean_positive <- mean(positive_values$values)
se_positive <- sd(positive_values$values) / sqrt(nrow(positive_values))

se_positive_P <- sd(hfai.allsites.df[hfai.allsites.df$treatment == "P" & hfai.allsites.df$values > 0, "values"]) / sqrt(nrow(hfai.allsites.df[hfai.allsites.df$treatment == "P" & hfai.allsites.df$values > 0, ]))


mean_positive_AB <- mean(hfai.allsites.df[hfai.allsites.df$treatment == "AB" & hfai.allsites.df$values > 0, "values"])
# mean positive P = 13.7
# mean positive AB = 14.2

print(hfai.allsites.df[hfai.allsites.df$treatment=="P" & hfai.allsites.df$values>0, "values"])
print(hfai.allsites.df[hfai.allsites.df$treatment=="AB" & hfai.allsites.df$values>0, "values"])

# Calculate the mean of positive "P" values relative to all positive values
mean_p_relative_to_all <- mean(p_positive_subset$values) / mean(hfai.allsites.df[hfai.allsites.df$values > 0, ]$values)


View(hfai.allsites.df)

hfai.allsites.df %>%
  filter(values > 0) %>%
  summarise(mean = mean(values), se = sd(values) / sqrt(n()))

hfai.allsites.df %>%
  filter(values < 0) %>%
  summarise(mean = mean(values))

hfai.allsites.df%>%
  filter(treatment == "AB")%>%
  filter(values > 0)%>%
n_distinct()
# P 34 > 0; AB 19 > 0; P 16 < 0; AB 31 < 0

# Subset the dataframe to include only positive values for each treatment
positive_values <- hfai.allsites.df[hfai.allsites.df$values > 0, "values"]

positive_values_P <- hfai.allsites.df[hfai.allsites.df$treatment == "P" & hfai.allsites.df$values > 0, "values"]
positive_values_AB <- hfai.allsites.df[hfai.allsites.df$treatment == "AB" & hfai.allsites.df$values > 0, "values"]

# Perform a t-test for each treatment against 0
t_test<- t.test(positive_values, mu = 0)
t_test_P <- t.test(positive_values_P, mu = 0)
t_test_AB <- t.test(positive_values_AB, mu = 0)

# Calculate Cohen's d for each treatment
cohen_d_P <- t_test_P$estimate / sd(positive_values_P)
cohen_d_AB <- t_test_AB$estimate / sd(positive_values_AB)
# Filter the dataframe for positive values (both "P" and "AB")
positive_values_subset <- hfai.allsites.df[hfai.allsites.df$values > 0, ]


# Calculate the proportion of positive values for the "P" treatment relative to all positive values
proportion_positive_P <- sum(hfai.allsites.df$treatment == "P" & hfai.allsites.df$values > 0) /
                        sum(hfai.allsites.df$values > 0)
proportion_positive_AB <- sum(hfai.allsites.df$treatment == "AB" & hfai.allsites.df$values > 0) /
                        sum(hfai.allsites.df$values > 0)

positive_values_subset%>%
  filter(treatment == "AB")%>%
  n_distinct()
hfai.allsites.df%>%
  filter(values > 0)%>%
  n_distinct()
#53 positive
# 47 negative


# Perform a t-test
t_test_present_result <- t.test(positive_all_df$values, mu = 0)
t_test_present_result

# Format the t-test result
t_test_label <- paste("p =", formatC(t_test_present_result$p.value, format = "e", digits = 2), ",", "t-test =", round(t_test_present_result$statistic, 2))
# Format the t-test result
p_value <- t_test_present_result$p.value
formatted_p_value <- if (p_value < 0.0001) {
  "p < 0.0001"
} else {
  paste("p =", formatC(p_value, digits = 1))
}

#### proportion that are positive

# Filter the dataframe for "AB" and "P" treatments
ab_p_subset <- hfai.allsites.df[hfai.allsites.df$treatment %in% c("AB", "P"), ]

# Further filter to include only rows where values are greater than 0
positive_values_subset <- ab_p_subset[ab_p_subset$values > 0, ]
View(positive_values_subset)
# Calculate the proportion of values greater than 0 for "AB"
prop_positive_ab <- sum(positive_values_subset$treatment == "AB") / nrow(positive_values_subset)

# Calculate the proportion of values greater than 0 for "P"
prop_positive_p <- sum(positive_values_subset$treatment == "P") / nrow(positive_values_subset)


##### proportion that are negative
# Further filter to include only rows where values are greater than 0
neg_values_subset <- ab_p_subset[ab_p_subset$values < 0, ]

# Calculate the proportion of values greater than 0 for "AB"
prop_neg_ab <- sum(neg_values_subset$treatment == "AB") / nrow(ab_p_subset$values)


# Calculate the proportion of values greater than 0 for "P"
prop_neg_p <- sum(neg_values_subset$treatment == "P") / nrow(ab_p_subset$values)

# Calculate mean and SE for 'AB' and 'P' values individually
mean_ab <- mean(positive_values_subset[positive_values_subset$treatment == "AB", "values"])

# Calculate the weighted mean for group P
weighted_mean_P <- sum(positive_values_subset[positive_values_subset$treatment == "P", "values"]) / sum(positive_values_subset$treatment == "P")

# Calculate the weighted mean for group AB
weighted_mean_AB <- sum(positive_values_subset[positive_values_subset$treatment == "AB", "values"]) / sum(positive_values_subset$treatment == "AB")

# Print the results
weighted_mean_P
weighted_mean_AB

med_ab <- median(positive_values_subset[positive_values_subset$treatment == "AB", "values"])
se_ab <- sd(positive_values_subset[positive_values_subset$treatment == "AB", "values"]) / sqrt(sum(positive_values_subset$treatment == "AB"))

mean_p <- mean(positive_values_subset[positive_values_subset$treatment == "P", "values"])
med_p <- median(positive_values_subset[positive_values_subset$treatment == "P", "values"])
se_p <- sd(positive_values_subset[positive_values_subset$treatment == "P", "values"]) / sqrt(sum(positive_values_subset$treatment == "P"))


# Perform a t-test
t_test_result <- t.test(positive_values_subset[positive_values_subset$treatment == "P", "values"], mu = 0)
t_test_result <- t.test(positive_values_subset[positive_values_subset$treatment == "AB", "values"], mu = 0)

d <- (mean_p - mean_ab) / sqrt((se_p^2 + se_ab^2) / 2)
d
#positive and negative values
total_positive <- sum(hfai.allsites.df$values > 0)
total_negative <- sum(hfai.allsites.df$values < 0)
# Calculate the percentage of positive values relative to negative
percentage_positive <- (total_positive / (total_positive + total_negative)) * 100
percentage_negative <- (total_negative / (total_negative + total_positive)) * 100

# 53% positive
# 47% negative

#both P and AB treatments
hfai.allsites.plot=ggplot(hfai.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  geom_bar(stat = "identity", position ="dodge", alpha = 0.3) +
  scale_fill_viridis_d()+
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.x = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 44.5, linetype = "dashed", color = "red")+
  scale_y_continuous(
    limits = c(-65, 255),
     breaks = seq(-65, 250, by = 65),
    expand = c(0, 0)
  )

hfai.allsites.plot
min(hfai.allsites.df$values) 
max(hfai.allsites.df$values)


ggplot(hfai.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  geom_bar(stat = "identity", position ="dodge", alpha = 0.3) +
  scale_fill_manual(values = c("grey", "grey1"))+
  coord_flip()+
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 44.5, linetype = "dashed", color = "red")+
  scale_y_continuous(
    limits = c(-65, 255),
     breaks = seq(-65, 250, by = 65),
    expand = c(0, 0)
  )

hfai.allsites.plot_flipped=ggplot(hfai.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  geom_bar(stat = "identity", position ="dodge", alpha = 0.4) +
  scale_fill_manual(values = c("grey", "grey0"))+
  coord_flip()+
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 46.5, linetype = "dashed", color = "red")
hfai.allsites.plot_flipped

max(hfai.allsites.df[hfai.allsites.df$treatment =="AB" & hfai.allsites.df$values >0, "values"])

ggsave("figures/hfai.allsites.plot_flipped.pdf", hfai.allsites.plot_flipped, width = 14, height = 10)


```

------------------------END--------------------------------
Everything below isnt used



















This is also the proper dataframe
generate **present**, site a (HOK 2)
```{r}
mass.hok2.present.df <- mass.hok2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values_z <- numeric()
rml_values_mass <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_z_loss <- mass.hok2.present.df$z_score[
      mass.hok2.present.df$tree_id == "HOK" & mass.hok2.present.df$tree_num == tree_num
    ]
    sek_z_loss <- mass.hok2.present.df$z_score[
      mass.hok2.present.df$tree_id == "SEK" & mass.hok2.present.df$tree_num == tree_num
    ]
    
    rml.z <- ifelse(
      ayres == "Aa", hok_z_loss / (hok_z_loss + sek_z_loss),
      ifelse(ayres == "Ba", sek_z_loss / (sek_z_loss + hok_z_loss), NA)
    )
    
    rml_values_z <- c(rml_values_z, rml.z * 100)
    
    hok_mass_loss <- mass.hok2.present.df$perc.mass_loss[
      mass.hok2.present.df$tree_id == "HOK" & mass.hok2.present.df$tree_num == tree_num
    ]
    sek_mass_loss <- mass.hok2.present.df$perc.mass_loss[
      mass.hok2.present.df$tree_id == "SEK" & mass.hok2.present.df$tree_num == tree_num
    ]
    
    rml.mass <- ifelse(
      ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
      ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss), NA)
    )
    
    rml_values_mass <- c(rml_values_mass, rml.mass * 100)
  }
}

# Add the rml values to the dataframe
mass.hok2.present.df$rml.z <- rml_values_z
mass.hok2.present.df$rml.mass <- rml_values_mass

# Create a new column combining tree_num and ayres with a period in between
mass.hok2.present.df$ayres.tree <- paste(mass.hok2.present.df$ayres, mass.hok2.present.df$tree_num, sep = ".")

# Create a new column based on Ayres HFAI equation 
mass.hok2.present.df$ayres.eq <- ifelse(
  mass.hok2.present.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""),
  paste("Armla", tree_num_values, sep = "."),
  ifelse(
    mass.hok2.present.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""),
    paste("Brmla", tree_num_values, sep = "."),
    NA
  )
)


```


This is also the proper dataframe
generate **present**, site a (SEK 2)
```{r}
mass.sek2.present.df <- mass.sek2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values_z <- numeric()
rml_values_mass <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_z_loss <- mass.sek2.present.df$z_score[
      mass.sek2.present.df$tree_id == "HOK" & mass.sek2.present.df$tree_num == tree_num
    ]
    sek_z_loss <- mass.sek2.present.df$z_score[
      mass.sek2.present.df$tree_id == "SEK" & mass.sek2.present.df$tree_num == tree_num
    ]
    
    rml.z <- ifelse(
      ayres == "Ab", hok_z_loss / (hok_z_loss + sek_z_loss),
      ifelse(ayres == "Bb", sek_z_loss / (sek_z_loss + hok_z_loss), NA)
    )
    
    rml_values_z <- c(rml_values_z, rml.z * 100)
    
    hok_mass_loss <- mass.sek2.present.df$perc.mass_loss[
      mass.sek2.present.df$tree_id == "HOK" & mass.sek2.present.df$tree_num == tree_num
    ]
    sek_mass_loss <- mass.sek2.present.df$perc.mass_loss[
      mass.sek2.present.df$tree_id == "SEK" & mass.sek2.present.df$tree_num == tree_num
    ]
    
    rml.mass <- ifelse(
      ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
      ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss), NA)
    )
    
    rml_values_mass <- c(rml_values_mass, rml.mass * 100)
  }
}

# Add the rml values to the dataframe
mass.sek2.present.df$rml.z <- rml_values_z
mass.sek2.present.df$rml.mass <- rml_values_mass

# Create a new column combining tree_num and ayres with a period in between
mass.sek2.present.df$ayres.tree <- paste(mass.sek2.present.df$ayres, mass.sek2.present.df$tree_num, sep = ".")

# Create a new column based on Ayres HFAI equation 
mass.sek2.present.df$ayres.eq <- ifelse(
  mass.sek2.present.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""),
  paste("Armlb", tree_num_values, sep = "."),
  ifelse(
    mass.sek2.present.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""),
    paste("Brmlb", tree_num_values, sep = "."),
    NA
  )
)


```



NEW: use this! merging HOK 1 and SEK 1 dataframe
merge **present**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site1.df=rbind(armla.brmla.df, brmlb.armlb.df)


# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Create an empty dataframe to store the results
result_df <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla..", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla..", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb..", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb..", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.present.site1.df[mass.present.site1.df$names == armla_name, "values"] +
      mass.present.site1.df[mass.present.site1.df$names == brmlb_name, "values"])/2)/ (
      (mass.present.site1.df[mass.present.site1.df$names == armlb_name, "values"] +
      mass.present.site1.df[mass.present.site1.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
    result_name <- LETTERS[(X - 1) * 5 + Y]
    
    # Get the treatment associated with the current index
  treatment <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Armla.", i), ]$treatment
  

    # Append the name and value to the result dataframe
    result_df <- rbind(result_df, data.frame(names = result_name, values = result, treatment = treatment))
  }
}


# View the resulting dataframe
View(result_df)
```


NEW: use this! merging HOK 2 and SEK 2 dataframe
merge **present**, merging HOK 2 and SEK 2 dataframe
```{r}
#merge both present site 1 dataframes
mass.present.site2.df=rbind(armla.brmla.df_2, brmlb.armlb.df_2)

# Create an empty dataframe to store the results
result_df_2 <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla..", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla..", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb..", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb..", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result_2 <- 
((
  (mass.present.site2.df[mass.present.site2.df$names == armla_name, "values"] +
      mass.present.site2.df[mass.present.site2.df$names == brmlb_name, "values"])/2)/ (
      (mass.present.site2.df[mass.present.site2.df$names == armlb_name, "values"] +
      mass.present.site2.df[mass.present.site2.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
result_name_2 <- paste0(LETTERS[(X - 1) * 5 + Y], "_2")

 # Get the treatment associated with the current index
  treatment <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  

    # Append the name and value to the result dataframe
    result_df_2 <- rbind(result_df_2, data.frame(names = result_name_2, values = result_2, treatment = treatment))
  }
}


# View the resulting dataframe
View(result_df_2)

#combine sites 1 and 2
hfai.allsites.df=rbind(result_df, result_df_2)
View(hfai.allsites.df)
```






This is the proper dataframe
merge **present**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site1.df=rbind(mass.hok1.present.df, mass.sek1.present.df)

# Create an empty data frame to store the results
hfai.df <- data.frame(hfai.z = numeric(0), hfai.mass = numeric(0), label = character(0), treatment = character(0))

# Create an empty list to store the data frames
hfai_dfs <- list()

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla.z <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Armla.", i), ]$rml.z
  brmlb.z <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml.z
  armlb.z <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml.z
  brmla.z <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml.z
  
 print(paste("armla.z:", armla.z))
  print(paste("brmlb.z:", brmlb.z))
  print(paste("armlb.z:", armlb.z))
  print(paste("brmla.z:", brmla.z))
  
  armla.mass <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Armla.", i), ]$rml.mass
  brmlb.mass <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml.mass
  armlb.mass <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml.mass
  brmla.mass <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml.mass
 
  
print(paste("armla.mass:", armla.mass))
  print(paste("brmlb.mass:", brmlb.mass))
  print(paste("armlb.mass:", armlb.mass))
  print(paste("brmla.mass:", brmla.mass))


  result.z <- (((armla.z + brmlb.z) / (armlb.z + brmla.z)) * 100) - 100
  result.mass <- (((armla.mass + brmlb.mass) / (armlb.mass + brmla.mass)) * 100) - 100

  print(paste("Length of result.z:", length(result.z)))
  print(paste("Length of result.mass:", length(result.mass)))
  
  # Get the treatment associated with the current index
  treatment <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Create a data frame with the results, label, and treatment
  hfai_df <- data.frame(hfai.z = result.z, hfai.mass = result.mass, label = LETTERS[i], treatment = treatment)
  
  # Append the data frame to the list
  hfai_dfs[[i]] <- hfai_df
}


# Combine the list of data frames into a single data frame
hfai.df <- do.call(rbind, hfai_dfs)

#HFAI (this is a stop and check to verify that values are correct!)
(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```

This is the proper dataframe
merge **present**, site a and b (aka HOK 2 and SEK 2) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site2.df=rbind(mass.hok2.present.df, mass.sek2.present.df)

# Calculate the index of the last label in the existing hfai.df
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Create an empty list to store the new data frames
hfai_dfs <- list()

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla.z <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armla.", i), ]$rml.z
  brmlb.z <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml.z
  armlb.z <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml.z
  brmla.z <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml.z
  
  armla.mass <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armla.", i), ]$rml.mass
  brmlb.mass <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml.mass
  armlb.mass <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml.mass
  brmla.mass <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml.mass
  
  result.z <- (((armla.z + brmlb.z) / (armlb.z + brmla.z)) * 100) - 100
  result.mass <- (((armla.mass + brmlb.mass) / (armlb.mass + brmla.mass)) * 100) - 100

  # Get the treatment associated with the current index
  treatment <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Create a data frame with the new results, label, and treatment
  hfai_df <- data.frame(hfai.z = result.z, hfai.mass = result.mass, label = next_label, treatment = treatment)
  
  # Append the data frame to the list
  hfai_dfs[[i]] <- hfai_df
}

# Combine the list of new data frames with the existing hfai.df
hfai.df <- rbind(hfai.df, do.call(rbind, hfai_dfs))




```

This is the proper plot
another plot check
```{r}
######## using mass loss

hfai.mass.plot=ggplot(hfai.df, aes(x = reorder(label, hfai.mass), y = hfai.mass, fill = label)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  scale_fill_viridis_d(guide = "none")+
  xlab("reciprocated pairs") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.x = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  )

print(hfai.mass.plot)

hfai.mass.plot

ggsave("figures/hfai.mass.plot.pdf", hfai.mass.plot, width = 14, height = 10)
```

Full Pairwise comparisons for HFAI

```{r}

# Initialize an empty matrix to store the results
armla_matrix <- matrix(NA, nrow = length(tree_num_values), ncol = length(tree_num_values))
brmla_matrix <- matrix(NA, nrow = length(tree_num_values), ncol = length(tree_num_values))

# Loop through the combinations and calculate Armla and Brmla values
for (i in 1:length(tree_num_values)) {
  for (j in 1:length(tree_num_values)) {
    ayres_num <- tree_num_values[i]
    tree_num <- tree_num_values[j]
    
    for (k in 1:length(tree_num_values)) {
      ba_num <- tree_num_values[k]
      
      armla <- ifelse(
        ayres_num == "Aa", mass.hok1.present.df$perc.mass_loss[
          mass.hok1.present.df$ayres.eq == paste("Armla.", ayres_num, ".", ba_num) &
            mass.hok1.present.df$tree_num == tree_num
        ],
        NA
      )
      
      armla_value <- ifelse(
        ayres_num == "Aa", armla / (armla + mass.hok1.present.df$perc.mass_loss[
          mass.hok1.present.df$ayres.eq == paste("Brmla.", ba_num, ".", ayres_num) &
            mass.hok1.present.df$tree_num == tree_num
        ]) * 100,
        NA
      )
      
      armla_matrix[i, k] <- armla_value
      
      brmla <- ifelse(
        ayres_num == "Ba", mass.hok1.present.df$perc.mass_loss[
          mass.hok1.present.df$ayres.eq == paste("Brmla.", ayres_num, ".", ba_num) &
            mass.hok1.present.df$tree_num == tree_num
        ],
        NA
      )
      
      brmla_value <- ifelse(
        ayres_num == "Ba", brmla / (brmla + mass.hok1.present.df$perc.mass_loss[
          mass.hok1.present.df$ayres.eq == paste("Armla.", ba_num, ".", ayres_num) &
            mass.hok1.present.df$tree_num == tree_num
        ]) * 100,
        NA
      )
      
      brmla_matrix[i, k] <- brmla_value
    }
  }
}

# Add the calculated values to the dataframe
for (i in 1:length(tree_num_values)) {
  for (j in 1:length(tree_num_values)) {
    mass.hok1.present.df[paste("Armla.", tree_num_values[i], ".", tree_num_values[j], sep = "")] <- armla_matrix[i, j]
    mass.hok1.present.df[paste("Brmla.", tree_num_values[i], ".", tree_num_values[j], sep = "")] <- brmla_matrix[i, j]
  }
}

# Define the columns to pivot
columns_to_pivot <- names(mass.hok1.present.df)[grepl("^Armla\\.|^Brmla\\.", names(mass.hok1.present.df))]

# Pivot the DataFrame to long format
mass.hok1.present.df_long <- mass.hok1.present.df %>%
  pivot_longer(
    cols = all_of(columns_to_pivot),
    names_to = "Name",
    values_to = "Value"
  ) %>%
  mutate(Name = sub("\\d\\.\\d$", "", Name))  # Remove the last part of the column names

# View the resulting DataFrame
head(mass.hok1.present.df_long)




```









```{r}

##this is the proper plot
########### using z loss
hfai.z.plot=ggplot(hfai.df[hfai.df$hfai.z > -800,], aes(x = reorder(label, hfai.z), y = hfai.z, fill = label)) +
  geom_bar(stat = "identity", fill = "white", color = "black") +
  geom_col_pattern(
    aes(pattern = treatment),
    colour = "black",
    pattern_fill = "black",
    pattern_angle = 45,
    pattern_density = 0.1,
    pattern_spacing = 0.01,
    position = position_dodge2(preserve = 'single'),
    alpha = 0.2  # Adjust the alpha value to control transparency
  ) +
  scale_pattern_manual(
    values = c("none", "stripe"),
    guide = guide_legend(override.aes = list(fill = "grey70")) # <- make lighter
  ) +
  xlab("") +
  ylab("Home-field advantage (% faster decomposition at home)") +
coord_flip(ylim = c(-800,300)) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
  )

hfai.z.plot=ggplot(hfai.df[hfai.df$hfai.z > -800,], aes(x = reorder(label, hfai.z), y = hfai.z, fill = label)) +
  geom_bar(stat = "identity", fill = "white", color = "black") +
  geom_col_pattern(
    aes(pattern = treatment),
    colour = "black",
    pattern_fill = "black",
    pattern_angle = 45,
    pattern_density = 0.1,
    pattern_spacing = 0.01,
    position = position_dodge2(preserve = 'single'),
    alpha = 0.2  # Adjust the alpha value to control transparency
  ) +
  scale_pattern_manual(
    values = c("none", "stripe"),
    guide = guide_legend(override.aes = list(fill = "grey70")) # <- make lighter
  ) +
  xlab("") +
  ylab("Home-field advantage (% faster decomposition at home)") +
coord_cartesian(ylim = c(-800,300)) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
  )

hfai.z.plot

ggsave("figures/hfai.z.plot.pdf", hfai.z.plot, width = 10, height = 6)

```



This is the proper dataframe with for loop
generate **absent**, site a (HOK 1)
```{r}
########################
mass.hok1.absent.df <- mass.hok1.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values_z <- numeric()
rml_values_mass <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_z_loss <- mass.hok1.absent.df$z_score[
      mass.hok1.absent.df$tree_id == "HOK" & mass.hok1.absent.df$tree_num == tree_num
    ]
    sek_z_loss <- mass.hok1.absent.df$z_score[
      mass.hok1.absent.df$tree_id == "SEK" & mass.hok1.absent.df$tree_num == tree_num
    ]
    
    rml.z <- ifelse(
      ayres == "Aa", hok_z_loss / (hok_z_loss + sek_z_loss),
      ifelse(ayres == "Ba", sek_z_loss / (sek_z_loss + hok_z_loss), NA)
    )
    
    rml_values_z <- c(rml_values_z, rml.z * 100)
    
    hok_mass_loss <- mass.hok1.absent.df$perc.mass_loss[
      mass.hok1.absent.df$tree_id == "HOK" & mass.hok1.absent.df$tree_num == tree_num
    ]
    sek_mass_loss <- mass.hok1.absent.df$perc.mass_loss[
      mass.hok1.absent.df$tree_id == "SEK" & mass.hok1.absent.df$tree_num == tree_num
    ]
    
    rml.mass <- ifelse(
      ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
      ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss), NA)
    )
    
    rml_values_mass <- c(rml_values_mass, rml.mass * 100)
  }
}

# Add the rml values to the dataframe
mass.hok1.absent.df$rml.z <- rml_values_z
mass.hok1.absent.df$rml.mass <- rml_values_mass

# Create a new column combining tree_num and ayres with a period in between
mass.hok1.absent.df$ayres.tree <- paste(mass.hok1.absent.df$ayres, mass.hok1.absent.df$tree_num, sep = ".")

# Create a new column based on Ayres HFAI equation 
mass.hok1.absent.df$ayres.eq <- ifelse(
  mass.hok1.absent.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""),
  paste("Armla", tree_num_values, sep = "."),
  ifelse(
    mass.hok1.absent.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""),
    paste("Brmla", tree_num_values, sep = "."),
    NA
  )
)

############### STOP ######################
##check the calculations (Armla.1-Armla.5)
((mass.hok1.present.df[mass.hok1.present.df$ayres.tree == "Aa.1",]$z_score) /
  (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == "Aa.1",]$z_score +
   mass.hok1.present.df[mass.hok1.present.df$ayres.tree == "Ba.1",]$z_score)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$z_score)) * 100


(
  (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == "Aa.3",]$perc.mass_loss) /
  (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == "Aa.3",]$perc.mass_loss +
   mass.hok1.present.df[mass.hok1.present.df$ayres.tree == "Ba.3",]$perc.mass_loss)
  
  * 100) 

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$z_score)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$z_score)) * 100


##check the calculations (Brmla.1-Brmla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$z_score)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$z_score)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$z_score)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$z_score)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$z_score)) * 100


```


This is also the proper dataframe
generate **absent**, site b (SEK 1)
```{r}
mass.sek1.absent.df <- mass.sek1.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values_z <- numeric()
rml_values_mass <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_z_loss <- mass.sek1.absent.df$z_score[
      mass.sek1.absent.df$tree_id == "HOK" & mass.sek1.absent.df$tree_num == tree_num
    ]
    sek_z_loss <- mass.sek1.present.df$z_score[
      mass.sek1.absent.df$tree_id == "SEK" & mass.sek1.absent.df$tree_num == tree_num
    ]
    
    rml.z <- ifelse(
      ayres == "Ab", hok_z_loss / (hok_z_loss + sek_z_loss),
      ifelse(ayres == "Bb", sek_z_loss / (sek_z_loss + hok_z_loss), NA)
    )
    
    rml_values_z <- c(rml_values_z, rml.z * 100)
    
    hok_mass_loss <- mass.sek1.absent.df$perc.mass_loss[
      mass.sek1.absent.df$tree_id == "HOK" & mass.sek1.absent.df$tree_num == tree_num
    ]
    sek_mass_loss <- mass.sek1.absent.df$perc.mass_loss[
      mass.sek1.absent.df$tree_id == "SEK" & mass.sek1.absent.df$tree_num == tree_num
    ]
    
    rml.mass <- ifelse(
      ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
      ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss), NA)
    )
    
    rml_values_mass <- c(rml_values_mass, rml.mass * 100)
  }
}

# Add the rml values to the dataframe
mass.sek1.absent.df$rml.z <- rml_values_z
mass.sek1.absent.df$rml.mass <- rml_values_mass

# Create a new column combining tree_num and ayres with a period in between
mass.sek1.absent.df$ayres.tree <- paste(mass.sek1.absent.df$ayres, mass.sek1.absent.df$tree_num, sep = ".")

# Create a new column based on Ayres HFAI equation 
mass.sek1.absent.df$ayres.eq <- ifelse(
  mass.sek1.absent.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""),
  paste("Armlb", tree_num_values, sep = "."),
  ifelse(
    mass.sek1.absent.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""),
    paste("Brmlb", tree_num_values, sep = "."),
    NA
  )
)


```



This is also the proper dataframe
generate **absent**, site a (HOK 2)
```{r}
mass.hok2.absent.df <- mass.hok2.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values_z <- numeric()
rml_values_mass <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_z_loss <- mass.hok2.absent.df$z_score[
      mass.hok2.absent.df$tree_id == "HOK" & mass.hok2.absent.df$tree_num == tree_num
    ]
    sek_z_loss <- mass.hok2.absent.df$z_score[
      mass.hok2.absent.df$tree_id == "SEK" & mass.hok2.absent.df$tree_num == tree_num
    ]
    
    rml.z <- ifelse(
      ayres == "Aa", hok_z_loss / (hok_z_loss + sek_z_loss),
      ifelse(ayres == "Ba", sek_z_loss / (sek_z_loss + hok_z_loss), NA)
    )
    
    rml_values_z <- c(rml_values_z, rml.z * 100)
    
    hok_mass_loss <- mass.hok2.absent.df$perc.mass_loss[
      mass.hok2.absent.df$tree_id == "HOK" & mass.hok2.absent.df$tree_num == tree_num
    ]
    sek_mass_loss <- mass.hok2.absent.df$perc.mass_loss[
      mass.hok2.absent.df$tree_id == "SEK" & mass.hok2.absent.df$tree_num == tree_num
    ]
    
    rml.mass <- ifelse(
      ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
      ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss), NA)
    )
    
    rml_values_mass <- c(rml_values_mass, rml.mass * 100)
  }
}

# Add the rml values to the dataframe
mass.hok2.absent.df$rml.z <- rml_values_z
mass.hok2.absent.df$rml.mass <- rml_values_mass

# Create a new column combining tree_num and ayres with a period in between
mass.hok2.absent.df$ayres.tree <- paste(mass.hok2.absent.df$ayres, mass.hok2.absent.df$tree_num, sep = ".")

# Create a new column based on Ayres HFAI equation 
mass.hok2.absent.df$ayres.eq <- ifelse(
  mass.hok2.absent.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""),
  paste("Armla", tree_num_values, sep = "."),
  ifelse(
    mass.hok2.absent.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""),
    paste("Brmla", tree_num_values, sep = "."),
    NA
  )
)


```


This is also the proper dataframe
generate **absent**, site a (SEK 2)
```{r}
mass.sek2.absent.df <- mass.sek2.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values_z <- numeric()
rml_values_mass <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_z_loss <- mass.sek2.absent.df$z_score[
      mass.sek2.absent.df$tree_id == "HOK" & mass.sek2.absent.df$tree_num == tree_num
    ]
    sek_z_loss <- mass.sek2.absent.df$z_score[
      mass.sek2.absent.df$tree_id == "SEK" & mass.sek2.absent.df$tree_num == tree_num
    ]
    
    rml.z <- ifelse(
      ayres == "Ab", hok_z_loss / (hok_z_loss + sek_z_loss),
      ifelse(ayres == "Bb", sek_z_loss / (sek_z_loss + hok_z_loss), NA)
    )
    
    rml_values_z <- c(rml_values_z, rml.z * 100)
    
    hok_mass_loss <- mass.sek2.absent.df$perc.mass_loss[
      mass.sek2.absent.df$tree_id == "HOK" & mass.sek2.absent.df$tree_num == tree_num
    ]
    sek_mass_loss <- mass.sek2.absent.df$perc.mass_loss[
      mass.sek2.absent.df$tree_id == "SEK" & mass.sek2.absent.df$tree_num == tree_num
    ]
    
    rml.mass <- ifelse(
      ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
      ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss), NA)
    )
    
    rml_values_mass <- c(rml_values_mass, rml.mass * 100)
  }
}

# Add the rml values to the dataframe
mass.sek2.absent.df$rml.z <- rml_values_z
mass.sek2.absent.df$rml.mass <- rml_values_mass

# Create a new column combining tree_num and ayres with a period in between
mass.sek2.absent.df$ayres.tree <- paste(mass.sek2.absent.df$ayres, mass.sek2.absent.df$tree_num, sep = ".")

# Create a new column based on Ayres HFAI equation 
mass.sek2.absent.df$ayres.eq <- ifelse(
  mass.sek2.absent.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""),
  paste("Armlb", tree_num_values, sep = "."),
  ifelse(
    mass.sek2.absent.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""),
    paste("Brmlb", tree_num_values, sep = "."),
    NA
  )
)


```

This is the proper dataframe
merge **absent**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.absent.site1.df=rbind(mass.hok1.absent.df, mass.sek1.absent.df)

# Calculate the index of the last label in the existing hfai.df
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Create an empty list to store the new data frames
hfai_dfs <- list()

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla.z <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Armla.", i), ]$rml.z
  brmlb.z <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml.z
  armlb.z <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml.z
  brmla.z <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml.z
  
  armla.mass <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Armla.", i), ]$rml.mass
  brmlb.mass <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml.mass
  armlb.mass <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml.mass
  brmla.mass <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml.mass
  
  result.z <- (((armla.z + brmlb.z) / (armlb.z + brmla.z)) * 100) - 100
  result.mass <- (((armla.mass + brmlb.mass) / (armlb.mass + brmla.mass)) * 100) - 100

  # Get the treatment associated with the current index
  treatment <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Create a data frame with the new results, label, and treatment
  hfai_df <- data.frame(hfai.z = result.z, hfai.mass = result.mass, label = next_label, treatment = treatment)
  
  # Append the data frame to the list
  hfai_dfs[[i]] <- hfai_df
}

# Combine the list of new data frames with the existing hfai.df
hfai.df <- rbind(hfai.df, do.call(rbind, hfai_dfs))




```


This is the proper dataframe
merge **absent**, site a and b (aka HOK 2 and SEK 2) dataframes 
```{r}
#merge both present site 1 dataframes
mass.absent.site2.df=rbind(mass.hok2.absent.df, mass.sek2.absent.df)

# Calculate the index of the last label in the existing hfai.df
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Create an empty list to store the new data frames
hfai_dfs <- list()

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla.z <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Armla.", i), ]$rml.z
  brmlb.z <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml.z
  armlb.z <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml.z
  brmla.z <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml.z
  
  armla.mass <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Armla.", i), ]$rml.mass
  brmlb.mass <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml.mass
  armlb.mass <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml.mass
  brmla.mass <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml.mass
  
  result.z <- (((armla.z + brmlb.z) / (armlb.z + brmla.z)) * 100) - 100
  result.mass <- (((armla.mass + brmlb.mass) / (armlb.mass + brmla.mass)) * 100) - 100

  # Get the treatment associated with the current index
  treatment <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Create a data frame with the new results, label, and treatment
  hfai_df <- data.frame(hfai.z = result.z, hfai.mass = result.mass, label = next_label, treatment = treatment)
  
  # Append the data frame to the list
  hfai_dfs[[i]] <- hfai_df
}

# Combine the list of new data frames with the existing hfai.df
hfai.df <- rbind(hfai.df, do.call(rbind, hfai_dfs))




```



This is the proper plot
another plot check
```{r}
######## using mass loss

hfai.mass.all.plot=ggplot(hfai.df, aes(x = reorder(label, hfai.mass), y = hfai.mass, fill = treatment)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  scale_fill_viridis_d()+
  xlab("reciprocated pairs") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.x = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  )

hfai.mass.all.plot


ggsave("figures/hfai.mass.all.plot.pdf", hfai.mass.all.plot, width = 14, height = 10)
```










generate **absent**, site a (HOK 1)
```{r}
mass.ab.hoko1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="HOK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.hoko1.df <- mass.ab.hoko1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "HOK" & mass.ab.hoko1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "SEK" & mass.ab.hoko1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}

##check the calculations (Armla.1-Armla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss)) * 100


##check the calculations (Brmla.1-Brmla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss)) * 100

# Add the rml values to the dataframe
mass.ab.hoko1.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.hoko1.df$ayres.tree <- paste(mass.ab.hoko1.df$ayres, mass.ab.hoko1.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.hoko1.df$ayres.eq <- ifelse(mass.ab.hoko1.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **absent**, site b (SEK 1)
```{r}
mass.ab.sek1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="SEK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.sek1.df <- mass.ab.sek1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)


###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "HOK" & mass.ab.sek1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "SEK" & mass.ab.sek1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.sek1.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.sek1.df$ayres.tree <- paste(mass.ab.sek1.df$ayres, mass.ab.sek1.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.sek1.df$ayres.eq <- ifelse(mass.ab.sek1.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.p.sek1.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **asbent**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.ab.site1.df=rbind(mass.ab.hoko1.df, mass.ab.sek1.df)

# Calculate the last letter index based on the existing data frame
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = next_label, treatment = treatment))
}


#HFAI (this is a stop and check to verify that values are correct!)
(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```

plot check
```{r}

ggplot(hfai.df, aes(x = label, y = hfai)) +
  geom_bar(stat = "identity",
           show.legend = FALSE) + # Remove the legend
  facet_grid(~treatment)+
  xlab("Group") +
  ylab("Value")+
  coord_flip()




```



generate **present**, site a (HOK 2)
```{r}
#now create a new column to have some reference to the Ayres et al. (2009) paper (species A= hoko trees, species B = sekiu trees)
mass.hok2.present.df <- mass.hok2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 2-1 with SEK 2-1; this follows similar logic as Ayres et al. (2009) Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.hok2.present.df$z_score[mass.hok2.present.df$tree_id == "HOK" & mass.hok2.present.df$tree_num == tree_num]
    sek_mass_loss <- mass.hok2.present.df$z_score[mass.hok2.present.df$tree_id == "SEK" & mass.hok2.present.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.hok2.present.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.hok2.present.df$ayres.tree <- paste(mass.hok2.present.df$ayres, mass.hok2.present.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.hok2.present.df$ayres.eq <- ifelse(mass.hok2.present.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.hok2.present.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **present**, site b (SEK 2)
```{r}
#now create a new column to have some reference to the Ayres et al. (2009) paper (species A= hoko trees, species B = sekiu trees)
mass.sek2.present.df <- mass.sek2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)


###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 2-1 with SEK 2-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.sek2.present.df$z_score[mass.sek2.present.df$tree_id == "HOK" & mass.sek2.present.df$tree_num == tree_num]
    sek_mass_loss <- mass.sek2.present.df$z_score[mass.sek2.present.df$tree_id == "SEK" & mass.p.sek2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.sek2.present.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.sek2.present.df$ayres.tree <- paste(mass.sek2.present.df$ayres, mass.sek2.present.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.sek2.present.df$ayres.eq <- ifelse(mass.sek2.present.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.sek2.present.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **present**, site a and b (aka HOK 2 and SEK 2) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site2.df=rbind(mass.hok2.present.df, mass.sek2.present.df)

# Calculate the last letter index based on the existing data frame
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = next_label, treatment = treatment))
}

#HFAI (this is a stop and check to verify that values are correct!)
(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```


## Now we'll tackle absent:
generate **absent**, site a (HOK 2)
```{r}
mass.ab.hoko2.df=mass.df%>%
  filter(treatment == "AB" & deploy=="HOK" & site_num=="2")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.hoko2.df <- mass.ab.hoko2.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.hoko2.df$mass_loss[mass.ab.hoko2.df$tree_id == "HOK" & mass.ab.hoko2.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.hoko2.df$mass_loss[mass.ab.hoko2.df$tree_id == "SEK" & mass.ab.hoko2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.hoko2.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.hoko2.df$ayres.tree <- paste(mass.ab.hoko2.df$ayres, mass.ab.hoko2.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.hoko2.df$ayres.eq <- ifelse(mass.ab.hoko2.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **absent**, site b (SEK 1)
```{r}
mass.ab.sek2.df=mass.df%>%
  filter(treatment == "AB" & deploy=="SEK" & site_num=="2")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.sek2.df <- mass.ab.sek2.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.sek2.df$mass_loss[mass.ab.sek2.df$tree_id == "HOK" & mass.ab.sek2.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.sek2.df$mass_loss[mass.ab.sek2.df$tree_id == "SEK" & mass.ab.sek2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.sek2.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.sek2.df$ayres.tree <- paste(mass.ab.sek2.df$ayres, mass.ab.sek2.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.sek2.df$ayres.eq <- ifelse(mass.ab.sek2.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.p.sek1.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **asbent**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.ab.site2.df=rbind(mass.ab.hoko2.df, mass.ab.sek2.df)

# Calculate the last letter index based on the existing data frame
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = next_label, treatment = treatment))
}

#HFAI (this is a stop and check to verify that values are correct!)
(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```




















```{r}
########### Next we'll try out absent, hoko 1
mass.ab.hoko1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="HOK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (Home= Aa, Away= Bb)
mass.ab.hoko1.df <- mass.ab.hoko1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "HOK" & mass.ab.hoko1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "SEK" & mass.ab.hoko1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.hoko1.df$rml <- rml_values







########### Next we'll try out absent, sekiu 1
mass.ab.sek1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="SEK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (Home= Aa, Away= Bb)
mass.ab.sek1.df <- mass.ab.sek1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "HOK" & mass.ab.sek1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "SEK" & mass.ab.sek1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.sek1.df$rml <- rml_values


#merge dataframes
mass.site1.df=rbind(mass.p.hoko1.df, mass.ab.hoko1.df, mass.p.sek1.df, mass.ab.sek1.df )



```




Ayres HFA index
```{r}
#the calculations for arml and brml are different, dependent on "relative to"

##Site 1: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site a (Hoko) 
site1.armla.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="1" & mass.df$deploy == "HOK", ])
#mass loss calculation
site1.armla.df= site1.armla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.armla.df <- site1.armla.df %>%
  mutate(rml = ((mass_loss[tree_id == "HOK" & deploy == "HOK"]) / 
                 (mass_loss[tree_id == "HOK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "HOK"])) * 100)

##Site 2: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site a (Hoko) 
site2.armla.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="2" & mass.df$deploy == "HOK", ])
#mass loss calculation
site2.armla.df= site2.armla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.armla.df <- site2.armla.df %>%
  mutate(rml = (mass_loss[tree_id == "HOK" & deploy == "HOK"] / 
                 (mass_loss[tree_id == "HOK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "HOK"])) * 100)

##Site 1: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site b (Sekiu) 
site1.armlb.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="2" & mass.df$deploy == "SEK", ])
#mass loss calculation
site1.armlb.df= site1.armlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.armlb.df <- site1.armlb.df %>%
  mutate(rml = (mass_loss[tree_id == "HOK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "HOK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "SEK"])) * 100)

##Site 2: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site b (Sekiu) 
site2.armlb.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="2" & mass.df$deploy == "SEK", ])
#mass loss calculation
site2.armlb.df= site2.armlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.armlb.df <- site2.armlb.df %>%
  mutate(rml = (mass_loss[tree_id == "HOK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "HOK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "SEK"])) * 100)

##Site 1: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site a (Hoko) 
site1.brmla.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="1" & mass.df$deploy == "HOK", ])
#mass loss calculation
site1.brmla.df= site1.brmla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.brmla.df <- site1.brmla.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "HOK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "HOK"])) * 100)

##Site 2: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site a (Hoko) 
site2.brmla.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="2" & mass.df$deploy == "HOK", ])
#mass loss calculation
site2.brmla.df= site2.brmla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.brmla.df <- site2.brmla.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "HOK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "HOK"])) * 100)

##Site 1: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site b (Sekiu) 
site1.brmlb.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="1" & mass.df$deploy == "SEK", ])
#mass loss calculation
site1.brmlb.df= site1.brmlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.brmlb.df <- site1.brmlb.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "SEK"])) * 100)

##Site 2: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site b (Sekiu) 
site2.brmlb.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="2" & mass.df$deploy == "SEK", ])
#mass loss calculation
site2.brmlb.df= site2.brmlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.brmlb.df <- site2.brmlb.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "SEK"])) * 100)

#merge dataframes
site1.hfa_index.df=rbind(site1.armla.df, site1.armlb.df, site1.brmla.df, site1.brmlb.df)
#HFA index calculation
site1.hfa_index.df <- site1.hfa_index.df %>%
  mutate(hfai = (((site1.armla.df$rml + site1.brmlb.df$rml) / 2) / ((site1.armlb.df$rml + site1.brmla.df$rml) / 2)) * 100 - 100)

#merge dataframes
site2.hfa_index.df=rbind(site2.armla.df, site2.armlb.df, site2.brmla.df, site2.brmlb.df)
#HFA index calculation
site2.hfa_index.df <- site2.hfa_index.df %>%
  mutate(hfai = (((site2.armla.df$rml + site2.brmlb.df$rml) / 2) / ((site2.armlb.df$rml + site2.brmla.df$rml) / 2)) * 100 - 100)

# Create the horizontal bar plot
ggplot(hfa_index.df, aes(x = hfai, y = interaction(deploy, hfa))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(title = "HFA Index Comparison",
       x = "HFA Index") +
  theme_minimal() +
  theme(legend.position = "none") +
  geom_vline(xintercept = 0, color = "black", size = 1) +
  scale_fill_manual(values = c("a" = "blue", "b" = "red")) 



```


Alder secondary metabolites from LC-MS
### Import LC-MS/MS data
```{r}

#installing and calling the necessary packages:
library("mixOmics")
library(ggpubr)
library('devtools')
library(vegan)

### load negative mode
neg_data <- read_csv("/Users/Cody1/Desktop/Projects/RAD Alder/LC-MS/RAD_negative_mode_11_3_2023_quant_copy.csv") %>% arrange(Sample)
neg_metadata <- read_csv("/Users/Cody1/Desktop/Projects/RAD Alder/LC-MS/RAD_prelim_metadata_neg.csv") %>% arrange(Sample)
all(neg_data$Sample == neg_metadata$Sample)

# Generate negative mode PCA raw
neg_PCA_whole <- mixOmics::pca(column_to_rownames(neg_data, var = "Sample"), ncomp = 2, scale = TRUE)
neg_PCA_whole_scores<- data.frame(PCA_whole_neg$variates$X, neg_metadata)

#### negative PCA plot for treatment (can also change for tree ID)
neg_PCA_whole_scores%>%
  ggscatter(x = "PC1", y = "PC2", color = "Attribute_type", alpha = 0.6, 
            title = paste("PCA RAW -", "Attribute_type", sep = " "),
            xlab = paste("PC1 (", round(neg_PCA_whole$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(neg_PCA_whole$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = neg_PCA_whole_scores %>% group_by(Attribute_type) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = Attribute_type), size = 4, shape = 8) +
  stat_ellipse(aes(PC1, PC2, color = Attribute_type), geom = "polygon", level = 0.95, alpha = 0.2) +  # Add ellipses
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()

# PERMANOVA
neg_dist_metabolites <- vegdist(column_to_rownames(neg_data, var = "Sample"), method = "euclidean")
neg_permanova <- adonis2(neg_dist_metabolites ~ neg_PCA_whole_scores$Attribute_type, neg_PCA_whole_scores, na.action = na.omit)


###################################

### load positive mode
pos_data <- read_csv("/Users/Cody1/Desktop/Projects/RAD Alder/LC-MS/RAD_positive_mode_11_3_2023_quant_copy.csv") %>% arrange(Sample)
pos_metadata <- read_csv("/Users/Cody1/Desktop/Projects/RAD Alder/LC-MS/RAD_prelim_metadata_pos.csv") %>% arrange(Sample)
all(pos_data$Sample == pos_metadata$Sample)

# Generate positive mode PCA raw
pos_PCA_whole <- mixOmics::pca(column_to_rownames(pos_data, var = "Sample"), ncomp = 2, scale = TRUE)
pos_PCA_whole_scores <- data.frame(pos_PCA_whole$variates$X, pos_metadata)

#### positive PCA plot for treatment (can also change for tree ID)
pos_PCA_whole_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = "Attribute_type", alpha = 0.6, 
            title = paste("PCA RAW -", "Attribute_type", sep = " "),
            xlab = paste("PC1 (", round(pos_PCA_whole$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(pos_PCA_whole$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = pos_PCA_whole_scores %>% group_by(Attribute_type) %>%
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = Attribute_type), size = 4, shape = 8)+
  stat_ellipse(aes(PC1, PC2, color = as.factor(Attribute_type)), geom = "polygon", level = 0.95, alpha = 0.2) +  # Add ellipses
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()

pos_PCA_whole_scores$Attribute_Sitenum=as.factor(pos_PCA_whole_scores$Attribute_Sitenum)

pos_PCA_whole_scores %>%
  ggscatter(
    x = "PC1", y = "PC2",
    color = "Attribute_TreeID",
    shape = "Attribute_Sitenum",
    alpha = 0.6,
    title = paste("PCA RAW -", "Attribute_Sitenum", sep = " "),
    xlab = paste("PC1 (", round(pos_PCA_whole$prop_expl_var$X[1] * 100, digits = 1), "%)", sep = ""),
    ylab = paste("PC2 (", round(pos_PCA_whole$prop_expl_var$X[2] * 100, digits = 1), "%)", sep = ""),
    ggtheme = theme_classic()
  ) +geom_point(
    data = pos_PCA_whole_scores %>% 
      group_by(Attribute_TreeID, Attribute_Sitenum) %>% 
      summarise_at(vars(matches("PC")), mean),
    aes(PC1, PC2, color = Attribute_TreeID, shape = Attribute_Sitenum),
    alpha = 0.6
  )+
  stat_ellipse(aes(PC1, PC2, color = Attribute_TreeID, shape = Attribute_Sitenum), geom = "polygon", level = 0.95, alpha = 0.1) +  # Add ellipses
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()


```

Negative mode stats
```{r}

# Generate PCA relative abundance
neg_data_ra <- (column_to_rownames(neg_data, var = "Sample"))/rowSums(column_to_rownames(neg_data, var = "Sample"))
neg_PCA_whole_ra <- mixOmics::pca(neg_data_ra, ncomp = 2, scale = TRUE)
neg_PCA_whole_ra_scores <- data.frame(neg_PCA_whole_ra$variates$X, neg_metadata)

#plot
neg_PCA_whole_ra_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = "Attribute_TreeID", alpha = 0.6, 
            title = paste("PCA RA -", "Attribute_TreeID", sep = " "),
            xlab = paste("PC1 (", round(neg_PCA_whole_ra$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(neg_PCA_whole_ra$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = neg_PCA_whole_ra_scores %>% group_by(Attribute_TreeID) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = Attribute_TreeID), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) +
  stat_ellipse(aes(PC1, PC2, color = Attribute_TreeID), geom = "polygon", level = 0.95, alpha = 0.2) +  # Add ellipses
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()

neg_PCA_whole_ra_scores$Attribute_Sitenum=as.factor(neg_PCA_whole_ra_scores$Attribute_Sitenum)


# plot with site

neg_PCA_whole_ra_scores %>%
  ggscatter(
    x = "PC1", y = "PC2",
    color = "Attribute_TreeID",
    shape = "Attribute_Sitenum",
    alpha = 0.6,
    title = paste("PCA RAW -", "Attribute_Sitenum", sep = " "),
    xlab = paste("PC1 (", round(neg_PCA_whole_ra$prop_expl_var$X[1] * 100, digits = 1), "%)", sep = ""),
    ylab = paste("PC2 (", round(neg_PCA_whole_ra$prop_expl_var$X[2] * 100, digits = 1), "%)", sep = ""),
    ggtheme = theme_classic()
  ) +geom_point(
    data = neg_PCA_whole_ra_scores %>% 
      group_by(Attribute_TreeID, Attribute_Sitenum) %>% 
      summarise_at(vars(matches("PC")), mean),
    aes(PC1, PC2, color = Attribute_TreeID, shape = Attribute_Sitenum),
    alpha = 0.6
  )+
  stat_ellipse(aes(PC1, PC2, color = Attribute_TreeID, shape = Attribute_Sitenum), geom = "polygon", level = 0.95, alpha = 0.1) +  # Add ellipses
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()
library(ggforce)

neg_PCA_whole_ra_scores %>%
  ggscatter(
    x = "PC1", y = "PC2",
    color = "Attribute_TreeID",
    shape = "Attribute_Sitenum",
    alpha = 0.6,
    title = paste("PCA RAW -", "Attribute_Sitenum", sep = " "),
    xlab = paste("PC1 (", round(neg_PCA_whole_ra$prop_expl_var$X[1] * 100, digits = 1), "%)", sep = ""),
    ylab = paste("PC2 (", round(neg_PCA_whole_ra$prop_expl_var$X[2] * 100, digits = 1), "%)", sep = ""),
    ggtheme = theme_classic()
  ) +
  geom_point(
    data = neg_PCA_whole_ra_scores %>% 
      group_by(Attribute_TreeID, Attribute_Sitenum) %>% 
      summarise_at(vars(matches("PC")), mean),
    aes(PC1, PC2, color = Attribute_TreeID, shape = Attribute_Sitenum),
    alpha = 0.6
  ) +
  geom_encircle(
    data = neg_PCA_whole_ra_scores %>% 
      group_by(Attribute_TreeID, Attribute_Sitenum) %>% 
      summarise_at(vars(matches("PC")), mean),
    aes(PC1, PC2, color = Attribute_TreeID),
    expand = 0.02, # Adjust the expansion factor as needed
    linetype = "dashed",
    alpha = 0.7
  ) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()


# PERMANOVA
neg_dist_metabolites_ra <- vegdist(neg_data_ra, method = "euclidean")
neg_permanova_ra <- adonis2(neg_dist_metabolites_ra ~ neg_PCA_whole_ra_scores$Attribute_type, neg_PCA_whole_ra_scores, na.action = na.omit)

############################

#by tree ID
adonis2(neg_dist_metabolites_ra ~ neg_PCA_whole_ra_scores$Attribute_TreeID, neg_PCA_whole_ra_scores, na.action = na.omit)

### Create PLSDA for AB vs No AB in HF diet mice
# Pairwise PLS_DA
data_clr_CT_MPS <- data_clr %>% dplyr::filter(metadata$Attribute_type != "H2O2")
metadata_clr_CT_MPS <- metadata %>% dplyr::filter(metadata$Attribute_type != "H2O2")

# Create PLSDA for AB vs No AB in HF diet mice
PLSDA_CT_MPS <- mixOmics::plsda(data_clr_CT_MPS, metadata_clr_CT_MPS$Attribute_type, ncomp = 3, scale = TRUE)
PLSDA_CT_MPS_scores <- data.frame(PLSDA_CT_MPS$variates$X, metadata_clr_CT_MPS)

numeric_matrix <- data.matrix(neg_data)
neg_data_numeric <- sapply(neg_data, function(x) as.numeric(as.character(x)))

neg_PLSDA<- mixOmics::plsda(neg_data_numeric, neg_metadata$Attribute_type, ncomp = 3, scale = TRUE)
neg_PLSDA_scores <- data.frame(neg_PLSDA$variates$X, neg_metadata)

neg_PLSDA_scores %>%
  ggscatter(x = "comp1", y = "comp2", color = "Attribute_type", alpha = 0.6, title = "PLS-DA AB vs P",
            xlab = paste("Component 1 (", round(neg_PLSDA$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("Component 2 (", round(neg_PLSDA$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            legend.title = "Group", ggtheme = theme_classic()) +
  geom_point(data = neg_PLSDA_scores %>% group_by(Attribute_type) %>% 
               summarise_at(vars(matches("comp")), mean), aes(comp1, comp2, color = Attribute_type), size = 3, shape = 8) +
  theme(plot.title = element_text(size = 10),axis.title = element_text(size = 8),
        axis.text = element_text(size = 5)) + coord_fixed()



perf_neg_plsda <- mixOmics::perf(neg_PLSDA, validation = "Mfold", progressBar = FALSE, auc = TRUE, folds = 4, nrepeat = 100) 

plot(perf_plsda_CT_MPS, legend = FALSE)

neg_VIPs <- as.data.frame(mixOmics::vip(neg_PLSDA))
neg_VIPs_filter <- dplyr::filter(neg_VIPs, neg_VIPs$comp1 > 1)
neg_VIPs_filter$ID <- rownames(neg_VIPs_filter)
neg_VIPs_select <- neg_VIPs_filter %>% dplyr::select(ID, comp1)
neg_VIPs_Load <- neg_VIPs_select %>% left_join(neg_loadings, by = c("ID" = "rowname")) %>% arrange(desc(comp1))
neg_VIPs_Load_top <- VIPs_CT_MPS_Load %>% head(., 50)

Loadings_CT_MPS_plot <- plotLoadings(neg_loadings, plot = FALSE, contrib = "max") %>%
  arrange(GroupContrib) %>%
  rownames_to_column(var = "Feature") %>%
  dplyr::filter(Feature %in% VIPs_CT_MPS_Load_top$ID) %>%
  ggdotchart(x = "Feature", y = "importance",
             color = "GroupContrib", sorting = "ascending",  
             dot.size = 2, add = "segments",                  
             add.params = list(color = "lightgray", size = 0.1),
             title = "Loadings PLS-DA - CT v MPS") + geom_hline(yintercept = 0, linetype = 2, color = "lightgray")





```

Positive mode stats
```{r}
# PERMANOVA
pos_dist_metabolites <- vegdist(column_to_rownames(pos_data, var = "Sample"), method = "euclidean")
pos_permanova <- adonis2(pos_dist_metabolites ~ pos_PCA_whole_scores$Attribute_type, pos_PCA_whole_scores, na.action = na.omit)

# Generate PCA relative abundance
pos_data_ra <- (column_to_rownames(pos_data, var = "Sample"))/rowSums(column_to_rownames(pos_data, var = "Sample"))
pos_PCA_whole_ra <- mixOmics::pca(pos_data_ra, ncomp = 2, scale = TRUE)
pos_PCA_whole_ra_scores <- data.frame(pos_PCA_whole_ra$variates$X, pos_metadata)

PCA_whole_ra_scores_pos %>%
  ggscatter(x = "PC1", y = "PC2", color = "Attribute_TreeID", alpha = 0.6, 
            title = paste("PCA RA -", "Attribute_TreeID", sep = " "),
            xlab = paste("PC1 (", round(PCA_whole_ra_pos$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_whole_ra_pos$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_whole_ra_scores_pos %>% group_by(Attribute_TreeID) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = Attribute_TreeID), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed() +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) +
  stat_ellipse(aes(PC1, PC2, color = Attribute_TreeID), geom = "polygon", level = 0.95, alpha = 0.2)   # Add ellipses

# PERMANOVA
pos_dist_metabolites_ra <- vegdist(pos_data_ra, method = "euclidean")
pos_permanova_ra <- adonis2(pos_dist_metabolites_ra ~ pos_PCA_whole_ra_scores$Attribute_type, pos_PCA_whole_ra_scores, na.action = na.omit)

#by tree ID
adonis2(pos_dist_metabolites_ra ~ PCA_whole_ra_scores_pos$Attribute_TreeID, PCA_whole_ra_scores_pos, na.action = na.omit)


```

