---
title: "RAD Alder 2023"
author: "Cody Spiegel"
date: "`r Sys.Date()`"
output: word_document
---

### Project overview
For Project ‘RAD’ Alder, gamma irradiated vs. non-irradiated leaf packs deployed at “home” or “away” sites tests the role of phyllosphere microbial communities on home-field advantage patterns in leaf litter decomposition. We utilized a reciprocal transplant experiment between the Hoko and Sekiu rivers on the Olympic Peninsula under four treatment conditions: phyllosphere present-home (PH), phyllosphere present-away (PAW), phyllosphere absent-home (ABH), phyllosphere absent-away (ABAW) to ask the following questions: 

1) How does dispersal from the regional species pool (i.e., source communities) influence community composition on the leaf packs?
2) How does microbial community composition (alpha and beta diversity) and function differ between the leaf packs, as well as change over the course of decomposition?
3) How does alder stoichiometric and secondary chemical composition differ between the leaf packs, as well as change over the course of decomposition?
4) How does microbial metabolization differ between the leaf packs, as well as change over the course of decomposition?
5) How does particulate leaf mass loss differ between the leaf packs? 


**DATA SETS**  
This data set focuses on:  

(1) **metabolism assay** 
(2) **Leaf mass loss** using % mass loss  
(3) **Alder C and N isotopic composotion**
(4) **Alder secondary chemistry (mass spec)** 

Load primary packages and figure formatting
```{r setup chunk, setup, include = FALSE, cache=FALSE, message=FALSE, warning=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

# use pacman to load all the packages you are missing!
pacman::p_load('knitr', 'lme4', 'lmerTest', 'tidyverse', 'magrittr', 'effects', 'plyr', 'dplyr', 
               'plotrix', 'car',"gridExtra", "cowplot", "tools", "mgcv", "gratia", "tidymv")

Fig.formatting<-(theme_classic()) +
  theme(text=element_text(size=25),
        legend.text.align = 0,
        panel.border = element_rect(fill=NA, colour = "black", size=1),
        aspect.ratio=1, 
        axis.ticks.length=unit(0.25, "cm"),
        axis.text.y=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20), 
        axis.text.x=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), angle = 45, colour="black", size=20)) +
  theme(legend.key.size = unit(0.4, "cm"), legend.position = "right") +
  theme(aspect.ratio=1.3) 




```


All metabolite decomp days
```{r}
meta.day1.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_1_metabolites_raw_corrected.csv")


meta.day3.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_3_metabolites_raw_corrected.csv")


meta.day5.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_5_metabolites_raw_corrected.csv")


meta.day7.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_7_metabolites_raw_corrected.csv")


meta.day11.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_11_metabolites_raw_corrected.csv")


meta.day17.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_17_metabolites_raw_corrected.csv")


#merge
meta.df= rbind(meta.day1.raw.df, meta.day3.raw.df, meta.day5.raw.df, meta.day7.raw.df, meta.day11.raw.df, meta.day17.raw.df)

#z score by deployment
# Assuming your data frame is named 'df'
meta.df=meta.df %>%
  filter(time == "960")%>%
  group_by(deploy, decomp_day) %>%
  mutate(z_score = (od_corrected_b_n - mean(od_corrected_b_n)) / sd(od_corrected_b_n))

meta.df$treatment=as.factor(meta.df$treatment)
meta.df$tree_id= as.factor(meta.df$tree_id)
meta.df$deploy = as.factor(meta.df$deploy)
meta.df$metabolite= as.factor(meta.df$metabolite)
meta.df$hfa=as.factor(meta.df$hfa)
meta.df$decomp_day=as.factor(meta.df$decomp_day)
meta.df$time = as.numeric(meta.df$time)



#By z-scoring the raw OD values, I can effectively assess the relative changes in microbial metabolism across the wells, regardless of their absolute starting values.

#z-scoring is a reasonable choice for this situation, as it standardizes the data within each well and facilitates comparisons. Note: z-scoring assumes a roughly symmetric distribution, but the raw OD values are right skewed.

```


metabolite assay: z score by metabolite and deployment 
```{r}
#z score
##hok 1
meta.hok1.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metabolite=="HOK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])
meta.hok1.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metaboltie=="SEK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])


meta.hok1.hok.df$z_score= (meta.hok1.hok.df$od_corrected_b_n - mean(meta.hok1.hok.df$od_corrected_b_n)) / sd(meta.hok1.hok.df$od_corrected_b_n)
meta.hok1.sek.df$z_score= (meta.hok1.sek.df$od_corrected_b_n - mean(meta.hok1.sek.df$od_corrected_b_n)) / sd(meta.hok1.sek.df$od_corrected_b_n)

##sek 1
meta.sek1.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metabolite=="SEK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])
meta.sek1.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metaboltie=="HOK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])

meta.sek1.sek.df$z_score= (meta.sek1.sek.df$od_corrected_b_n - mean(meta.sek1.sek.df$od_corrected_b_n)) / sd(meta.sek1.sek.df$od_corrected_b_n)
meta.sek1.hok.df$z_score= (meta.sek1.hok.df$od_corrected_b_n - mean(meta.sek1.hok.df$od_corrected_b_n)) / sd(meta.sek1.hok.df$od_corrected_b_n)

#hok 2 
meta.hok2.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metabolite=="HOK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])
meta.hok2.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metaboltie=="SEK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])

meta.hok2.hok.df$z_score= (meta.hok2.hok.df$od_corrected_b_n - mean(meta.hok2.hok.df$od_corrected_b_n)) / sd(meta.hok2.hok.df$od_corrected_b_n)
meta.hok2.sek.df$z_score= (meta.hok2.sek.df$od_corrected_b_n - mean(meta.hok2.sek.df$od_corrected_b_n)) / sd(meta.hok2.sek.df$od_corrected_b_n)


##sek 2
meta.sek2.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metabolite=="SEK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])
meta.sek2.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metaboltie=="HOK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])

meta.sek2.sek.df$z_score= (meta.sek2.sek.df$od_corrected_b_n - mean(meta.sek2.sek.df$od_corrected_b_n)) / sd(meta.sek2.sek.df$od_corrected_b_n)
meta.sek2.hok.df$z_score= (meta.sek2.hok.df$od_corrected_b_n - mean(meta.sek2.hok.df$od_corrected_b_n)) / sd(meta.sek2.hok.df$od_corrected_b_n)

# Assuming you have datasets named meta.day1.raw.df, meta.day3.raw.df, ..., meta.day17.raw.df
# Put them into a list
metabolite_list <- list(meta.day1.raw.df, meta.day3.raw.df, meta.day5.raw.df, meta.day7.raw.df, meta.day11.raw.df, meta.day17.raw.df)

# Define the conditions for subsetting
meta.time.points <- c("30", "120", "960", "1440", "2880", "4320")
# Define the actual day numbers corresponding to each dataset
actual_day_numbers <- c(1, 3, 5, 7, 11, 17)

# Create an empty list to store subsets
subset_list <- list()

# Loop through datasets
for (i in seq_along(metabolite_list)) {
  # Access each dataset from the metabolite_list
  meta.merge <- metabolite_list[[i]]
  
  # Create a list to store subsets for each day
  day_subsets <- list()
  
  # Loop through time points
  for (time_point in meta.time.points) {
    # Subset the data based on specific conditions
    subset_data <- meta.merge[meta.merge$deploy == "HOK" &
                             meta.merge$metabolite == "HOK" &
                             meta.merge$site_num == "1" & 
                             meta.merge$time == time_point, ]
    
    # Calculate the z-scores within each subset_data
    subset_data$z_score <- (subset_data$od_corrected_b_n - mean(subset_data$od_corrected_b_n)) / sd(subset_data$od_corrected_b_n)
    
    # Store the subset with z-scores in the day_subsets list
    day_subsets[[time_point]] <- subset_data
  }
  
  # Merge all time points for each day into one dataframe
  merged_data <- bind_rows(day_subsets)
  
  # Store the merged dataframe in the subset_list
  day_number <- actual_day_numbers[i]
  subset_list[[paste0("day", day_number)]] <- merged_data
}

# Rename the list elements to indicate the day
day_names <- paste0("day", actual_day_numbers)
names(subset_list) <- day_names

#rename into new dataframes
meta.day1.df=subset_list$day1
meta.day3.df=subset_list$day3
meta.day5.df=subset_list$day5
meta.day7.df=subset_list$day7
meta.day11.df=subset_list$day11
meta.day17.df=subset_list$day17

#merge into one big dataframe
meta.df= rbind(meta.day1.df, meta.day3.df, meta.day5.df, meta.day7.df, meta.day11.df, meta.day17.df)

#create new HFA code (if deployed at home or away, fed a home or away metabolite)
meta.df$group <- 
  
ifelse(
  meta.df$tree_id == "HOK" & meta.df$deploy == "HOK" & meta.df$metabolite == "HOK", "home-home",
  
  ifelse(
    meta.df$tree_id == "HOK" & meta.df$deploy == "HOK" & meta.df$metabolite == "SEK", "home-away",
    
    ifelse(
      meta.df$tree_id == "SEK" & meta.df$deploy == "SEK" & meta.df$metabolite == "SEK", "home-home",
      
      ifelse(
      meta.df$tree_id == "SEK" & meta.df$deploy == "SEK" & meta.df$metabolite == "HOK", "home-away",
      
      ifelse(
      meta.df$tree_id == "HOK" & meta.df$deploy == "SEK" & meta.df$metabolite == "HOK", "away-home-away",
      
      ifelse(
      meta.df$tree_id == "HOK" & meta.df$deploy == "SEK" & meta.df$metabolite == "SEK", "away-home-home",
      
      ifelse(
      meta.df$tree_id == "SEK" & meta.df$deploy == "HOK" & meta.df$metabolite == "SEK", "away-home-away",
      
     ifelse(
      meta.df$tree_id == "SEK" & meta.df$deploy == "HOK" & meta.df$metabolite == "HOK", "away-home-home",
        NA  # Default value if none of the conditions are met
      )
    )
  )
)
)
)
)
)

# Group by the factors of interest
z_scored_data <- meta.df %>%
  group_by(deploy, metabolite, decomp_day) %>%
  mutate(
    mean_OD = mean(od_corrected_b_n),           # Calculate mean for the group
    sd_OD = sd(od_corrected_b_n),               # Calculate standard deviation for the group
    z_score = (od_corrected_b_n - mean_OD) / sd_OD  # Calculate z-score
  )
z_scored_data


```


Day 1: plots
```{r}
#color blind palette
library(viridis) 

meta.df %>% 
  ggplot( aes(x = decomp_day, linetype= treatment, color = group, y = z_score)) +
  geom_line(aes(group = interaction(hfa, treatment, site_num, metabolite, deploy, tree_num,tree_id))) + 
  geom_point(aes(group = tree_num)) +   
  theme_test()+
  facet_wrap(deploy~group)


```


```{r}
day1.z.30.plot=ggplot(meta.df[meta.df$time=="30",], aes(x = de, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
   scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day1.z.120.plot=ggplot(meta.day1.df[meta.day1.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.960.plot=ggplot(meta.day1.df[meta.day1.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.1440.plot=ggplot(meta.day1.df[meta.day1.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.2880.plot=ggplot(meta.day1.df[meta.day1.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day1.z.4320.plot=ggplot(meta.day1.df[meta.day1.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Create the grid with modified plots
extract.legend <- get_legend(
  day1.z.4320.plot + theme(legend.box.margin = margin(1, 1, 1, 1, "cm"),
                       legend.key.size = unit(50, "pt"),
                       legend.text = element_text(size=20),
                       legend.title= element_blank())
)


# Combine plots
all.z.day1.plot <- plot_grid(
  day1.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


all.z.day1.plot
# Set the overall size of the combined plot
ggsave("figures/all.z.day1.plot.pdf", all.z.day1.plot, width = 49, height = 10)



```


Day 3: plots
```{r}

day3.z.30.plot=ggplot(meta.day3.df[meta.day3.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
  scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day3.z.120.plot=ggplot(meta.day3.df[meta.day3.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
     scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.960.plot=ggplot(meta.day3.df[meta.day3.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
     scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.1440.plot=ggplot(meta.day3.df[meta.day3.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.2880.plot=ggplot(meta.day3.df[meta.day3.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day3.z.4320.plot=ggplot(meta.day3.df[meta.day3.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


# Combine plots
all.z.day3.plot <- plot_grid(
  day3.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day3.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day3.plot.pdf", all.z.day3.plot, width = 49, height = 10)



```



Day 5: plots
```{r}

day5.z.30.plot=ggplot(meta.day5.df[meta.day5.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day5.z.120.plot=ggplot(meta.day5.df[meta.day5.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


day5.z.960.plot=ggplot(meta.day5.df[meta.day5.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



day5.z.1440.plot=ggplot(meta.day5.df[meta.day5.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
          scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



day5.z.2880.plot=ggplot(meta.day5.df[meta.day5.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day5.z.4320.plot=ggplot(meta.day5.df[meta.day5.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Combine plots
all.z.day5.plot <- plot_grid(
  day5.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day5.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day5.plot.pdf", all.z.day5.plot, width = 49, height = 10)



```



Day 7: plots
```{r}

day7.z.30.plot=ggplot(meta.day7.df[meta.day7.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
   scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day7.z.30.plot

day7.z.120.plot=ggplot(meta.day7.df[meta.day7.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
     scale_fill_viridis(discrete = TRUE) + 
 geom_hline(data = meta.day7.df[meta.day7.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.120.plot

day7.z.960.plot=ggplot(meta.day7.df[meta.day7.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.960.plot


day7.z.1440.plot=ggplot(meta.day7.df[meta.day7.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.1440.plot

day7.z.2880.plot=ggplot(meta.day7.df[meta.day7.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.2880.plot

day7.z.4320.plot=ggplot(meta.day7.df[meta.day7.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


# Combine plots
all.z.day7.plot <- plot_grid(
  day7.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day7.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day7.plot.pdf", all.z.day7.plot, width = 49, height = 12)



```



Day 11: z scored
```{r}
day11.z.30.plot=ggplot(meta.day11.df[meta.day11.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day11.z.30.plot

day11.z.120.plot=ggplot(meta.day11.df[meta.day11.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day11.z.120.plot

day11.z.960.plot=ggplot(meta.day11.df[meta.day11.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.960.plot


day11.z.1440.plot=ggplot(meta.day11.df[meta.day11.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.1440.plot

day11.z.2880.plot=ggplot(meta.day11.df[meta.day11.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.2880.plot

day11.z.4320.plot=ggplot(meta.day11.df[meta.day11.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


# Combine plots
all.z.day11.plot <- plot_grid(
  day11.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day11.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)

# Set the overall size of the combined plot
ggsave("figures/all.z.day11.plot.pdf", all.z.day11.plot, width = 49, height = 12)



```





Day 17: z scored
```{r}

day17.z.30.plot=ggplot(meta.day17.df[meta.day17.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day17.z.30.plot

day17.z.120.plot=ggplot(meta.day17.df[meta.day17.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day17.z.120.plot

day17.z.960.plot=ggplot(meta.day17.df[meta.day17.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day17.z.960.plot


day17.z.1440.plot=ggplot(meta.day17.df[meta.day17.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


day17.z.1440.plot

day17.z.2880.plot=ggplot(meta.day17.df[meta.day17.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day17.z.2880.plot

day17.z.4320.plot=ggplot(meta.day17.df[meta.day17.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


# Combine plots
all.z.day17.plot <- plot_grid(
  day17.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day17.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
   day17.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14,14,14, 14, 3)
)

# Set the overall size of the combined plot
ggsave("figures/all.z.day17.plot.pdf", all.z.day17.plot, width = 49, height = 12)


```

All metabolite assays plots
```{r}

all.z.days.plot <- plot_grid(
  all.z.day1.plot,
  all.z.day3.plot,
  all.z.day5.plot,
  all.z.day7.plot,
  all.z.day11.plot,
  all.z.day17.plot,
  nrow = 6
)



all.z.days.plot
ggsave("figures/all.z.days.plot.pdf", all.z.days.plot, width = 49, height = 30)

```


Metabolite analysis
```{r}

#make a big metabolite dataframe

merged_df=rbind(meta.day1.raw.df,meta.day3.raw.df,meta.day5.raw.df,meta.day7.raw.df, meta.day11.raw.df)
class(merged_df$time)
merged_df$treatment=as.factor(as.character(merged_df$treatment))
merged_df$tree_id=as.factor(as.character(merged_df$tree_id))
merged_df$tree_num=as.factor(as.integer(merged_df$tree_num))
merged_df$deploy=as.factor(as.character(merged_df$deploy))
merged_df$site_num=as.factor(as.integer(merged_df$site_num))
merged_df$metabolite=as.factor(as.character(merged_df$metabolite))
merged_df$hfa=as.factor(as.character(merged_df$hfa))
merged_df$time=as.numeric(as.integer(merged_df$time))

class(merged_df$metabolite)

# Fit the linear mixed-effects model with nested random effects and fixed effect
model <- lmer(z_score ~ treatment + hfa + treatment * hfa +
              (1 | deploy/tree_id/metabolite),
              data = merged_df[merged_df$decomp_day == "7" & merged_df$time=="4320", ],
              na.action = na.exclude)
anova(model)






```
16 hr metabolite plots
```{r}

day1.16hr.plot=ggplot(meta.day1.df[meta.day1.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 1")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting +
   theme(legend.position = "none")

day1.16hr.plot

day3.16hr.plot=ggplot(meta.day3.df[meta.day3.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 3")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting +
   theme(legend.position = "none")


day5.16hr.plot=ggplot(meta.day5.df[meta.day5.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "Treatment", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 5")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting+
   theme(legend.position = "none")

day5.16hr.plot

day7.16hr.plot=ggplot(meta.day7.df[meta.day7.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "Treatment", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 5")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting+
   theme(legend.position = "none")

day11.16hr.plot=ggplot(meta.day11.df[meta.day11.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 11")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting+
  theme(legend.position = "none")

day11.16hr.plot

all.16hr.z.plots=plot_grid(
  day1.16hr.plot,
  day3.16hr.plot,
  day5.16hr.plot,
  day11.16hr.plot,
  ncol=5,
  extract.legend,
  rel_widths = c(8,8,8,8,3))


all.day16.z.plots
ggsave("figures/all.16hr.z.plots.pdf", all.16hr.z.plots, width = 26, height = 8)


```
Background information:

The Home Field Advantage Index (HFAI) derived from Ayres et al. (2009) uses the relative mass loss values (rml) of species A and B at different sites. The HFAI therefore represents the difference in the rate of litter decomposition between two different sites ("Hoko" and "Sekiu") for both species A and B. Species A and B, in this case, are alder trees of different origin. For example, species A is Hoko alder and species B is Sekiu alder. There are 10 alder trees per river, representing biological replicates. So, the HFAI quantifies the *advantage* or *disadvantage* of each "species" in their respective home environment compared to an away environment.

Interpreting HFAI values involves considering their sign and magnitude:

**Positive** HFAI Values:
Positive values indicate that both species A and B experience faster decomposition rates at their home site compared to the away site. This suggests that both species are better adapted to their home environment in terms of litter decomposition.

**Negative** HFAI Values:
Negative values indicate that both species A and B experience slower decomposition rates at their home site compared to the away site. This suggests that both species might not be well-suited for decomposition in their own environment and may decompose more efficiently in a different environment.

**Magnitude** of HFAI:
The magnitude of the HFAI values indicates the strength of the advantage or disadvantage. Larger absolute values (positive or negative) suggest a greater difference in decomposition rates between the home and away sites.


Load mass data 
```{r}
library(stringr)
library(purrr)
mass.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/leaf_mass.csv")
###check for any white spaces clean up
columns_with_whitespace <- map_lgl(mass.df, ~ any(str_detect(.x, "^\\s+|\\s+$")))
columns_with_whitespace

#correction factor adjusting for blotted and fresh dry mass
mass.df= mass.df%>%
  mutate(final_correct = (0.941*(final_mass)-0.00337))

# calculate % mass loss 
mass.df <- mass.df %>%
  mutate(perc.mass_loss = ((initial_mass - final_correct) / initial_mass)*100)

# Calculate the constant C, adding 1 for safety
C <- abs(min(mass.df$perc.mass_loss)) + 1  

#create new column with perc.mass_loss + constant
mass.df <- mass.df %>%
  mutate(perc.mass_loss = perc.mass_loss + C) %>%
  mutate(prop.mass_loss = perc.mass_loss/100) %>%
mutate(arcsine_trans = asin(sqrt(prop.mass_loss)))


#subset P treatment only for "true" HFA  
mass.present.df=mass.df %>%
  filter(treatment == "P")

#subset AB treatment only for HFA  
mass.absent.df=mass.df %>%
  filter(treatment == "AB")


```

**Subset dataframes to z score by deployment site (all HOK 1, all HOK 2, etc...)**

Z scoring controls for site variability, and allows for a better comparison of treatments. But how exactly?

Z-scoring is a statistical measure that quantifies how far away a particular data point is from the mean of a dataset in terms of standard deviations (i.e., the spread of the data points from the mean). It's a way to standardize data and make it more comparable across different scales or distributions. Here's how z-scoring works:

1) calculate the mean and standard deviation: for the variable you want to z-score (in this case mass loss by deployment site), calculate its mean (average) and standard deviation. These values provide information about the central tendency (i.e., the center/midpoint of the data or a value around which the data clusters) and the spread of the data.

2) subtract the mean: for each data point, subtract the mean of the variable. This step shifts the data so that the new mean becomes 0.

3) divide by the standard deviation: after subtracting the mean, divide the result by the standard deviation. This step scales the data based on the spread of the original data, so that the new data points are expressed in terms of standard deviations from the mean.

Mathematically, the z-score (Z) for a data point (X) with mean (μ) and standard deviation (σ) is calculated using the formula:

Z = (X - μ) / σ

The resulting z-score indicates how many standard deviations the original data point is away from the mean. If the z-score is positive, it means the data point is above the mean, and if it's negative, it's below the mean.

Note: z-scoring assumes a roughly normal distribution of the data.

Below, the percent mass loss for each deployment site is z scored. Additionally, the larger dataframe has been subset to only include P treatments so that the "true" HFA can be calculated.
```{r}
#check for distribution of mass loss
hist(mass.df$perc.mass_loss)
  #roughly normal...that's good!

#####Begin z scoring
#HOK 1 subset
mass.hok1.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.hok1.df=mass.hok1.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.hok1.df=mass.hok1.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))


#HOK 1 P subset
mass.hok1.present.df = data.frame(mass.present.df[mass.present.df$deploy == "HOK" & mass.present.df$site_num == "1", ])
# Calculate z-scores 
mass.hok1.present.df=mass.hok1.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.hok1.present.df=mass.hok1.present.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))



#HOK 2 subset
mass.hok2.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.hok2.df=mass.hok2.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.hok2.df=mass.hok2.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))


#HOK 2 P subset
mass.hok2.present.df = data.frame(mass.present.df[mass.present.df$deploy == "HOK" & mass.present.df$site_num == "2", ])
# Calculate z-scores 
mass.hok2.present.df=mass.hok2.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.hok2.present.df=mass.hok2.present.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))



#SEK 1 subset
mass.sek1.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.sek1.df=mass.sek1.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.sek1.df=mass.sek1.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))


#SEK 1 P subset
mass.sek1.present.df = data.frame(mass.present.df[mass.present.df$deploy == "SEK" & mass.present.df$site_num == "1", ])
# Calculate z-scores 
mass.sek1.present.df=mass.sek1.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.sek1.present.df=mass.sek1.present.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))


#SEK 2 subset
mass.sek2.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.sek2.df=mass.sek2.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.sek2.df=mass.sek2.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))


#SEK 2 P subset
mass.sek2.present.df = data.frame(mass.present.df[mass.present.df$deploy == "SEK" & mass.present.df$site_num == "2", ])
# Calculate z-scores 
mass.sek2.present.df=mass.sek2.present.df%>%
  mutate(z_score= (perc.mass_loss - mean(perc.mass_loss))/sd(perc.mass_loss))
#z score arcsine_trans 
mass.sek2.present.df=mass.sek2.present.df%>%
  mutate(arcsine_z_score= (arcsine_trans - mean(arcsine_trans))/sd(arcsine_trans))



##merge all sites
mass.deploy.merged.df=rbind(mass.hok1.df, mass.hok2.df, mass.sek1.df, mass.sek2.df)
mass.deploy.merged.df$treatment=as.factor(as.character(mass.deploy.merged.df$treatment))
mass.deploy.merged.df$tree_id=as.factor(as.character(mass.deploy.merged.df$tree_id))
mass.deploy.merged.df$tree_id=as.factor(as.character(mass.deploy.merged.df$tree_id))
mass.deploy.merged.df$deploy=as.factor(as.character(mass.deploy.merged.df$deploy))
mass.deploy.merged.df$hfa=as.factor(as.character(mass.deploy.merged.df$hfa))

#merge all P treatment sites 
mass.present.deploy.merged.df=rbind(mass.hok1.present.df, mass.hok2.present.df, mass.sek1.present.df, mass.sek2.present.df)
mass.present.deploy.merged.df$treatment=as.factor(as.character(mass.present.deploy.merged.df$treatment))
mass.present.deploy.merged.df$tree_id=as.factor(as.character(mass.present.deploy.merged.df$tree_id))
mass.present.deploy.merged.df$tree_id=as.factor(as.character(mass.present.deploy.merged.df$tree_id))
mass.present.deploy.merged.df$deploy=as.factor(as.character(mass.present.deploy.merged.df$deploy))
mass.present.deploy.merged.df$hfa=as.factor(as.character(mass.present.deploy.merged.df$hfa))

```

Let's first look at the P only to see if HFA is there
```{r}
###P treatment only to see HFA effects
#create a proper "site" column for the figure

#plot
ggplot(mass.present.deploy.merged.df, aes(x = hfa, y = z_score)) +
  geom_boxplot(alpha= 0.3) +
  geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.2)) +
  labs(x = "", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_x_discrete(labels = c("away", "home")) +
  scale_shape_manual(values = c(7,10), name= "site") +
  geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

```


Analysis for **present only** dataframe
To be analyze the mass loss data, arcsine square root transforming proportional mass loss is preferred. This are outlined below...

perform lmer and add result to P treatment only plot
Lmer 3 (see below) is the model that best fits the data
```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
#relevel HFA factors
levels(mass.present.deploy.merged.df$hfa)
mass.present.deploy.merged.df$hfa=factor(mass.present.deploy.merged.df$hfa, levels = c("H", "A"))
levels(mass.present.deploy.merged.df$hfa)
names(mass.deploy.merged.df)

#models:
lm1 <- lm(z_score ~  hfa, data = mass.present.deploy.merged.df)
lm2 <- lm(z_score ~ factor(hfa, levels = c("A", "H")), data = mass.present.deploy.merged.df)
lmer1 <- lmer(z_score ~ hfa + (1 | tree_id), data= mass.present.deploy.merged.df)
lmer2 <- lmer(z_score ~ tree_id * deploy + (1 | tree_num), data= mass.present.deploy.merged.df)

lmer3 <- lmer(z_score ~ hfa + (1 | tree_num) + (1 | tree_id), data= mass.present.deploy.merged.df)

lmer4 <- lmer(z_score ~ hfa + (1 | tree_id /  tree_num), data= mass.present.deploy.merged.df)
summary(lmer5)

# lmer3 is best fit
lmer3 <- lmer(z_score ~ hfa + (1 | tree_num) + (1 | tree_id), data= mass.present.deploy.merged.df)

lmer10 <- lmer(arcsine_trans_scaled ~ hfa + (1 | deploy) + (1 | tree_num) + (1 | tree_id), data= mass.present.deploy.merged.df)

summary(lmer10)
plot(residuals(lmer5))
plot(fitted(lmer5), residuals(lmer5))
shapiro.test(residuals(lmer5))

qqnorm(residuals(lmer5))
qqline(residuals(lmer5))

AIC(lmer3, lmer10)


```


Generic t-test for P only treatment
```{r}

# the t test will use z score values
#combined site
mass.present.deploy.merged.df$combined_site <- paste("site", mass.present.deploy.merged.df$site_num, sep = "")

stat.test.z <- mass.present.deploy.merged.df %>% 
  t_test(z_score ~ hfa, var.equal = T) %>%
  add_significance()

stat.test.arc <- mass.present.deploy.merged.df %>% 
  t_test(arcsine_trans~ hfa, var.equal = T) %>%
  add_significance()

stat.test.z

#check levels
levels(mass.present.deploy.merged.df$hfa)
mass.present.deploy.merged.df$hfa=factor(mass.present.deploy.merged.df$hfa, levels = c("H", "A"))
levels(mass.present.deploy.merged.df$hfa)


# Create the box plot with original elements as ggplot from above
leaf.z.presentonly.plot= ggboxplot(
  mass.present.deploy.merged.df, x = "hfa", y = "z_score", 
  ylab = "mass loss (z score)", xlab = "", fill="hfa", alpha= 0.4) + 
  scale_fill_manual(values = c("grey", "grey0"), guide = "none")+
  stat_pvalue_manual(stat.test.arc, y.position = 2.5, tip.length = c(0.16, 0.06)) +
  labs(subtitle = get_test_label(stat.test.arc, detailed = T)) +
    scale_x_discrete(labels = c("home", "away"))+
    geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.2))+
    scale_shape_manual(values = c(7,10), name= "site")+
    geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16),
          plot.background = element_blank()) 

leaf.z.presentonly.plot

# Create the box plot with original elements as ggplot from above
leaf.mass.presentonly.plot= ggboxplot(
  mass.present.deploy.merged.df, x = "hfa", y = "perc.mass_loss", 
  ylab = "mass loss (%)", xlab = "", fill="hfa", alpha= 0.4) + 
  scale_fill_manual(values = c("grey", "grey0"), guide = "none")+
  stat_pvalue_manual(stat.test.arc, y.position = 60, tip.length = c(0.06, 0.16)) +
  labs(subtitle = get_test_label(stat.test.arc, detailed = T)) +
    scale_x_discrete(labels = c("home", "away"))+
    geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.1))+
    scale_shape_manual(values = c(7,10), name= "site")+
    geom_hline(data = mass.present.deploy.merged.df, aes(yintercept = mean(perc.mass_loss), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16),
          plot.background = element_blank()) 

leaf.mass.presentonly.plot

ggsave("figures/leaf.z.presentonly.plot.pdf", leaf.z.presentonly.plot, width = 8, height = 6)


```
Analysis for **all treatment** dataframe
There are two ways to go about model exploration: using z scored mass loss or arcsine square root transforming percent mass loss. These are outlined below...

first, plot to visualize the 4 treatments
```{r}
# Combine treatment and hfa into one group
mass.deploy.merged.df$combined_group <- paste(mass.deploy.merged.df$treatment, mass.deploy.merged.df$hfa, sep = "-")

# Create a custom order for the combined groups
combined_group_order <- c("P-H", "P-A","AB-H", "AB-A")
custom_palette <- c("#2CA02C",  "#D62728", "#1F77B4","#FF7F0E")


# Plot with combined groups but no combo site
ggplot(mass.deploy.merged.df, aes(x = combined_group, y = z_score, fill = combined_group)) +
  geom_boxplot(alpha= 0.3) +
  scale_x_discrete(limits = combined_group_order)+
  geom_point(aes(shape = as.factor(deploy)), position = position_jitter(width = 0.2), alpha = 0.7) +
  labs(x = "treatment", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_fill_manual(values = custom_palette, guide = "none") +
  scale_shape_manual(values = c(15, 16), name ="site") +
  geom_hline(data = mass.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting



```

model exploration of all treatments: ph, pa, abh, aba; lmer 7 is best fit
```{r}
lmer4 <- lmer(arcsine_z_score ~ treatment*hfa + (1|tree_num) + (1|tree_id), data=mass.deploy.merged.df)
lmer5 <- lmer(arcsine_z_score ~ combined_group + (1|tree_num) + (1|tree_id), data=mass.deploy.merged.df)
lmer6 <- lmer(z_score ~ combined_group + (1|tree_num) + (1|tree_id), data=mass.deploy.merged.df)
lmer7 <- lmer(z_score ~ treatment*hfa + (1|tree_num) + (1|tree_id), data=mass.deploy.merged.df)

AIC(lmer4, lmer7)
summary(lmer7)
# Type III ANOVA for fixed effects
Anova(lmer7, type="III")

library(emmeans)
# Post-hoc tests for interactions (if significant)
interaction_emm <- emmeans(lmer7, specs = ~treatment:hfa)
contrast(interaction_emm, method = "pairwise")

# Fit your linear mixed-effects model
lmer7 <- lmer(z_score ~ treatment * hfa + (1 | tree_num) + (1 | tree_id), data = mass.deploy.merged.df)


# Calculate Cohen's d for specific group comparisons
cohen_d <- with(mass.deploy.merged.df, {
  d1 <- (mean(z_score[treatment == "P" & hfa == "H"]) - mean(z_score[treatment == "P" & hfa == "A"])) / sqrt((var(z_score[treatment == "P" & hfa == "H"]) + var(z_score[treatment == "P" & hfa == "A"])) / 2)
  d2 <- (mean(z_score[treatment == "AB" & hfa == "H"]) - mean(z_score[treatment == "AB" & hfa == "A"])) / sqrt((var(z_score[treatment == "AB" & hfa == "H"]) + var(z_score[treatment == "AB" & hfa == "A"])) / 2)
  d3 <- (mean(z_score[treatment == "P" & hfa == "H"]) - mean(z_score[treatment == "AB" & hfa == "H"])) / sqrt((var(z_score[treatment == "P" & hfa == "H"]) + var(z_score[treatment == "AB" & hfa == "H"])) / 2)
  d4 <- (mean(z_score[treatment == "P" & hfa == "A"]) - mean(z_score[treatment == "AB" & hfa == "A"])) / sqrt((var(z_score[treatment == "P" & hfa == "A"]) + var(z_score[treatment == "AB" & hfa == "A"])) / 2)
  data.frame(Group = c("P-H vs. P-A", "AB-H vs. AB-A", "P-H vs. AB-H", "P-A vs. AB-A"),
             Cohen_d = c(d1, d2, d3, d4))
})

# Print the Cohen's d values
print(cohen_d)

install.packages("MuMIn")
library(MuMIn)
r.squaredGLMM(lmer7)
```


This is the proper all treatment dataframe for nonparametric test
```{r}
install.packages("dunn.test")
library(dunn.test)
kw1 <- kruskal.test(z_score ~ combined_group, data = mass.deploy.merged.df)
kw1
# Perform post-hoc Dunn's test
posthoc <- dunn_test(z_score ~ combined_group, data = mass.deploy.merged.df, method = "bonferroni")# Create a function to format p-values and add stars for significance
format_p_value <- function(p_value) {
  if (p_value < 0.001) {
    return("***")
  } else if (p_value < 0.01) {
    return("**")
  } else if (p_value < 0.05) {
    return("*")
  } else {
    return("ns")
  }
}

# Apply the formatting function to the p-values
dunn.test.z.res$formatted_p_value <- sapply(dunn.test.z.res$p.adj, format_p_value)
##this plot works
# Create a custom order for the combined groups
combined_group_order <- c("P-H", "P-A","AB-H", "AB-A")
library(ggpubr)

custom_palette <- c("#D62728","#FF7F0E", "#2CA02C", "#1F77B4")
levels(as.factor(mass.deploy.merged.df$combined_group))

leaf.mass.all.plot=ggboxplot(
  mass.deploy.merged.df, x = "combined_group", y = "z_score", 
  ylab = "mass loss (z score)", xlab = "", fill="combined_group", alpha= 0.4) + 
    scale_x_discrete(limits = combined_group_order)+
  scale_fill_viridis_d( guide = "none")+
  stat_pvalue_manual(dunn.test.z.res, step.increase = 0.03, step.group.by = "group2", tip.length = 0.01) +
    labs( x= "treatment", y="mass loss (z score)") +
  geom_point(aes(shape = deploy), position = position_jitter(width = 0.2), alpha = 0.7) +
  scale_shape_manual(values = c(15, 16), name ="site") +
     geom_hline(data = mass.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16),
          plot.background = element_blank(),
          axis.text.x = element_text(angle = 22, hjust = 0.5, vjust = 0.5)) 

leaf.mass.all.plot

count(mass.deploy.merged.df$combined_group)

range(mass.deploy.merged.df$perc.mass_loss)

ggsave("figures/leaf.mass.all.plot.pdf", leaf.mass.all.plot, width = 8, height = 6)

```


Analysis for **absent only** dataframe
There are two ways to go about model exploration: using z scored mass loss or arcsine square root transforming percent mass loss. These are outlined below...

perform t-test and add result to AB treatment only plot
```{r}
library(tidyverse)
library(ggpubr)
library(rstatix)
#plot
mass.absent.deploy.merged.df$combined_site <- paste("site", mass.absent.deploy.merged.df$site_num, sep = "")

leaf.mass.absentonly.plot=ggplot(mass.absent.deploy.merged.df, aes(x = hfa, y = z_score)) +
  geom_boxplot(alpha= 0.3) +
  geom_point(aes(shape = as.factor(combined_site)), position = position_jitter(width = 0.2)) +
  labs(x = "", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_x_discrete(labels = c("away", "home")) +
  scale_fill_manual(values = custom_palette) +
  scale_shape_manual(values = c(7,10), name= "site") +
  geom_hline(data = mass.absent.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

leaf.mass.absentonly.plot
# Create a histogram
hist_data <- hist(mass.absent.deploy.merged.df$z_score, plot = FALSE)

# Create a density plot
density_data <- density(mass.absent.deploy.merged.df$z_score)

# Plot histogram with density overlay
hist(mass.absent.deploy.merged.df$z_score, freq = F, main = "")
lines(density_data, col = "red")
  #normal

shapiro.test(mass.absent.deploy.merged.df$z_score)
  # p= 0.39, normal

# Q-Q Plots
qqnorm(mass.absent.deploy.merged.df$z_score, main = "Q-Q Plot")
qqline(mass.absent.deploy.merged.df$z_score)
  #normal


# Levene's Test
leveneTest(z_score ~ hfa, data = mass.absent.deploy.merged.df)
  # p= 0.63, variances between H and A are equal (i.e., the homogeneity assumption of the variance is met)


# the t test will use z score values
stat.ab.test.z <- mass.absent.deploy.merged.df %>% 
  t_test(z_score ~ hfa, var.equal = T) %>%
  add_significance()
stat.ab.test.z

# Create the box plot with original elements as ggplot from above
ggboxplot(
  mass.absent.deploy.merged.df, x = "hfa", y = "z_score", 
  ylab = "mass loss (z score)", xlab = "") + 
  stat_pvalue_manual(stat.ab.test.z, tip.length = 0) +
  labs(subtitle = get_test_label(stat.ab.test.z, detailed = TRUE)) +
    scale_x_discrete(labels = c("away", "home"))+
    geom_point(aes(shape = as.factor(combined_site)), position = position_jitter(width = 0.2))+
    scale_shape_manual(values = c(7,10), name= "site")+
    geom_hline(data = mass.absent.deploy.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
    theme(legend.position = "right",
          legend.text = element_text(size=16),
          legend.title = element_text(size = 16),
          text = element_text(size =16)) 

leaf.mass.presentonly.plot

ggsave("figures/leaf.mass.presentonly.plot.pdf", leaf.mass.presentonly.plot, width = 10, height = 8)


```


**Exploration of nonparametric models**
Non-parametric tests are more robust to deviations from normality and heteroscedasticity, but they do have their own assumptions. For example, the Mann-Whitney U test assumes that the distributions of the two groups are similar in shape. The Kruskal-Wallis test assumes that the distributions of all groups are similar and that the differences in the distributions are due to location shifts. 


The Kruskal-Wallis test can be an alternative. This test is an extension of the Wilcoxon rank-sum test for comparing more than two independent groups.

```{r}
# Perform Kruskal-Wallis test
kw_test <- kruskal.test(z_score ~ hfa, data = mass.present.deploy.merged.df)

# Calculate effect size (Eta-squared) from Kruskal-Wallis test
H <- kw_test$statistic
df <- kw_test$parameter
N <- sum(table(mass.present.deploy.merged.df$hfa))
eta_squared <- (H - df) / (N - df)

# Print effect size
cat("Effect size (Eta-squared) for Kruskal-Wallis test:", eta_squared, "\n")

#The calculated effect size (Eta-squared) for the Kruskal-Wallis test is -0.01813634. Eta-squared is a measure of the proportion of variance explained by a factor in non-parametric tests, similar to how R-squared works in linear regression. The negative value suggests a very small effect, but it's important to note that Eta-squared can range from 0 to 1, with higher values indicating larger effects. In this case, the small negative value indicates that the factor (hfa) is not explaining much variance in the data.


```


----------------- START OF HFAI CALCULATIONS ----------------------------------

Calculating HFA Index (HFAI) from Ayres et al. (2009):

There are two equations from which HFAI is dervied: the relative mass loss (rml) of species decomposing at home and away site, and the scaling equation which allows for interpretaibility of the home-field advantage. Below is a brief summary of the components found in the derivaitons:

Aa: This represents the percent mass loss of leaf litter from species A (HOK) when it's decomposing at site a (HOK, or the area dominated by species A). This value is given as a percentage.

Ba: This represents the percent mass loss of leaf litter from species B (SEK) when it's decomposing at site a This value is also given as a percentage.

(Aa + Ba): This is the sum of the percent mass loss of leaf litter from both species A and B when they're decomposing at site a, or HOK and SEK leaf packs at HOK.

(Aa / (Aa + Ba) * 100): This is the full equation that calculates the proportion of the mass loss contributed by species A at site a, relative to the **total mass** loss by both species A and B combined at that site. This percentage value indicates how much of the total mass loss at site a is due to species A.

A similar equation is calculated for species B decomposing at site a. The variables are substituted accordingly.

Armla and Armlb; Brlma, Brmlb: These are the relative mass loss values for species A and B, respectively, when they are decomposing at site a (dominated by species A) or site b (dominated by species B). These values are expressed as percentages of total mass loss at their respective sites.

(((ARMLa + BRMLb)^2 / (ARMLb + BRMLa)^2) * 100)-100: This is the full equation that calculates the squared ratio of the combined relative mass loss of species A at site a and species B at site b (i.e., home), divided by the squared ratio of the combined relative mass loss of species A at site b and species B at site a (i.e., away).

100 is subtracted from the value obtained. This is done to measure the percentage difference in mass loss at the home site (site dominated by the species) compared to the away site (site dominated by the other species). If the value is positive, it indicates that the litter decomposes faster at its home site, showing a home-field advantage. If the value is negative, it suggests that decomposition is slower at the home site, indicating a home-field disadvantage.

The equation calculates the Home-Field Advantage Index (HFAI) by comparing the relative mass loss of litter from two species at their home and away sites. The result is expressed as a percentage, indicating how much faster (or slower) the decomposition occurs at the home site compared to the away site.


NEW: use this! This is a now a proper test dataframe with for loop!

NOTICE: THIS IS FOR SITE 1:

generate **present**, site a (HOK 1)
```{r}
########################
mass.hok1.present.df <- mass.hok1.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Create a new column combining tree_num and ayres with a period in between
mass.hok1.present.df$ayres.tree <- paste(mass.hok1.present.df$ayres, mass.hok1.present.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Armla
names_vector_armla <- character(0)
values_vector_armla <- numeric(0)
treatment_vector_armla <- character(0)


# Initialize empty vectors to store names and values for Brmla
names_vector_brmla <- character(0)
values_vector_brmla <- numeric(0)
treatment_vector_brmla <- character(0)

# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Aa_column <- paste("Aa.", X, sep = "")
    Ba_column <- paste("Ba.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    armla_result <- (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Aa_column, "perc.mass_loss"] / 
                     (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Aa_column, "perc.mass_loss"] + 
                      mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Ba_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    brmla_result <- (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Ba_column, "perc.mass_loss"] / 
                     (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Ba_column, "perc.mass_loss"] + 
                      mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Aa_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    armla_name <- paste("Armla", X, Y, sep = ".")
    brmla_name <- paste("Brmla", X, Y, sep = ".")

     # Extract the treatment from mass.hok1.present.df for the current combination
    treatment_name <- mass.hok1.present.df[mass.hok1.present.df$ayres.tree == Aa_column, "treatment"]
    

    # Append the names and values to the respective vectors for Armla
    names_vector_armla <- c(names_vector_armla, armla_name)
    values_vector_armla <- c(values_vector_armla, armla_result)
    treatment_vector_armla <- c(treatment_vector_armla, treatment_name)

    
    # Append the names and values to the respective vectors for Brmla
    names_vector_brmla <- c(names_vector_brmla, brmla_name)
    values_vector_brmla <- c(values_vector_brmla, brmla_result)
    treatment_vector_brmla <- c(treatment_vector_brmla, treatment_name)


  }
}

# Create dataframes for Armla and Brmla
armla_df <- data.frame(names = names_vector_armla, values = values_vector_armla, treatment= treatment_vector_armla)
brmla_df <- data.frame(names = names_vector_brmla, values = values_vector_brmla, treatment=treatment_vector_brmla)

#merge the dataframes
armla.brmla.df_1=rbind(armla_df, brmla_df)
armla.brmla.df_1
```


NEW: use this! This is a now a proper test dataframe with for loop!
generate **absent**, site a (HOK 1)
```{r}
########################
mass.hok1.absent.df <- mass.hok1.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Create a new column combining tree_num and ayres with a period in between
mass.hok1.absent.df$ayres.tree <- paste(mass.hok1.absent.df$ayres, mass.hok1.absent.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values<- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Armla
names_vector_armla<- character(0)
values_vector_armla<- numeric(0)
treatment_vector_armla <- character(0)


# Initialize empty vectors to store names and values for Brmla
names_vector_brmla <- character(0)
values_vector_brmla <- numeric(0)
treatment_vector_brmla <- character(0)

# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Aa_column <- paste("Aa.", X, sep = "")
    Ba_column <- paste("Ba.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    armla_result <- (mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Aa_column, "perc.mass_loss"] / 
                     (mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Aa_column, "perc.mass_loss"] + 
                      mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Ba_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    brmla_result <- (mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Ba_column, "perc.mass_loss"] / 
                     (mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Ba_column, "perc.mass_loss"] + 
                      mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Aa_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    armla_name <- paste("Armla", X, Y, sep = ".")
    brmla_name <- paste("Brmla", X, Y, sep = ".")
    
    # Extract the treatment from mass.hok1.present.df for the current combination
    treatment_name <- mass.hok1.absent.df[mass.hok1.absent.df$ayres.tree == Aa_column, "treatment"]
    
    
    # Append the names and values to the respective vectors for Armla
    names_vector_armla <- c(names_vector_armla, armla_name)
    values_vector_armla<- c(values_vector_armla, armla_result)
        treatment_vector_armla <- c(treatment_vector_armla, treatment_name)

    
    # Append the names and values to the respective vectors for Brmla
    names_vector_brmla <- c(names_vector_brmla, brmla_name)
    values_vector_brmla <- c(values_vector_brmla, brmla_result)
        treatment_vector_brmla <- c(treatment_vector_brmla, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
armla_ab_df <- data.frame(names = names_vector_armla, values = values_vector_armla, treatment = treatment_vector_armla)

brmla_ab_df <- data.frame(names = names_vector_brmla, values = values_vector_brmla, treatment = treatment_vector_brmla)

# View the resulting dataframes
print(armla_df)
print(brmla_df)

armla.brmla.absent.df_1=rbind(armla_ab_df, brmla_ab_df)
armla.brmla.absent.df_1

```


NEW: use this! This is a now a proper test dataframe with for loop!
generate **present**, site a (SEK 1)
```{r}
mass.sek1.present.df <- mass.sek1.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  # Create a new column combining tree_num and ayres with a period in between
mass.sek1.present.df$ayres.tree <- paste(mass.sek1.present.df$ayres, mass.sek1.present.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Brmlb
names_vector_brmlb <- character(0)
values_vector_brmlb <- numeric(0)
treatment_vector_brmlb <- character(0)

# Initialize empty vectors to store names and values for Armlb
names_vector_armlb <- character(0)
values_vector_armlb <- numeric(0)
treatment_vector_armlb <- character (0)

# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Bb_column <- paste("Bb.", X, sep = "")
    Ab_column <- paste("Ab.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    brmlb_result <- (mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Bb_column, "perc.mass_loss"] / 
                     (mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Bb_column, "perc.mass_loss"] + 
                      mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Ab_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    armlb_result <- (mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Ab_column, "perc.mass_loss"] / 
                     (mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Ab_column, "perc.mass_loss"] + 
                      mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Bb_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    brmlb_name <- paste("Brmlb", X, Y, sep = ".")
    armlb_name <- paste("Armlb", X, Y, sep = ".")
    
    # Extract the treatment from mass.hok1.present.df for the current combination
    treatment_name <- mass.sek1.present.df[mass.sek1.present.df$ayres.tree == Ab_column, "treatment"]
    
    
    # Append the names and values to the respective vectors for Brmlb
    names_vector_brmlb <- c(names_vector_brmlb, brmlb_name)
    values_vector_brmlb <- c(values_vector_brmlb, brmlb_result)
    treatment_vector_brmlb <- c(treatment_vector_brmlb, treatment_name)
    
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_armlb <- c(names_vector_armlb, armlb_name)
    values_vector_armlb <- c(values_vector_armlb, armlb_result)
        treatment_vector_armlb <- c(treatment_vector_armlb, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
brmlb_df <- data.frame(names = names_vector_brmlb, values = values_vector_brmlb, treatment = treatment_vector_brmlb)
armlb_df <- data.frame(names = names_vector_armlb, values = values_vector_armlb, treatment = treatment_vector_armlb)

# View the resulting dataframes
print(brmlb_df)
print(armlb_df)

brmlb.armlb.df_1=rbind(brmlb_df, armlb_df)

brmlb.armlb.df_1
```

NEW: use this! This is a now a proper test dataframe with for loop!
generate **absent**, site a (SEK 1)
```{r}
mass.sek1.absent.df <- mass.sek1.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  # Create a new column combining tree_num and ayres with a period in between
mass.sek1.absent.df$ayres.tree <- paste(mass.sek1.absent.df$ayres, mass.sek1.absent.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values_3 <- 1:5
Y_values_3 <- 1:5

# Initialize empty vectors to store names and values for Brmlb
names_vector_brmlb <- character(0)
values_vector_brmlb <- numeric(0)
treatment_vector_brmlb <- character (0)

# Initialize empty vectors to store names and values for Armlb
names_vector_armlb <- character(0)
values_vector_armlb <- numeric(0)
treatment_vector_armlb <- character (0)


# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values_3) {
  for (Y in Y_values_3) {
    Bb_column <- paste("Bb.", X, sep = "")
    Ab_column <- paste("Ab.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    brmlb_result <- (mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Bb_column, "perc.mass_loss"] / 
                     (mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Bb_column, "perc.mass_loss"] + 
                      mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Ab_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    armlb_result <- (mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Ab_column, "perc.mass_loss"] / 
                     (mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Ab_column, "perc.mass_loss"] + 
                      mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Bb_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    brmlb_name <- paste("Brmlb", X, Y, sep = ".")
    armlb_name <- paste("Armlb", X, Y, sep = ".")
    
     # Extract the treatment from mass.hok1.present.df for the current combination
    treatment_name <- mass.sek1.absent.df[mass.sek1.absent.df$ayres.tree == Ab_column, "treatment"]
    
    
    
    # Append the names and values to the respective vectors for Brmlb
    names_vector_brmlb <- c(names_vector_brmlb, brmlb_name)
    values_vector_brmlb <- c(values_vector_brmlb, brmlb_result)
    treatment_vector_brmlb <- c(treatment_vector_brmlb, treatment_name)
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_armlb <- c(names_vector_armlb, armlb_name)
    values_vector_armlb <- c(values_vector_armlb, armlb_result)
        treatment_vector_armlb <- c(treatment_vector_armlb, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
brmlb_ab_df <- data.frame(names = names_vector_brmlb, values = values_vector_brmlb, treatment = treatment_vector_brmlb)
armlb_ab_df <- data.frame(names = names_vector_armlb, values = values_vector_armlb, treatment = treatment_vector_armlb)


brmlb.armlb.absent.df_1=rbind(brmlb_ab_df, armlb_ab_df)
brmlb.armlb.absent.df_1

```


NEW: use this! merging HOK 1 and SEK 1 dataframe
merge **present**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site1.df=rbind(armla.brmla.df_1, brmlb.armlb.df_1)
mass.present.site1.df
# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Create an empty dataframe to store the results
result_df_1 <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla.", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla.", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb.", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb.", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.present.site1.df[mass.present.site1.df$names == armla_name, "values"] +
      mass.present.site1.df[mass.present.site1.df$names == brmlb_name, "values"])/2)/ (
      (mass.present.site1.df[mass.present.site1.df$names == armlb_name, "values"] +
      mass.present.site1.df[mass.present.site1.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
    result_name <- LETTERS[(X - 1) * 5 + Y]
    
  # Get the treatment for the current combination
    treatment_name <- mass.present.site1.df[mass.present.site1.df$names == armla_name, "treatment"]
 
# Append the name and value to the result dataframe
    result_df_1 <- rbind(result_df_1, data.frame(names = result_name, values = result, treatment= treatment_name))  }
}


# View the resulting dataframe
result_df_1
```


NEW: use this! merging HOK 1 and SEK 1 dataframe
merge **absent**, merging HOK 1 and SEK 1 dataframe
```{r}
#merge both present site 1 dataframes
mass.absent.site1.df=rbind(armla.brmla.absent.df_1, brmlb.armlb.absent.df_1)

# Create an empty dataframe to store the results
result_ab_df_1 <- data.frame(names = character(0), values = numeric(0), treatment = character (0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla.", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla.", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb.", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb.", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.absent.site1.df[mass.absent.site1.df$names == armla_name, "values"] +
      mass.absent.site1.df[mass.absent.site1.df$names == brmlb_name, "values"])/2)/ (
      (mass.absent.site1.df[mass.absent.site1.df$names == armlb_name, "values"] +
      mass.absent.site1.df[mass.absent.site1.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
result_name <- paste0(LETTERS[(X - 1) * 5 + Y], "_1")

# Get the treatment for the current combination
    treatment_name <- mass.absent.site1.df[mass.absent.site1.df$names == armla_name, "treatment"]
    

    # Append the name and value to the result dataframe
    result_ab_df_1 <- rbind(result_ab_df_1, data.frame(names = result_name, values = result, treatment= treatment_name))
  }
}


# View the resulting dataframe
result_ab_df_1


```


combine P and AB; plot check for P and AB site 1 
```{r}
#combine sites 1 and 2
hfai.allsite1.df=rbind(result_df_1, result_ab_df_1)

ggplot(hfai.allsite1.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  geom_bar(stat = "identity", position= "dodge", alpha = 0.5) +
  scale_fill_viridis_d()+
  xlab("reciprocated pairs") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.x = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  )


hfai.allsite1.df$values %>%
  n_distinct()



```


NOTICE: THIS IS NOW SITE 2:

NEW: use this! This is a now a proper test dataframe with for loop!
generate **present**, site a (HOK 2)
```{r}
########################
mass.hok2.present.df <- mass.hok2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)
# Create a new column combining tree_num and ayres with a period in between
mass.hok2.present.df$ayres.tree <- paste(mass.hok2.present.df$ayres, mass.hok2.present.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Armla
names_vector_armla <- character(0)
values_vector_armla <- numeric(0)
treatment_vector_armla<- character(0)

# Initialize empty vectors to store names and values for Brmla
names_vector_brmla <- character(0)
values_vector_brmla<- numeric(0)
treatment_vector_brmla<- character(0)


# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Aa_column <- paste("Aa.", X, sep = "")
    Ba_column <- paste("Ba.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    armla_result <- (mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Aa_column, "perc.mass_loss"] / 
                     (mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Aa_column, "perc.mass_loss"] + 
                      mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Ba_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    brmla_result <- (mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Ba_column, "perc.mass_loss"] / 
                     (mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Ba_column, "perc.mass_loss"] + 
                      mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Aa_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    armla_name <- paste("Armla", X, Y, sep = ".")
    brmla_name <- paste("Brmla", X, Y, sep = ".")
    
    # Get the treatment for the current combination
    treatment_name <- mass.hok2.present.df[mass.hok2.present.df$ayres.tree == Aa_column, "treatment"]
    
    # Append the names and values to the respective vectors for Armla
    names_vector_armla <- c(names_vector_armla, armla_name)
    values_vector_armla <- c(values_vector_armla, armla_result)
    treatment_vector_armla<- c(treatment_vector_armla, treatment_name)

    # Append the names and values to the respective vectors for Brmla
    names_vector_brmla <- c(names_vector_brmla, brmla_name)
    values_vector_brmla <- c(values_vector_brmla, brmla_result)
        treatment_vector_brmla<- c(treatment_vector_brmla, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
armla_df_2 <- data.frame(names = names_vector_armla, values = values_vector_armla, treatment = treatment_vector_armla)
brmla_df_2 <- data.frame(names = names_vector_brmla, values = values_vector_brmla, treatment = treatment_vector_brmla)



# View the resulting dataframes
print(armla_df_2)
print(brmla_df_2)

armla.brmla.df_2=rbind(armla_df_2, brmla_df_2)


```


NEW: use this! This is a now a proper test dataframe with for loop!
generate **absent**, site a (HOK 2)
```{r}
########################
mass.hok2.absent.df <- mass.hok2.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Create a new column combining tree_num and ayres with a period in between
mass.hok2.absent.df$ayres.tree <- paste(mass.hok2.absent.df$ayres, mass.hok2.absent.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Armla
names_vector_armla <- character(0)
values_vector_armla <- numeric(0)
treatment_vector_armla <- character(0)

# Initialize empty vectors to store names and values for Brmla
names_vector_brmla <- character(0)
values_vector_brmla <- numeric(0)
treatment_vector_brmla <- character(0)


# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Aa_column <- paste("Aa.", X, sep = "")
    Ba_column <- paste("Ba.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    armla_result <- (mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Aa_column, "perc.mass_loss"] / 
                     (mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Aa_column, "perc.mass_loss"] + 
                      mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Ba_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    brmla_result <- (mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Ba_column, "perc.mass_loss"] / 
                     (mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Ba_column, "perc.mass_loss"] + 
                      mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Aa_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    armla_name <- paste("Armla", X, Y, sep = ".")
    brmla_name <- paste("Brmla", X, Y, sep = ".")
    
    # create treatment 
    
    treatment_name <- mass.hok2.absent.df[mass.hok2.absent.df$ayres.tree == Aa_column, "treatment"]
    
    # Append the names and values to the respective vectors for Armla
    names_vector_armla <- c(names_vector_armla, armla_name)
    values_vector_armla <- c(values_vector_armla, armla_result)
    treatment_vector_armla <- c(treatment_vector_armla, treatment_name)
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_brmla <- c(names_vector_brmla, brmla_name)
    values_vector_brmla <- c(values_vector_brmla, brmla_result)
        treatment_vector_brmla <- c(treatment_vector_brmla, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
armla_ab_df_2 <- data.frame(names = names_vector_armla, values = values_vector_armla, treatment = treatment_vector_armla)
brmla_ab_df_2 <- data.frame(names = names_vector_brmla, values = values_vector_brmla, treatment = treatment_vector_brmla)

# View the resulting dataframes
print(armla_ab_df_2)
print(armla_ab_df_2)

armla.brmla.absent.df_2=rbind(armla_ab_df_2, brmla_ab_df_2)
armla.brmla.absent.df_2

```

NEW: use this! This is a now a proper test dataframe with for loop!
generate **present**, site a (SEK 2)
```{r}
mass.sek2.present.df <- mass.sek2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  # Create a new column combining tree_num and ayres with a period in between
mass.sek2.present.df$ayres.tree <- paste(mass.sek2.present.df$ayres, mass.sek2.present.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Brmlb
names_vector_brmlb <- character(0)
values_vector_brmlb <- numeric(0)
treatment_vector_brmlb <- character(0)


# Initialize empty vectors to store names and values for Armlb
names_vector_armlb <- character(0)
values_vector_armlb <- numeric(0)
treatment_vector_armlb <- character(0)

# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Bb_column <- paste("Bb.", X, sep = "")
    Ab_column <- paste("Ab.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    brmlb_result <- (mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Bb_column, "perc.mass_loss"] / 
                     (mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Bb_column, "perc.mass_loss"] + 
                      mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Ab_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    armlb_result <- (mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Ab_column, "perc.mass_loss"] / 
                     (mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Ab_column, "perc.mass_loss"] + 
                      mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Bb_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    brmlb_name <- paste("Brmlb", X, Y, sep = ".")
    armlb_name <- paste("Armlb", X, Y, sep = ".")
    
    # create treatment
    treatment_name <- mass.sek2.present.df[mass.sek2.present.df$ayres.tree == Bb_column, "treatment"]

    
    # Append the names and values to the respective vectors for Brmlb
    names_vector_brmlb <- c(names_vector_brmlb, brmlb_name)
    values_vector_brmlb <- c(values_vector_brmlb, brmlb_result)
    treatment_vector_brmlb<- c(treatment_vector_brmlb, treatment_name)
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_armlb <- c(names_vector_armlb, armlb_name)
    values_vector_armlb <- c(values_vector_armlb, armlb_result)
        treatment_vector_armlb<- c(treatment_vector_armlb, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
brmlb_df_2 <- data.frame(names = names_vector_brmlb, values = values_vector_brmlb, treatment = treatment_vector_brmlb)
armlb_df_2 <- data.frame(names = names_vector_armlb, values = values_vector_armlb, treatment = treatment_vector_armlb)

# View the resulting dataframes
print(brmlb_df_2)
print(armlb_df_2)

brmlb.armlb.df_2=rbind(brmlb_df_2, armlb_df_2)
brmlb.armlb.df_2

```

NEW: use this! This is a now a proper test dataframe with for loop!
generate **absent**, site a (SEK 2)
```{r}
mass.sek2.absent.df <- mass.sek2.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  # Create a new column combining tree_num and ayres with a period in between
mass.sek2.absent.df$ayres.tree <- paste(mass.sek2.absent.df$ayres, mass.sek2.absent.df$tree_num, sep = ".")

# Define the range for X and Y (1 to 5)
X_values<- 1:5
Y_values <- 1:5

# Initialize empty vectors to store names and values for Brmlb
names_vector_brmlb <- character(0)
values_vector_brmlb <- numeric(0)
treatment_vector_brmlb<- character(0)

# Initialize empty vectors to store names and values for Armlb
names_vector_armlb <- character(0)
values_vector_armlb <- numeric(0)
treatment_vector_armlb<- character(0)


# Loop through all combinations of X and Y for Aa and Ba
for (X in X_values) {
  for (Y in Y_values) {
    Bb_column <- paste("Bb.", X, sep = "")
    Ab_column <- paste("Ab.", Y, sep = "")
    
    # Calculate the result for Armla for the current combination
    brmlb_result <- (mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Bb_column, "perc.mass_loss"] / 
                     (mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Bb_column, "perc.mass_loss"] + 
                      mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Ab_column, "perc.mass_loss"])) * 100
    
    # Calculate the result for Brmla for the current combination
    armlb_result <- (mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Ab_column, "perc.mass_loss"] / 
                     (mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Ab_column, "perc.mass_loss"] + 
                      mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Bb_column, "perc.mass_loss"])) * 100
    
    # Create a new name for the results
    brmlb_name <- paste("Brmlb", X, Y, sep = ".")
    armlb_name <- paste("Armlb", X, Y, sep = ".")
    
    # create treatment 
    treatment_name <- mass.sek2.absent.df[mass.sek2.absent.df$ayres.tree == Bb_column, "treatment"]
    
    
    # Append the names and values to the respective vectors for Brmlb
    names_vector_brmlb <- c(names_vector_brmlb, brmlb_name)
    values_vector_brmlb <- c(values_vector_brmlb, brmlb_result)
    treatment_vector_brmlb<- c(treatment_vector_brmlb, treatment_name)
    
    # Append the names and values to the respective vectors for Brmla
    names_vector_armlb <- c(names_vector_armlb, armlb_name)
    values_vector_armlb <- c(values_vector_armlb, armlb_result)
        treatment_vector_armlb<- c(treatment_vector_armlb, treatment_name)

  }
}

# Create dataframes for Armla and Brmla
brmlb_ab_df_2 <- data.frame(names = names_vector_brmlb, values = values_vector_brmlb, treatment = treatment_vector_brmlb)
armlb_ab_df_2 <- data.frame(names = names_vector_armlb, values = values_vector_armlb, treatment = treatment_vector_armlb)


brmlb.armlb.absent.df_2=rbind(brmlb_ab_df_2, armlb_ab_df_2)
brmlb.armlb.absent.df_2

```

NEW: use this! merging HOK 2 and SEK 2 dataframe
merge **present**, site a and b (aka HOK 2 and SEK 2) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site2.df=rbind(armla.brmla.df_2, brmlb.armlb.df_2)
mass.present.site2.df
# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Create an empty dataframe to store the results
result_df_2 <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla.", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla.", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb.", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb.", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.present.site2.df[mass.present.site2.df$names == armla_name, "values"] +
      mass.present.site2.df[mass.present.site2.df$names == brmlb_name, "values"])/2)/ (
      (mass.present.site2.df[mass.present.site2.df$names == armlb_name, "values"] +
      mass.present.site2.df[mass.present.site2.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
    result_name <- paste0(LETTERS[(X - 1) * 5 + Y], "_2")
    
  # Get the treatment for the current combination
    treatment_name <- mass.present.site2.df[mass.present.site2.df$names == armla_name, "treatment"]
 
# Append the name and value to the result dataframe
    result_df_2 <- rbind(result_df_2, data.frame(names = result_name, values = result, treatment= treatment_name))  }
}


# View the resulting dataframe

print(result_df_2)



```


NEW: use this! merging HOK 2 and SEK 2 dataframe
merge **absent**, merging HOK 2 and SEK 2 dataframe
```{r}
#merge both present site 1 dataframes
mass.absent.site2.df=rbind(armla.brmla.absent.df_2, brmlb.armlb.absent.df_2)
mass.absent.site2.df
# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Create an empty dataframe to store the results
result_ab_df_2 <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla.", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla.", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb.", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb.", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.absent.site2.df[mass.absent.site2.df$names == armla_name, "values"] +
      mass.absent.site2.df[mass.absent.site2.df$names == brmlb_name, "values"])/2)/ (
      (mass.absent.site2.df[mass.absent.site2.df$names == armlb_name, "values"] +
      mass.absent.site2.df[mass.absent.site2.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
    result_name <- paste0(LETTERS[(X - 1) * 5 + Y], "_3")
    
  # Get the treatment for the current combination
    treatment_name <- mass.absent.site2.df[mass.absent.site2.df$names == armla_name, "treatment"]
 
# Append the name and value to the result dataframe
    result_ab_df_2 <- rbind(result_ab_df_2, data.frame(names = result_name, values = result, treatment= treatment_name))  }
}


# View the resulting dataframe
result_ab_df_2

```


NEW: use this! merging all dataframes (Sites 1 and 2; includes P and AB)
merge all sites; HFAI plots
```{r}
#combine sites 1 and 2 including P and AB
hfai.allsites.df=rbind(result_df_1, result_df_2, result_ab_df_1, result_ab_df_2)
hfai.allsites.df$treatment=as.factor(hfai.allsites.df$treatment)


####################  P ONLY ###################
#filter P treatment only for HFA effect
hfai.present.allsites.df=hfai.allsites.df%>%
  filter(treatment == "P")


# Filter the dataframe to include only positive cases
positive_present_df <- hfai.present.allsites.df[hfai.present.allsites.df$values > 0, ]

# Calculate the mean and standard error (SE)
mean_hfa <- mean(positive_present_df$values)
se_hfa <- sd(positive_present_df$values) / sqrt(nrow(positive_present_df))
# Perform a t-test
t_test_present_result <- t.test(positive_present_df$values, mu = 0)
t_test_present_result



# Format the t-test result
t_test_label <- paste("p =", formatC(t_test_present_result$p.value, format = "e", digits = 2), ",", "t-test =", round(t_test_present_result$statistic, 2))
# Format the t-test result
p_value <- t_test_present_result$p.value
formatted_p_value <- if (p_value < 0.0001) {
  "p < 0.0001"
} else {
  paste("p =", formatC(p_value, digits = 1))
}


#positive and negative values
total_positive <- sum(hfai.present.allsites.df$values > 0)
total_negative <- sum(hfai.present.allsites.df$values < 0)
# Calculate the percentage of positive values relative to negative
percentage_positive <- (total_positive / (total_positive + total_negative)) * 100
percentage_negative <- (total_negative / (total_negative + total_positive)) * 100
# 68% positive
# 32% negative

#just P treatment
#just P treatment with coord_flip
hfai.present.allsites.plot=ggplot(hfai.present.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  scale_fill_manual(values = "grey")+
  geom_bar(stat = "identity", position ="dodge", alpha = 0.6) +
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.x = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 14.5, linetype = "dashed", color = "red") +
scale_y_continuous(limits = c(-20, 70), breaks = seq(-20, 80, by = 20))
hfai.present.allsites.plot

#just P treatment with coord_flip
hfai.present.allsites.plot_flipped=ggplot(hfai.present.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  scale_fill_manual(values = "grey")+
  geom_bar(stat = "identity", position ="dodge", alpha = 0.6) +
  coord_flip()+
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 14.5, linetype = "dashed", color = "red") +
scale_y_continuous(limits = c(-20, 70), breaks = seq(-20, 70, by = 10))



hfai.present.allsites.plot_flipped=ggplot(hfai.present.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  scale_fill_manual(values = "grey")+
  geom_bar(stat = "identity", position ="dodge", alpha = 0.6) +
  coord_flip()+
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 16.5, linetype = "dashed", color = "red")

hfai.present.allsites.plot_flipped
ggsave("figures/hfai.present.allsites.plot_flipped.pdf", hfai.present.allsites.plot_flipped, width = 14, height = 10)


```


```{r}
######################### BOTH P and AB #########################################

# Filter the dataframe to include only positive cases

mean_positive_P <- mean(hfai.allsites.df[hfai.allsites.df$treatment == "P" & hfai.allsites.df$values > 0, "values"])

positive_values <- hfai.allsites.df %>%
  filter(values > 0)

mean_positive <- mean(positive_values$values)
se_positive <- sd(positive_values$values) / sqrt(nrow(positive_values))

se_positive_P <- sd(hfai.allsites.df[hfai.allsites.df$treatment == "P" & hfai.allsites.df$values > 0, "values"]) / sqrt(nrow(hfai.allsites.df[hfai.allsites.df$treatment == "P" & hfai.allsites.df$values > 0, ]))


mean_positive_AB <- mean(hfai.allsites.df[hfai.allsites.df$treatment == "AB" & hfai.allsites.df$values > 0, "values"])
# mean positive P = 13.7
# mean positive AB = 14.2

print(hfai.allsites.df[hfai.allsites.df$treatment=="P" & hfai.allsites.df$values>0, "values"])
print(hfai.allsites.df[hfai.allsites.df$treatment=="AB" & hfai.allsites.df$values>0, "values"])

# Calculate the mean of positive "P" values relative to all positive values
mean_p_relative_to_all <- mean(p_positive_subset$values) / mean(hfai.allsites.df[hfai.allsites.df$values > 0, ]$values)


View(hfai.allsites.df)

hfai.allsites.df %>%
  filter(values > 0) %>%
  summarise(mean = mean(values), se = sd(values) / sqrt(n()))

hfai.allsites.df %>%
  filter(values < 0) %>%
  summarise(mean = mean(values))

hfai.allsites.df%>%
  filter(treatment == "AB")%>%
  filter(values > 0)%>%
n_distinct()
# P 34 > 0; AB 19 > 0; P 16 < 0; AB 31 < 0

# Subset the dataframe to include only positive values for each treatment
positive_values <- hfai.allsites.df[hfai.allsites.df$values > 0, "values"]

positive_values_P <- hfai.allsites.df[hfai.allsites.df$treatment == "P" & hfai.allsites.df$values > 0, "values"]
positive_values_AB <- hfai.allsites.df[hfai.allsites.df$treatment == "AB" & hfai.allsites.df$values > 0, "values"]

# Perform a t-test for each treatment against 0
t_test<- t.test(positive_values, mu = 0)
t_test_P <- t.test(positive_values_P, mu = 0)
t_test_AB <- t.test(positive_values_AB, mu = 0)

# Calculate Cohen's d for each treatment
cohen_d_P <- t_test_P$estimate / sd(positive_values_P)
cohen_d_AB <- t_test_AB$estimate / sd(positive_values_AB)
# Filter the dataframe for positive values (both "P" and "AB")
positive_values_subset <- hfai.allsites.df[hfai.allsites.df$values > 0, ]


# Calculate the proportion of positive values for the "P" treatment relative to all positive values
proportion_positive_P <- sum(hfai.allsites.df$treatment == "P" & hfai.allsites.df$values > 0) /
                        sum(hfai.allsites.df$values > 0)
proportion_positive_AB <- sum(hfai.allsites.df$treatment == "AB" & hfai.allsites.df$values > 0) /
                        sum(hfai.allsites.df$values > 0)

positive_values_subset%>%
  filter(treatment == "AB")%>%
  n_distinct()
hfai.allsites.df%>%
  filter(values > 0)%>%
  n_distinct()
#53 positive
# 47 negative


# Perform a t-test
t_test_present_result <- t.test(positive_all_df$values, mu = 0)
t_test_present_result

# Format the t-test result
t_test_label <- paste("p =", formatC(t_test_present_result$p.value, format = "e", digits = 2), ",", "t-test =", round(t_test_present_result$statistic, 2))
# Format the t-test result
p_value <- t_test_present_result$p.value
formatted_p_value <- if (p_value < 0.0001) {
  "p < 0.0001"
} else {
  paste("p =", formatC(p_value, digits = 1))
}

#### proportion that are positive

# Filter the dataframe for "AB" and "P" treatments
ab_p_subset <- hfai.allsites.df[hfai.allsites.df$treatment %in% c("AB", "P"), ]

# Further filter to include only rows where values are greater than 0
positive_values_subset <- ab_p_subset[ab_p_subset$values > 0, ]
View(positive_values_subset)
# Calculate the proportion of values greater than 0 for "AB"
prop_positive_ab <- sum(positive_values_subset$treatment == "AB") / nrow(positive_values_subset)

# Calculate the proportion of values greater than 0 for "P"
prop_positive_p <- sum(positive_values_subset$treatment == "P") / nrow(positive_values_subset)


##### proportion that are negative
# Further filter to include only rows where values are greater than 0
neg_values_subset <- ab_p_subset[ab_p_subset$values < 0, ]

# Calculate the proportion of values greater than 0 for "AB"
prop_neg_ab <- sum(neg_values_subset$treatment == "AB") / nrow(ab_p_subset$values)


# Calculate the proportion of values greater than 0 for "P"
prop_neg_p <- sum(neg_values_subset$treatment == "P") / nrow(ab_p_subset$values)

# Calculate mean and SE for 'AB' and 'P' values individually
mean_ab <- mean(positive_values_subset[positive_values_subset$treatment == "AB", "values"])

# Calculate the weighted mean for group P
weighted_mean_P <- sum(positive_values_subset[positive_values_subset$treatment == "P", "values"]) / sum(positive_values_subset$treatment == "P")

# Calculate the weighted mean for group AB
weighted_mean_AB <- sum(positive_values_subset[positive_values_subset$treatment == "AB", "values"]) / sum(positive_values_subset$treatment == "AB")

# Print the results
weighted_mean_P
weighted_mean_AB

med_ab <- median(positive_values_subset[positive_values_subset$treatment == "AB", "values"])
se_ab <- sd(positive_values_subset[positive_values_subset$treatment == "AB", "values"]) / sqrt(sum(positive_values_subset$treatment == "AB"))

mean_p <- mean(positive_values_subset[positive_values_subset$treatment == "P", "values"])
med_p <- median(positive_values_subset[positive_values_subset$treatment == "P", "values"])
se_p <- sd(positive_values_subset[positive_values_subset$treatment == "P", "values"]) / sqrt(sum(positive_values_subset$treatment == "P"))


# Perform a t-test
t_test_result <- t.test(positive_values_subset[positive_values_subset$treatment == "P", "values"], mu = 0)
t_test_result <- t.test(positive_values_subset[positive_values_subset$treatment == "AB", "values"], mu = 0)

d <- (mean_p - mean_ab) / sqrt((se_p^2 + se_ab^2) / 2)
d
#positive and negative values
total_positive <- sum(hfai.allsites.df$values > 0)
total_negative <- sum(hfai.allsites.df$values < 0)
# Calculate the percentage of positive values relative to negative
percentage_positive <- (total_positive / (total_positive + total_negative)) * 100
percentage_negative <- (total_negative / (total_negative + total_positive)) * 100

# 53% positive
# 47% negative

#both P and AB treatments
hfai.allsites.plot=ggplot(hfai.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  geom_bar(stat = "identity", position ="dodge", alpha = 0.3) +
  scale_fill_viridis_d()+
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.x = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 44.5, linetype = "dashed", color = "red")+
  scale_y_continuous(
    limits = c(-65, 255),
     breaks = seq(-65, 250, by = 65),
    expand = c(0, 0)
  )

hfai.allsites.plot
min(hfai.allsites.df$values) 
max(hfai.allsites.df$values)


ggplot(hfai.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  geom_bar(stat = "identity", position ="dodge", alpha = 0.3) +
  scale_fill_manual(values = c("grey", "grey1"))+
  coord_flip()+
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 44.5, linetype = "dashed", color = "red")+
  scale_y_continuous(
    limits = c(-65, 255),
     breaks = seq(-65, 250, by = 65),
    expand = c(0, 0)
  )

hfai.allsites.plot_flipped=ggplot(hfai.allsites.df, aes(x = reorder(names, values), y = values, fill = treatment)) +
  geom_bar(stat = "identity", position ="dodge", alpha = 0.4) +
  scale_fill_manual(values = c("grey", "grey0"))+
  coord_flip()+
  xlab("") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank())+  
  geom_vline(xintercept = 46.5, linetype = "dashed", color = "red")
hfai.allsites.plot_flipped

max(hfai.allsites.df[hfai.allsites.df$treatment =="AB" & hfai.allsites.df$values >0, "values"])

ggsave("figures/hfai.allsites.plot_flipped.pdf", hfai.allsites.plot_flipped, width = 14, height = 10)


```

------------------------END--------------------------------
Everything below isnt used



















This is also the proper dataframe
generate **present**, site a (HOK 2)
```{r}
mass.hok2.present.df <- mass.hok2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values_z <- numeric()
rml_values_mass <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_z_loss <- mass.hok2.present.df$z_score[
      mass.hok2.present.df$tree_id == "HOK" & mass.hok2.present.df$tree_num == tree_num
    ]
    sek_z_loss <- mass.hok2.present.df$z_score[
      mass.hok2.present.df$tree_id == "SEK" & mass.hok2.present.df$tree_num == tree_num
    ]
    
    rml.z <- ifelse(
      ayres == "Aa", hok_z_loss / (hok_z_loss + sek_z_loss),
      ifelse(ayres == "Ba", sek_z_loss / (sek_z_loss + hok_z_loss), NA)
    )
    
    rml_values_z <- c(rml_values_z, rml.z * 100)
    
    hok_mass_loss <- mass.hok2.present.df$perc.mass_loss[
      mass.hok2.present.df$tree_id == "HOK" & mass.hok2.present.df$tree_num == tree_num
    ]
    sek_mass_loss <- mass.hok2.present.df$perc.mass_loss[
      mass.hok2.present.df$tree_id == "SEK" & mass.hok2.present.df$tree_num == tree_num
    ]
    
    rml.mass <- ifelse(
      ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
      ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss), NA)
    )
    
    rml_values_mass <- c(rml_values_mass, rml.mass * 100)
  }
}

# Add the rml values to the dataframe
mass.hok2.present.df$rml.z <- rml_values_z
mass.hok2.present.df$rml.mass <- rml_values_mass

# Create a new column combining tree_num and ayres with a period in between
mass.hok2.present.df$ayres.tree <- paste(mass.hok2.present.df$ayres, mass.hok2.present.df$tree_num, sep = ".")

# Create a new column based on Ayres HFAI equation 
mass.hok2.present.df$ayres.eq <- ifelse(
  mass.hok2.present.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""),
  paste("Armla", tree_num_values, sep = "."),
  ifelse(
    mass.hok2.present.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""),
    paste("Brmla", tree_num_values, sep = "."),
    NA
  )
)


```


This is also the proper dataframe
generate **present**, site a (SEK 2)
```{r}
mass.sek2.present.df <- mass.sek2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values_z <- numeric()
rml_values_mass <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_z_loss <- mass.sek2.present.df$z_score[
      mass.sek2.present.df$tree_id == "HOK" & mass.sek2.present.df$tree_num == tree_num
    ]
    sek_z_loss <- mass.sek2.present.df$z_score[
      mass.sek2.present.df$tree_id == "SEK" & mass.sek2.present.df$tree_num == tree_num
    ]
    
    rml.z <- ifelse(
      ayres == "Ab", hok_z_loss / (hok_z_loss + sek_z_loss),
      ifelse(ayres == "Bb", sek_z_loss / (sek_z_loss + hok_z_loss), NA)
    )
    
    rml_values_z <- c(rml_values_z, rml.z * 100)
    
    hok_mass_loss <- mass.sek2.present.df$perc.mass_loss[
      mass.sek2.present.df$tree_id == "HOK" & mass.sek2.present.df$tree_num == tree_num
    ]
    sek_mass_loss <- mass.sek2.present.df$perc.mass_loss[
      mass.sek2.present.df$tree_id == "SEK" & mass.sek2.present.df$tree_num == tree_num
    ]
    
    rml.mass <- ifelse(
      ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
      ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss), NA)
    )
    
    rml_values_mass <- c(rml_values_mass, rml.mass * 100)
  }
}

# Add the rml values to the dataframe
mass.sek2.present.df$rml.z <- rml_values_z
mass.sek2.present.df$rml.mass <- rml_values_mass

# Create a new column combining tree_num and ayres with a period in between
mass.sek2.present.df$ayres.tree <- paste(mass.sek2.present.df$ayres, mass.sek2.present.df$tree_num, sep = ".")

# Create a new column based on Ayres HFAI equation 
mass.sek2.present.df$ayres.eq <- ifelse(
  mass.sek2.present.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""),
  paste("Armlb", tree_num_values, sep = "."),
  ifelse(
    mass.sek2.present.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""),
    paste("Brmlb", tree_num_values, sep = "."),
    NA
  )
)


```



NEW: use this! merging HOK 1 and SEK 1 dataframe
merge **present**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site1.df=rbind(armla.brmla.df, brmlb.armlb.df)


# Define the range for X and Y (1 to 5)
X_values <- 1:5
Y_values <- 1:5

# Create an empty dataframe to store the results
result_df <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla..", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla..", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb..", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb..", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result <- 
((
  (mass.present.site1.df[mass.present.site1.df$names == armla_name, "values"] +
      mass.present.site1.df[mass.present.site1.df$names == brmlb_name, "values"])/2)/ (
      (mass.present.site1.df[mass.present.site1.df$names == armlb_name, "values"] +
      mass.present.site1.df[mass.present.site1.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
    result_name <- LETTERS[(X - 1) * 5 + Y]
    
    # Get the treatment associated with the current index
  treatment <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Armla.", i), ]$treatment
  

    # Append the name and value to the result dataframe
    result_df <- rbind(result_df, data.frame(names = result_name, values = result, treatment = treatment))
  }
}


# View the resulting dataframe
View(result_df)
```


NEW: use this! merging HOK 2 and SEK 2 dataframe
merge **present**, merging HOK 2 and SEK 2 dataframe
```{r}
#merge both present site 1 dataframes
mass.present.site2.df=rbind(armla.brmla.df_2, brmlb.armlb.df_2)

# Create an empty dataframe to store the results
result_df_2 <- data.frame(names = character(0), values = numeric(0), treatment = character(0))

# Loop through all combinations of X and Y for Armla, Brmla, Brmlb, and Armlb
for (X in X_values) {
  for (Y in Y_values) {
    armla_name <- paste("Armla..", X, ".", Y, sep = "")
    brmla_name <- paste("Brmla..", X, ".", Y, sep = "")
    brmlb_name <- paste("Brmlb..", X, ".", Y, sep = "")
    armlb_name <- paste("Armlb..", X, ".", Y, sep = "")
    
    # Calculate the result for the current combination
    result_2 <- 
((
  (mass.present.site2.df[mass.present.site2.df$names == armla_name, "values"] +
      mass.present.site2.df[mass.present.site2.df$names == brmlb_name, "values"])/2)/ (
      (mass.present.site2.df[mass.present.site2.df$names == armlb_name, "values"] +
      mass.present.site2.df[mass.present.site2.df$names == brmla_name, "values"])/2
    )*100
)-100
     # Create a new name (letter) for the result
result_name_2 <- paste0(LETTERS[(X - 1) * 5 + Y], "_2")

 # Get the treatment associated with the current index
  treatment <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  

    # Append the name and value to the result dataframe
    result_df_2 <- rbind(result_df_2, data.frame(names = result_name_2, values = result_2, treatment = treatment))
  }
}


# View the resulting dataframe
View(result_df_2)

#combine sites 1 and 2
hfai.allsites.df=rbind(result_df, result_df_2)
View(hfai.allsites.df)
```






This is the proper dataframe
merge **present**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site1.df=rbind(mass.hok1.present.df, mass.sek1.present.df)

# Create an empty data frame to store the results
hfai.df <- data.frame(hfai.z = numeric(0), hfai.mass = numeric(0), label = character(0), treatment = character(0))

# Create an empty list to store the data frames
hfai_dfs <- list()

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla.z <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Armla.", i), ]$rml.z
  brmlb.z <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml.z
  armlb.z <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml.z
  brmla.z <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml.z
  
 print(paste("armla.z:", armla.z))
  print(paste("brmlb.z:", brmlb.z))
  print(paste("armlb.z:", armlb.z))
  print(paste("brmla.z:", brmla.z))
  
  armla.mass <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Armla.", i), ]$rml.mass
  brmlb.mass <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml.mass
  armlb.mass <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml.mass
  brmla.mass <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml.mass
 
  
print(paste("armla.mass:", armla.mass))
  print(paste("brmlb.mass:", brmlb.mass))
  print(paste("armlb.mass:", armlb.mass))
  print(paste("brmla.mass:", brmla.mass))


  result.z <- (((armla.z + brmlb.z) / (armlb.z + brmla.z)) * 100) - 100
  result.mass <- (((armla.mass + brmlb.mass) / (armlb.mass + brmla.mass)) * 100) - 100

  print(paste("Length of result.z:", length(result.z)))
  print(paste("Length of result.mass:", length(result.mass)))
  
  # Get the treatment associated with the current index
  treatment <- mass.present.site1.df[mass.present.site1.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Create a data frame with the results, label, and treatment
  hfai_df <- data.frame(hfai.z = result.z, hfai.mass = result.mass, label = LETTERS[i], treatment = treatment)
  
  # Append the data frame to the list
  hfai_dfs[[i]] <- hfai_df
}


# Combine the list of data frames into a single data frame
hfai.df <- do.call(rbind, hfai_dfs)

#HFAI (this is a stop and check to verify that values are correct!)
(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```

This is the proper dataframe
merge **present**, site a and b (aka HOK 2 and SEK 2) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site2.df=rbind(mass.hok2.present.df, mass.sek2.present.df)

# Calculate the index of the last label in the existing hfai.df
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Create an empty list to store the new data frames
hfai_dfs <- list()

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla.z <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armla.", i), ]$rml.z
  brmlb.z <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml.z
  armlb.z <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml.z
  brmla.z <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml.z
  
  armla.mass <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armla.", i), ]$rml.mass
  brmlb.mass <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml.mass
  armlb.mass <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml.mass
  brmla.mass <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml.mass
  
  result.z <- (((armla.z + brmlb.z) / (armlb.z + brmla.z)) * 100) - 100
  result.mass <- (((armla.mass + brmlb.mass) / (armlb.mass + brmla.mass)) * 100) - 100

  # Get the treatment associated with the current index
  treatment <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Create a data frame with the new results, label, and treatment
  hfai_df <- data.frame(hfai.z = result.z, hfai.mass = result.mass, label = next_label, treatment = treatment)
  
  # Append the data frame to the list
  hfai_dfs[[i]] <- hfai_df
}

# Combine the list of new data frames with the existing hfai.df
hfai.df <- rbind(hfai.df, do.call(rbind, hfai_dfs))




```

This is the proper plot
another plot check
```{r}
######## using mass loss

hfai.mass.plot=ggplot(hfai.df, aes(x = reorder(label, hfai.mass), y = hfai.mass, fill = label)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  scale_fill_viridis_d(guide = "none")+
  xlab("reciprocated pairs") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.x = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  )

print(hfai.mass.plot)

hfai.mass.plot

ggsave("figures/hfai.mass.plot.pdf", hfai.mass.plot, width = 14, height = 10)
```

Full Pairwise comparisons for HFAI

```{r}

# Initialize an empty matrix to store the results
armla_matrix <- matrix(NA, nrow = length(tree_num_values), ncol = length(tree_num_values))
brmla_matrix <- matrix(NA, nrow = length(tree_num_values), ncol = length(tree_num_values))

# Loop through the combinations and calculate Armla and Brmla values
for (i in 1:length(tree_num_values)) {
  for (j in 1:length(tree_num_values)) {
    ayres_num <- tree_num_values[i]
    tree_num <- tree_num_values[j]
    
    for (k in 1:length(tree_num_values)) {
      ba_num <- tree_num_values[k]
      
      armla <- ifelse(
        ayres_num == "Aa", mass.hok1.present.df$perc.mass_loss[
          mass.hok1.present.df$ayres.eq == paste("Armla.", ayres_num, ".", ba_num) &
            mass.hok1.present.df$tree_num == tree_num
        ],
        NA
      )
      
      armla_value <- ifelse(
        ayres_num == "Aa", armla / (armla + mass.hok1.present.df$perc.mass_loss[
          mass.hok1.present.df$ayres.eq == paste("Brmla.", ba_num, ".", ayres_num) &
            mass.hok1.present.df$tree_num == tree_num
        ]) * 100,
        NA
      )
      
      armla_matrix[i, k] <- armla_value
      
      brmla <- ifelse(
        ayres_num == "Ba", mass.hok1.present.df$perc.mass_loss[
          mass.hok1.present.df$ayres.eq == paste("Brmla.", ayres_num, ".", ba_num) &
            mass.hok1.present.df$tree_num == tree_num
        ],
        NA
      )
      
      brmla_value <- ifelse(
        ayres_num == "Ba", brmla / (brmla + mass.hok1.present.df$perc.mass_loss[
          mass.hok1.present.df$ayres.eq == paste("Armla.", ba_num, ".", ayres_num) &
            mass.hok1.present.df$tree_num == tree_num
        ]) * 100,
        NA
      )
      
      brmla_matrix[i, k] <- brmla_value
    }
  }
}

# Add the calculated values to the dataframe
for (i in 1:length(tree_num_values)) {
  for (j in 1:length(tree_num_values)) {
    mass.hok1.present.df[paste("Armla.", tree_num_values[i], ".", tree_num_values[j], sep = "")] <- armla_matrix[i, j]
    mass.hok1.present.df[paste("Brmla.", tree_num_values[i], ".", tree_num_values[j], sep = "")] <- brmla_matrix[i, j]
  }
}

# Define the columns to pivot
columns_to_pivot <- names(mass.hok1.present.df)[grepl("^Armla\\.|^Brmla\\.", names(mass.hok1.present.df))]

# Pivot the DataFrame to long format
mass.hok1.present.df_long <- mass.hok1.present.df %>%
  pivot_longer(
    cols = all_of(columns_to_pivot),
    names_to = "Name",
    values_to = "Value"
  ) %>%
  mutate(Name = sub("\\d\\.\\d$", "", Name))  # Remove the last part of the column names

# View the resulting DataFrame
head(mass.hok1.present.df_long)




```









```{r}

##this is the proper plot
########### using z loss
hfai.z.plot=ggplot(hfai.df[hfai.df$hfai.z > -800,], aes(x = reorder(label, hfai.z), y = hfai.z, fill = label)) +
  geom_bar(stat = "identity", fill = "white", color = "black") +
  geom_col_pattern(
    aes(pattern = treatment),
    colour = "black",
    pattern_fill = "black",
    pattern_angle = 45,
    pattern_density = 0.1,
    pattern_spacing = 0.01,
    position = position_dodge2(preserve = 'single'),
    alpha = 0.2  # Adjust the alpha value to control transparency
  ) +
  scale_pattern_manual(
    values = c("none", "stripe"),
    guide = guide_legend(override.aes = list(fill = "grey70")) # <- make lighter
  ) +
  xlab("") +
  ylab("Home-field advantage (% faster decomposition at home)") +
coord_flip(ylim = c(-800,300)) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
  )

hfai.z.plot=ggplot(hfai.df[hfai.df$hfai.z > -800,], aes(x = reorder(label, hfai.z), y = hfai.z, fill = label)) +
  geom_bar(stat = "identity", fill = "white", color = "black") +
  geom_col_pattern(
    aes(pattern = treatment),
    colour = "black",
    pattern_fill = "black",
    pattern_angle = 45,
    pattern_density = 0.1,
    pattern_spacing = 0.01,
    position = position_dodge2(preserve = 'single'),
    alpha = 0.2  # Adjust the alpha value to control transparency
  ) +
  scale_pattern_manual(
    values = c("none", "stripe"),
    guide = guide_legend(override.aes = list(fill = "grey70")) # <- make lighter
  ) +
  xlab("") +
  ylab("Home-field advantage (% faster decomposition at home)") +
coord_cartesian(ylim = c(-800,300)) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
  )

hfai.z.plot

ggsave("figures/hfai.z.plot.pdf", hfai.z.plot, width = 10, height = 6)

```



This is the proper dataframe with for loop
generate **absent**, site a (HOK 1)
```{r}
########################
mass.hok1.absent.df <- mass.hok1.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values_z <- numeric()
rml_values_mass <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_z_loss <- mass.hok1.absent.df$z_score[
      mass.hok1.absent.df$tree_id == "HOK" & mass.hok1.absent.df$tree_num == tree_num
    ]
    sek_z_loss <- mass.hok1.absent.df$z_score[
      mass.hok1.absent.df$tree_id == "SEK" & mass.hok1.absent.df$tree_num == tree_num
    ]
    
    rml.z <- ifelse(
      ayres == "Aa", hok_z_loss / (hok_z_loss + sek_z_loss),
      ifelse(ayres == "Ba", sek_z_loss / (sek_z_loss + hok_z_loss), NA)
    )
    
    rml_values_z <- c(rml_values_z, rml.z * 100)
    
    hok_mass_loss <- mass.hok1.absent.df$perc.mass_loss[
      mass.hok1.absent.df$tree_id == "HOK" & mass.hok1.absent.df$tree_num == tree_num
    ]
    sek_mass_loss <- mass.hok1.absent.df$perc.mass_loss[
      mass.hok1.absent.df$tree_id == "SEK" & mass.hok1.absent.df$tree_num == tree_num
    ]
    
    rml.mass <- ifelse(
      ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
      ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss), NA)
    )
    
    rml_values_mass <- c(rml_values_mass, rml.mass * 100)
  }
}

# Add the rml values to the dataframe
mass.hok1.absent.df$rml.z <- rml_values_z
mass.hok1.absent.df$rml.mass <- rml_values_mass

# Create a new column combining tree_num and ayres with a period in between
mass.hok1.absent.df$ayres.tree <- paste(mass.hok1.absent.df$ayres, mass.hok1.absent.df$tree_num, sep = ".")

# Create a new column based on Ayres HFAI equation 
mass.hok1.absent.df$ayres.eq <- ifelse(
  mass.hok1.absent.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""),
  paste("Armla", tree_num_values, sep = "."),
  ifelse(
    mass.hok1.absent.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""),
    paste("Brmla", tree_num_values, sep = "."),
    NA
  )
)

############### STOP ######################
##check the calculations (Armla.1-Armla.5)
((mass.hok1.present.df[mass.hok1.present.df$ayres.tree == "Aa.1",]$z_score) /
  (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == "Aa.1",]$z_score +
   mass.hok1.present.df[mass.hok1.present.df$ayres.tree == "Ba.1",]$z_score)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$z_score)) * 100


(
  (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == "Aa.3",]$perc.mass_loss) /
  (mass.hok1.present.df[mass.hok1.present.df$ayres.tree == "Aa.3",]$perc.mass_loss +
   mass.hok1.present.df[mass.hok1.present.df$ayres.tree == "Ba.3",]$perc.mass_loss)
  
  * 100) 

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$z_score)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$z_score)) * 100


##check the calculations (Brmla.1-Brmla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$z_score)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$z_score)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$z_score)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$z_score)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$z_score) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$z_score +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$z_score)) * 100


```


This is also the proper dataframe
generate **absent**, site b (SEK 1)
```{r}
mass.sek1.absent.df <- mass.sek1.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values_z <- numeric()
rml_values_mass <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_z_loss <- mass.sek1.absent.df$z_score[
      mass.sek1.absent.df$tree_id == "HOK" & mass.sek1.absent.df$tree_num == tree_num
    ]
    sek_z_loss <- mass.sek1.present.df$z_score[
      mass.sek1.absent.df$tree_id == "SEK" & mass.sek1.absent.df$tree_num == tree_num
    ]
    
    rml.z <- ifelse(
      ayres == "Ab", hok_z_loss / (hok_z_loss + sek_z_loss),
      ifelse(ayres == "Bb", sek_z_loss / (sek_z_loss + hok_z_loss), NA)
    )
    
    rml_values_z <- c(rml_values_z, rml.z * 100)
    
    hok_mass_loss <- mass.sek1.absent.df$perc.mass_loss[
      mass.sek1.absent.df$tree_id == "HOK" & mass.sek1.absent.df$tree_num == tree_num
    ]
    sek_mass_loss <- mass.sek1.absent.df$perc.mass_loss[
      mass.sek1.absent.df$tree_id == "SEK" & mass.sek1.absent.df$tree_num == tree_num
    ]
    
    rml.mass <- ifelse(
      ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
      ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss), NA)
    )
    
    rml_values_mass <- c(rml_values_mass, rml.mass * 100)
  }
}

# Add the rml values to the dataframe
mass.sek1.absent.df$rml.z <- rml_values_z
mass.sek1.absent.df$rml.mass <- rml_values_mass

# Create a new column combining tree_num and ayres with a period in between
mass.sek1.absent.df$ayres.tree <- paste(mass.sek1.absent.df$ayres, mass.sek1.absent.df$tree_num, sep = ".")

# Create a new column based on Ayres HFAI equation 
mass.sek1.absent.df$ayres.eq <- ifelse(
  mass.sek1.absent.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""),
  paste("Armlb", tree_num_values, sep = "."),
  ifelse(
    mass.sek1.absent.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""),
    paste("Brmlb", tree_num_values, sep = "."),
    NA
  )
)


```



This is also the proper dataframe
generate **absent**, site a (HOK 2)
```{r}
mass.hok2.absent.df <- mass.hok2.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values_z <- numeric()
rml_values_mass <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_z_loss <- mass.hok2.absent.df$z_score[
      mass.hok2.absent.df$tree_id == "HOK" & mass.hok2.absent.df$tree_num == tree_num
    ]
    sek_z_loss <- mass.hok2.absent.df$z_score[
      mass.hok2.absent.df$tree_id == "SEK" & mass.hok2.absent.df$tree_num == tree_num
    ]
    
    rml.z <- ifelse(
      ayres == "Aa", hok_z_loss / (hok_z_loss + sek_z_loss),
      ifelse(ayres == "Ba", sek_z_loss / (sek_z_loss + hok_z_loss), NA)
    )
    
    rml_values_z <- c(rml_values_z, rml.z * 100)
    
    hok_mass_loss <- mass.hok2.absent.df$perc.mass_loss[
      mass.hok2.absent.df$tree_id == "HOK" & mass.hok2.absent.df$tree_num == tree_num
    ]
    sek_mass_loss <- mass.hok2.absent.df$perc.mass_loss[
      mass.hok2.absent.df$tree_id == "SEK" & mass.hok2.absent.df$tree_num == tree_num
    ]
    
    rml.mass <- ifelse(
      ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
      ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss), NA)
    )
    
    rml_values_mass <- c(rml_values_mass, rml.mass * 100)
  }
}

# Add the rml values to the dataframe
mass.hok2.absent.df$rml.z <- rml_values_z
mass.hok2.absent.df$rml.mass <- rml_values_mass

# Create a new column combining tree_num and ayres with a period in between
mass.hok2.absent.df$ayres.tree <- paste(mass.hok2.absent.df$ayres, mass.hok2.absent.df$tree_num, sep = ".")

# Create a new column based on Ayres HFAI equation 
mass.hok2.absent.df$ayres.eq <- ifelse(
  mass.hok2.absent.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""),
  paste("Armla", tree_num_values, sep = "."),
  ifelse(
    mass.hok2.absent.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""),
    paste("Brmla", tree_num_values, sep = "."),
    NA
  )
)


```


This is also the proper dataframe
generate **absent**, site a (SEK 2)
```{r}
mass.sek2.absent.df <- mass.sek2.absent.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values_z <- numeric()
rml_values_mass <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_z_loss <- mass.sek2.absent.df$z_score[
      mass.sek2.absent.df$tree_id == "HOK" & mass.sek2.absent.df$tree_num == tree_num
    ]
    sek_z_loss <- mass.sek2.absent.df$z_score[
      mass.sek2.absent.df$tree_id == "SEK" & mass.sek2.absent.df$tree_num == tree_num
    ]
    
    rml.z <- ifelse(
      ayres == "Ab", hok_z_loss / (hok_z_loss + sek_z_loss),
      ifelse(ayres == "Bb", sek_z_loss / (sek_z_loss + hok_z_loss), NA)
    )
    
    rml_values_z <- c(rml_values_z, rml.z * 100)
    
    hok_mass_loss <- mass.sek2.absent.df$perc.mass_loss[
      mass.sek2.absent.df$tree_id == "HOK" & mass.sek2.absent.df$tree_num == tree_num
    ]
    sek_mass_loss <- mass.sek2.absent.df$perc.mass_loss[
      mass.sek2.absent.df$tree_id == "SEK" & mass.sek2.absent.df$tree_num == tree_num
    ]
    
    rml.mass <- ifelse(
      ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
      ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss), NA)
    )
    
    rml_values_mass <- c(rml_values_mass, rml.mass * 100)
  }
}

# Add the rml values to the dataframe
mass.sek2.absent.df$rml.z <- rml_values_z
mass.sek2.absent.df$rml.mass <- rml_values_mass

# Create a new column combining tree_num and ayres with a period in between
mass.sek2.absent.df$ayres.tree <- paste(mass.sek2.absent.df$ayres, mass.sek2.absent.df$tree_num, sep = ".")

# Create a new column based on Ayres HFAI equation 
mass.sek2.absent.df$ayres.eq <- ifelse(
  mass.sek2.absent.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""),
  paste("Armlb", tree_num_values, sep = "."),
  ifelse(
    mass.sek2.absent.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""),
    paste("Brmlb", tree_num_values, sep = "."),
    NA
  )
)


```

This is the proper dataframe
merge **absent**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.absent.site1.df=rbind(mass.hok1.absent.df, mass.sek1.absent.df)

# Calculate the index of the last label in the existing hfai.df
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Create an empty list to store the new data frames
hfai_dfs <- list()

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla.z <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Armla.", i), ]$rml.z
  brmlb.z <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml.z
  armlb.z <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml.z
  brmla.z <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml.z
  
  armla.mass <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Armla.", i), ]$rml.mass
  brmlb.mass <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml.mass
  armlb.mass <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml.mass
  brmla.mass <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml.mass
  
  result.z <- (((armla.z + brmlb.z) / (armlb.z + brmla.z)) * 100) - 100
  result.mass <- (((armla.mass + brmlb.mass) / (armlb.mass + brmla.mass)) * 100) - 100

  # Get the treatment associated with the current index
  treatment <- mass.absent.site1.df[mass.absent.site1.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Create a data frame with the new results, label, and treatment
  hfai_df <- data.frame(hfai.z = result.z, hfai.mass = result.mass, label = next_label, treatment = treatment)
  
  # Append the data frame to the list
  hfai_dfs[[i]] <- hfai_df
}

# Combine the list of new data frames with the existing hfai.df
hfai.df <- rbind(hfai.df, do.call(rbind, hfai_dfs))




```


This is the proper dataframe
merge **absent**, site a and b (aka HOK 2 and SEK 2) dataframes 
```{r}
#merge both present site 1 dataframes
mass.absent.site2.df=rbind(mass.hok2.absent.df, mass.sek2.absent.df)

# Calculate the index of the last label in the existing hfai.df
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Create an empty list to store the new data frames
hfai_dfs <- list()

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla.z <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Armla.", i), ]$rml.z
  brmlb.z <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml.z
  armlb.z <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml.z
  brmla.z <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml.z
  
  armla.mass <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Armla.", i), ]$rml.mass
  brmlb.mass <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml.mass
  armlb.mass <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml.mass
  brmla.mass <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml.mass
  
  result.z <- (((armla.z + brmlb.z) / (armlb.z + brmla.z)) * 100) - 100
  result.mass <- (((armla.mass + brmlb.mass) / (armlb.mass + brmla.mass)) * 100) - 100

  # Get the treatment associated with the current index
  treatment <- mass.absent.site2.df[mass.absent.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Create a data frame with the new results, label, and treatment
  hfai_df <- data.frame(hfai.z = result.z, hfai.mass = result.mass, label = next_label, treatment = treatment)
  
  # Append the data frame to the list
  hfai_dfs[[i]] <- hfai_df
}

# Combine the list of new data frames with the existing hfai.df
hfai.df <- rbind(hfai.df, do.call(rbind, hfai_dfs))




```



This is the proper plot
another plot check
```{r}
######## using mass loss

hfai.mass.all.plot=ggplot(hfai.df, aes(x = reorder(label, hfai.mass), y = hfai.mass, fill = treatment)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  scale_fill_viridis_d()+
  xlab("reciprocated pairs") +
  ylab("home-field advantage (% faster decomposition at home)") +
  theme_minimal() +
  theme(text = element_text(size =20), 
    axis.text.x = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  )

hfai.mass.all.plot


ggsave("figures/hfai.mass.all.plot.pdf", hfai.mass.all.plot, width = 14, height = 10)
```










generate **absent**, site a (HOK 1)
```{r}
mass.ab.hoko1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="HOK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.hoko1.df <- mass.ab.hoko1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "HOK" & mass.ab.hoko1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "SEK" & mass.ab.hoko1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}

##check the calculations (Armla.1-Armla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss)) * 100


##check the calculations (Brmla.1-Brmla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss)) * 100

# Add the rml values to the dataframe
mass.ab.hoko1.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.hoko1.df$ayres.tree <- paste(mass.ab.hoko1.df$ayres, mass.ab.hoko1.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.hoko1.df$ayres.eq <- ifelse(mass.ab.hoko1.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **absent**, site b (SEK 1)
```{r}
mass.ab.sek1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="SEK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.sek1.df <- mass.ab.sek1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)


###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "HOK" & mass.ab.sek1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "SEK" & mass.ab.sek1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.sek1.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.sek1.df$ayres.tree <- paste(mass.ab.sek1.df$ayres, mass.ab.sek1.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.sek1.df$ayres.eq <- ifelse(mass.ab.sek1.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.p.sek1.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **asbent**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.ab.site1.df=rbind(mass.ab.hoko1.df, mass.ab.sek1.df)

# Calculate the last letter index based on the existing data frame
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = next_label, treatment = treatment))
}


#HFAI (this is a stop and check to verify that values are correct!)
(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```

plot check
```{r}

ggplot(hfai.df, aes(x = label, y = hfai)) +
  geom_bar(stat = "identity",
           show.legend = FALSE) + # Remove the legend
  facet_grid(~treatment)+
  xlab("Group") +
  ylab("Value")+
  coord_flip()




```



generate **present**, site a (HOK 2)
```{r}
#now create a new column to have some reference to the Ayres et al. (2009) paper (species A= hoko trees, species B = sekiu trees)
mass.hok2.present.df <- mass.hok2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 2-1 with SEK 2-1; this follows similar logic as Ayres et al. (2009) Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.hok2.present.df$z_score[mass.hok2.present.df$tree_id == "HOK" & mass.hok2.present.df$tree_num == tree_num]
    sek_mass_loss <- mass.hok2.present.df$z_score[mass.hok2.present.df$tree_id == "SEK" & mass.hok2.present.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.hok2.present.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.hok2.present.df$ayres.tree <- paste(mass.hok2.present.df$ayres, mass.hok2.present.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.hok2.present.df$ayres.eq <- ifelse(mass.hok2.present.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.hok2.present.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **present**, site b (SEK 2)
```{r}
#now create a new column to have some reference to the Ayres et al. (2009) paper (species A= hoko trees, species B = sekiu trees)
mass.sek2.present.df <- mass.sek2.present.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)


###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 2-1 with SEK 2-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.sek2.present.df$z_score[mass.sek2.present.df$tree_id == "HOK" & mass.sek2.present.df$tree_num == tree_num]
    sek_mass_loss <- mass.sek2.present.df$z_score[mass.sek2.present.df$tree_id == "SEK" & mass.p.sek2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.sek2.present.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.sek2.present.df$ayres.tree <- paste(mass.sek2.present.df$ayres, mass.sek2.present.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.sek2.present.df$ayres.eq <- ifelse(mass.sek2.present.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.sek2.present.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **present**, site a and b (aka HOK 2 and SEK 2) dataframes 
```{r}
#merge both present site 1 dataframes
mass.present.site2.df=rbind(mass.hok2.present.df, mass.sek2.present.df)

# Calculate the last letter index based on the existing data frame
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.present.site2.df[mass.present.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = next_label, treatment = treatment))
}

#HFAI (this is a stop and check to verify that values are correct!)
(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```


## Now we'll tackle absent:
generate **absent**, site a (HOK 2)
```{r}
mass.ab.hoko2.df=mass.df%>%
  filter(treatment == "AB" & deploy=="HOK" & site_num=="2")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.hoko2.df <- mass.ab.hoko2.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.hoko2.df$mass_loss[mass.ab.hoko2.df$tree_id == "HOK" & mass.ab.hoko2.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.hoko2.df$mass_loss[mass.ab.hoko2.df$tree_id == "SEK" & mass.ab.hoko2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.hoko2.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.hoko2.df$ayres.tree <- paste(mass.ab.hoko2.df$ayres, mass.ab.hoko2.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.hoko2.df$ayres.eq <- ifelse(mass.ab.hoko2.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **absent**, site b (SEK 1)
```{r}
mass.ab.sek2.df=mass.df%>%
  filter(treatment == "AB" & deploy=="SEK" & site_num=="2")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.sek2.df <- mass.ab.sek2.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.sek2.df$mass_loss[mass.ab.sek2.df$tree_id == "HOK" & mass.ab.sek2.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.sek2.df$mass_loss[mass.ab.sek2.df$tree_id == "SEK" & mass.ab.sek2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.sek2.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.sek2.df$ayres.tree <- paste(mass.ab.sek2.df$ayres, mass.ab.sek2.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.sek2.df$ayres.eq <- ifelse(mass.ab.sek2.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.p.sek1.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **asbent**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.ab.site2.df=rbind(mass.ab.hoko2.df, mass.ab.sek2.df)

# Calculate the last letter index based on the existing data frame
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = next_label, treatment = treatment))
}

#HFAI (this is a stop and check to verify that values are correct!)
(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```




















```{r}
########### Next we'll try out absent, hoko 1
mass.ab.hoko1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="HOK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (Home= Aa, Away= Bb)
mass.ab.hoko1.df <- mass.ab.hoko1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "HOK" & mass.ab.hoko1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "SEK" & mass.ab.hoko1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.hoko1.df$rml <- rml_values







########### Next we'll try out absent, sekiu 1
mass.ab.sek1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="SEK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (Home= Aa, Away= Bb)
mass.ab.sek1.df <- mass.ab.sek1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "HOK" & mass.ab.sek1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "SEK" & mass.ab.sek1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.sek1.df$rml <- rml_values


#merge dataframes
mass.site1.df=rbind(mass.p.hoko1.df, mass.ab.hoko1.df, mass.p.sek1.df, mass.ab.sek1.df )



```




Ayres HFA index
```{r}
#the calculations for arml and brml are different, dependent on "relative to"

##Site 1: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site a (Hoko) 
site1.armla.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="1" & mass.df$deploy == "HOK", ])
#mass loss calculation
site1.armla.df= site1.armla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.armla.df <- site1.armla.df %>%
  mutate(rml = ((mass_loss[tree_id == "HOK" & deploy == "HOK"]) / 
                 (mass_loss[tree_id == "HOK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "HOK"])) * 100)

##Site 2: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site a (Hoko) 
site2.armla.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="2" & mass.df$deploy == "HOK", ])
#mass loss calculation
site2.armla.df= site2.armla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.armla.df <- site2.armla.df %>%
  mutate(rml = (mass_loss[tree_id == "HOK" & deploy == "HOK"] / 
                 (mass_loss[tree_id == "HOK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "HOK"])) * 100)

##Site 1: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site b (Sekiu) 
site1.armlb.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="2" & mass.df$deploy == "SEK", ])
#mass loss calculation
site1.armlb.df= site1.armlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.armlb.df <- site1.armlb.df %>%
  mutate(rml = (mass_loss[tree_id == "HOK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "HOK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "SEK"])) * 100)

##Site 2: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site b (Sekiu) 
site2.armlb.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="2" & mass.df$deploy == "SEK", ])
#mass loss calculation
site2.armlb.df= site2.armlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.armlb.df <- site2.armlb.df %>%
  mutate(rml = (mass_loss[tree_id == "HOK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "HOK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "SEK"])) * 100)

##Site 1: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site a (Hoko) 
site1.brmla.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="1" & mass.df$deploy == "HOK", ])
#mass loss calculation
site1.brmla.df= site1.brmla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.brmla.df <- site1.brmla.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "HOK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "HOK"])) * 100)

##Site 2: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site a (Hoko) 
site2.brmla.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="2" & mass.df$deploy == "HOK", ])
#mass loss calculation
site2.brmla.df= site2.brmla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.brmla.df <- site2.brmla.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "HOK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "HOK"])) * 100)

##Site 1: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site b (Sekiu) 
site1.brmlb.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="1" & mass.df$deploy == "SEK", ])
#mass loss calculation
site1.brmlb.df= site1.brmlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.brmlb.df <- site1.brmlb.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "SEK"])) * 100)

##Site 2: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site b (Sekiu) 
site2.brmlb.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="2" & mass.df$deploy == "SEK", ])
#mass loss calculation
site2.brmlb.df= site2.brmlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.brmlb.df <- site2.brmlb.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "SEK"])) * 100)

#merge dataframes
site1.hfa_index.df=rbind(site1.armla.df, site1.armlb.df, site1.brmla.df, site1.brmlb.df)
#HFA index calculation
site1.hfa_index.df <- site1.hfa_index.df %>%
  mutate(hfai = (((site1.armla.df$rml + site1.brmlb.df$rml) / 2) / ((site1.armlb.df$rml + site1.brmla.df$rml) / 2)) * 100 - 100)

#merge dataframes
site2.hfa_index.df=rbind(site2.armla.df, site2.armlb.df, site2.brmla.df, site2.brmlb.df)
#HFA index calculation
site2.hfa_index.df <- site2.hfa_index.df %>%
  mutate(hfai = (((site2.armla.df$rml + site2.brmlb.df$rml) / 2) / ((site2.armlb.df$rml + site2.brmla.df$rml) / 2)) * 100 - 100)

# Create the horizontal bar plot
ggplot(hfa_index.df, aes(x = hfai, y = interaction(deploy, hfa))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(title = "HFA Index Comparison",
       x = "HFA Index") +
  theme_minimal() +
  theme(legend.position = "none") +
  geom_vline(xintercept = 0, color = "black", size = 1) +
  scale_fill_manual(values = c("a" = "blue", "b" = "red")) 



```







