---
title: "RAD Alder 2023"
author: "Cody Spiegel"
date: "`r Sys.Date()`"
output: word_document
---

### Project overview
For Project ‘RAD’ Alder, gamma irradiated vs. non-irradiated leaf packs deployed at “home” or “away” sites tests the role of phyllosphere microbial communities on home-field advantage patterns in leaf litter decomposition. We utilized a reciprocal transplant experiment between the Hoko and Sekiu rivers on the Olympic Peninsula under four treatment conditions: phyllosphere present-home (PH), phyllosphere present-away (PAW), phyllosphere absent-home (ABH), phyllosphere absent-away (ABAW) to ask the following questions: 

1) How does dispersal from the regional species pool (i.e., source communities) influence community composition on the leaf packs?
2) How does microbial community composition (alpha and beta diversity) and function differ between the leaf packs, as well as change over the course of decomposition?
3) How does alder stoichiometric and secondary chemical composition differ between the leaf packs, as well as change over the course of decomposition?
4) How does microbial metabolization differ between the leaf packs, as well as change over the course of decomposition?
5) How does particulate leaf mass loss differ between the leaf packs? 


**DATA SETS**  
This data set focuses on:  

(1) **metabolism assay** 
(2) **Leaf mass loss** using % mass loss  
(3) **Alder C and N isotopic composotion**
(4) **Alder secondary chemistry (mass speci)** 

```{r setup chunk, setup, include = FALSE, cache=FALSE, message=FALSE, warning=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

# use pacman to load all the packages you are missing!
pacman::p_load('knitr', 'lme4', 'lmerTest', 'tidyverse', 'magrittr', 'effects', 'plyr', 'dplyr', 
               'plotrix', 'car',"gridExtra", "cowplot", "tools", "mgcv", "gratia", "tidymv")

Fig.formatting<-(theme_classic()) +
  theme(text=element_text(size=25),
        legend.text.align = 0,
        panel.border = element_rect(fill=NA, colour = "black", size=1),
        aspect.ratio=1, 
        axis.ticks.length=unit(0.25, "cm"),
        axis.text.y=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=20), 
        axis.text.x=element_text(
        margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), angle = 45, colour="black", size=20)) +
  theme(legend.key.size = unit(0.4, "cm"), legend.position = "right") +
  theme(aspect.ratio=1.3) 




```


All metabolite decomp days
```{r}
meta.day1.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_1_metabolites_raw_corrected.csv")


meta.day3.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_3_metabolites_raw_corrected.csv")


meta.day5.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_5_metabolites_raw_corrected.csv")


meta.day7.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_7_metabolites_raw_corrected.csv")


meta.day11.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_11_metabolites_raw_corrected.csv")


meta.day17.raw.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/assay/day_17_metabolites_raw_corrected.csv")



#By z-scoring the raw OD values, I can effectively assess the relative changes in microbial metabolism across the wells, regardless of their absolute starting values.

#z-scoring is a reasonable choice for this situation, as it standardizes the data within each well and facilitates comparisons. Note: z-scoring assumes a roughly symmetric distribution, but the raw OD values are right skewed.

```


metabolite assay: z score by metabolite and deployment 
```{r}
#z score
##hok 1
meta.hok1.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metabolite=="HOK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])
meta.hok1.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metaboltie=="SEK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])


meta.hok1.hok.df$z_score= (meta.hok1.hok.df$od_corrected_b_n - mean(meta.hok1.hok.df$od_corrected_b_n)) / sd(meta.hok1.hok.df$od_corrected_b_n)
meta.hok1.sek.df$z_score= (meta.hok1.sek.df$od_corrected_b_n - mean(meta.hok1.sek.df$od_corrected_b_n)) / sd(meta.hok1.sek.df$od_corrected_b_n)

##sek 1
meta.sek1.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metabolite=="SEK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])
meta.sek1.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metaboltie=="HOK" & meta.day1.raw.df$site_num == "1" & meta.day1.raw.df$time=="30", ])

meta.sek1.sek.df$z_score= (meta.sek1.sek.df$od_corrected_b_n - mean(meta.sek1.sek.df$od_corrected_b_n)) / sd(meta.sek1.sek.df$od_corrected_b_n)
meta.sek1.hok.df$z_score= (meta.sek1.hok.df$od_corrected_b_n - mean(meta.sek1.hok.df$od_corrected_b_n)) / sd(meta.sek1.hok.df$od_corrected_b_n)

#hok 2 
meta.hok2.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metabolite=="HOK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])
meta.hok2.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "HOK" & meta.day1.raw.df$metaboltie=="SEK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])

meta.hok2.hok.df$z_score= (meta.hok2.hok.df$od_corrected_b_n - mean(meta.hok2.hok.df$od_corrected_b_n)) / sd(meta.hok2.hok.df$od_corrected_b_n)
meta.hok2.sek.df$z_score= (meta.hok2.sek.df$od_corrected_b_n - mean(meta.hok2.sek.df$od_corrected_b_n)) / sd(meta.hok2.sek.df$od_corrected_b_n)


##sek 2
meta.sek2.sek.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metabolite=="SEK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])
meta.sek2.hok.df<- data.frame(meta.day1.raw.df[meta.day1.raw.df$deploy == "SEK" & meta.day1.raw.df$metaboltie=="HOK" & meta.day1.raw.df$site_num == "2" & meta.day1.raw.df$time=="30", ])

meta.sek2.sek.df$z_score= (meta.sek2.sek.df$od_corrected_b_n - mean(meta.sek2.sek.df$od_corrected_b_n)) / sd(meta.sek2.sek.df$od_corrected_b_n)
meta.sek2.hok.df$z_score= (meta.sek2.hok.df$od_corrected_b_n - mean(meta.sek2.hok.df$od_corrected_b_n)) / sd(meta.sek2.hok.df$od_corrected_b_n)

# Assuming you have datasets named meta.day1.raw.df, meta.day3.raw.df, ..., meta.day17.raw.df
# Put them into a list
metabolite_list <- list(meta.day1.raw.df, meta.day3.raw.df, meta.day5.raw.df, meta.day7.raw.df, meta.day11.raw.df, meta.day17.raw.df)

# Define the conditions for subsetting
meta.time.points <- c("30", "120", "960", "1440", "2880", "4320")
# Define the actual day numbers corresponding to each dataset
actual_day_numbers <- c(1, 3, 5, 7, 11, 17)

# Create an empty list to store subsets
subset_list <- list()

# Loop through datasets
for (i in seq_along(metabolite_list)) {
  # Access each dataset from the metabolite_list
  meta.merge <- metabolite_list[[i]]
  
  # Create a list to store subsets for each day
  day_subsets <- list()
  
  # Loop through time points
  for (time_point in meta.time.points) {
    # Subset the data based on specific conditions
    subset_data <- meta.merge[meta.merge$deploy == "HOK" &
                             meta.merge$metabolite == "HOK" &
                             meta.merge$site_num == "1" & 
                             meta.merge$time == time_point, ]
    
    # Calculate the z-scores within each subset_data
    subset_data$z_score <- (subset_data$od_corrected_b_n - mean(subset_data$od_corrected_b_n)) / sd(subset_data$od_corrected_b_n)
    
    # Store the subset with z-scores in the day_subsets list
    day_subsets[[time_point]] <- subset_data
  }
  
  # Merge all time points for each day into one dataframe
  merged_data <- bind_rows(day_subsets)
  
  # Store the merged dataframe in the subset_list
  day_number <- actual_day_numbers[i]
  subset_list[[paste0("day", day_number)]] <- merged_data
}

# Rename the list elements to indicate the day
day_names <- paste0("day", actual_day_numbers)
names(subset_list) <- day_names

#rename into new dataframes
meta.day1.df=subset_list$day1
meta.day3.df=subset_list$day3
meta.day5.df=subset_list$day5
meta.day7.df=subset_list$day7
meta.day11.df=subset_list$day11
meta.day17.df=subset_list$day17


```


Day 1: plots
```{r}
#color blind palette
library(viridis) 

day1.z.30.plot=ggplot(meta.day1.df[meta.day1.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
   scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day1.z.120.plot=ggplot(meta.day1.df[meta.day1.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.960.plot=ggplot(meta.day1.df[meta.day1.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.1440.plot=ggplot(meta.day1.df[meta.day1.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day1.z.2880.plot=ggplot(meta.day1.df[meta.day1.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day1.z.4320.plot=ggplot(meta.day1.df[meta.day1.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Create the grid with modified plots
extract.legend <- get_legend(
  day1.z.4320.plot + theme(legend.box.margin = margin(1, 1, 1, 1, "cm"),
                       legend.key.size = unit(50, "pt"),
                       legend.text = element_text(size=20),
                       legend.title= element_blank())
)


# Combine plots
all.z.day1.plot <- plot_grid(
  day1.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day1.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day1.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


all.z.day1.plot
# Set the overall size of the combined plot
ggsave("figures/all.z.day1.plot.pdf", all.z.day1.plot, width = 49, height = 10)



```


Day 3: plots
```{r}

day3.z.30.plot=ggplot(meta.day3.df[meta.day3.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
  scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day3.z.120.plot=ggplot(meta.day3.df[meta.day3.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
     scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.960.plot=ggplot(meta.day3.df[meta.day3.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
     scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.1440.plot=ggplot(meta.day3.df[meta.day3.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day3.z.2880.plot=ggplot(meta.day3.df[meta.day3.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day3.z.4320.plot=ggplot(meta.day3.df[meta.day3.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


# Combine plots
all.z.day3.plot <- plot_grid(
  day3.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day3.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day3.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day3.plot.pdf", all.z.day3.plot, width = 49, height = 10)



```



Day 5: plots
```{r}

day5.z.30.plot=ggplot(meta.day5.df[meta.day5.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day5.z.120.plot=ggplot(meta.day5.df[meta.day5.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


day5.z.960.plot=ggplot(meta.day5.df[meta.day5.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



day5.z.1440.plot=ggplot(meta.day5.df[meta.day5.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
          scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



day5.z.2880.plot=ggplot(meta.day5.df[meta.day5.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


day5.z.4320.plot=ggplot(meta.day5.df[meta.day5.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 



# Combine plots
all.z.day5.plot <- plot_grid(
  day5.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day5.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day5.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day5.plot.pdf", all.z.day5.plot, width = 49, height = 10)



```



Day 7: plots
```{r}

day7.z.30.plot=ggplot(meta.day7.df[meta.day7.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
   scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day7.z.30.plot

day7.z.120.plot=ggplot(meta.day7.df[meta.day7.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
     scale_fill_viridis(discrete = TRUE) + 
 geom_hline(data = meta.day7.df[meta.day7.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.120.plot

day7.z.960.plot=ggplot(meta.day7.df[meta.day7.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.960.plot


day7.z.1440.plot=ggplot(meta.day7.df[meta.day7.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.1440.plot

day7.z.2880.plot=ggplot(meta.day7.df[meta.day7.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.2880.plot

day7.z.4320.plot=ggplot(meta.day7.df[meta.day7.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
      scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


# Combine plots
all.z.day7.plot <- plot_grid(
  day7.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day7.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day7.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)


# Set the overall size of the combined plot
ggsave("figures/all.z.day7.plot.pdf", all.z.day7.plot, width = 49, height = 12)



```



Day 11: z scored
```{r}
day11.z.30.plot=ggplot(meta.day11.df[meta.day11.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day11.z.30.plot

day11.z.120.plot=ggplot(meta.day11.df[meta.day11.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day11.z.120.plot

day11.z.960.plot=ggplot(meta.day11.df[meta.day11.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.960.plot


day11.z.1440.plot=ggplot(meta.day11.df[meta.day11.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day7.z.1440.plot

day11.z.2880.plot=ggplot(meta.day11.df[meta.day11.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day5.z.2880.plot

day11.z.4320.plot=ggplot(meta.day11.df[meta.day11.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


# Combine plots
all.z.day11.plot <- plot_grid(
  day11.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day11.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day11.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14, 14, 14, 14,3)
)

# Set the overall size of the combined plot
ggsave("figures/all.z.day11.plot.pdf", all.z.day11.plot, width = 49, height = 12)



```





Day 17: z scored
```{r}

day17.z.30.plot=ggplot(meta.day17.df[meta.day17.df$time=="30",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
        scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="30",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("30 min")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day17.z.30.plot

day17.z.120.plot=ggplot(meta.day17.df[meta.day17.df$time=="120",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+ 
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="120",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
  ggtitle("2 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day17.z.120.plot

day17.z.960.plot=ggplot(meta.day17.df[meta.day17.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("16 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 

day17.z.960.plot


day17.z.1440.plot=ggplot(meta.day17.df[meta.day17.df$time=="1440",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
         scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="1440",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("24 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting 


day17.z.1440.plot

day17.z.2880.plot=ggplot(meta.day17.df[meta.day17.df$time=="2880",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="2880",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("48 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting

day17.z.2880.plot

day17.z.4320.plot=ggplot(meta.day17.df[meta.day17.df$time=="4320",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
       scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day17.df[meta.day17.df$time=="4320",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("72 hours")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting


# Combine plots
all.z.day17.plot <- plot_grid(
  day17.z.30.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25)),
  day17.z.120.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.960.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.1440.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  day17.z.2880.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
   day17.z.4320.plot + theme(legend.position = "none", plot.margin = margin(1, 1, 1, 1, "cm"), plot.title = element_text(size= 25), axis.title.y = element_blank()),
  extract.legend,
  ncol = 7,
  rel_widths = c(14, 14, 14,14,14, 14, 3)
)

# Set the overall size of the combined plot
ggsave("figures/all.z.day17.plot.pdf", all.z.day17.plot, width = 49, height = 12)


```

All metabolite assays plots
```{r}

all.z.days.plot <- plot_grid(
  all.z.day1.plot,
  all.z.day3.plot,
  all.z.day5.plot,
  all.z.day7.plot,
  all.z.day11.plot,
  all.z.day17.plot,
  nrow = 6
)



all.z.days.plot
ggsave("figures/all.z.days.plot.pdf", all.z.days.plot, width = 49, height = 30)

```


Metabolite analysis
```{r}

#make a big metabolite dataframe

merged_df=rbind(meta.day1.raw.df,meta.day3.raw.df,meta.day5.raw.df,meta.day7.raw.df, meta.day11.raw.df)
class(merged_df$time)
merged_df$treatment=as.factor(as.character(merged_df$treatment))
merged_df$tree_id=as.factor(as.character(merged_df$tree_id))
merged_df$tree_num=as.factor(as.integer(merged_df$tree_num))
merged_df$deploy=as.factor(as.character(merged_df$deploy))
merged_df$site_num=as.factor(as.integer(merged_df$site_num))
merged_df$metabolite=as.factor(as.character(merged_df$metabolite))
merged_df$hfa=as.factor(as.character(merged_df$hfa))
merged_df$time=as.numeric(as.integer(merged_df$time))

class(merged_df$metabolite)

# Fit the linear mixed-effects model with nested random effects and fixed effect
model <- lmer(z_score ~ treatment + hfa + treatment * hfa +
              (1 | deploy/tree_id/metabolite),
              data = merged_df[merged_df$decomp_day == "7" & merged_df$time=="4320", ],
              na.action = na.exclude)
anova(model)






```
16 hr metabolite plots
```{r}

day1.16hr.plot=ggplot(meta.day1.df[meta.day1.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "", y = "Optical density (z-scored)") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day1.df[meta.day1.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 1")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting +
   theme(legend.position = "none")

day1.16hr.plot

day3.16hr.plot=ggplot(meta.day3.df[meta.day3.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day3.df[meta.day3.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 3")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting +
   theme(legend.position = "none")


day5.16hr.plot=ggplot(meta.day5.df[meta.day5.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "Treatment", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day5.df[meta.day5.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 5")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting+
   theme(legend.position = "none")

day5.16hr.plot

day7.16hr.plot=ggplot(meta.day7.df[meta.day7.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "Treatment", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day7.df[meta.day7.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 5")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting+
   theme(legend.position = "none")

day11.16hr.plot=ggplot(meta.day11.df[meta.day11.df$time=="960",], aes(x = treatment, y = z_score, fill = factor(hfa))) +
  geom_boxplot() +
  coord_cartesian(ylim = c(-1.5,2))+
  labs(x = "", y ="") +
  theme(legend.position = "right")+
    scale_fill_viridis(discrete = TRUE) +
 geom_hline(data = meta.day11.df[meta.day11.df$time=="960",],
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted")+
    ggtitle("Day 11")+
  theme(plot.title = element_text(size = 12))+
  Fig.formatting+
  theme(legend.position = "none")

day11.16hr.plot

all.16hr.z.plots=plot_grid(
  day1.16hr.plot,
  day3.16hr.plot,
  day5.16hr.plot,
  day11.16hr.plot,
  ncol=5,
  extract.legend,
  rel_widths = c(8,8,8,8,3))


all.day16.z.plots
ggsave("figures/all.16hr.z.plots.pdf", all.16hr.z.plots, width = 26, height = 8)


```
Background information:

The Home Field Advantage Index (HFAI) derived from Ayres et al. (2009) uses the relative mass loss values (rml) of species A and B at different sites. The HFAI therefore represents the difference in the rate of litter decomposition between two different sites ("Hoko" and "Sekiu") for both species A and B. Species A and B, in this case, are alder trees of different origin. For example, species A is Hoko alder and species B is Sekiu alder. There are 10 alder trees per river, representing biological replicates. So, the HFAI quantifies the *advantage* or *disadvantage* of each "species" in their respective home environment compared to an away environment.

Interpreting HFAI values involves considering their sign and magnitude:

**Positive** HFAI Values:
Positive values indicate that both species A and B experience faster decomposition rates at their home site compared to the away site. This suggests that both species are better adapted to their home environment in terms of litter decomposition.

**Negative** HFAI Values:
Negative values indicate that both species A and B experience slower decomposition rates at their home site compared to the away site. This suggests that both species might not be well-suited for decomposition in their own environment and may decompose more efficiently in a different environment.

**Magnitude** of HFAI:
The magnitude of the HFAI values indicates the strength of the advantage or disadvantage. Larger absolute values (positive or negative) suggest a greater difference in decomposition rates between the home and away sites.


Load mass data (Ayres et al.(2009) HFA index)
```{r}
library(stringr)
library(purrr)
mass.df=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/leaf_mass.csv")
mass.df2=read.csv("/Users/Cody1/Desktop/GitHub/RAD Alder/data/leaf_mass2.csv")
###check for any white spaces clean up
columns_with_whitespace <- map_lgl(mass.df2, ~ any(str_detect(.x, "^\\s+|\\s+$")))
columns_with_whitespace

#correction factor adjusting for blotted and fresh dry mass
mass.df= mass.df%>%
  mutate(final_correct = (0.941*(final_mass)-0.00337))

# calculate % mass loss 
mass.df <- mass.df %>%
  mutate(mass_loss = ((initial_mass - final_correct) / initial_mass)*100)

# calculate abs mass loss 
mass.df <- mass.df %>%
  mutate(abs.mass_loss = abs(initial_mass - final_correct))


##create subset of all positive values
mass.df=(mass.df[mass.df$mass_loss > 0,])
mass.df2=(mass.df2[mass.df2$mass_loss > 0,])


#calculate a k constant
mass.df <- mass.df %>%
  mutate(k = -log(1 - mass_loss / 100) / (2 * 19))

```


Subset dataframes to z score by deployment site
```{r}

#P HOK 1 subset
mass.hok1.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.hok1.df=mass.hok1.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#HOK 2 subset
mass.hok2.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.hok2.df=mass.hok2.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#SEK 1 subset
mass.sek1.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.sek1.df=mass.sek1.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#SEK 2 subset
mass.sek2.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.sek2.df=mass.sek2.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

##merge all sites
mass.deploy.merged.df=rbind(mass.hok1.df, mass.hok2.df, mass.sek1.df, mass.sek2.df)
mass.deploy.merged.df$treatment=as.factor(as.character(mass.deploy.merged.df$treatment))
mass.deploy.merged.df$tree_id=as.factor(as.character(mass.deploy.merged.df$tree_id))
mass.deploy.merged.df$tree_id=as.factor(as.character(mass.deploy.merged.df$tree_id))
mass.deploy.merged.df$deploy=as.factor(as.character(mass.deploy.merged.df$deploy))
mass.deploy.merged.df$hfa=as.factor(as.character(mass.deploy.merged.df$hfa))

```

leaf % mass loss with z score by deployment site plot
```{r}
# Combine treatment and hfa into one group
mass.deploy.merged.df$combined_group <- paste(mass.deploy.merged.df$treatment, mass.deploy.merged.df$hfa, sep = "-")
mass.deploy.merged.df$combined_site <- paste("site", mass.deploy.merged.df$site_num, sep = "")

# Create a custom order for the combined groups
combined_group_order <- c("AB-A", "AB-H", "P-A","P-H")

# Plot with combined groups but no combo site
leaf.mass.combo.group.zsite.plot=ggplot(mass.deploy.merged.df, aes(x = combined_group, y = z_score, fill = combined_group)) +
  geom_boxplot(alpha= 0.3) +
  scale_x_discrete(limits = combined_group_order)+
  geom_point(aes(shape = combined_group), position = position_jitter(width = 0.2), alpha = 0.5) +
  labs(x = "treatment", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_fill_manual(values = custom_palette, guide = "none") +
  scale_shape_manual(values = c(21, 22, 23, 24), guide = "none") +
  geom_hline(data = mass.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

# Plot with combined groups but no combo site
leaf.mass.combo.all.zsite.plot=ggplot(mass.deploy.merged.df, aes(x = combined_group, y = z_score, fill = interaction(combined_site, combined_group))) +
  geom_boxplot(alpha= 0.3) +
  scale_x_discrete(limits = combined_group_order)+
  geom_point(aes(shape = interaction(combined_site, combined_group)), position = position_jitter(width = 0.2), alpha = 0.5) +
  labs(x = "treatment", y = "mass loss (z score)") +
  theme(legend.position = "right") +
  scale_fill_manual(values = custom_palette, name = "") +
  scale_shape_manual(values = c(0, 1, 2, 3, 4, 5, 6, 7), name = "") +
  geom_hline(data = mass.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting


#plots
# Define your custom color palette
custom_palette <- c("#1F77B4", "#FF7F0E")


#z score plot
leaf.mass.z.plot=
  ggplot(mass.deploy.merged.df, aes(x = treatment, y = z_score, fill = hfa)) +
  geom_boxplot() +
  geom_point(aes(x = treatment, shape = hfa), position = position_jitter(width = 0.2), alpha = 0.5) +  # Adding jittered data points
  labs(x = "treatment", y = "mass loss (z score)") +
  Fig.formatting +
  theme(legend.position = "right") +
scale_fill_manual(values = custom_palette) + 
  scale_shape_manual(values = c(16, 17)) +  # Defining shapes for data points
  geom_hline(data = mass.merged.df,
             aes(yintercept = mean(z_score),
                 linetype = "Average"),
             color = "red",
             linetype = "dotted") +
  theme(plot.title = element_text(size = 12)) +
  Fig.formatting

ggsave("figures/leaf.mass.combo.all.zsite.plot.pdf", leaf.mass.combo.all.zsite.plot, width = 8, height = 6)

```

Subset dataframes to z score by deployment site and treatment
```{r}
#P HOK 1 subset
mass.p.hok1.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$treatment == "P" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.p.hok1.df=mass.p.hok1.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#AB HOK 1 subset
mass.ab.hok1.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$treatment == "AB" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.ab.hok1.df=mass.ab.hok1.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#P HOK 2 subset
mass.p.hok2.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$treatment == "P" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.p.hok2.df=mass.p.hok2.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#AB HOK 2 subset
mass.ab.hok2.df <- data.frame(mass.df[mass.df$deploy == "HOK" & mass.df$treatment == "AB" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.ab.hok2.df=mass.ab.hok2.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#P SEK 1 subset
mass.p.sek1.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$treatment == "P" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.p.sek1.df=mass.p.sek1.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#AB SEK 1 subset
mass.ab.sek1.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$treatment == "AB" & mass.df$site_num == "1", ])
# Calculate z-scores 
mass.ab.sek1.df=mass.ab.sek1.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#P SEK 2 subset
mass.p.sek2.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$treatment == "P" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.p.sek2.df=mass.p.sek2.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))

#AB SEK 2 subset
mass.ab.sek2.df <- data.frame(mass.df[mass.df$deploy == "SEK" & mass.df$treatment == "AB" & mass.df$site_num == "2", ])
# Calculate z-scores 
mass.ab.sek2.df=mass.ab.sek2.df%>%
  mutate(z_score= (mass_loss - mean(mass_loss))/sd(mass_loss))



##merge all sites
mass.deploy.treat.merged.df=rbind(mass.p.hok1.df, mass.ab.hok1.df, mass.p.hok2.df, mass.ab.hok2.df, mass.p.sek1.df, mass.ab.sek1.df, mass.p.sek2.df, mass.ab.sek2.df)
mass.deploy.treat.merged.df$treatment=as.factor(as.character(mass.deploy.treat.merged.df$treatment))
mass.deploy.treat.merged.df$tree_id=as.factor(as.character(mass.deploy.treat.merged.df$tree_id))
mass.deploy.treat.merged.df$tree_id=as.factor(as.character(mass.deploy.treat.merged.df$tree_id))
mass.deploy.treat.merged.df$deploy=as.factor(as.character(mass.deploy.treat.merged.df$deploy))
mass.deploy.treat.merged.df$hfa=as.factor(as.character(mass.deploy.treat.merged.df$hfa))

```


leaf % mass loss as z score by deployment site AND treatment
```{r}
# Combine treatment and hfa into one group
mass.deploy.treat.merged.df$combined_group <- paste(mass.deploy.treat.merged.df$treatment, mass.deploy.treat.merged.df$hfa, sep = "-")
mass.deploy.treat.merged.df$combined_site <- paste("site", mass.deploy.treat.merged.df$site_num, sep = "")


# Create a custom order for the combined groups
combined_group_order <- c("AB-A", "AB-H", "P-A","P-H")


custom_palette <- c("#1F77B4", "#FF7F0E", "#2CA02C", "#D62728")

custom_palette <- c("#1F77B4", "#FF7F0E", "#2CA02C", "#D62728", "#9467BD", "#8C564B", "#E377C2", "#7F7F7F")


# Plot with combined groups and site 
leaf.mass.combo.site.plot=ggplot(mass.deploy.treat.merged.df, aes(x = combined_group, y = z_score, fill= interaction(combined_site, combined_group))) +
  geom_boxplot(alpha = 0.3) +
  scale_x_discrete(limits = combined_group_order) +
  geom_point(aes(shape = interaction(combined_site, combined_group)), position = position_jitter(width = 0.2), alpha = 0.5) +
  labs(x = "treatment", y = "mass loss (z score)") +
  theme(legend.position = "none",
axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(0,1,2,3,4,5,6,7), name = "") +
   scale_fill_manual(values = custom_palette, name = "")+
  geom_hline(data = mass.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  Fig.formatting
leaf.mass.combo.site.plot

# Plot with combined groups but no combo site 
leaf.mass.combo.group.plot=ggplot(mass.deploy.treat.merged.df, aes(x = combined_group, y = z_score, fill= combined_group)) +
  geom_boxplot(alpha = 0.3) +
  scale_x_discrete(limits = combined_group_order) +
  geom_point(aes(shape = combined_group), position = position_jitter(width = 0.2), alpha = 0.5) +
  labs(x = "treatment", y = "mass loss (z score)") +
  theme(legend.position = "none",
axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(21, 22, 23, 24), name = "") +
   scale_fill_manual(values = custom_palette, name = "")+
  geom_hline(data = mass.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  Fig.formatting

leaf.mass.combo.group.plot

#trying to shade with different shapes
ggplot(mass.deploy.treat.merged.df, aes(x = combined_group, y = z_score, fill = combined_group)) +
  geom_boxplot_pattern(
    aes(pattern = combined_group),
    colour = "black",
    pattern_fill = "black",
    pattern_angle = 45,
    pattern_density = 0.1,
    pattern_spacing = 0.01,
    position = position_dodge2(preserve = 'single'),
    alpha = 0.2  # Adjust the alpha value to control transparency
  ) +
    geom_point(aes(shape = interaction(site_num, combined_group)), position = position_jitter(width = 0.2), alpha = 0.5)+
  geom_boxplot(alpha = 0.3) +
  scale_x_discrete(limits = combined_group_order) +
  labs(x = "Treatment Group", y = "Mass Loss (z score)") +
  theme(legend.position = "bottom",  # Move legend to the bottom
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16, 17, 18, 19, 20, 21, 22, 23)) +
  scale_fill_manual(values = custom_palette) +
  geom_hline(data = mass.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  guides(fill = guide_legend(override.aes = list(shape = c(16, 17, 18, 19, 20, 21, 22, 23)))) +  # Specify the shape order
  Fig.formatting






#basic HFA
ggplot(mass.deploy.treat.merged.df, aes(x = combined_group, y = z_score)) +
  geom_boxplot(alpha = 0.3) +
  scale_x_discrete(limits = combined_group_order) +
  geom_point(aes(shape = interaction(site_num, hfa)), position = position_jitter(width = 0.2), alpha = 0.5) +
  labs(x = "Treatment Group", y = "Mass Loss (z score)") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.title = "Treatment") +
  scale_shape_manual(values = c(16, 17, 18, 19)) +
  geom_hline(data = mass.merged.df, aes(yintercept = mean(z_score), linetype = "Average"), color = "red", linetype = "dotted") +
  Fig.formatting


 leaf.mass.z.plot
ggsave("figures/leaf.mass.combo.group.plot.pdf", leaf.mass.combo.group.plot, width = 8, height = 6)



```


maybe try z scoring by site and using the z score percent mass loss for HFAI?


generate **present**, site a (HOK 1)
```{r}
mass.p.hoko1.df=mass.df%>%
  filter(treatment == "P" & deploy=="HOK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.p.hoko1.df <- mass.p.hoko1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.p.hoko1.df$mass_loss[mass.p.hoko1.df$tree_id == "HOK" & mass.p.hoko1.df$tree_num == tree_num]
    sek_mass_loss <- mass.p.hoko1.df$mass_loss[mass.p.hoko1.df$tree_id == "SEK" & mass.p.hoko1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}

##check the calculations (Armla.1-Armla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss)) * 100


##check the calculations (Brmla.1-Brmla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss)) * 100


# Add the rml values to the dataframe
mass.p.hoko1.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.p.hoko1.df$ayres.tree <- paste(mass.p.hoko1.df$ayres, mass.p.hoko1.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.p.hoko1.df$ayres.eq <- ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **present**, site b (SEK 1)
```{r}
mass.p.sek1.df=mass.df%>%
  filter(treatment == "P" & deploy=="SEK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.p.sek1.df <- mass.p.sek1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)


###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.p.sek1.df$mass_loss[mass.p.sek1.df$tree_id == "HOK" & mass.p.sek1.df$tree_num == tree_num]
    sek_mass_loss <- mass.p.sek1.df$mass_loss[mass.p.sek1.df$tree_id == "SEK" & mass.p.sek1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}

##check the calculations (Armla.1-Armla.5)
((mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.1",]$mass_loss) /
  (mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.1",]$mass_loss +
   mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Ba.1",]$mass_loss)) * 100

((mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.2",]$mass_loss) /
  (mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.2",]$mass_loss +
   mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Ba.2",]$mass_loss)) * 100


((mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.3",]$mass_loss) /
  (mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.3",]$mass_loss +
   mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Ba.3",]$mass_loss)) * 100

((mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.4",]$mass_loss) /
  (mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.4",]$mass_loss +
   mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Ba.4",]$mass_loss)) * 100


((mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.5",]$mass_loss) /
  (mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Aa.5",]$mass_loss +
   mass.p.sek1.df[mass.p.sek1.df$ayres.tree == "Ba.5",]$mass_loss)) * 100


##check the calculations (Brmla.1-Brmla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss)) * 100

# Add the rml values to the dataframe
mass.p.sek1.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.p.sek1.df$ayres.tree <- paste(mass.p.sek1.df$ayres, mass.p.sek1.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.p.sek1.df$ayres.eq <- ifelse(mass.p.sek1.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.p.sek1.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **present**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.p.site1.df=rbind(mass.p.hoko1.df, mass.p.sek1.df)

# Create an empty data frame to store the results
hfai.df <- data.frame(hfai = numeric(0))
# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.p.site1.df[mass.p.site1.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.p.site1.df[mass.p.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.p.site1.df[mass.p.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.p.site1.df[mass.p.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.p.site1.df[mass.p.site1.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = LETTERS[i], treatment = treatment))
}

#HFAI (this is a stop and check to verify that values are correct!)
(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.p.site1.df[mass.p.site1.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.p.site1.df[mass.p.site1.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```

## Now we'll tackle absent:
generate **absent**, site a (HOK 1)
```{r}
mass.ab.hoko1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="HOK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.hoko1.df <- mass.ab.hoko1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "HOK" & mass.ab.hoko1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "SEK" & mass.ab.hoko1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}

##check the calculations (Armla.1-Armla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss)) * 100

((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss)) * 100


##check the calculations (Brmla.1-Brmla.5)
((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.1",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.1",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.2",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.2",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.3",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.3",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.4",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.4",]$mass_loss)) * 100


((mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss) /
  (mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Ba.5",]$mass_loss +
   mass.p.hoko1.df[mass.p.hoko1.df$ayres.tree == "Aa.5",]$mass_loss)) * 100

# Add the rml values to the dataframe
mass.ab.hoko1.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.hoko1.df$ayres.tree <- paste(mass.ab.hoko1.df$ayres, mass.ab.hoko1.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.hoko1.df$ayres.eq <- ifelse(mass.ab.hoko1.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **absent**, site b (SEK 1)
```{r}
mass.ab.sek1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="SEK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.sek1.df <- mass.ab.sek1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)


###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "HOK" & mass.ab.sek1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "SEK" & mass.ab.sek1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.sek1.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.sek1.df$ayres.tree <- paste(mass.ab.sek1.df$ayres, mass.ab.sek1.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.sek1.df$ayres.eq <- ifelse(mass.ab.sek1.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.p.sek1.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **asbent**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.ab.site1.df=rbind(mass.ab.hoko1.df, mass.ab.sek1.df)

# Calculate the last letter index based on the existing data frame
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.ab.site1.df[mass.ab.site1.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = next_label, treatment = treatment))
}


#HFAI (this is a stop and check to verify that values are correct!)
(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.ab.site1.df[mass.ab.site1.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```

plot check
```{r}

ggplot(hfai.df, aes(x = label, y = hfai)) +
  geom_bar(stat = "identity",
           show.legend = FALSE) + # Remove the legend
  facet_grid(~treatment)+
  xlab("Group") +
  ylab("Value")+
  coord_flip()




```

generate **present**, site a (HOK 2)
```{r}
mass.p.hoko2.df=mass.df%>%
  filter(treatment == "P" & deploy=="HOK" & site_num=="2")
#now create a new column to have some reference to the Ayres et al. (2009) paper (species A= hoko trees, species B = sekiu trees)
mass.p.hoko2.df <- mass.p.hoko2.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 2-1 with SEK 2-1; this follows similar logic as Ayres et al. (2009) Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.p.hoko2.df$mass_loss[mass.p.hoko2.df$tree_id == "HOK" & mass.p.hoko2.df$tree_num == tree_num]
    sek_mass_loss <- mass.p.hoko2.df$mass_loss[mass.p.hoko2.df$tree_id == "SEK" & mass.p.hoko2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.p.hoko2.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.p.hoko2.df$ayres.tree <- paste(mass.p.hoko2.df$ayres, mass.p.hoko2.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.p.hoko2.df$ayres.eq <- ifelse(mass.p.hoko2.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **present**, site b (SEK 2)
```{r}
mass.p.sek2.df=mass.df%>%
  filter(treatment == "P" & deploy=="SEK" & site_num=="2")
#now create a new column to have some reference to the Ayres et al. (2009) paper (species A= hoko trees, species B = sekiu trees)
mass.p.sek2.df <- mass.p.sek2.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)


###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 2-1 with SEK 2-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.p.sek2.df$mass_loss[mass.p.sek2.df$tree_id == "HOK" & mass.p.sek2.df$tree_num == tree_num]
    sek_mass_loss <- mass.p.sek2.df$mass_loss[mass.p.sek2.df$tree_id == "SEK" & mass.p.sek2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.p.sek2.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.p.sek2.df$ayres.tree <- paste(mass.p.sek2.df$ayres, mass.p.sek2.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.p.sek2.df$ayres.eq <- ifelse(mass.p.sek2.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.p.sek1.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **present**, site a and b (aka HOK 2 and SEK 2) dataframes 
```{r}
#merge both present site 1 dataframes
mass.p.site2.df=rbind(mass.p.hoko2.df, mass.p.sek2.df)

# Calculate the last letter index based on the existing data frame
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.p.site2.df[mass.p.site2.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.p.site2.df[mass.p.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.p.site2.df[mass.p.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.p.site2.df[mass.p.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.p.site2.df[mass.p.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = next_label, treatment = treatment))
}

#HFAI (this is a stop and check to verify that values are correct!)
(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.p.site2.df[mass.p.site2.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.p.site2.df[mass.p.site2.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```

## Now we'll tackle absent:
generate **absent**, site a (HOK 2)
```{r}
mass.ab.hoko2.df=mass.df%>%
  filter(treatment == "AB" & deploy=="HOK" & site_num=="2")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.hoko2.df <- mass.ab.hoko2.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))
  #since we're at site a, the notation follows species A at a (Aa) or species B at a (Ba)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.hoko2.df$mass_loss[mass.ab.hoko2.df$tree_id == "HOK" & mass.ab.hoko2.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.hoko2.df$mass_loss[mass.ab.hoko2.df$tree_id == "SEK" & mass.ab.hoko2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.hoko2.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.hoko2.df$ayres.tree <- paste(mass.ab.hoko2.df$ayres, mass.ab.hoko2.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.hoko2.df$ayres.eq <- ifelse(mass.ab.hoko2.df$ayres.tree %in% paste("Aa.", tree_num_values, sep = ""), 
                                     paste("Armla", tree_num_values, sep = "."), 
                                     ifelse(mass.p.hoko1.df$ayres.tree %in% paste("Ba.", tree_num_values, sep = ""), 
                                            paste("Brmla", tree_num_values, sep = "."), 
                                            NA))
  #Armla = species A rml at a 
  #Brmla = species B rml at a
```

generate **absent**, site b (SEK 1)
```{r}
mass.ab.sek2.df=mass.df%>%
  filter(treatment == "AB" & deploy=="SEK" & site_num=="2")
#now create a new column to have some reference to the Ayres paper (species A= hoko trees, species B = sekiu trees)
mass.ab.sek2.df <- mass.ab.sek2.df %>%
  mutate(ayres = ifelse(hfa == "H", "Bb", ifelse(hfa == "A", "Ab", NA)))
  #since we're at site b now, the notation follows species A at b (Ab) or species B at b (Bb)

# Define the values you want to loop through
ayres_values <- c("Ab", "Bb")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.sek2.df$mass_loss[mass.ab.sek2.df$tree_id == "HOK" & mass.ab.sek2.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.sek2.df$mass_loss[mass.ab.sek2.df$tree_id == "SEK" & mass.ab.sek2.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Ab", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Bb", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.sek2.df$rml <- rml_values
# Create a new column combining tree_num and ayres with a period in between
mass.ab.sek2.df$ayres.tree <- paste(mass.ab.sek2.df$ayres, mass.ab.sek2.df$tree_num, sep = ".")
# Create a new column based on Ayres HFAI equation 
mass.ab.sek2.df$ayres.eq <- ifelse(mass.ab.sek2.df$ayres.tree %in% paste("Ab.", tree_num_values, sep = ""), 
                                     paste("Armlb", tree_num_values, sep = "."), 
                                     ifelse(mass.p.sek1.df$ayres.tree %in% paste("Bb.", tree_num_values, sep = ""), 
                                            paste("Brmlb", tree_num_values, sep = "."), 
                                            NA))
  #Armlb = species A rml at b
  #Brmlb= species B rml at b
```

merge **asbent**, site a and b (aka HOK 1 and SEK 1) dataframes 
```{r}
#merge both present site 1 dataframes
mass.ab.site2.df=rbind(mass.ab.hoko2.df, mass.ab.sek2.df)

# Calculate the last letter index based on the existing data frame
last_label <- tail(hfai.df$label, 1)
last_letter_index <- match(last_label, LETTERS)

# Loop through each set of equations and calculate the values
for (i in 1:5) {
  armla <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Armla.", i), ]$rml
  brmlb <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Brmlb.", i), ]$rml
  armlb <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Armlb.", i), ]$rml
  brmla <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Brmla.", i), ]$rml
  
  result <- (((armla + brmlb / 2) / (armlb + brmla / 2)) * 100) - 100
  
  # Get the treatment associated with the current index
  treatment <- mass.ab.site2.df[mass.ab.site2.df$ayres.eq == paste0("Armla.", i), ]$treatment
  
  # Determine the next label based on the last label index
  next_label <- LETTERS[last_letter_index + i]
  
  # Append the result, label, and treatment to the data frame
  hfai.df <- rbind(hfai.df, data.frame(hfai = result, label = next_label, treatment = treatment))
}

#HFAI (this is a stop and check to verify that values are correct!)
(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.1",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.1",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.1",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.1",]$rml) / 2)) * 100 - 100
    

(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.2",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.2",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.2",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.2",]$rml) / 2)) * 100 - 100
    
(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.3",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.3",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.3",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.3",]$rml) / 2)) * 100 - 100

(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.4",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.4",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.4",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.4",]$rml) / 2)) * 100 - 100

(((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armla.5",]$rml) + 
  (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmlb.5",]$rml) / 2) /
  ((mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Armlb.5",]$rml) + 
   (mass.ab.site2.df[mass.ab.site2.df$ayres.eq=="Brmla.5",]$rml) / 2)) * 100 - 100
    

```

another plot check
```{r}
ggplot(hfai.df, aes(x = reorder(label, hfai), y = hfai, fill = treatment)) +
  geom_bar(stat = "identity") +
  xlab("") +
  ylab("Home-field advantage (% faster decomposition at home)") +
  coord_flip() +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  )

# Filter out outliers
filtered.hfai.data <- hfai.df %>%
  filter(hfai >= -300, hfai <= 300)

#final plot looks good
  #no color, but shading with slight opacity
hfai.plot=ggplot(filtered.hfai.data, aes(x = reorder(label, hfai), y = hfai)) +
  geom_bar(stat = "identity", fill = "white", color = "black") +
  geom_col_pattern(
    aes(pattern = treatment),
    colour = "black",
    pattern_fill = "black",
    pattern_angle = 45,
    pattern_density = 0.1,
    pattern_spacing = 0.01,
    position = position_dodge2(preserve = 'single'),
    alpha = 0.2  # Adjust the alpha value to control transparency
  ) +
  scale_pattern_manual(
    values = c("none", "stripe"),
    guide = guide_legend(override.aes = list(fill = "grey70")) # <- make lighter
  ) +
  xlab("") +
  ylab("Home-field advantage (% faster decomposition at home)") +
coord_flip() +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank()
  )

hfai.plot
ggsave("figures/hfai.plot.pdf", hfai.plot, width = 26, height = 8)

```


















```{r}
########### Next we'll try out absent, hoko 1
mass.ab.hoko1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="HOK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (Home= Aa, Away= Bb)
mass.ab.hoko1.df <- mass.ab.hoko1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "HOK" & mass.ab.hoko1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.hoko1.df$mass_loss[mass.ab.hoko1.df$tree_id == "SEK" & mass.ab.hoko1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.hoko1.df$rml <- rml_values







########### Next we'll try out absent, sekiu 1
mass.ab.sek1.df=mass.df%>%
  filter(treatment == "AB" & deploy=="SEK" & site_num=="1")
#now create a new column to have some reference to the Ayres paper (Home= Aa, Away= Bb)
mass.ab.sek1.df <- mass.ab.sek1.df %>%
  mutate(ayres = ifelse(hfa == "H", "Aa", ifelse(hfa == "A", "Ba", NA)))

###tree numbers are used for pairwise relative mass loss (RML) calculation (e.g., HOK 1-1 with SEK 1-1; this follows similar logic as Ayres Species A to B comparison)

# Define the values you want to loop through
ayres_values <- c("Aa", "Ba")
tree_num_values <- c("1", "2", "3", "4", "5")

# Initialize an empty vector to store the results
rml_values <- numeric()

# Loop through the combinations and calculate rml values
for (ayres in ayres_values) {
  for (tree_num in tree_num_values) {
    hok_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "HOK" & mass.ab.sek1.df$tree_num == tree_num]
    sek_mass_loss <- mass.ab.sek1.df$mass_loss[mass.ab.sek1.df$tree_id == "SEK" & mass.ab.sek1.df$tree_num == tree_num]
    
    rml <- ifelse(ayres == "Aa", hok_mass_loss / (hok_mass_loss + sek_mass_loss),
                  ifelse(ayres == "Ba", sek_mass_loss / (sek_mass_loss + hok_mass_loss),
                         NA))
    
    rml_values <- c(rml_values, rml * 100)
  }
}
# Add the rml values to the dataframe
mass.ab.sek1.df$rml <- rml_values


#merge dataframes
mass.site1.df=rbind(mass.p.hoko1.df, mass.ab.hoko1.df, mass.p.sek1.df, mass.ab.sek1.df )



```




Ayres HFA index
```{r}
#the calculations for arml and brml are different, dependent on "relative to"

##Site 1: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site a (Hoko) 
site1.armla.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="1" & mass.df$deploy == "HOK", ])
#mass loss calculation
site1.armla.df= site1.armla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.armla.df <- site1.armla.df %>%
  mutate(rml = ((mass_loss[tree_id == "HOK" & deploy == "HOK"]) / 
                 (mass_loss[tree_id == "HOK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "HOK"])) * 100)

##Site 2: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site a (Hoko) 
site2.armla.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="2" & mass.df$deploy == "HOK", ])
#mass loss calculation
site2.armla.df= site2.armla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.armla.df <- site2.armla.df %>%
  mutate(rml = (mass_loss[tree_id == "HOK" & deploy == "HOK"] / 
                 (mass_loss[tree_id == "HOK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "HOK"])) * 100)

##Site 1: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site b (Sekiu) 
site1.armlb.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="2" & mass.df$deploy == "SEK", ])
#mass loss calculation
site1.armlb.df= site1.armlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.armlb.df <- site1.armlb.df %>%
  mutate(rml = (mass_loss[tree_id == "HOK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "HOK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "SEK"])) * 100)

##Site 2: Species A (Hoko genotypes) relative mass loss (RML) to species B (Sekiu genotypes) at site b (Sekiu) 
site2.armlb.df <- data.frame(mass.df[(mass.df$tree_id == "HOK" | mass.df$tree_id == "SEK") & mass.df$site_num=="2" & mass.df$deploy == "SEK", ])
#mass loss calculation
site2.armlb.df= site2.armlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.armlb.df <- site2.armlb.df %>%
  mutate(rml = (mass_loss[tree_id == "HOK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "HOK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "SEK" & deploy == "SEK"])) * 100)

##Site 1: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site a (Hoko) 
site1.brmla.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="1" & mass.df$deploy == "HOK", ])
#mass loss calculation
site1.brmla.df= site1.brmla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.brmla.df <- site1.brmla.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "HOK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "HOK"])) * 100)

##Site 2: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site a (Hoko) 
site2.brmla.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="2" & mass.df$deploy == "HOK", ])
#mass loss calculation
site2.brmla.df= site2.brmla.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.brmla.df <- site2.brmla.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "HOK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "HOK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "HOK"])) * 100)

##Site 1: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site b (Sekiu) 
site1.brmlb.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="1" & mass.df$deploy == "SEK", ])
#mass loss calculation
site1.brmlb.df= site1.brmlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site1.brmlb.df <- site1.brmlb.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "SEK"])) * 100)

##Site 2: Species B (Sekiu genotypes) relative mass loss (RML) to species A (Hoko genotypes) at site b (Sekiu) 
site2.brmlb.df <- data.frame(mass.df[(mass.df$tree_id == "SEK" | mass.df$tree_id == "HOK") & mass.df$site_num=="2" & mass.df$deploy == "SEK", ])
#mass loss calculation
site2.brmlb.df= site2.brmlb.df%>%
    mutate(mass_loss = ((initial_mass- final_correct) / initial_mass) * 100)
#RML calculation
site2.brmlb.df <- site2.brmlb.df %>%
  mutate(rml = (mass_loss[tree_id == "SEK" & deploy == "SEK"] / 
                 (mass_loss[tree_id == "SEK" & deploy == "SEK"] + 
                  mass_loss[tree_id == "HOK" & deploy == "SEK"])) * 100)

#merge dataframes
site1.hfa_index.df=rbind(site1.armla.df, site1.armlb.df, site1.brmla.df, site1.brmlb.df)
#HFA index calculation
site1.hfa_index.df <- site1.hfa_index.df %>%
  mutate(hfai = (((site1.armla.df$rml + site1.brmlb.df$rml) / 2) / ((site1.armlb.df$rml + site1.brmla.df$rml) / 2)) * 100 - 100)

#merge dataframes
site2.hfa_index.df=rbind(site2.armla.df, site2.armlb.df, site2.brmla.df, site2.brmlb.df)
#HFA index calculation
site2.hfa_index.df <- site2.hfa_index.df %>%
  mutate(hfai = (((site2.armla.df$rml + site2.brmlb.df$rml) / 2) / ((site2.armlb.df$rml + site2.brmla.df$rml) / 2)) * 100 - 100)

# Create the horizontal bar plot
ggplot(hfa_index.df, aes(x = hfai, y = interaction(deploy, hfa))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  labs(title = "HFA Index Comparison",
       x = "HFA Index") +
  theme_minimal() +
  theme(legend.position = "none") +
  geom_vline(xintercept = 0, color = "black", size = 1) +
  scale_fill_manual(values = c("a" = "blue", "b" = "red")) 



```







